### 2018-02-09 scipy-1.0/paper.tex
Don't use Lipsum in the paper
Ubuntu LaTeX base doesn't have it.

\documentclass[11pt]{article}

\begin{document}

One day this will be a wonderful paper.

\end{document}

### 2018-03-05 scipy-1.0/paper.tex
Add document structure and Scientific Reports template.

@@
-\documentclass[11pt]{article}
+\documentclass[fleqn,10pt]{wlscirep}
+\title{SciPy - some descriptive title here}
 
+\author[1]{Pauli Virtanen}
+\author[2,*]{Ralf Gommers}
+\author[3]{TODO}
+\affil[1]{Affiliation, department, city, postcode, country}
+\affil[2]{Affiliation, department, city, postcode, country}
+\affil[2]{Affiliation, department, city, postcode, country}
+
+\affil[*]{ralf.gommers@gmail.com}
+
+\keywords{Scientific computing, Python, Mathematics}
+
+\begin{abstract}
+TODO. Abstract must be under 200 words and not include subheadings or citations.
+\end{abstract}
 \begin{document}
 
-One day this will be a wonderful paper.
+\flushbottom
+\maketitle
+\thispagestyle{empty}
+
+\section*{Introduction}
+
+\textit{The Introduction section expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.}
+
+\textbf{History}
+
+\textbf{Project goals and scope}
+
+\textbf{Current status (maturity, users)}
+
+\section*{Architecture and implementation choices}
+
+\subsection*{Submodule organisation}
+
+\subsection*{Common infrastructure}
+
+\subsection*{Language preferences}
+
+\subsection*{API and ABI evolution}
+
+\section*{Key technical improvements}
+
+Here we describe key technical improvements made in the last three years.
+
+\subsection*{Data structures}
+\textbf{cKDTree}
+\textbf{Sparse matrices}
+
+\subsection*{Unified bindings to compiled code}
+LowLevelCallable
+
+\subsection*{Cython bindings for BLAS, LAPACK and special}
+
+\subsection*{Numerical optimization}
+
+\subsection*{Statistical distributions}
+
+\subsection*{Polynomial interpolators}
+
+\subsection*{Test and benchmark suite}
+
+\section*{Project organisation and community}
+
+\textbf{Governance}
+
+\textbf{Roadmap}
+
+\textbf{Community beyond the SciPy library}
+
+\textbf{Maintainers and contributors}
+
+\section*{Discussion}
+
+\textit{The Discussion should be succinct and must not contain subheadings.}
+
+\textbf{Impact now}
+
+\textbf{Future development}.
+\textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}
+
+\bibliography{references}
+
+\noindent Use the cite command for an inline citation, e.g. \cite{behnel2011cython}.
+
+\section*{Acknowledgements (not compulsory)}
+
+Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
+
+\section*{Author contributions statement}
+
+Must include all authors, identified by initials, for example:
+A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
+
+\section*{Additional information}
+
+To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing financial interests} (mandatory statement).
+
+The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing financial interests statement} on behalf of all authors of the paper.
+This statement must be included in the submitted article file.
 

### 2018-03-27 scipy-1.0/scipy-optimize.tex
ENH: added some introductory material for scipy.optimize section and an introduction to the root finding problem

\newcommand{\RR}{\ensuremath{\mathbb{R}}}
The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
\begin{enumerate}
\item root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
\item general, linear optimization problems (\texttt{linprog}),
\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
\item general, global optimization problems (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
\end{enumerate}
The following subsections summarize SciPy's functionality in these areas, highlighting the recent additions in SciPy 1.0.

\subsubsection{Root Finding}
The general ``root finding'' problem is to find a root $\mathbf{x} \in \RR^m$ of $\mathbf{f}: \RR^m \rightarrow \RR^m$, that is, to solve
\begin{equation}
\mathbf{f}(\mathbf{x}) = \mathbf{0}
\end{equation}
for a solution $\mathbf{x}$.\footnote{Equivalently the problem is to simultaneously find the roots $x_i \in \RR$ of several scalar functions $f_i : \RR \rightarrow \RR$, that is, to solve $f_i(x_0, x_1, \dots, x_{m-1}) = 0$ for $x_i$, $i \in \{0, 1, \dots {m-1}\}$.} The function \texttt{scipy.optimize.root} provides a common interface to several algorithms for solving problems of this type. For the special case\footnote{that is, to solve a single scalar equation $f(x) = 0$ for a single scalar variable $x$} $m = 1$, any one of several specialized functions \texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, or \texttt{newton} may provide improved performance or accuracy. (Have there been any recent improvements? Do we want to summarize the methods as @antonior92 has done for $minimize$? Is this a good template for the beginning of the following subsections?)

\subsubsection{Linear Sum Assignment}

\subsubsection{Least Squares}

\subsubsection{Linear Optimization}

\subsubsection{Scalar, Nonlinear Optimization}

\subsubsection{Multivariable, Nonlinear, Local Optimization}

\subsubsection{Global Optimization}

### 2018-03-27 scipy-1.0/scipy-optimize.tex
ENH: added additional questions to scipy optimize Root Finding subsubsection

@@
 \begin{enumerate}
-\item root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
+\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
 \item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
@@
 Do we want to summarize the methods as @antonior92 has done for $minimize$?
+Do we have to explain that these methods only provide \emph{one} solution, and that they are iterative based on a user-provided guess?
+Do we have to explain the notion of tolerance?
 Is this a good template for the beginning of the following subsections?)

### 2018-03-31 scipy-1.0/scipy-optimize.tex
ENH: removed most subsubsections from scipy.optimize subsection; added text about linprog

@@
 \end{enumerate}
-The following subsections summarize SciPy's functionality in these areas, highlighting the recent additions in SciPy 1.0.
+Documentation of SciPy's functionality in each these areas can be found in (cite SciPy documentation), but here we highlight recent additions through SciPy 1.0.
 
-\subsubsection{Root Finding}
-The general ``root finding'' problem is to find a root $\mathbf{x} \in \RR^m$ of $\mathbf{f}: \RR^m \rightarrow \RR^m$, that is, to solve
-\begin{equation}
-\mathbf{f}(\mathbf{x}) = \mathbf{0}
-\end{equation}
-for a solution $\mathbf{x}$.\footnote{Equivalently the problem is to simultaneously find the roots $x_i \in \RR$ of several scalar functions $f_i : \RR \rightarrow \RR$, that is, to solve $f_i(x_0, x_1, \dots, x_{m-1}) = 0$ for $x_i$, $i \in \{0, 1, \dots {m-1}\}$.} The function \texttt{scipy.optimize.root} provides a common interface to several algorithms for solving problems of this type.
+%\subsubsection{Root Finding}
+%The general ``root finding'' problem is to find a root $\mathbf{x} \in \RR^m$ of $\mathbf{f}: \RR^m \rightarrow \RR^m$, that is, to solve
+%\begin{equation}
+%\mathbf{f}(\mathbf{x}) = \mathbf{0}
+%\end{equation}
+%for a solution $\mathbf{x}$.\footnote{Equivalently the problem is to simultaneously find the roots $x_i \in \RR$ of several scalar functions $f_i : \RR \rightarrow \RR$, that is, to solve $f_i(x_0, x_1, \dots, x_{m-1}) = 0$ for $x_i$, $i \in \{0, 1, \dots {m-1}\}$.} The function \texttt{scipy.optimize.root} provides a common interface to several algorithms for solving problems of this type.
 For the special case\footnote{that is, to solve a single scalar equation $f(x) = 0$ for a single scalar variable $x$} $m = 1$, any one of several specialized functions \texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, or \texttt{newton} may provide improved performance or accuracy.
@@
 
-\subsubsection{Linear Sum Assignment}
-
-\subsubsection{Least Squares}
-
 \subsubsection{Linear Optimization}
 
-\subsubsection{Scalar, Nonlinear Optimization}
+A new interior-point optimizer for continuous linear programming problems, \texttt{linprog} with \texttt{method='interior-point'}, was released with SciPy 1.0.
+Implementing the core algorithm of the commercial solver MOSEK \cite{andersen2000mosek}, it solves all of the 90+ NETLIB LP benchmark problems \cite{netlib} tested.
+Unlike some interior point methods, this homogenous self-dual formulation provides certificates of infeasibility or unboundedness as appropriate.
+
+A presolve routine based on \cite{andersen1995presolving} solves trivial problems and otherwise performs problem simplifications, such as bound tightening and removal of fixed variables, and one of several routines for eliminating redundant equality constraints is automatically chosen to reduce the chance of numerical difficulties caused by singular matrices.
+Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers --- often for the same system with multiple right hand sides due to the predictor-corrector approach --- provide speed sufficient for problems with tens of thousands of variables and constraints.
+
+Compared to the previously implemented simplex method, the new interior-point method is faster for all but the smallest problems, and is suitable for solving medium- and large-sized problems on which the existing simplex implementation fails.
+However, the interior point method typically returns a solution near the center of an optimal face, yet basic solutions are often preferred for sensitivity analysis and for use in mixed integer programming algorithms.
+This motivates the need for a crossover routine or a new implementation of the simplex method for sparse problems in a future release, either of which would require an improved sparse linear system solver with efficient support for rank-one updates
 
@@
 
-\subsubsection{Global Optimization}
+\subsubsection{Other Improvements}
+
+\paragraph{Nonlinear Root Finding}
+Mention df-sane?
+
+\paragraph{Linear Sum Assignment}
+(Should this be included?)
+
+\paragraph{Curve Fitting}
+As covariant errors are commonplace in scientific data used to fit parametric models, the \texttt{curve\_fit} routine in SciPy 1.0 now accepts a covariance matrix of errors in the data to be fitted.
+
+\paragraph{Global Optimization}

### 2018-03-31 scipy-1.0/scipy-optimize.tex

@@
 
-\subsubsection{Linear Optimization}
+\subsubsection*{Linear Optimization}
 
@@
 
-\subsubsection{Multivariable, Nonlinear, Local Optimization}
+\subsubsection*{Nonlinear Optimization}
 
-\subsubsection{Other Improvements}
+The \texttt{minimize} function provides methods for solving general nonlinear optimization
+problems. It unifies several methods with a common interface in order to facilitate
+interchanging between solvers with different characteristics.  Table~\ref{tab:minimize} summarizes
+all the methods available for solving minimization problems of the type:
+\begin{equation}
+  \label{eq:minimization-prob}
+  \min_x f(x)
+\end{equation}
+where $f$ is a multivariate function $f: \mathbb{R}^n \rightarrow \mathbb{R}$ and $x \in \mathbb{R}^n$.
+\texttt{COBYLA} and \texttt{SLSQP} methods allow nonlinear constraints to be included in the optimization
+problems, \texttt{L-BFGS-B} and \texttt{TNC} allow box constraints $x^l \le x \le x^u$,
+and the remaining methods are unconstrained optimization methods.
 
-\paragraph{Nonlinear Root Finding}
-Mention df-sane?
+\begin{table}[]
+  \centering
+  \caption{Optimization methods from \texttt{minimize}.  The field \textit{version added} gives when the algorithm
+    was included in SciPy. Algorithm with \textit{version added} equals to 0.6* were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether the first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires the second-order derivatives from
+    the user, it computes an approximation internally and uses it to accelerate the method convergence.
+    \textit{Iterative Hessian factorization} indicates algorithms that factorize the Hessian in an iterative way,
+    which does not require the explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution. The convergence rates mentioned at the table
+    are: linear (L), superlinear (S) and quadratic (Q). When the convergence rate is denoted by S$^*$ the algorithm
+    has a superlinear rate for the parameters used in SciPy, but could  achieve a quadratic convergence rate for other parameter choices.
+    The \textit{global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$). The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
+    \hline
+    Equality constraints &&&&&&&\cmark&&&&& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize}
+\end{table}
 
-\paragraph{Linear Sum Assignment}
-(Should this be included?)
+The available methods implement a variety of strategies for solving optimization problems.
+Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
+noisy objective functions. When the derivatives are available, however, other methods offer faster convergence rates
+and have the convergence to a local solution proved.
 
-\paragraph{Curve Fitting}
-As covariant errors are commonplace in scientific data used to fit parametric models, the \texttt{curve\_fit} routine in SciPy 1.0 now accepts a covariance matrix of errors in the data to be fitted.
+Available derivative-based methods start with an initial guess and refine it iteratively, seeking
+a local solution to the optimization problem. There are algorithms implementing two competing strategies:
+line-search and trust-region. The so-called line-search (LS) methods at each iteration choose a direction
+and find, approximately, the optimal point in the given direction. Trust-region (TR) methods, on the other
+hand, choose a region for which the local approximation is valid and find, approximately, the  best point
+inside this \textit{trust-region}.
 
-\paragraph{Global Optimization}
+An example of trust-region method introduced in the latest releases is \texttt{trust-exact},
+which a  provides a fast convergence rate by solving  almost exactly the trust-region subproblem.
+Its downside is it requires the Hessian to be stored and factored every iteration, which may preclude
+the solution of very large problems.
+
+Some of the available optimization methods are appropriate for large-scale problems
+(e.g.~problems with more than 1000 variables).
+\texttt{CG} does not use second derivatives and is adequate for large-scale problems due to the low storage requirements.
+\texttt{L-BFGS-B} maintain a simple and compact representation of the Hessian and is also adequate for large-scale problems.
+Finally, \texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use the second order derivatives to achieve faster
+convergence but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way,
+which does not require the Hessian to be stored.
+
+The method \texttt{trust-krylov}, introduced in the latest release, is as a good compromise between
+\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
+subproblem than \texttt{trust-ncg} but does not require the Hessian to be stored and factored as in \texttt{trust-exact}.

### 2018-04-02 scipy-1.0/paper.tex
Add stub / brainstorms for the cKDTree module.
ENH: First draft of benchmark suite section.

@@
 \documentclass[fleqn,10pt]{wlscirep}
-\title{SciPy - some descriptive title here}
+
+% Packages
+\usepackage[super]{nth}
+\usepackage{rotating}
+\usepackage{makecell}
+\usepackage{pifont}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+
+% New Commands
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+
+\title {SciPy - some descriptive title here}
 
@@
 \subsection*{Numerical optimization}
+\input{scipy-optimize}
 
@@
 \subsection*{Test and benchmark suite}
+
+    \subsubsection*{Benchmark suite}
+    The airspeed velocity (asv) library enables benchmarking Python packages over their lifetimes, and the performance of the SciPy
+    code base was monitored with asv starting in February of 2015 (PR \#4501). In addition to ensuring that unit tests are passing (see above),
+    confirming that performance generally remains constant or improves over the commit hash history of the project allows us to objectively
+    measure that our code base is improving, to empower scientific applications.
+
+    Consider the following asv benchmark results, spanning roughly nine years of project history, that demonstrate the gradual performance
+    improvements in \texttt{scipy.spatial.cKDTree.query()} (nearest-neighbor search) performance, using command:
+
+    \texttt{python run.py run -e -s 800 --bench "\textbackslash btime\_query\textbackslash b" "02de46a546..b3ddb2c"}
+
+    \begin{figure}[!ht]
+        \centering
+        \includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+        \caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query() over a roughly nine year commit history time frame.
+The results are based on Python 2.7 performance on the master branch of the project using numpy 1.8.2 and Cython versions 0.27.3 and 0.21.1 (for improved backward compatibility).
+Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility / sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
+    \end{figure}
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2018-04-09 scipy-1.0/scipy-optimize.tex
MAINT: Pinned the location of Table 1 to prevent interjection into asv subsection.
MAINT: Formatting fixes such that optimize and asv sections are not interleaved with other sections.
MAINT: global optimizer

@@
 \item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
-\item general, global optimization problems (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
+\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
 \end{enumerate}
@@
 interchanging between solvers with different characteristics.  Table~\ref{tab:minimize} summarizes
-all the methods available for solving minimization problems of the type:
+the methods available for solving minimization problems of the type:
 \begin{equation}
@@
 
-\begin{table}[]
+The available methods implement a variety of strategies for solving optimization problems.
+Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
+noisy objective functions. When the derivatives are available, however, other methods offer faster convergence rates
+and have the convergence to a local solution proved.
+
+Available derivative-based methods start with an initial guess and refine it iteratively, seeking
+a local solution to the optimization problem. There are algorithms implementing two competing strategies:
+line-search and trust-region. The so-called line-search (LS) methods at each iteration choose a direction
+and find, approximately, the optimal point in the given direction. Trust-region (TR) methods, on the other
+hand, choose a region for which the local approximation is valid and find, approximately, the  best point
+inside this \textit{trust-region}.
+
+An example of trust-region method introduced in the latest releases is \texttt{trust-exact},
+which a  provides a fast convergence rate by solving  almost exactly the trust-region subproblem.
+Its downside is it requires the Hessian to be stored and factored every iteration, which may preclude
+the solution of very large problems.
+
+Some of the available optimization methods are appropriate for large-scale problems
+(e.g.~problems with more than 1000 variables).
+\texttt{CG} does not use second derivatives and is adequate for large-scale problems due to the low storage requirements.
+\texttt{L-BFGS-B} maintain a simple and compact representation of the Hessian and is also adequate for large-scale problems.
+Finally, \texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use the second order derivatives to achieve faster
+convergence but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way,
+which does not require the Hessian to be stored.
+
+The method \texttt{trust-krylov}, introduced in the latest release, is a good compromise between
+\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
+subproblem than \texttt{trust-ncg} but does not require the Hessian to be stored and factored as in \texttt{trust-exact}.
+
+\begin{table}[H]
   \centering
@@
 
-The available methods implement a variety of strategies for solving optimization problems.
-Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
-noisy objective functions. When the derivatives are available, however, other methods offer faster convergence rates
-and have the convergence to a local solution proved.
-
-Available derivative-based methods start with an initial guess and refine it iteratively, seeking
-a local solution to the optimization problem. There are algorithms implementing two competing strategies:
-line-search and trust-region. The so-called line-search (LS) methods at each iteration choose a direction
-and find, approximately, the optimal point in the given direction. Trust-region (TR) methods, on the other
-hand, choose a region for which the local approximation is valid and find, approximately, the  best point
-inside this \textit{trust-region}.
-
-An example of trust-region method introduced in the latest releases is \texttt{trust-exact},
-which a  provides a fast convergence rate by solving  almost exactly the trust-region subproblem.
-Its downside is it requires the Hessian to be stored and factored every iteration, which may preclude
-the solution of very large problems.
-
-Some of the available optimization methods are appropriate for large-scale problems
-(e.g.~problems with more than 1000 variables).
-\texttt{CG} does not use second derivatives and is adequate for large-scale problems due to the low storage requirements.
-\texttt{L-BFGS-B} maintain a simple and compact representation of the Hessian and is also adequate for large-scale problems.
-Finally, \texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use the second order derivatives to achieve faster
-convergence but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way,
-which does not require the Hessian to be stored.
-
-The method \texttt{trust-krylov}, introduced in the latest release, is as a good compromise between
-\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
-subproblem than \texttt{trust-ncg} but does not require the Hessian to be stored and factored as in \texttt{trust-exact}.
+% Some problems may require the use of global optimizers to find the minimum of the objective function over the set of all parameter values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+% \texttt{Basinhopping} \cite{Wales1997} uses random perturbation of the parameter vector, with local minimization.
+The acceptance of the new coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
+% \texttt{differential\textunderscore evolution} uses

### 2018-04-10 scipy-1.0/scipy-optimize.tex
MAINT: complete global optimization blurn

@@
 
-% Some problems may require the use of global optimizers to find the minimum of the objective function over the set of all parameter values, especially if the objective function possesses many local minima that other solvers could get stuck in.
-% \texttt{Basinhopping} \cite{Wales1997} uses random perturbation of the parameter vector, with local minimization.
+Some problems may require the use of global optimizers to find the minimum of the objective function over the set of all parameter values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+\texttt{Brute} performs a grid search over the entire parameter range, with the best grid point undergoing further polishing by another solver.
+However, this approach becomes increasingly inefficient as the problem dimensionality increases, or if the objective function has many extrema.
+\texttt{Basinhopping} \cite{Wales1997} uses random perturbation of the parameter vector, with local minimization.
 The acceptance of the new coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
-% \texttt{differential\textunderscore evolution} uses
+\texttt{differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic solver that improves by gradual evolution of a population of candidate solutions.
+In each iteration trial candidates are generated by combination of candidates from the existing population; if the trial candidates represent an improvement then the population is updated.
+\texttt{differential\textunderscore evolution} does not require that the objective function be differentiable.
+%  basinhopping may need differentiable objective for the local minimization.
+The SciPy benchmark suite possesses a comprehensive range of global optimization problems (196), of varying difficulties, that can be used to evaluate the performance of a particular solver.
+This suite is used to decide whether new solvers have good enough performance to be added to the package.

### 2018-04-11 scipy-1.0/scipy-optimize.tex
ENH: optimization improvements: attempted to better coordinate treatment of linear and nonlinear optimization

--- 2018-03-31 scipy-1.0/scipy-optimize.tex
@@
 
-A presolve routine based on \cite{andersen1995presolving} solves trivial problems and otherwise performs problem simplifications, such as bound tightening and removal of fixed variables, and one of several routines for eliminating redundant equality constraints is automatically chosen to reduce the chance of numerical difficulties caused by singular matrices.
+A presolve routine \cite{andersen1995presolving} solves trivial problems and otherwise performs problem simplifications, such as bound tightening and removal of fixed variables, and one of several routines for eliminating redundant equality constraints is automatically chosen to reduce the chance of numerical difficulties caused by singular matrices.
 Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers --- often for the same system with multiple right hand sides due to the predictor-corrector approach --- provide speed sufficient for problems with tens of thousands of variables and constraints.
@@
 
-The \texttt{minimize} function provides methods for solving general nonlinear optimization
+The \texttt{minimize} function provides methods for solving nonlinear optimization
 problems. It unifies several methods with a common interface in order to facilitate
-interchanging between solvers with different characteristics.  Table~\ref{tab:minimize} summarizes
-all the methods available for solving minimization problems of the type:
-\begin{equation}
-  \label{eq:minimization-prob}
-  \min_x f(x)
-\end{equation}
-where $f$ is a multivariate function $f: \mathbb{R}^n \rightarrow \mathbb{R}$ and $x \in \mathbb{R}^n$.
-\texttt{COBYLA} and \texttt{SLSQP} methods allow nonlinear constraints to be included in the optimization
-problems, \texttt{L-BFGS-B} and \texttt{TNC} allow box constraints $x^l \le x \le x^u$,
-and the remaining methods are unconstrained optimization methods.
+interchanging between solvers with different characteristics, summarized in Table~\ref{tab:minimize}.
+%% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).
+The reader can easily find the standard of meaning of `nonlinear programming', so I believe it should be removed.
+If we decide that it is appropriate to include, then we should add the entire problem statement - including constraints - and we should also include the entire linear programming problem statement in the previous section.
+% summarizes all the methods available for solving minimization problems of the type:
+%\begin{equation}
+%  \label{eq:minimization-prob}
+%  \min_x f(x)
+%\end{equation}
+%where $f$ is a multivariate function $f: \mathbb{R}^n \rightarrow \mathbb{R}$ and $x \in \mathbb{R}^n$.
+%%  Without the problem statement this sentence isn't really appropriate.
+%\texttt{COBYLA} and \texttt{SLSQP} methods support nonlinear constraints, \texttt{L-BFGS-B} and \texttt{TNC} allow box constraints $x^l \le x \le x^u$,
+%and the remaining methods are for unconstrained problems.
 
-\begin{table}[]
+% The available methods implement a variety of solution strategies.
+Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
+noisy objective functions. When derivatives are available, however, other methods offer faster convergence rates
+and have proven convergence properties. 
+The derivative-based methods start with an initial guess and refine it iteratively, seeking
+a local solution to the optimization problem. Each of these algorithms implements one of two complementary strategies:
+line-search or trust-region. At each iteration, the line-search (LS) methods choose a direction
+and approximate the optimal point in the given direction. Trust-region (TR) methods, on the other
+hand, choose a ``trust-region'' region for which a local model is valid and approximate the best point
+inside.
+
+One recently-added trust-region method is \texttt{trust-exact},
+which achieves fast convergence by solving the trust-region subproblem almost exactly.
+Unfortunately, it requires the Hessian to be stored and factored every iteration, which may preclude
+the solution of large problems (e.g.~problems with more than 1000 variables).
+Other methods are appropriate for large-scale problems, including
+\texttt{CG}, which does not use second derivatives and thus has low memory requirements, and \texttt{L-BFGS-B}, which maintains a simple and compact representation of the Hessian.
+\texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use second order derivatives to achieve faster
+convergence, but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way that does not require the Hessian to be stored.
+The method \texttt{trust-krylov}, introduced in the latest release, is a good compromise between
+\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
+subproblem than \texttt{trust-ncg}, but it does not require the Hessian to be stored and factored as in \texttt{trust-exact}.
+
+While most \texttt{minimize} development prior to SciPy 1.0 has focused on unconstrained minimization, methods \texttt{SLSQP} and \texttt{COBYLA}\footnote{We note that any equality constraint $c_{\text{eq}}(x) = 0$ can be expressed as two inequality constraints $c_{\text{eq}}(x) \leq 0$ and $-c_{\text{eq}}(x) \leq 0$} support fully general nonlinear programming.
+However, neither provides special treatment for sparsity.
+Accordingly, ongoing development efforts for \texttt{minimize} are focused on the addition of nonlinear programming algorithms for large, sparse problems.
+
+\begin{table}[H]
   \centering
-  \caption{Optimization methods from \texttt{minimize}.  The field \textit{version added} gives when the algorithm
-    was included in SciPy. Algorithm with \textit{version added} equals to 0.6* were added in version 0.6 or before.
+  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
     The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
     (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
-    indicates whether the first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
-    with $\sim$ the algorithm does not requires the second-order derivatives from
-    the user, it computes an approximation internally and uses it to accelerate the method convergence.
-    \textit{Iterative Hessian factorization} indicates algorithms that factorize the Hessian in an iterative way,
-    which does not require the explicit matrix factorization or storage of the Hessian.
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
     \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
-    iterate is sufficiently close to the solution. The convergence rates mentioned at the table
-    are: linear (L), superlinear (S) and quadratic (Q). When the convergence rate is denoted by S$^*$ the algorithm
-    has a superlinear rate for the parameters used in SciPy, but could  achieve a quadratic convergence rate for other parameter choices.
-    The \textit{global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
     point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$). The table also indicates which algorithms
@@
 \end{table}
-
-The available methods implement a variety of strategies for solving optimization problems.
-Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
-noisy objective functions. When the derivatives are available, however, other methods offer faster convergence rates
-and have the convergence to a local solution proved.
-
-Available derivative-based methods start with an initial guess and refine it iteratively, seeking
-a local solution to the optimization problem. There are algorithms implementing two competing strategies:
-line-search and trust-region. The so-called line-search (LS) methods at each iteration choose a direction
-and find, approximately, the optimal point in the given direction. Trust-region (TR) methods, on the other
-hand, choose a region for which the local approximation is valid and find, approximately, the  best point
-inside this \textit{trust-region}.
-
-An example of trust-region method introduced in the latest releases is \texttt{trust-exact},
-which a  provides a fast convergence rate by solving  almost exactly the trust-region subproblem.
-Its downside is it requires the Hessian to be stored and factored every iteration, which may preclude
-the solution of very large problems.
-
-Some of the available optimization methods are appropriate for large-scale problems
-(e.g.~problems with more than 1000 variables).
-\texttt{CG} does not use second derivatives and is adequate for large-scale problems due to the low storage requirements.
-\texttt{L-BFGS-B} maintain a simple and compact representation of the Hessian and is also adequate for large-scale problems.
-Finally, \texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use the second order derivatives to achieve faster
-convergence but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way,
-which does not require the Hessian to be stored.
-
-The method \texttt{trust-krylov}, introduced in the latest release, is as a good compromise between
-\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
-subproblem than \texttt{trust-ncg} but does not require the Hessian to be stored and factored as in \texttt{trust-exact}.

### 2018-04-11 scipy-1.0/subpackages.tex

The SciPy library is organized as a collection of subpackages.  Here we
summarize the scope and capabilities of each subpackage.

\begin{description}

\item[\texttt{cluster}] ~ \newline
    The \texttt{cluster} subpackage contains two submodules:
    \texttt{cluster.vq} provides vector quantization and k-means algorithms,
    and \texttt{cluster.hierarchy} provides functions for hierarchical and
    agglomerative clustering.
\item[\texttt{constants}] ~ \newline
    Physical and mathematical constants, including the CODATA recommended
    values of the fundamental physical constants\cite{CODATA2014}.
\item[\texttt{fftpack}] ~ \newline
    Fast Fourier Transform routines.  In addition to the FFT, the subpackage
    includes functions for the discrete sine and cosine transforms and for
    pseudo-differential operators.
\item[\texttt{integrate}] ~ \newline
    The \texttt{integrate} subpackage provides tools for the numerical
    computation of single and multiple definite integrals, and for the
    solution of ordinary differential equations, including initial value
    problems and two-point boundary value problems.
\item[\texttt{interpolate}] ~ \newline
    The \texttt{interpolate} subpackage contains spline functions and
    classes, one-dimensional and multi-dimensional (univariate and
    multivariate) interpolation classes, Lagrange and Taylor polynomial
    interpolators, and wrappers for FITPACK and DFITPACK functions.
\item[\texttt{io}] ~ \newline
    A collection of functions and classes for reading and writing MATLAB, IDL,
    Matrix Market, Fortran, NetCDF, Harwell-Boeing, WAV and ARFF data files. 
\item[\texttt{linalg}] ~ \newline
    Linear algebra functions, including:
    elementary functions of a matrix, such as the trace, determinant, norm and
    condition number;
    basic solver for $Ax = b$;
    specialized solvers for Toeplitz matrices, circulant matrices, triangular
    matrices and other structured matrices; least squares solver and
    pseudo-inverse calculations; eigenvalue and eigenvector calculations
    (basic and generalized); matrix decompositions, including Cholesky, Schur,
    Hessenberg, $LU$, $LDL^{\intercal}$, $QR$, $QZ$, singular value, and polar;
    and functions to create specialized matrices, such as diagonal, Toeplitz,
    Hankel, companion, Hilbert, and more.
\item[\texttt{ndimage}] ~ \newline
    This package contains various functions for multi-dimensional image
    processing, including: convolution and assorted linear and nonlinear
    filters (Gaussian filter, median filter, Sobel filter, etc.);
    interpolation; region labeling and processing; and mathematical morphology
    functions.
\item[\texttt{misc}] ~ \newline
    A collection of functions that did not fit into the other subpackages.
    While this subpackage still exists in version 1.0.0, effort is continuing
    to deprecate the contents of this subpackage and eventually remove it.
\item[\texttt{odr}] ~ \newline
    Orthogonal distance regression, including Python wrappers for the Fortran
    library ODRPACK.
\item[\texttt{optimize}] ~ \newline
    The package includes:
    implementations of many minimization algorithms; a general linear
    programming solver; a routine for least-squares curve fitting; and a
    collection of general nonlinear solvers for root-finding.
\item[\texttt{signal}] ~ \newline
    The \texttt{signal} subpackage focuses on signal processing (plus some
    basic linear systems theory).  Functionality includes:
    convolution and correlation; splines; filtering and filter design;
    continuous and discrete time linear systems; waveform generation;
    window functions; wavelet computations; peak finding; and spectral
    analysis.  
\item[\texttt{sparse}] ~ \newline
    This package includes implementations of several representations of
    sparse matrices.  It contains two subpackages, 
    \texttt{scipy.sparse.linalg} and \texttt{scipy.sparse.csgraph}.

    \texttt{scipy.sparse.linalg} provides a collection of linear algebra
    routines that work with sparse matrices, including linear equation
    solvers, eigenvalue decomposition, singular value decomposition
    and LU factorization.

    \texttt{scipy.sparse.csgraph} provides a collections of graph algorithms
    for which the graph is represented using a sparse matrix.  Algorithms
    include connected components, shortest path, minimum spanning tree
    and more.
\item[\texttt{spatial}] ~ \newline
    This subpackage provides spatial data structures and algorithms,
    including the k-d tree, Delaunay triangulation, convex hulls and Voronoi
    diagrams.  The subpackage \texttt{scipy.spatial.distance} provides
    a large collection of distance functions, along with functions for
    computing the distance between all pairs of vectors in a given collection
    of points or between all pairs from two collections of points.
\item[\texttt{special}] ~ \newline
    The name comes the class of functions traditionally known as \emph{special
    functions}, but over time, the module has grown to include functions
    beyond the classical special functions.  A more appropriate characterization
    of this subpackage is simply \emph{useful functions}.
    It includes: a large collection of the classical special functions
    such as Airy, Bessel, etc.; orthogonal families of polynomials;
    the Gamma function, and functions related to it;
    functions for computing the PDF, CDF and quantile function for several
    probability distributions;
    information theory functions;
    combinatorial functions \texttt{comb} and \texttt{factorial};
    and more.
\item[\texttt{stats}] ~ \newline
    The \texttt{stats} subpackage provides: a large collection of continuous
    and discrete \emph{probability distributions}, each with methods to compute
    the PDF or PMF, CDF, moments and other statistics, generation of random
    variates, and more;
    \emph{statistical tests}, including Pearson's correlation, Spearman's rank-order
    correlations, Kendal's tau, chi-squared test and its generalization as the
    Cressie-Read power divergence, contingency table tests including Fisher's
    exact test and Mood's median test, and many more;
    and assorted \emph{transformations and statistics} of data.
\end{description}

### 2018-04-11 scipy-1.0/scipy-optimize.tex

@@
 \end{table}
+
+Some problems may require the use of global optimizers to find the minimum of the objective function over the set of all parameter values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+\texttt{Brute} performs a grid search over the entire parameter range, with the best grid point undergoing further polishing by another solver.
+However, this approach becomes increasingly inefficient as the problem dimensionality increases, or if the objective function has many extrema.
+\texttt{Basinhopping} \cite{Wales1997} uses random perturbation of the parameter vector, with local minimization.
+The acceptance of the new coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
+\texttt{differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic solver that improves by gradual evolution of a population of candidate solutions.
+In each iteration trial candidates are generated by combination of candidates from the existing population; if the trial candidates represent an improvement then the population is updated.
+\texttt{differential\textunderscore evolution} does not require that the objective function be differentiable.
+%  basinhopping may need differentiable objective for the local minimization.
+The SciPy benchmark suite possesses a comprehensive range of global optimization problems (196), of varying difficulties, that can be used to evaluate the performance of a particular solver.
+This suite is used to decide whether new solvers have good enough performance to be added to the package.

### 2018-04-12 scipy-1.0/paper.tex
MAINT: Pinned the location of Table 1 to prevent interjection into asv subsection.
MAINT: add paragraph on `asv continuous`

@@
 \usepackage{amsmath}
+\usepackage{float}
 
@@
 \subsection*{Submodule organisation}
+\input{subpackages}
 
@@
 
-    Consider the following asv benchmark results, spanning roughly nine years of project history, that demonstrate the gradual performance
-    improvements in \texttt{scipy.spatial.cKDTree.query()} (nearest-neighbor search) performance, using command:
+    Consider the following asv benchmark results, spanning roughly nine years of project history. These demonstrate the gradual performance
+    improvements in a nearest-neighbor search, \texttt{scipy.spatial.cKDTree.query()} using the command:
 
@@
 
-    \begin{figure}[!ht]
+   Any pull request can be compared against the \texttt{master} branch with the command \texttt{asv continuous master new-feature}.
+This will provide a benchmark against the master branch and the branch a new feature is implemented in.
+More features are available in the documentation, including arguments to select which benchmarks to run.
+
+    \begin{figure}[H]
         \centering
@@
 The results are based on Python 2.7 performance on the master branch of the project using numpy 1.8.2 and Cython versions 0.27.3 and 0.21.1 (for improved backward compatibility).
-Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility / sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
+Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility and sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
     \end{figure}

### 2018-04-12 scipy-1.0/paper.tex
BUG: restructure doc flow (reference figure)

@@
 
-    Consider the following asv benchmark results, spanning roughly nine years of project history. These demonstrate the gradual performance
-    improvements in a nearest-neighbor search, \texttt{scipy.spatial.cKDTree.query()} using the command:
+    Consider the asv benchmark results shown in Figure~\ref{fig:asv-bench}, spanning roughly nine years of project history. These demonstrate the gradual performance
+    improvements in a nearest-neighbor search through \texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
 
@@
 
-   Any pull request can be compared against the \texttt{master} branch with the command \texttt{asv continuous master new-feature}.
-This will provide a benchmark against the master branch and the branch a new feature is implemented in.
-More features are available in the documentation, including arguments to select which benchmarks to run.
-
-    \begin{figure}[H]
+        \begin{figure}
         \centering
@@
 Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility and sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
-    \end{figure}
+        \label{fig:asv-bench}
+    \end{figure
+
+   Any pull request can be compared against the \texttt{master} branch with the command \texttt{asv continuous master new-feature}.
+This will provide a benchmark against the master branch and the branch a new feature is implemented in.
+More features are available in the documentation, including arguments to select which benchmarks to run.
 

### 2018-04-12 scipy-1.0/paper.tex
BUG: missing end bracket

@@
     \subsubsection*{Benchmark suite}
+
     The airspeed velocity (asv) library enables benchmarking Python packages over their lifetimes, and the performance of the SciPy
@@
 
-    Consider the asv benchmark results shown in Figure~\ref{fig:asv-bench}, spanning roughly nine years of project history. These demonstrate the gradual performance
+    Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning roughly nine years of project history. These demonstrate the gradual performance
     improvements in a nearest-neighbor search through \texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
@@
 Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility and sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
-        \label{fig:asv-bench}
-    \end{figure
+        \label{fig:asvbench}
+    \end{figure}
 

### 2018-04-12 scipy-1.0/scipy-optimize.tex
ENH: optimize improvements: attempted to coordinate global and local minimization sections

--- 2018-04-11 scipy-1.0/scipy-optimize.tex
@@
 \subsubsection*{Nonlinear Optimization}
+\paragraph{Local Minimization}
 
-The \texttt{minimize} function provides methods for solving nonlinear optimization
+The \texttt{minimize} function provides methods for finding local minima of nonlinear optimization
 problems. It unifies several methods with a common interface in order to facilitate
@@
 
-While most \texttt{minimize} development prior to SciPy 1.0 has focused on unconstrained minimization, methods \texttt{SLSQP} and \texttt{COBYLA}\footnote{We note that any equality constraint $c_{\text{eq}}(x) = 0$ can be expressed as two inequality constraints $c_{\text{eq}}(x) \leq 0$ and $-c_{\text{eq}}(x) \leq 0$} support fully general nonlinear programming.
-However, neither provides special treatment for sparsity.
+While most \texttt{minimize} development prior to SciPy 1.0 has focused on unconstrained minimization, 
+% methods \texttt{SLSQP} and \texttt{COBYLA}\footnote{We note that any equality constraint $c_{\text{eq}}(x) = 0$ can be expressed as two inequality constraints $c_{\text{eq}}(x) \leq 0$ and $-c_{\text{eq}}(x) \leq 0$}
+\texttt{SLSQP} supports fully general nonlinear programming.
+However, it does not provide special treatment for sparsity, limiting its utility to small problems.
 Accordingly, ongoing development efforts for \texttt{minimize} are focused on the addition of nonlinear programming algorithms for large, sparse problems.
@@
     \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$). The table also indicates which algorithms
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a local or global minimum. The table also indicates which algorithms
     can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
@@
 \end{table}
+
+\paragraph{Global Minimization}
+As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values.
+% , especially if the objective function possesses many local minima that other solvers could get stuck in.
+The \texttt{brute} function performs a grid search over the entire domain, and the best grid point is ``polished'' by a local minimizer.
+However, this approach becomes prohibitively expensive as the problem dimensionality increases.
+% As far as I understand how brute works, the run time does not depend on the number of extrema, as it only polishes the best one
+% , or if the objective function has many extrema. 
+To address this, the
+\texttt{basinhopping} \cite{Wales1997} function uses random perturbation of the decision variables in conjunction with local minimization.
+The acceptance of the perturbed coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
+\texttt{differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic solver that improves by gradual evolution of a population of candidate solutions.
+In each iteration, trial candidates are generated by combination of candidates from the existing population; if the trial candidates represent an improvement then the population is updated.
+% Is the following a unique feature of differential_evolution?
+If so, state it (although it seems that neither `brute`or `basinhopping` inherently require the a differentiable problem).
+If not, why mention it?
+% \texttt{differential\textunderscore evolution} does not require that the objective function be differentiable.
+%  basinhopping may need differentiable objective for the local minimization.
+Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems
+% They're ``comprehensive''
+% of varying difficulties, 
+% that can be used to evaluate the performance of the solvers. 
+%This suite is used to decide
+for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits inclusion in the package. 
+% have good enough performance to be added to the package.

### 2018-04-15 scipy-1.0/scipy-optimize.tex
ENH: optimize: reverted some changes as requested from last commit

@@
     \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a local or global minimum. The table also indicates which algorithms
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
     can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
@@
 \paragraph{Global Minimization}
-As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values.
-% , especially if the objective function possesses many local minima that other solvers could get stuck in.
+As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+% e.g.
+if there is concern that `minimize` could get stuck in an unacceptable local minimum?
 The \texttt{brute} function performs a grid search over the entire domain, and the best grid point is ``polished'' by a local minimizer.

### 2018-04-16 scipy-1.0/paper.tex
MAINT: add position flag for figure
ENH: update asv benchmarks with additional version of Cython.

@@
 
-    \begin{figure}
+    \begin{figure}[H]
         \centering
@@
         \caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query() over a roughly nine year commit history time frame.
-The results are based on Python 2.7 performance on the master branch of the project using numpy 1.8.2 and Cython versions 0.27.3 and 0.21.1 (for improved backward compatibility).
+The results are based on Python 2.7 performance on the master branch of the project using numpy 1.8.2 and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward compatibility).
 Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility and sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}

### 2018-05-22 scipy-1.0/paper.tex
ENH: add draft of test suite section.

@@
 
+    \subsubsection*{Test suite}
+    The SciPy test suite is orchestrated by a continuous integration matrix
+    that includes posix and Windows (32/64-bit) platforms managed by Travis CI and appveyor,
+    respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and include
+    code linting with pyflakes and pycodestyle. There are more than $13000$ unit
+    tests in the test suite, which is written for usage with the pytest runner, and
+    with a 76.5 \% coverage of approximately $204000$ lines
+    of code. Documentation for the code is automatically built and hosted by the
+    CircleCI service to facilitate evaluation of documentation changes / integrity.
+    Our full test suite also passes with PyPy3, a just-in-time compiled version
+    of the Python language.
+    % NOTE: is there a citation for PyPy?
+
 \section*{Project organisation and community}

### 2018-05-27 scipy-1.0/paper.tex
ENH: draft of 'impact now' Discussion Section.

@@
 \textbf{Impact now}
+SciPy has a strong developer community and a massive user base. GitHub
+traffic metrics report roughly 20,000 unique visitors to the source
+website between May 14 and May 27, 2018, with 721 unique copies ("clones")
+of the code base over that roughly two-week period. The developer community
+consists of 610 unique contributors of source code, with $>19,000$ commits
+accepted into the code base (GitHub page data).
+
+% need some better usage stats sources?
+% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
+From the user side, there have been at least 15 million downloads
+of the SciPy manylinux wheels between the start of 2016 and May of 2018
+(njs blog). The recommended manuscript for citing SciPy,
+published online 17 years ago, has been cited > 3000
+times. Some of the most prominent citing articles include the IPython
+publication, a heavily-cited Nature Biotechnology article, a dark
+matter annihlitation study in the field of cosmology, and many of the
+other major downstream software libraries in the Python scientific
+computing stack.
 

### 2018-05-27 scipy-1.0/paper.tex
MAINT: minor formatting fixes for 'impact now' section.

@@
 traffic metrics report roughly 20,000 unique visitors to the source
-website between May 14 and May 27, 2018, with 721 unique copies ("clones")
+website between May 14 and May 27, 2018, with 721 unique copies (``clones")
 of the code base over that roughly two-week period. The developer community
@@
 (njs blog). The recommended manuscript for citing SciPy,
-published online 17 years ago, has been cited > 3000
+published online 17 years ago, has been cited $>3000$
 times. Some of the most prominent citing articles include the IPython

### 2018-05-30 scipy-1.0/paper.tex
Drafted a short governance section.

--- 2018-05-22 scipy-1.0/paper.tex
@@
 
+SciPy adopted an official Governance Document on August 3, 2017 (see PR\#6953). In summary,
+we have a Benevolant Dictator for Life (BDFL) in Pauli Virtanen, and a (currently) 18
+member Steering Council chaired by Ralf Gommers that i.e., decides on who gets commit rights.
+
+SciPy adopted an official Code of Conduct on October 24, 2017 (see PR\#7963), drawing
+inspiration from the Apache Foundation Code of Conduct, the Contributor Covenant,
+the Jupyter Code of Conduct, and the Open Source Guides - Code of Conduct. In short,
+we stated five specific guidelines: \emph{be open} (to anyone participating in our community),
+\emph{be empathetic and patient} (in resolving conflicts), \emph{be collaborative} (we depend
+on each other to build up the tools in the library), \emph{be inquisitive} (identifying issues
+early on can be helpful to everyone), and \emph{be careful with wording} (no harassment or
+exclusionary behavior). We outline a broad diversity inclusion statement, and provide
+instructions for contacting the three members of our Code of Conduct Committee or an
+external representative on the NumFOCUS Board of Directors.
+
 \textbf{Roadmap}

### 2018-05-31 scipy-1.0/paper.tex
First round of revisions for PR 29.

@@
 on each other to build up the tools in the library), \emph{be inquisitive} (identifying issues
-early on can be helpful to everyone), and \emph{be careful with wording} (no harassment or
-exclusionary behavior). We outline a broad diversity inclusion statement, and provide
+early on can be helpful to everyone), and \emph{be careful with wording} (do not harass or exclude).
+We outline a broad diversity inclusion statement, and provide
 instructions for contacting the three members of our Code of Conduct Committee or an

### 2018-06-02 scipy-1.0/paper.tex
Second round of revisions for PR 29.

@@
 
-SciPy adopted an official Governance Document on August 3, 2017 (see PR\#6953). In summary,
-we have a Benevolant Dictator for Life (BDFL) in Pauli Virtanen, and a (currently) 18
-member Steering Council chaired by Ralf Gommers that i.e., decides on who gets commit rights.
+SciPy adopted an official Governance Document on August 3,
+2017\cite{SciPyProjectGovernance}.
+We have a Benevolant Dictator for Life (BDFL), Pauli Virtanen, leading the
+project, following a detailed discussion of the governance model with the
+community on the mailing list.  In practice, the overruling authority of the BDFL
+is anticipated to be used only when the Steering Council cannot agree on a matter.
+The BDFL is expected to consider any strong indications that they should step down;
+although the BDFL may appoint a successor it is generally expected that the Steering
+Council will be consulted on the matter.
 
-SciPy adopted an official Code of Conduct on October 24, 2017 (see PR\#7963), drawing
-inspiration from the Apache Foundation Code of Conduct, the Contributor Covenant,
-the Jupyter Code of Conduct, and the Open Source Guides - Code of Conduct. In short,
+The (currently) 18 member Steering Council is chaired by Ralf Gommers and performs a number of
+tasks including nominating and voting on new Council Members, and actively
+contributing to the progress of the project (could be code review but also
+outreach activities, etc.). Generally, 1 year of sustained and substantial
+contributions (not restricted to source code contributions) that improve
+the project are the prerequiste for nomination of new Council Members. Council
+Members do not get special treatment for the vast majority of the day to day
+project activities (code submission / review), though it is anticipated that their
+experience with the project will serve as helpful guidance in many matters.
+Council Members have commit rights to the project, but will typically only
+incorporate changes when there are no substantive community objections. The
+Chair of the Steering Council is responsible for starting more formal
+technical reviews of the direction of the project, ensuring that the
+composition of the Council remains current, and communicating / summarizing
+any Council activities performed privately to the broader community. We also
+describe in some detail how institutional partners may contribute to the
+project, but no financial weight from grants or other sources
+may be leveraged to circumvent or overrule the technical direction of the
+project decided upon by the Steering Committee. The Chair does not have
+a fixed term, but is nominated by the Steering Committee and is expected
+to yield to substantive claims that stepping down is the appropriate action.
+
+SciPy adopted an official Code of Conduct on October 24, 2017
+\cite{SciPyCodeOfConduct}, drawing
+inspiration from the Apache Foundation Code of
+Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
+\cite{ContributorConvenant},
+the Jupyter Code of Conduct \cite{Jupyter_COC}, and the Open Source Guides -
+Code of Conduct\cite{OSG_COC}. In short,
 we stated five specific guidelines: \emph{be open} (to anyone participating in our community),
@@
 early on can be helpful to everyone), and \emph{be careful with wording} (do not harass or exclude).
-We outline a broad diversity inclusion statement, and provide
+We outlined a broad diversity inclusion statement, and provide
 instructions for contacting the three members of our Code of Conduct Committee or an

### 2018-06-18 scipy-1.0/paper.tex
Drafted a subsection for LowLevelCallable.

--- 2018-05-31 scipy-1.0/paper.tex
@@
 \subsection*{Unified bindings to compiled code}
-LowLevelCallable
+\subsubsection*{LowLevelCallable}
+As of SciPy version 0.19, it is possible for users to wrap low-level
+functions in a \texttt{scipy.LowLevelCallable()} object that reduces
+the overhead for calling compiled C functions directly from Python.
+Supported low-level functions include \texttt{PyCapsule} objects,
+ctypes function pointers, and cffi function pointers. The low level
+function signature must be consistent with the expectations of the
+routine it is passed to. For example, the documentation for
+\texttt{scipy.ndimage.generic\_filter} defines two acceptable C callback
+function signatures that may be used to produce functions that operate
+on each element of image data with low overhead. The C code may be
+generated using numba or Cython, for example, as long as the function
+call signature matches the specifications. Furthermore, it is even
+possible to generate a low-level callback function automatically
+from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
 
@@
 
-SciPy adopted an official Governance Document on August 3, 2017 (see PR\#6953). In summary,
-we have a Benevolant Dictator for Life (BDFL) in Pauli Virtanen, and a (currently) 18
-member Steering Council chaired by Ralf Gommers that i.e., decides on who gets commit rights.
-
-SciPy adopted an official Code of Conduct on October 24, 2017 (see PR\#7963), drawing
-inspiration from the Apache Foundation Code of Conduct, the Contributor Covenant,
-the Jupyter Code of Conduct, and the Open Source Guides - Code of Conduct. In short,
-we stated five specific guidelines: \emph{be open} (to anyone participating in our community),
-\emph{be empathetic and patient} (in resolving conflicts), \emph{be collaborative} (we depend
-on each other to build up the tools in the library), \emph{be inquisitive} (identifying issues
-early on can be helpful to everyone), and \emph{be careful with wording} (do not harass or exclude).
-We outline a broad diversity inclusion statement, and provide
-instructions for contacting the three members of our Code of Conduct Committee or an
-external representative on the NumFOCUS Board of Directors.
-
 \textbf{Roadmap}

### 2018-06-27 scipy-1.0/paper.tex

--- 2018-06-02 scipy-1.0/paper.tex
@@
 \textbf{Impact now}
+SciPy has a strong developer community and a massive user base. GitHub
+traffic metrics report roughly 20,000 unique visitors to the source
+website between May 14 and May 27, 2018, with 721 unique copies (``clones")
+of the code base over that roughly two-week period. The developer community
+consists of 610 unique contributors of source code, with $>19,000$ commits
+accepted into the code base (GitHub page data).
+
+% need some better usage stats sources?
+% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
+From the user side, there have been at least 15 million downloads
+of the SciPy manylinux wheels between the start of 2016 and May of 2018
+(njs blog). The recommended manuscript for citing SciPy,
+published online 17 years ago, has been cited $>3000$
+times. Some of the most prominent citing articles include the IPython
+publication, a heavily-cited Nature Biotechnology article, a dark
+matter annihlitation study in the field of cosmology, and many of the
+other major downstream software libraries in the Python scientific
+computing stack.
 

### 2018-06-27 scipy-1.0/ckdtree.tex

The cKDTree module is rewritten in C++ with templated classes.

The support to a box of periodic boundary condition is added. A periodic
condition is typically used in computer simulations of physical processes.

\subsubsection{k-NN}

The API for querying neighbours is enhanced, to allow querying a subset of
k nearest neighbours.

\subsubsection{sparse distance matrix}

generating approximated sparse distance matrix

\subsubsection{Pair-Counting}

The cKDTree module implements the dual tree counting algorithm in (cite).

An improvement to the pair counting algorithm is added to improve the scale
with the number of bins.

\textbf{Add a Figure to show the scaling, before and after.}

The cKDTree can now be augmented by weights. Weighted paircount is essential
in many scientific applications, e.g. computing correlation function of galaxies.

\textbf{perhaps give an example or some formula.}

(cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)

The main purpose of the paircounting algorithm in cKDTree is to provide a readily
available tool. The KDTree based algorithm is not necessarily the fastest algorithm
depending on the application. E.g. for low number densities a chaining mesh based algorithm
can be easily many times faster than a tree based algorithm.

### 2018-06-27 scipy-1.0/paper.tex

@@
 \textbf{cKDTree}
+\input{ckdtree}
 \textbf{Sparse matrices}

### 2018-06-27 scipy-1.0/paper.tex
ENH: add some authors and a title for manuscript.
Remove duplicated author
ENH: first draft of abstract for SciPy paper.

@@
 \begin{abstract}
-TODO. Abstract must be under 200 words and not include subheadings or citations.
+SciPy is an open source scientific computing library for the Python programming language.
+SciPy 1.0 was released in late 2017, about 16 years after the original
+version 0.1 release. SciPy
+has become a \emph{de facto} standard in the community, with more than 600
+unique code contributors and millions of downloads per year, despite
+the absence of an official project publication, until now. The library
+includes functionality spanning clustering, Fourier transforms,
+integration, interpolation, file I/O, linear algebra, image processing,
+orthogonal distance regression, minimization algorithms, signal processing,
+sparse matrix handling, computational geometry, and statistics. In this
+work, we highlight some recent technical developments and provide
+an overview of the capabilities and development practices of the
+SciPy library.
 \end{abstract}

### 2018-06-28 scipy-1.0/paper.tex
Pixel vs point model

--- 2018-06-27 scipy-1.0/paper.tex
@@
 % need some better usage stats sources?
+%
+% note: @stefanv has pypinfo set up, so can grab all the info shown at
+% https://github.com/ofek/pypinfo
+%
 % SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
@@
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}
+
+\subsection{ndimage pixel vs point}
+
+\begin{itemize}
+  \item A pixel is not a little square
+    (http://alvyray.com/Memos/CG/Microsoft/6\_pixel.pdf)
+  \item Match edges vs match centers
+    (http://entropymine.com/imageworsener/matching/)
+\end{itemize}
+
+There are two competing models for representing image measurements:
+the first considers an image as a collection of pixels, square blocks
+of size 1x1 with centers offset by (0.5, 0.5).  This model corresponds
+closely with a camera sensor, that integrates light as it falls on
+(more-or-less) square elements, producing a digital readout for each.
+
+Another view is that of an image as a grid of point measurements.
+This model explicitly reject two notions encouraged by the pixel
+model: that our signal is best represented as a step-wise functio, and
+that a single measurement represents knowledge of the surrounding
+square area element.  Instead, we think of the image as a smoothly
+varying function, of which we have measurements at grid positions.
+
+While the representations may not seem very different, they have
+significant implications for image processing algorithms such as
+rescaling.
+
+{\tt ndimage} currently follows the pixel model, but we think that the
+data point model will serve us better in the long run:
+
+\begin{enumerate}
+  \item It makes no assumption about the sampling mechanism used
+    (camera, scanner, laser, etc.).
+  \item NumPy coordinates will map directly to image coordinates,
+    without a 0.5 offset.
+  \item It makes it easier to reason about extrapolation.  E.g., the current
+    dubious notion of ``reflection'', in which the image is padded
+    {\tt d c b a | a b c d | d c b a} becomes nonsensical, and leaves
+    only the reliable notion of mirroring, {\tt d c b | a b c d | c b a}.
+\end{enumerate}
+
+The second model, while consistent, does introduce some challenges
+w.r.t. traditional image manipulation (see attached figure).
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{static/image_pixel_model}
+    \caption{Caption goes here.}
+    \label{fig:image-pixel-model}
+\end{figure}
 

### 2018-06-28 scipy-1.0/paper.tex
ENH: drafted roadmap section of manuscript.

--- 2018-06-27 scipy-1.0/paper.tex
@@
 \usepackage{float}
+\usepackage{pgfplots}
+
+\pgfplotsset{compat=newest}
 
@@
 \newcommand{\xmark}{\ding{55}}%
+\newcommand{\code}[1]{\texttt{#1}}
 
-\title {SciPy - some descriptive title here}
+\title{SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python}
 
@@
 \author[2,*]{Ralf Gommers}
-\author[3]{TODO}
+\author[3,4]{Tyler Reddy}
+\author[5]{Anne Archibald}
+\author[6]{Andrew Nelson}
+\author[7]{Charles Harris}
+\author[8]{CJ Carey}
+\author[9]{Denis Laxalde}
+\author[10]{Eric Larson}
+\author[11]{Eric Moore}
+\author[12]{Eric Quintero}
+\author[13]{Evgeni Burovski}
+\author[14]{Jaime Fernández del Río}
+\author[15]{Josef Perktold}
+\author[16]{Josh Wilson}
+\author[17]{Matthew Brett}
+\author[18]{Nikolay Mayorov}
+\author[19]{Warren Weckesser}
+\author[20]{Matt Haberland}
+\author[21]{Scott Sievert}
+\author[22]{Yu Feng}
+\author[23]{Antonio Horta Ribeiro}
+
 \affil[1]{Affiliation, department, city, postcode, country}
@@
 \affil[2]{Affiliation, department, city, postcode, country}
+\affil[3]{UC Berkeley Institute for Data Science,
+          Berkeley, CA, 94720, USA}
+\affil[4]{Los Alamos National Laboratory,
+	  Theoretical Division 6,
+          Los Alamos, NM, 87545, USA}
+\affil[5]{Affiliation, department, city, postcode, country}
+\affil[6]{Affiliation, department, city, postcode, country}
+\affil[7]{Affiliation, department, city, postcode, country}
+\affil[8]{Affiliation, department, city, postcode, country}
+\affil[9]{Affiliation, department, city, postcode, country}
+\affil[10]{Affiliation, department, city, postcode, country}
+\affil[11]{Affiliation, department, city, postcode, country}
+\affil[12]{Affiliation, department, city, postcode, country}
+\affil[13]{Affiliation, department, city, postcode, country}
+\affil[14]{Affiliation, department, city, postcode, country}
+\affil[15]{Affiliation, department, city, postcode, country}
+\affil[16]{Affiliation, department, city, postcode, country}
+\affil[17]{Affiliation, department, city, postcode, country}
+\affil[18]{Affiliation, department, city, postcode, country}
+\affil[19]{Affiliation, department, city, postcode, country}
+\affil[20]{Affiliation, department, city, postcode, country}
+\affil[21]{Affiliation, department, city, postcode, country}
+\affil[22]{Affiliation, department, city, postcode, country}
+\affil[23]{Affiliation, department, city, postcode, country}
 
@@
 
+% NOTE: some usage stats below are from https://libraries.io/pypi/scipy
+% as well as github metrics; normally one doesn't put citations directly
+% in abstract though (depends on field / journal)
 \begin{abstract}
@@
 version 0.1 release. SciPy
-has become a \emph{de facto} standard in the community, with more than 600
-unique code contributors and millions of downloads per year, despite
-the absence of an official project publication, until now. The library
-includes functionality spanning clustering, Fourier transforms,
+has become a \emph{de facto} standard for leveraging scientific algorithms
+in the Python programming language, with more than 600
+unique code contributors, millions of downloads per year, 161 dependent
+packages, and 28700 dependent respositories.
+The library includes functionality spanning clustering, Fourier transforms,
 integration, interpolation, file I/O, linear algebra, image processing,
@@
 sparse matrix handling, computational geometry, and statistics. In this
-work, we highlight some recent technical developments and provide
-an overview of the capabilities and development practices of the
-SciPy library.
+work, we provide an overview of the capabilities and development practices of the
+SciPy library and highlight some recent technical developments.
 \end{abstract}
@@
 
-\subsection*{Language preferences}
+\subsection*{Language choices}
+Python, Cython, Fortran, C and C++ are the programming languages
+used to implement scientific algorithms in the SciPy library. An analysis
+of our code base using the \texttt{linguist} library provides a detailed
+breakdown as \% composition by programming language in SciPy
+(Figure~\ref{fig:linguist}).
+
+    \begin{figure}[H]
+        \centering
+
+        \begin{tikzpicture}[font=\sffamily]
+        \begin{axis}[height=6cm, width=10cm,
+        ybar, ymax=60, ymin=0,bar width=1cm,
+        ylabel={\% of Scipy code base},
+        ytick={0,20,...,60},
+        axis lines*=left,
+        axis y line shift=0pt,
+        enlarge x limits=0.15,
+        symbolic x coords={py,f,c,cy,cpp},
+        xticklabels={Python,FORTRAN,C,Cython,C++},
+        ticklabel style={align=center,font=\sffamily\small,
+        /pgf/number format/assume math mode=true},
+        xtick=data,
+        nodes near coords,
+        nodes near coords style={/pgf/number format/assume math mode=true}
+        ]
+        \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
+        \end{axis}
+        \end{tikzpicture}
+        \caption{The breakdown of programming languages used in the
+	         SciPy library determined using the linguist library.
+		 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
+		 and Makefile are excluded for clarity and mostly
+		 provide supporting roles in tests, building, and
+		 documentation.}
+        \label{fig:linguist}
+    \end{figure}
+
+For implementing new functionality, we have a clear order of language preference.
+First Python, if performance is not an issue. If it is, then in order of
+preference: Cython, C, C++, Fortran. The main motivation for this is 
+maintainability: Cython has the highest abstraction level and most 
+Python developers will understand it. C is also widely known,
+and easier to deal with for the core development team
+than C++ and especially Fortran.
+
+While it is not surprising that Python is heavily used in SciPy, the
+usage distribution of the other languages warrants some discussion. Fortran
+is an extremely well-established scientific programming language, both
+for historical reasons and because of its continued excellent
+performance\cite{Koelbel:1993:HPF:562354}. We
+wrap the FFTPACK Fortran library for performing Fourier
+transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} since
+this library has been a standard in the field for 33 years and has a license
+that is compatible with our own. Likewise, we wrap the Fortran source
+for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial 
+value problem for ordinary differential equation systems for 30 years. For
+similar reasons we also wrap the Fortran libraries QUADPACK\cite{1983qspa.book.....P} (for numerical
+integration of one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103} (curve-fitting /
+interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal distance regression),
+MINPACK\cite{osti_6997568} (minimization of linear and nonlinear equations),
+ARPACK\cite{leh:sor:yan96} (solving large scale eigenvalue problems), 
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling Bessel Functions), and 
+CDFLIB\cite{CDFLIB_site} (evaluation of cumulative density functions).
+
+Rounding out the top three languages in SciPy is C, which is also extremely
+well-established over several decades\cite{Kernighan:1988:CPL:576122} of
+scientific computing. Alongside its objected-oriented relative C++, C
+can be leveraged in Python after being exposed using the Cython
+language. Cython has been described as a creole language that mixes
+the best parts of Python and lower-level C / C++
+paradigms\cite{behnel2011cython}. We thus often use Cython as a glue
+between well-established low-level scientific computing libraries and
+the Python interface offered by SciPy. We also use Cython to enable
+performance enhancements in Python code, especially for cases where heavily
+used inner loops benefit from a compiled code with static typing. Some of the C
+libraries that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842} 
+(iterative solving of the trust region problem), SuperLU\cite{li05,superlu_ug99} (solution of
+large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821} (computational
+geometry), and Cephes\cite{cephes_netlib} (mathematics algorithms).
+
+Therefore, the relative abundance of different programming languages
+in the SciPy library results from a combination of the usage of powerful performance
+enhancing languages that interact well with Python (i.e., Cython), and
+the usage of languages (and their libraries) that have proven reliable
+and performant over many decades. The position that SciPy occupies
+near the foundation of the scientific Python ecosystem is such that
+adoption of new languages or major dependencies is generally unlikely--our
+choices are strongly driven by long-term stability. GPU acceleration,
+new transpiling libraries, and the latest JIT compilation approaches
+(i.e., Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall outside the remit
+of the main SciPy library. That said, we have recently increased our
+efforts to support compatibility with some of these options, and having
+our full test suite pass with the PyPy JIT compiler\cite{Bolz:2009:TMP:1565824.1565827}
+is now a requirement in our development workflow.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Polynomial interpolators}
+\input{poly}
 
@@
 \textbf{Roadmap}
+The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
+maintained by the community that describes some of the major directions
+for improvement for the project, as well as specific limitations and
+matters that require assistance moving forward.
+
+In general, we encourage changes that improvie clarity in the API of the library,
+but strongly discourage breaking backwards compatibility, given our position near
+the base of the scientific Python computing stack.
+
+Unit test coverage for most recent code additions / adjustments is quite
+respectable, but some of the historical components of the code base
+would still benefit from increased test coverage. We are still striving
+to increase the number of high-quality SciPy usage tutorials, but our
+standard documentation for library features is already in excellent shape.
+
+The low-level Cython code in our library (which interacts with C-level
+code and exposes it for Python usage) could use some measure of modernization,
+including migration to typed memoryviews to handle NumPy arrays in place
+of the older syntax.
+
+Some specific submodules of SciPy would benefit from  attention moving forward.
+\texttt{fftpack} has some single-precision floating point errors, a caching bug,
+and would benefit from the addition of the Bluestein algorithm. There's also
+too much community code duplication with \texttt{numpy.fft}, which
+requires clarification of best choices and long-term plans.
+
+\texttt{scipy.integrate} now has a new ODE solver interface, and we should aim
+to gracefully deprecate the older API. Support for integrating complex-valued
+functions and integrating multiple intervals could be beneficial.
+
+\texttt{scipy.interpolate} could benefit from a set of new features including
+spline fitting routines with better user control; integration, differentiation,
+and arithmetic routines for splines; transparent tensor-product splines,
+Non-Uniform Rational B-Splines support; and mesh refinement and coarsening of
+B-splines and corresponding tensor products.
+
+\texttt{scipy.io} would benefit from PCM float support, but we otherwise
+direct users with more sophisticated needs to domain-specific audio
+libraries.
+
+For \texttt{scipy.linalg} we need to remove functions that are duplicate with 
+\texttt{numpy.linalg}, \texttt{get\_lapack\_funcs} should always use flapack, and
+we should continue to wrap more LAPACK functions. Ideally, we'd also
+add type-generic wrappers in the Cython BLAS and LAPACK interfaces, and
+transform many of the linear algebra routines into gufuncs.
+
+We're planning to remove \texttt{scipy.misc} as a public module, and
+rename \texttt{scipy.odr} to \texttt{scipy.regression} or
+\texttt{scipy.fitting}, and include \texttt{optimize.curve\_fit}. This 
+module will then provide a home for other fitting functionality.
+\texttt{scipy.optimize} would benefit from a few more good global optimizers 
+as well as large-scale optimizers.
+
+\texttt{scipy.ndimage} should be completely migrated to a data point
+model (values on a grid), instead of a pixel model (elements with centers),
+both of which have been used in the past, because the former is now
+proving more effective in practice. We also need to standardize the
+\texttt{scipy.ndimage.morphology} API.
+
+\texttt{scipy.signal} needs its overlap with \texttt{scipy.ndimage}
+to be reduced; B-splines functionality should be moved to \texttt{interpolate},
+\texttt{firwin2} can be merged into \texttt{firwin}; Second Order Sections need
+to be updated to match the capabilities in other routines, and wavelet-related
+routines are somewhat out of place.
+
+\texttt{scipy.sparse} is fairly mature now, but there are many data
+formats and the community continues to move forward here so we will aim
+to take advantage of their improvements while preserving backwards
+compatibility to the extent that it is possible. \texttt{sparse.linalg}
+may benefit from wrappers for PROPACK for faster sparse SVD computation.
+
+For \texttt{scipy.spatial}, KDTree will be removed, and cKDTree will be 
+renamed to KDTree in a backwards-compatible way. \texttt{distance\_wrap.c} 
+likely needs a rewrite in Cython. \texttt{scipy.special} needs precision
+improvements for hypergeometric functions, parabolic cylinder functions,
+and spheroidal wave functions.
 
@@
 of the SciPy manylinux wheels between the start of 2016 and May of 2018
-(njs blog). The recommended manuscript for citing SciPy,
-published online 17 years ago, has been cited $>3000$
+(njs blog). The recommended source for citing SciPy is simply a reference
+to the website / library itself\cite{SciPylib}, and has been cited $>3000$
 times. Some of the most prominent citing articles include the IPython
-publication, a heavily-cited Nature Biotechnology article, a dark
-matter annihlitation study in the field of cosmology, and many of the
+publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
+article\cite{pmid17287757}, a dark
+matter annihilitation study in the field of
+cosmology\cite{1475-7516-2012-08-007}, and many of the
 other major downstream software libraries in the Python scientific

### 2018-06-28 scipy-1.0/paper.tex
First round of revisions for manuscript PR 31 -- adjust title based on feedback.

--- 2018-06-27 scipy-1.0/paper.tex
@@
 
-\title {SciPy - some descriptive title here}
+\title {SciPy 1.0 - Fundamental Algorithms for Scientific Computing in Python}
 
@@
 \begin{abstract}
-SciPy is an open source scientific computing library for the Python programming language.
-SciPy 1.0 was released in late 2017, about 16 years after the original
-version 0.1 release. SciPy
-has become a \emph{de facto} standard in the community, with more than 600
-unique code contributors and millions of downloads per year, despite
-the absence of an official project publication, until now. The library
-includes functionality spanning clustering, Fourier transforms,
-integration, interpolation, file I/O, linear algebra, image processing,
-orthogonal distance regression, minimization algorithms, signal processing,
-sparse matrix handling, computational geometry, and statistics. In this
-work, we highlight some recent technical developments and provide
-an overview of the capabilities and development practices of the
-SciPy library.
+TODO. Abstract must be under 200 words and not include subheadings or citations.
 \end{abstract}

### 2018-06-28 scipy-1.0/paper.tex
First round of revisions for manuscript PR 32 -- improving the abstract draft based on feedback.

@@
 
-\title {SciPy 1.0 - Fundamental Algorithms for Scientific Computing in Python}
+\title {SciPy - some descriptive title here}
 
@@
 \author[2,*]{Ralf Gommers}
-\author[3,4]{Tyler Reddy}
-\author[5]{Anne Archibald}
-\author[6]{Andrew Nelson}
-\author[7]{Charles Harris}
-\author[8]{CJ Carey}
-\author[9]{Denis Laxalde}
-\author[10]{Eric Larson}
-\author[11]{Eric Moore}
-\author[12]{Eric Quintero}
-\author[13]{Evgeni Burovski}
-\author[14]{Jaime Fernández del Río}
-\author[15]{Josef Perktold}
-\author[16]{Josh Wilson}
-\author[17]{Matthew Brett}
-\author[18]{Nikolay Mayorov}
-\author[19]{Warren Weckesser}
-\author[20]{Matt Haberland}
-\author[21]{Scott Sievert}
-\author[22]{Yu Feng}
-\author[23]{Antonio Horta Ribeiro}
-
+\author[3]{TODO}
 \affil[1]{Affiliation, department, city, postcode, country}
@@
 \affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{UC Berkeley Institute for Data Science,
-          Berkeley, CA, 94720, USA}
-\affil[4]{Los Alamos National Laboratory,
-	  Theoretical Division 6,
-          Los Alamos, NM, 87545, USA}
-\affil[5]{Affiliation, department, city, postcode, country}
-\affil[6]{Affiliation, department, city, postcode, country}
-\affil[7]{Affiliation, department, city, postcode, country}
-\affil[8]{Affiliation, department, city, postcode, country}
-\affil[9]{Affiliation, department, city, postcode, country}
-\affil[10]{Affiliation, department, city, postcode, country}
-\affil[11]{Affiliation, department, city, postcode, country}
-\affil[12]{Affiliation, department, city, postcode, country}
-\affil[13]{Affiliation, department, city, postcode, country}
-\affil[14]{Affiliation, department, city, postcode, country}
-\affil[15]{Affiliation, department, city, postcode, country}
-\affil[16]{Affiliation, department, city, postcode, country}
-\affil[17]{Affiliation, department, city, postcode, country}
-\affil[18]{Affiliation, department, city, postcode, country}
-\affil[19]{Affiliation, department, city, postcode, country}
-\affil[20]{Affiliation, department, city, postcode, country}
-\affil[21]{Affiliation, department, city, postcode, country}
-\affil[22]{Affiliation, department, city, postcode, country}
-\affil[23]{Affiliation, department, city, postcode, country}
 
@@
 
+% NOTE: some usage stats below are from https://libraries.io/pypi/scipy
+% as well as github metrics; normally one doesn't put citations directly
+% in abstract though (depends on field / journal)
 \begin{abstract}
-TODO. Abstract must be under 200 words and not include subheadings or citations.
+SciPy is an open source scientific computing library for the Python programming language.
+SciPy 1.0 was released in late 2017, about 16 years after the original
+version 0.1 release. SciPy
+has become a \emph{de facto} standard for leveraging scientific algorithms
+in the Python programming language, with more than 600
+unique code contributors, millions of downloads per year, 161 dependent
+packages, and 28700 dependent respositories.
+The library includes functionality spanning clustering, Fourier transforms,
+integration, interpolation, file I/O, linear algebra, image processing,
+orthogonal distance regression, minimization algorithms, signal processing,
+sparse matrix handling, computational geometry, and statistics. In this
+work, we provide an overview of the capabilities and development practices of the
+SciPy library and highlight some recent technical developments.
 \end{abstract}

### 2018-06-28 scipy-1.0/paper.tex
First round of revisions for manuscript PR 34 based on reviewer comments for the draft roadmap.

--- 2018-06-28 scipy-1.0/paper.tex
@@
 
-In general, we encourage changes that improvie clarity in the API of the library,
+In general, we encourage changes that improve clarity in the API of the library,
 but strongly discourage breaking backwards compatibility, given our position near
@@
 
-\texttt{scipy.integrate} now has a new ODE solver interface, and we should aim
+\texttt{scipy.integrate} has a new ODE solver interface, and we should aim
 to gracefully deprecate the older API. Support for integrating complex-valued
@@
 
-\texttt{scipy.sparse} is fairly mature now, but there are many data
+While \texttt{scipy.sparse} is mature, there are many data
 formats and the community continues to move forward here so we will aim

### 2018-06-28 scipy-1.0/paper.tex

--- 2018-06-28 scipy-1.0/paper.tex
@@
 
-% NOTE: some usage stats below are from https://libraries.io/pypi/scipy
-% as well as github metrics; normally one doesn't put citations directly
-% in abstract though (depends on field / journal)
 \begin{abstract}
-SciPy is an open source scientific computing library for the Python programming language.
-SciPy 1.0 was released in late 2017, about 16 years after the original
-version 0.1 release. SciPy
-has become a \emph{de facto} standard for leveraging scientific algorithms
-in the Python programming language, with more than 600
-unique code contributors, millions of downloads per year, 161 dependent
-packages, and 28700 dependent respositories.
-The library includes functionality spanning clustering, Fourier transforms,
-integration, interpolation, file I/O, linear algebra, image processing,
-orthogonal distance regression, minimization algorithms, signal processing,
-sparse matrix handling, computational geometry, and statistics. In this
-work, we provide an overview of the capabilities and development practices of the
-SciPy library and highlight some recent technical developments.
+TODO. Abstract must be under 200 words and not include subheadings or citations.
 \end{abstract}

### 2018-06-28 scipy-1.0/paper.tex
Second round of revisions for manuscript PR 34 based on reviewer feedback.

--- 2018-06-28 scipy-1.0/paper.tex
@@
 \pgfplotsset{compat=newest}
+\usepackage{listings}
 
@@
 Unit test coverage for most recent code additions / adjustments is quite
-respectable, but some of the historical components of the code base
+respectable, with overall project test coverage holding steady at
+around 76\% for the last six months (Figure~\ref{fig:coverage}),
+but some of the historical components of the code base
 would still benefit from increased test coverage. We are still striving
-to increase the number of high-quality SciPy usage tutorials, but our
+to increase the number of high-quality SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}, but our
 standard documentation for library features is already in excellent shape.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/coverage-chart}
+\caption{\% coverage of the SciPy code base by unit tests over the six
+	 months preceding June 28, 2018 as reported by the \texttt{codecov}
+	 service. The analysis of lines of source code covered is performed
+	 automatically in our continuous integation suite using the
+	 \texttt{pytest-cov} and \texttt{gcov} libraries.}
+\label{fig:coverage}
+\end{figure}
 
@@
 code and exposes it for Python usage) could use some measure of modernization,
-including migration to typed memoryviews to handle NumPy arrays in place
-of the older syntax.
+including migration to typed memoryviews to handle NumPy arrays instead of
+the older syntax:
 
-Some specific submodules of SciPy would benefit from  attention moving forward.
+\begin{lstlisting}[language=Python]
+# example of a function that takes a single
+# 3 dimesional typed memoryview as an argument
+# this allows Cython-level handling of a NumPy
+# array object
+cpdef int func_new(int[:, :, :] arr):
+    pass
+
+# example of the old syntax for handling
+# the same scenario
+cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    pass
+
+\end{lstlisting}
+
 \texttt{fftpack} has some single-precision floating point errors, a caching bug,
@@
 
-For \texttt{scipy.linalg} we need to remove functions that are duplicate with 
-\texttt{numpy.linalg}, \texttt{get\_lapack\_funcs} should always use flapack, and
+For \texttt{scipy.linalg}, we need to remove functions that are duplicate with 
+\texttt{numpy.linalg} unless performance is superior in SciPy,
+\texttt{get\_lapack\_funcs} should always use flapack, and
 we should continue to wrap more LAPACK functions. Ideally, we'd also
@@
 rename \texttt{scipy.odr} to \texttt{scipy.regression} or
-\texttt{scipy.fitting}, and include \texttt{optimize.curve\_fit}. This 
+\texttt{scipy.fitting}, and relocate \texttt{optimize.curve\_fit}. This 
 module will then provide a home for other fitting functionality.
@@
 model (values on a grid), instead of a pixel model (elements with centers),
-both of which have been used in the past, because the former is now
+both of which have been used in the past, because the data point model is now
 proving more effective in practice. We also need to standardize the

### 2018-06-28 scipy-1.0/paper.tex
Formatting adjustments to title in response to reviewer comments in manuscript PR 31.

--- 2018-06-28 scipy-1.0/paper.tex
@@
 
-\title {SciPy - some descriptive title here}
+\title{SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python}
 
@@
 of the SciPy manylinux wheels between the start of 2016 and May of 2018
-(njs blog). The recommended source for citing SciPy is simply a reference
-to the website / library itself\cite{SciPylib}, and has been cited $>3000$
+(njs blog). The recommended manuscript for citing SciPy,
+published online 17 years ago, has been cited $>3000$
 times. Some of the most prominent citing articles include the IPython
-publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
-article\cite{pmid17287757}, a dark
-matter annihilitation study in the field of
-cosmology\cite{1475-7516-2012-08-007}, and many of the
+publication, a heavily-cited Nature Biotechnology article, a dark
+matter annihlitation study in the field of cosmology, and many of the
 other major downstream software libraries in the Python scientific

### 2018-06-28 scipy-1.0/paper.tex
Tidy text:
- Address edits from reviews
- Make a note where current thinking may be flawed (further research required)
- Add caption to figure (needs some work, but better than nothing)

--- 2018-06-28 scipy-1.0/paper.tex
@@
+\PassOptionsToPackage{dvipsnames}{xcolor}
 \documentclass[fleqn,10pt]{wlscirep}
@@
 \usepackage{float}
+
+\usepackage[dvipsnames]{xcolor}
+\usepackage{todonotes}
+\presetkeys%
+    {todonotes}%
+    {inline,linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum}{}
 
@@
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}
+
+\subsection*{ndimage pixel vs point}
+
+\begin{itemize}
+  \item A pixel is not a little square
+    (http://alvyray.com/Memos/CG/Microsoft/6\_pixel.pdf)
+  \item Match edges vs match centers
+    (http://entropymine.com/imageworsener/matching/)
+\end{itemize}
+
+There are two competing models for representing image measurements:
+the first considers an image as a collection of pixels, square blocks
+of size 1x1 with centers offset by (0.5, 0.5).  This model corresponds
+closely with a camera sensor, that integrates light as it falls on
+(more-or-less) square elements, producing a digital readout for each.
+
+Another view is that of an image as a grid of point measurements.
+This model explicitly rejects two notions encouraged by the pixel
+model: that our signal is best represented as a step-wise function, and
+that a single measurement represents knowledge of the surrounding
+square area element.  Instead, we think of the image as a smoothly
+varying function, of which we have measurements at grid positions.
+
+While the representations may not seem very different, they have
+significant implications for image processing algorithms such as
+rescaling.
+
+{\tt ndimage} currently follows the pixel model for operations like
+{\tt zoom} but the data point model for general geometric
+transformations and interpolation.  We think that consistently using
+the data point model will serve us better in the long run:
+
+\begin{enumerate}
+  \item It makes no assumption about the sampling mechanism used
+    (camera, scanner, laser, etc.).
+  \item NumPy coordinates map directly to measurement coordinates,
+    without a 0.5 offset. \todo{Will have to rethink this statement,
+      since this {\bf is} the current underlying model for interpolation}
+  \item It makes it easier to reason about extrapolation.  E.g., the current
+    dubious notion of ``reflection'', in which the image is padded
+    {\tt d c b a | a b c d | d c b a} becomes nonsensical, and leaves
+    only the reliable notion of mirroring, {\tt d c b | a b c d | c b a}.
+\end{enumerate}
+
+The second model, while consistent, does introduce some challenges
+w.r.t. traditional image manipulation (see attached figure).
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{static/image_pixel_model}
+    \caption{Upscaling an image 2x.  In the data point model, the
+      original data points remain in the image corners, which means it
+      contributes less to the center of the resulting image than to
+      the corners.  In the pixel model, the original pixels contribute
+      more evenly to both of these.}
+    \label{fig:image-pixel-model}
+\end{figure}
 

### 2018-06-29 scipy-1.0/paper.tex
Early draft of language choices subsection.

--- 2018-06-28 scipy-1.0/paper.tex
@@
 
-\subsection*{Language preferences}
+\subsection*{Language choices}
+Python / Cython, Fortran, C and C++ are the programming languages
+used to implement scientific algorithms in the SciPy library. An analysis
+of our code base using the \texttt{linguist} library provides a detailed
+breakdown as \% composition by programming language in SciPy
+(Figure~\ref{fig:linguist}).
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=\textwidth]{static/linguist}
+        \caption{The breakdown of programming languages used in the
+	         SciPy library determined using the linguist library.
+		 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
+		 and Makefile are excluded for clarity and mostly
+		 provide supporting roles in tests, building, and
+		 documentation.}
+        \label{fig:linguist}
+    \end{figure}
 

### 2018-07-01 scipy-1.0/paper.tex
Language choices subsection now distinguishes Cython & Python
* we now use a custom branch of linguist to parse Cython
and Python source components separately
* the SciPy library language composition Figure is now
generated using TikZ to facilitate modification

Co-authored-by: Ilhan Polat <ilhanpolat@gmail.com>

@@
 \subsection*{Language choices}
-Python / Cython, Fortran, C and C++ are the programming languages
+Python, Cython, Fortran, C and C++ are the programming languages
 used to implement scientific algorithms in the SciPy library. An analysis
@@
         \centering
-        \includegraphics[width=\textwidth]{static/linguist}
+
+        \begin{tikzpicture}[font=\sffamily]
+        \begin{axis}[height=6cm, width=10cm,
+        ybar, ymax=60, ymin=0,bar width=1cm,
+        ylabel={\% of Scipy code base},
+        ytick={0,20,...,60},
+        axis lines*=left,
+        axis y line shift=0pt,
+        enlarge x limits=0.15,
+        symbolic x coords={py,f,c,cy,cpp},
+        xticklabels={Python,FORTRAN,C,Cython,C++},
+        ticklabel style={align=center,font=\sffamily\small,
+        /pgf/number format/assume math mode=true},
+        xtick=data,
+        nodes near coords,
+        nodes near coords style={/pgf/number format/assume math mode=true}
+        ]
+        \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
+        \end{axis}
+        \end{tikzpicture}
         \caption{The breakdown of programming languages used in the

### 2018-07-06 scipy-1.0/paper.tex

--- 2018-06-29 scipy-1.0/paper.tex
@@
 
-\title{SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python}
+\title {SciPy - some descriptive title here}
 
@@
 
-\subsection*{Language choices}
-Python / Cython, Fortran, C and C++ are the programming languages
-used to implement scientific algorithms in the SciPy library. An analysis
-of our code base using the \texttt{linguist} library provides a detailed
-breakdown as \% composition by programming language in SciPy
-(Figure~\ref{fig:linguist}).
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[width=\textwidth]{static/linguist}
-        \caption{The breakdown of programming languages used in the
-	         SciPy library determined using the linguist library.
-		 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
-		 and Makefile are excluded for clarity and mostly
-		 provide supporting roles in tests, building, and
-		 documentation.}
-        \label{fig:linguist}
-    \end{figure}
+\subsection*{Language preferences}
 

### 2018-07-06 scipy-1.0/paper.tex
Drafted FORTRAN section of language choices description along with appropriate citations.

--- 2018-07-01 scipy-1.0/paper.tex
@@
     \end{figure}
+
+While it is not surprising that Python is heavily used in SciPy, the
+usage distribution of the other languages warrants some discussion. Fortran
+is an extremely well-established scientific programming language, both
+for historical reasons and because of its continued excellent
+performance\cite{Koelbel:1993:HPF:562354}. We
+wrap the FFTPACK Fortran library for performing Fourier
+transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} since
+this library has been a standard in the field for 33 years and has a license
+that is compatible with our own. Likewise, we wrap the Fortran source
+for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial 
+value problem for ordinary differential equation systems for 30 years. For
+similar reasons we also warp the Fortran libraries QUADPACK\cite{1983qspa.book.....P} (for numerical
+integration of one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103} (curve-fitting /
+interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal distance regression),
+MINPACK\cite{osti_6997568} (minimization of linear and nonlinear equations),
+ARPACK\cite{leh:sor:yan96} (solving large scale eigenvalue problems), 
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling Bessel Functions), and 
+CDFLIB\cite{CDFLIB_site} (evaluation of cumulative density functions).
 

### 2018-07-06 scipy-1.0/paper.tex
Added C language family section to language choices description along with appropriate references.

@@
 CDFLIB\cite{CDFLIB_site} (evaluation of cumulative density functions).
+
+Rounding out the top three languages in SciPy is C, which is also extremely
+well-established over several decades\cite{Kernighan:1988:CPL:576122} of
+scientific computing. Alongside its objected-oriented relative C++, C
+can be leveraged in Python after being exposed using the Cython
+language. Cython has been described as a creole language that mixes
+the best parts of Python and lower-level C / C++
+paradigms\cite{behnel2011cython}. We thus often use Cython as a glue
+between well-established low-level scientific computing libraries and
+the Python interface offered by SciPy. We also use Cython to enable
+performance enhancements in Python code, especially for cases where heavily
+used inner loops benefit from a compiled code with static typing. Some of the C
+libraries that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842} 
+(iterative solving of the trust region problem), SuperLU\cite{li05,superlu_ug99} (solution of
+large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821} (computational
+geometry), and Cephes\cite{cephes_netlib} (mathematics algorithms).
 

### 2018-07-06 scipy-1.0/paper.tex
Added a final summary paragraph to the language choices subsection along with appropriate references for discussion of JIT compilers.

@@
 
+Therefore, the relative abundance of different programming languages
+in the SciPy library results from a combination of the usage of powerful performance
+enhancing languages that interact well with Python (i.e., Cython), and
+the usage of languages (and their libraries) that have proven reliable
+and performant over many decades. The position that SciPy occupies
+near the foundation of the scientific Python ecosystem is such that
+adoption of new languages or major dependencies is generally unlikely--our
+choices are strongly driven by long-term stability. GPU acceleration,
+new transpiling libraries, and the latest JIT compilation approaches
+(i.e., Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall outside the remit
+of the main SciPy library. That said, we have recently increased our
+efforts to support compatibility with some of these options, and having
+our full test suite pass with the PyPy JIT compiler\cite{Bolz:2009:TMP:1565824.1565827}
+is now a requirement in our development workflow.
+
 \subsection*{API and ABI evolution}

### 2018-07-07 scipy-1.0/paper.tex
Early draft of tech improvements in scipy.stats in last 3 years.

@@
 \pgfplotsset{compat=newest}
+\usepackage{listings, textcomp}
+
+\lstset{ %
+  basicstyle=\ttfamily\footnotesize,  % size of fonts used for the code
+  breaklines=true,   % automatic line breaking only at whitespace
+  captionpos=b,   % sets the caption-position to bottom
+  commentstyle=\color{gray},  % comment style
+  keywordstyle=\color{blue},  % keyword style
+  stringstyle=\color{red},  % string literal style
+  upquote=true  %straight single quotes (requires textcomp)
+}
 
@@
 \newcommand{\xmark}{\ding{55}}%
+\newcommand{\code}[1]{\texttt{#1}}
 
@@
 
+For implementing new functionality, we have a clear order of language preference.
+First Python, if performance is not an issue. If it is, then in order of
+preference: Cython, C, C++, Fortran. The main motivation for this is 
+maintainability: Cython has the highest abstraction level and most 
+Python developers will understand it. C is also widely known,
+and easier to deal with for the core development team
+than C++ and especially Fortran.
+
 While it is not surprising that Python is heavily used in SciPy, the
@@
 value problem for ordinary differential equation systems for 30 years. For
-similar reasons we also warp the Fortran libraries QUADPACK\cite{1983qspa.book.....P} (for numerical
+similar reasons we also wrap the Fortran libraries QUADPACK\cite{1983qspa.book.....P} (for numerical
 integration of one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103} (curve-fitting /
@@
 \subsection*{API and ABI evolution}
+In general, we encourage changes that improve clarity in the API of the library,
+but strongly discourage breaking backwards compatibility, given our position near
+the base of the scientific Python computing stack.
 
@@
 \subsection*{Statistical distributions}
+Version 1.0 of SciPy saw the addition of the energy distance algorithm in
+\texttt{scipy.stats.energy\_distance}. Energy distance is used in a varity of applications ranging
+from genetic analysis to multivariate independence. We also added the
+first Wasserstein distance algorithm to the library in
+\texttt{scipy.stats.wasserstein\_distance}. The Wasserstein distance is often
+used in color digital image analysis.
 
+We now also provide the cumulative distribution function of the multivariate
+normal distribution, \\\texttt{scipy.stats.multivariate\_normal}, via the newly
+associated \texttt{logcdf} and \texttt{cdf} methods.
+
+Other recent additions to \texttt{scipy.stats} include the Argus distribution
+in \texttt{scipy.stats.argus} (used in particle physics), a class for
+generating a distribution from a histogram in
+\texttt{scipy.stats.rv\_histogram}, an implementation of the multinomial
+distribution in \texttt{scipy.stats.multinomial} (used for example in natural
+language processing), a weighted Kendall's tau statistic algorithm in
+\texttt{scipy.stats.weightedtau}, interquantile range handling in
+\texttt{scipy.stats.iqr}, the Burr type XII distribution is available in
+\texttt{scipy.stats.burr12}, the skew normal distribution has been implemented 
+as \texttt{scipy.stats.skewnorm}, a trepzoidal continuous random variable
+has been added in \texttt{scipy.stats.trapz}, and the matrix normal distribution
+is now available in \texttt{scipy.stats.matrix\_normal}.
+
+From a design standpoint, many functions in \texttt{scipy.stats} now support
+a \texttt{nan\_policy} keyword to allow user-level control over the algorithmic
+decision to propagate, omit, or raise an error in the case that \texttt{NaN}
+values are encountered.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Polynomial interpolators}
+\input{poly}
 
@@
     with a 76.5 \% coverage of approximately $204000$ lines
-    of code. Documentation for the code is automatically built and hosted by the
-    CircleCI service to facilitate evaluation of documentation changes / integrity.
+    of code holding steady for the last six months (Figure~\ref{fig:coverage}).
+    Some of the historical components of the code base would still benefit from
+    increased test coverage. Documentation for the code is automatically built and hosted 
+    by the CircleCI service to facilitate evaluation of documentation changes / integrity.
     Our full test suite also passes with PyPy3, a just-in-time compiled version
     of the Python language.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/coverage-chart}
+\caption{\% coverage of the SciPy code base by unit tests over the six
+	 months preceding June 28, 2018 as reported by the \texttt{codecov}
+	 service. The analysis of lines of source code covered is performed
+	 automatically in our continuous integation suite using the
+	 \texttt{pytest-cov} and \texttt{gcov} libraries.}
+\label{fig:coverage}
+\end{figure}
+
     % NOTE: is there a citation for PyPy?
@@
 \textbf{Roadmap}
+
+The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
+maintained by the community that describes some of the major directions
+for improvement for the project, as well as specific limitations and
+matters that require assistance moving forward.
+
+We are still striving to increase the number of SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}, but our
+standard documentation for library features is already in excellent shape.
+
+The low-level Cython code in our library (which interacts with C-level
+code and exposes it for Python usage) could use some measure of modernization,
+including migration to typed memoryviews to handle NumPy arrays instead of
+the older syntax:
+
+\begin{lstlisting}[language=Python]
+# example of a function that takes a single
+# 3 dimensional typed memoryview as an argument
+# this allows Cython-level handling of a NumPy
+# array object
+cpdef int func_new(int[:, :, :] arr):
+    pass
+
+# example of the old syntax for handling
+# the same scenario
+cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    pass
+
+\end{lstlisting}
+
+\texttt{fftpack} and {linalg} have too much functionality overlap with
+\texttt{numpy} equivalents.
+
+\texttt{scipy.interpolate} could benefit from a set of new features including
+spline fitting routines with better user control; arithmetic routines for splines;
+transparent tensor-product splines, Non-Uniform Rational B-Splines support;
+and mesh refinement and coarsening of B-splines and corresponding tensor products.
+Some spline functionality in \texttt{scipy.signal} should be migrated to \texttt{interpolate} 
+and Second Order Sections need to be updated to match the capabilities in other routines.
+
+For \texttt{scipy.linalg}, the major change needed is to support a more recent
+version of LAPACK. Users should be able access an increasing repertoire of
+LAPACK functions using either \texttt{get\_lapack\_funcs} or
+\texttt{sp.linalg.lapack.name\_of\_function}.
+
+\texttt{scipy.optimize} would benefit from a few more good global optimizers 
+as well as large-scale optimizers.
+
+\texttt{scipy.ndimage} should be completely migrated to a data point
+model (values on a grid), instead of a pixel model (elements with centers),
+both of which have been used in the past, because the data point model is now
+proving more effective in practice.
+
+While \texttt{scipy.sparse} is mature, there are many data
+formats and the community continues to move forward here so we will aim
+to take advantage of their improvements\cite{Sparse_Arrays} 
+while preserving backwards compatibility to the extent that it is possible.
+\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
+for faster sparse SVD computation.
+
+\texttt{scipy.spatial} will benefit from \texttt{spatial.transforms} with
+support for rotation matrices.
+
+\texttt{scipy.special} needs precision improvements for hypergeometric 
+functions, parabolic cylinder functions, and spheroidal wave functions.
 

### 2018-07-10 scipy-1.0/ckdtree.tex
Drafted ckdtree sparse distance matrix section.

@@
 
-The API for querying neighbours is enhanced, to allow querying a subset of
-k nearest neighbours.
+% TODO: please check / improve literature citations used here
+% what about the time complexity of the kNN implementation?
+In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
+parameter. This is an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
+because it generates and returns a data structure that may be only a fraction 
+of the size of the full neighbor list. A C++ implementation is provided that releases
+the Python global interpreter lock for the neighbor search process, uses
+a memory pool to allocate and automatically reclaim \texttt{structs}, and
+heapsort to sort the $k$ nearest neighbors distances, even in the case
+of periodic boundary conditions. When $k$ is a non-contiguous list of nearest
+neighbors to account for (i.e., \texttt{[1, 3, 4]}), the intervening distances
+between the nearest neighbor ($k = 1$) and the farthest neighbor requested
+are all interrogated, but only those neighbors that are requested are retained
+in memory. Thus, the maximum heap size for this algorithm is conceptually determined
+as \texttt{np.arange(max(k))}.
 
@@
 
-generating approximated sparse distance matrix
+In 2015 SciPy added support for generating approximated sparse distance matrices 
+between \texttt{KDTree} objects (sets of points organized into a binary space 
+partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
+The user provides a maximum distance value
+above which all distances between points in two provided \texttt{KDTree} objects
+are set to 0. The distance metric used between k-d trees is not constrained
+to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
+bounded by 1 and infinity is valid. By default, a dictionary of keys
+based sparse matrix is the returned data structure. Previously-published
+results had suggested that a hashing approach to sparse matrix assembly
+is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
+The dictionary object maps row and column indices to elements of the sparse matrix, which is
+efficient for matrix construction because all zero values are skipped in
+the dictionary mapping. The C++ level sparse matrix construction releases the Python
+global interpreter lock for increased performance. Once constructed, the
+dictionary of keys sparse matrix has the amortized constant time complexity 
+($O(1)$) for distance value retrieval that one would expect from a 
+dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
+operations on the sparse matrix, SciPy allows the dictionary of keys
+sparse matrix to be directly converted to the common CSR, CSC, and COO
+data structures.
 

### 2018-07-11 scipy-1.0/paper.tex
Revision of manuscript PR 37 based on reviewer feedback; added a few more citations to scipy.stats tech improvements section as well.

--- 2018-07-07 scipy-1.0/paper.tex
@@
 Version 1.0 of SciPy saw the addition of the energy distance algorithm in
-\texttt{scipy.stats.energy\_distance}. Energy distance is used in a varity of applications ranging
-from genetic analysis to multivariate independence. We also added the
+\texttt{scipy.stats.energy\_distance}. Energy distance is used in a variety
+of applications ranging from genetic analysis\cite{Hu2008DetectingIC} to 
+multivariate independence\cite{szekely2009}. We also added the
 first Wasserstein distance algorithm to the library in
 \texttt{scipy.stats.wasserstein\_distance}. The Wasserstein distance is often
-used in color digital image analysis.
+used in color digital image analysis \cite{192468}.
 
-We now also provide the cumulative distribution function of the multivariate
+We now provide the cumulative distribution function of the multivariate
 normal distribution, \\\texttt{scipy.stats.multivariate\_normal}, via the newly
-associated \texttt{logcdf} and \texttt{cdf} methods.
+associated \texttt{logcdf} and \texttt{cdf} methods. Other recent distributions 
+added to \texttt{scipy.stats} include the Argus distribution
+in \texttt{scipy.stats.argus} (used in particle physics\cite{ALBRECHT1990278}), a histogram-based
+distribution in \texttt{scipy.stats.rv\_histogram}, the multinomial
+distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
+language processing\cite{Griffiths5228}), the Burr type XII distribution in
+\texttt{scipy.stats.burr12}, the skew normal distribution 
+in \texttt{scipy.stats.skewnorm}, the trapezoidal distribution in 
+\texttt{scipy.stats.trapz}, and the matrix normal distribution
+in \texttt{scipy.stats.matrix\_normal}.
 
-Other recent additions to \texttt{scipy.stats} include the Argus distribution
-in \texttt{scipy.stats.argus} (used in particle physics), a class for
-generating a distribution from a histogram in
-\texttt{scipy.stats.rv\_histogram}, an implementation of the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used for example in natural
-language processing), a weighted Kendall's tau statistic algorithm in
-\texttt{scipy.stats.weightedtau}, interquantile range handling in
-\texttt{scipy.stats.iqr}, the Burr type XII distribution is available in
-\texttt{scipy.stats.burr12}, the skew normal distribution has been implemented 
-as \texttt{scipy.stats.skewnorm}, a trepzoidal continuous random variable
-has been added in \texttt{scipy.stats.trapz}, and the matrix normal distribution
-is now available in \texttt{scipy.stats.matrix\_normal}.
+There's also a new weighted Kendall's tau statistic algorithm in
+\texttt{scipy.stats.weightedtau}, and interquantile range handling in
+\texttt{scipy.stats.iqr}
 

### 2018-07-13 scipy-1.0/ckdtree.tex
First draft of cKDTree KNN algorithm / tech improvement section.

--- 2018-06-27 scipy-1.0/ckdtree.tex
@@
 
-The API for querying neighbours is enhanced, to allow querying a subset of
-k nearest neighbours.
+% TODO: please check / improve literature citations used here
+% what about the time complexity of the kNN implementation?
+In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
+parameter. This may be an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
+because it generates and returns a data structure that may be only a fraction 
+of the size of the full neighbor list. A C++ implementation is provided that releases
+the Python global interpreter lock for the neighbor search process, uses
+a memory pool to allocate and automatically reclaim \texttt{structs}, and
+heapsort to sort the $k$ nearest neighbors distances, even in the case
+of periodic boundary conditions. When $k$ is a non-contiguous list of nearest
+neighbors to account for (i.e., \texttt{[1, 3, 4]}), the intervening distances
+between the nearest neighbor ($k = 1$) and the farthest neighbor requested
+are all interrogated, but only those neighbors that are requested are retained
+in memory. Thus, the max heap size for this algorithm is conceptually determined
+as \texttt{np.arange(max(k))}.
 

### 2018-07-14 scipy-1.0/ckdtree.tex
First round of revisions based on reviewer feedback in manuscript PR 39.

@@
 In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
-parameter. This may be an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
+parameter. This is an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
 because it generates and returns a data structure that may be only a fraction 
@@
 are all interrogated, but only those neighbors that are requested are retained
-in memory. Thus, the max heap size for this algorithm is conceptually determined
+in memory. Thus, the maximum heap size for this algorithm is conceptually determined
 as \texttt{np.arange(max(k))}.

### 2018-07-14 scipy-1.0/poly.tex
ENH: first draft of the `interpolate` subsection

Traditionally, SciPy was heavily relying on the venerable \texttt{FITPACK} Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for univariate interpolation and approximation of data. FITPACK provides a collection of routines for calculating smoothing splines and automatic knot selection. One issue with FITPACK and its Python wrappers, \code{UnivariateSpline} and \code{splrep}/\code{splev} combo, is the monolithic design: specific algorithms for construction of interpolating and smoothing splines and knot selection are tightly coupled at the Fortran level. The library offers very limited degree of user control over the details of the spline fitting it performs. 
%
An alternative set of spline interpolation routines was implemented in the (currently deprecated) \code{splmake}/\code{spleval} combo, and used as a backend of the \code{interp1d} routine. To further complicate matters, the spline format used by \code{splmake}/\code{spleval} was incompatible with \code{splrep}/\code{splev}.
%
Routines for manipulating piecewise polynomials and B-splines were either not available or lacking in ease of use and/or performance characteristics.

Implementing a new, modular design of polynomial interpolators was spread over several releases. The goal of this effort was (i) to have a set of basic objects representing piecewise polynomials, (ii) to implement a collection of algorithms for constructing various interpolators, and (iii) to provide users with building blocks for constructing additional interpolators.

At the lowest level of the new design are classes which represent univariate piecewise polynomials: \code{PPoly}, \code{BPoly} and \code{BSpine}, which allow efficient vectorized evaluations, differentiation and integration.
\code{PPoly} represents piecewise polynomials in the power basis in terms of breakpoints and coefficients at each interval. \code{BPoly} is similar, and  represents piecewise polynomials in the Bernstein basis (which is suitable for \textit{e.g.,} constructing Bezier curves). \code{BSpline} represents spline curves, \textit{i.e.,} linear combinations of b-spline basis elements.
\cite{deBoor1978} 
The spline representation used by \code{BSpline} is directly compatible with FITPACK.

In the next layer, these polynomial classes are used for constructing several common ways of interpolating data: \code{CubicSpline} constructs a twice differentiable piecewise cubic function, \code{Akima1DInterpolator} and \code{PCHIPInterpolator} implement two classic prescriptions for constructing a $C^1$ continuous monotone shape-preserving interpolator. \cite{FritschCarlson1980, Akima1970}

Interpolation with known or estimated derivatives, \textit{i.e.,} cubic or higher-order Hermite splines, can be constructed via either subclassing \code{PPoly} or \code{BPoly}, or using the factory method \code{BPoly.from\_derivatives}.

For working with b-splines, there are several options. Using \code{UnivariateSpline} in new code is not recommended. The basic object to represent b-splines is \code{BSpline}, and its use should be preferred over direct manipulation of \code{tck} tuples of knots, coefficients and the spline degree. Given data points, the helper function \code{make\_interp\_spline} constructs a \code{BSpline} instance, representing a spline which interpolates the data. Interpolated values at intermediate points can be computed either via its \code{\_\_call\_\_} method, or via \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine accepts both \code{tck} tuples and \code{Bspline} instances.

Fitting \emph{smoothing} splines to data can currently only be done with FITPACK. Building alternative algorithms is certainly possible, but is still to be implemented.

### 2018-07-17 scipy-1.0/poly.tex
MAINT: address review comments in the `interpolate` subsection

@@
-Traditionally, SciPy was heavily relying on the venerable \texttt{FITPACK} Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for univariate interpolation and approximation of data.
-FITPACK provides a collection of routines for calculating smoothing splines and automatic knot selection.
-One issue with FITPACK and its Python wrappers, \code{UnivariateSpline} and \code{splrep}/\code{splev} combo, is the monolithic design: specific algorithms for construction of interpolating and smoothing splines and knot selection are tightly coupled at the Fortran level.
-The library offers very limited degree of user control over the details of the spline fitting it performs.
+Traditionally, SciPy relied heavily on the venerable \texttt{FITPACK}
+Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for
+univariate interpolation and approximation of data. FITPACK provides a
+collection of routines for calculating smoothing splines and automatic knot
+selection. One issue with FITPACK and its Python wrappers, \code{UnivariateSpline} and \code{splrep}/\code{splev} combo, is the monolithic design: specific
+algorithms for construction of interpolating and smoothing splines and knot
+selection are tightly coupled at the Fortran level. The library offers very
+limited degree of user control over the details of the spline fitting it performs. 
 %
-An alternative set of spline interpolation routines was implemented in the (currently deprecated) \code{splmake}/\code{spleval} combo, and used as a backend of the \code{interp1d} routine.
-To further complicate matters, the spline format used by \code{splmake}/\code{spleval} was incompatible with \code{splrep}/\code{splev}.
+An alternative set of spline interpolation routines was implemented in the
+(now deprecated) \code{splmake}/\code{spleval} combo, and used as a backend of
+the \code{interp1d} routine. To further complicate matters, the spline format
+used by \code{splmake}/\code{spleval} was incompatible with
+\code{splrep}/\code{splev}.
 %
-Routines for manipulating piecewise polynomials and B-splines were either not available or lacking in ease of use and/or performance characteristics.
+Routines for manipulating piecewise polynomials and b-splines were either not available or lacking in ease of use and/or performance characteristics.
 
-Implementing a new, modular design of polynomial interpolators was spread over several releases.
-The goal of this effort was (i) to have a set of basic objects representing piecewise polynomials, (ii) to implement a collection of algorithms for constructing various interpolators, and (iii) to provide users with building blocks for constructing additional interpolators.
+Implementing a new, modular design of polynomial interpolators was spread over
+several releases. The goals of this effort were to have a set of basic objects
+representing piecewise polynomials, to implement a collection of algorithms
+for constructing various interpolators, and to provide users with building
+blocks for constructing additional interpolators.
 
-At the lowest level of the new design are classes which represent univariate piecewise polynomials: \code{PPoly}, \code{BPoly} and \code{BSpine}, which allow efficient vectorized evaluations, differentiation and integration.
-\code{PPoly} represents piecewise polynomials in the power basis in terms of breakpoints and coefficients at each interval.
-\code{BPoly} is similar, and  represents piecewise polynomials in the Bernstein basis (which is suitable for \textit{e.g.,} constructing Bezier curves).
-\code{BSpline} represents spline curves, \textit{i.e.,} linear combinations of b-spline basis elements.
-\cite{deBoor1978} 
-The spline representation used by \code{BSpline} is directly compatible with FITPACK.
+At the lowest level of the new design are classes which represent univariate
+piecewise polynomials: \code{PPoly}, \code{BPoly} and \code{BSpline}, which allow
+efficient vectorized evaluations, differentiation, integration and root-finding.
+\code{PPoly} represents piecewise polynomials in the power basis in terms of
+breakpoints and coefficients at each interval. \code{BPoly} is similar, and
+represents piecewise polynomials in the Bernstein basis (which is suitable
+for e.g., constructing Bezier curves).
 
@@
 \cite{FritschCarlson1980, Akima1970}
-
-Interpolation with known or estimated derivatives, \textit{i.e.,} cubic or higher-order Hermite splines, can be constructed via either subclassing \code{PPoly} or \code{BPoly}, or using the factory method \code{BPoly.from\_derivatives}.
+%
+Interpolation with known or estimated derivatives, i.e., cubic or higher-order Hermite splines, can be constructed by either subclassing \code{PPoly} or \code{BPoly}, or using the factory method \code{BPoly.from\_derivatives}.
 
 For working with b-splines, there are several options.
-Using \code{UnivariateSpline} in new code is not recommended.
-The basic object to represent b-splines is \code{BSpline}, and its use should be preferred over direct manipulation of \code{tck} tuples of knots, coefficients and the spline degree.
-Given data points, the helper function \code{make\_interp\_spline} constructs a \code{BSpline} instance, representing a spline which interpolates the data.
-Interpolated values at intermediate points can be computed either via its \code{\_\_call\_\_} method, or via \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine accepts both \code{tck} tuples and \code{Bspline} instances.
+The basic object is \code{BSpline}, which represents spline curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}.
+The spline
+representation used by \code{BSpline} is directly compatible with FITPACK. We
+recommend that users work with \code{BSpline} objects instead of directly
+manipulating \code{tck} tuples of knots, coefficients and the spline degree.
+%
+Using \code{UnivariateSpline} in new code is not recommended. Given data points,
+the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
+instance, representing a spline which interpolates the data. Interpolated values
+at intermediate points can be computed either via its \code{\_\_call\_\_} method,
+or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
+accepts both \code{tck} tuples and \code{Bspline} instances.
 

### 2018-07-24 scipy-1.0/paper.tex
ENH: first draft of the `interpolate` subsection
Start rough draft of sparse matrices tech improvements section.

--- 2018-07-07 scipy-1.0/paper.tex
@@
 \author[23]{Antonio Horta Ribeiro}
+\author[24]{Ian Henriksen}
 
@@
 \affil[23]{Affiliation, department, city, postcode, country}
+\affil[24]{University of Texas at Austin,
+           Instite for Computational Engineering and Sciences,
+	   Austin, TX, 78712, USA}
 
@@
 
-\textbf{Project goals and scope}
+\textbf{Project scope}
+SciPy aims to provide fundamental algorithms for scientific computing. The
+breadth of its scope was derived from the Guide to Available Mathematical
+Software classification system (GAMS\cite{boisvert1991guide}). In areas
+that move relatively slowly, e.g. linear algebra, SciPy aims to provide
+complete coverage. In other areas it aims to provide fundamental building
+blocks while interacting well with other packages specialized in that area.
+For example, SciPy provides what one expect to find in a
+statistics textbook (probability distributions, hypothesis tests, frequency
+statistics, correlation functions, and more), while Statsmodels provides
+more advanced statistical estimators and inference methods, scikit-learn
+covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
+and probabilistic modelling.
+
+We use the following criteria to determine whether or not to include new
+functionality in SciPy:
+\begin{itemize}
+    \item The algorithm is of relevance to multiple fields of science.
+    \item The algorithm is demonstrably important.  E.g., it is classic
+    enough to be included in textbooks, or it is based on a peer-reviewed article
+    which has a significant number of citations.
+\end{itemize}
+
+In terms of software systems and architecture, SciPy's scope matches NumPy's:
+algorithms for in-memory computing on single machines, with support for a wide
+range of data types and process architectures. Distributed computing and support
+for graphics processing units (GPUs) is explicitly out of scope.
 
@@
 \subsection*{API and ABI evolution}
-In general, we encourage changes that improve clarity in the API of the library,
-but strongly discourage breaking backwards compatibility, given our position near
-the base of the scientific Python computing stack.
+The application programming interface (API) for SciPy consists of
+approximately 1500 functions and classes.
+Our policy for evolving the API over time is that new functionality can be added,
+while removing or changing existing functionality can only be done if the benefits
+of that exceeds the (often significant) costs to users, \textit{and} only
+after giving clear deprecation warnings to those users for at least one year.
+In general, we encourage changes that improve clarity in the API of the
+library but strongly discourage breaking backwards compatibility\footnote{
+    Backwards compatibility means that new versions of SciPy will keep user
+    code written for an older version working.}
+given our position near the base of the scientific Python computing stack.
+
+In addition to the Python API, SciPy has C and Cython interfaces in a number
+of submodules. Therefore, we have to also consider the application binary
+interface (ABI). This ABI has been stable for a long time, and we aim to
+evolve it only in a backwards compatible way.
 
@@
 \textbf{Sparse matrices}
+Over the last three years our sparse matrix handling internals have
+been rewritten and performance has been improved.
 
@@
 
+When writing algorithms or wrappers in Cython\cite{behnel2011cython}, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
+SciPy currently includes Cython wrappers\cite{blas-lapack-wrappers-scipy-2015} for many BLAS and LAPACK\cite{LAPACK} routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls{\textemdash}this can give performance gains of one or two orders of magnitude for many use cases.
+Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.
+Function pointers to wrappers for library routines can also be shared in this way, so that Cython modules using the interface provided by SciPy no longer have to worry about finding and linking against the original library.
+This lets other extensions avoid the complexity of finding and using the correct libraries on the various possible Fortran compiler ABI and OS combinations.
+The low-level Cython wrappers provide an OS and Fortran compiler independent interface for the libraries they wrap.
+
+To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
+The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
+The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
+The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
+
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
+In other words, wrappers are provided for routines that are included in the earliest supported LAPACK version and are not removed or changed in some incompatible way in some subsequent LAPACK release.
+The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers we provide.
+Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
+
 \subsection*{Numerical optimization}
@@
 \subsection*{Statistical distributions}
-Version 1.0 of SciPy saw the addition of the energy distance algorithm in
-\texttt{scipy.stats.energy\_distance}. Energy distance is used in a varity of applications ranging
-from genetic analysis to multivariate independence. We also added the
-first Wasserstein distance algorithm to the library in
-\texttt{scipy.stats.wasserstein\_distance}. The Wasserstein distance is often
-used in color digital image analysis.
-
-We now also provide the cumulative distribution function of the multivariate
-normal distribution, \\\texttt{scipy.stats.multivariate\_normal}, via the newly
-associated \texttt{logcdf} and \texttt{cdf} methods.
-
-Other recent additions to \texttt{scipy.stats} include the Argus distribution
-in \texttt{scipy.stats.argus} (used in particle physics), a class for
-generating a distribution from a histogram in
-\texttt{scipy.stats.rv\_histogram}, an implementation of the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used for example in natural
-language processing), a weighted Kendall's tau statistic algorithm in
-\texttt{scipy.stats.weightedtau}, interquantile range handling in
-\texttt{scipy.stats.iqr}, the Burr type XII distribution is available in
-\texttt{scipy.stats.burr12}, the skew normal distribution has been implemented 
-as \texttt{scipy.stats.skewnorm}, a trepzoidal continuous random variable
-has been added in \texttt{scipy.stats.trapz}, and the matrix normal distribution
-is now available in \texttt{scipy.stats.matrix\_normal}.
-
-From a design standpoint, many functions in \texttt{scipy.stats} now support
-a \texttt{nan\_policy} keyword to allow user-level control over the algorithmic
-decision to propagate, omit, or raise an error in the case that \texttt{NaN}
-values are encountered.
+The \texttt{scipy.stats} module provides provides more than 100 probability
+distributions and a framework to implement additional distributions. Given
+an expression for a probability distribution function (pdf),
+the framework will calculate moments, cumulative distribution function (cdf),
+differential entropy, and more. Those properties are calculated based on
+generic code, e.g., numerical integration of the pdf to obtain the cdf.
+Where possible, the generic code is replaced by explicit formulas or more
+accurate or efficient implementations. For example, most distributions have
+an explicit formula for the cdf, so it is used instead of integrating the
+pdf. The distributions framework also allows the user to draw random
+variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
+parameter, which is either a \texttt{numpy.random.RandomState} object or an
+integer used to generate such an object, to provide fine-grained user-level
+control over the random state of a workflow.
+Currently about 80 continuous univariate distributions, 12 discrete univariate
+distributions and 10 multivariate distributions are implemented.
+Key recent distributions added to \texttt{scipy.stats} include the histogram-based
+distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
+distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
+language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
+far the only multivariate discrete distribution available in SciPy.
 
@@
 formats and the community continues to move forward here so we will aim
-to take advantage of their improvements\cite{Sparse_Arrays} 
+to take advantage of their improvements\cite{abbasi2018sparse} 
 while preserving backwards compatibility to the extent that it is possible.
@@
 \textbf{Maintainers and contributors}
+The SciPy project has approximately 15 active maintainers - people who review
+others' contributions, and in general do everything needed to ensure that
+the software and the project move forward. Maintainers are critical to the
+health of the project\cite{eghbal2016}; their skills and efforts largely
+determine how fast the project moves forward, and they enable contributions
+from a much larger group. Accordingly, the project also has an additional
+~100 unique contributors for every 6-monthly release cycle. Anyone with the
+interest and skills can become a contributor or maintainer; the SciPy
+Developer Guide provides guidance on how to do that.
 
@@
 
-% need some better usage stats sources?
-% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
-From the user side, there have been at least 15 million downloads
-of the SciPy manylinux wheels between the start of 2016 and May of 2018
-(njs blog). The recommended source for citing SciPy is simply a reference
-to the website / library itself\cite{SciPylib}, and has been cited $>3000$
-times. Some of the most prominent citing articles include the IPython
-publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
-article\cite{pmid17287757}, a dark
-matter annihilitation study in the field of
-cosmology\cite{1475-7516-2012-08-007}, and many of the
-other major downstream software libraries in the Python scientific
-computing stack.
+From the user side, there were 13,096,468 PyPI downloads
+of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
+underestimate of total downloads given that we don't have access to the
+popular Anaconda package manager download statistics. The recommended 
+source for citing SciPy is simply a reference to the website / library
+itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
+prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
+scientific collaboration that lead to the observation of gravitational
+waves \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly
+with MacOS and in the Intel Distribution for Python\cite{intel-python}.
 
 \textbf{Future development}.
+Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}.
+This is due to the current interface only supporting two-dimensional objects.
+Even sub-arrays are two-dimensional, and dimensions greater than two aren't supported.
+The new package supports dimensions greater than two, and it bases itself around arrays rather than matrices (which is strictly more powerful).
+It also features improved inter-operability with NumPy and other community packages.
+
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}

### 2018-07-24 scipy-1.0/paper.tex
Third round of revisions for manuscript PR 34 based on reviewer feedback.

--- 2018-07-11 scipy-1.0/paper.tex
@@
 \pgfplotsset{compat=newest}
-\usepackage{listings, textcomp}
-
-\lstset{ %
-  basicstyle=\ttfamily\footnotesize,  % size of fonts used for the code
-  breaklines=true,   % automatic line breaking only at whitespace
-  captionpos=b,   % sets the caption-position to bottom
-  commentstyle=\color{gray},  % comment style
-  keywordstyle=\color{blue},  % keyword style
-  stringstyle=\color{red},  % string literal style
-  upquote=true  %straight single quotes (requires textcomp)
-}
+\usepackage{listings}
 
@@
 \subsection*{Statistical distributions}
-Version 1.0 of SciPy saw the addition of the energy distance algorithm in
-\texttt{scipy.stats.energy\_distance}. Energy distance is used in a variety
-of applications ranging from genetic analysis\cite{Hu2008DetectingIC} to 
-multivariate independence\cite{szekely2009}. We also added the
-first Wasserstein distance algorithm to the library in
-\texttt{scipy.stats.wasserstein\_distance}. The Wasserstein distance is often
-used in color digital image analysis \cite{192468}.
-
-We now provide the cumulative distribution function of the multivariate
-normal distribution, \\\texttt{scipy.stats.multivariate\_normal}, via the newly
-associated \texttt{logcdf} and \texttt{cdf} methods. Other recent distributions 
-added to \texttt{scipy.stats} include the Argus distribution
-in \texttt{scipy.stats.argus} (used in particle physics\cite{ALBRECHT1990278}), a histogram-based
-distribution in \texttt{scipy.stats.rv\_histogram}, the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
-language processing\cite{Griffiths5228}), the Burr type XII distribution in
-\texttt{scipy.stats.burr12}, the skew normal distribution 
-in \texttt{scipy.stats.skewnorm}, the trapezoidal distribution in 
-\texttt{scipy.stats.trapz}, and the matrix normal distribution
-in \texttt{scipy.stats.matrix\_normal}.
-
-There's also a new weighted Kendall's tau statistic algorithm in
-\texttt{scipy.stats.weightedtau}, and interquantile range handling in
-\texttt{scipy.stats.iqr}
-
-From a design standpoint, many functions in \texttt{scipy.stats} now support
-a \texttt{nan\_policy} keyword to allow user-level control over the algorithmic
-decision to propagate, omit, or raise an error in the case that \texttt{NaN}
-values are encountered.
 
@@
 # example of a function that takes a single
-# 3 dimensional typed memoryview as an argument
+# 3 dimesional typed memoryview as an argument
 # this allows Cython-level handling of a NumPy

### 2018-07-25 scipy-1.0/paper.tex

@@
 \pgfplotsset{compat=newest}
-\usepackage{listings}
+\usepackage{listings, textcomp}
+
+\lstset{ %
+  basicstyle=\ttfamily\footnotesize,  % size of fonts used for the code
+  breaklines=true,   % automatic line breaking only at whitespace
+  captionpos=b,   % sets the caption-position to bottom
+  commentstyle=\color{gray},  % comment style
+  keywordstyle=\color{blue},  % keyword style
+  stringstyle=\color{red},  % string literal style
+  upquote=true  %straight single quotes (requires textcomp)
+}
 
@@
 # example of a function that takes a single
-# 3 dimesional typed memoryview as an argument
+# 3 dimensional typed memoryview as an argument
 # this allows Cython-level handling of a NumPy

### 2018-07-25 scipy-1.0/paper.tex
Revised / condensed based on feedback in manuscript PR 37.

@@
 
+We now provide the cumulative distribution function (using the Fortran code by
+Alan Genz) of the multivariate normal distribution, \\\texttt{scipy.stats.multivariate\_normal},
+via the newly associated \texttt{logcdf} and \texttt{cdf} methods. Other key recent distributions 
+added to \texttt{scipy.stats} include the histogram-based
+distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
+distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
+language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
+far the only multivariate discrete distribution available in SciPy.
+
+From a design standpoint, many functions in \texttt{scipy.stats} now support
+a \texttt{nan\_policy} keyword to allow user-level control over the algorithmic
+decision to propagate, omit, or raise an error in the case that \texttt{NaN}
+values are encountered. Much of the functionality offered by
+\texttt{scipy.stats} for drawing random variates now accepts a
+\texttt{random\_state} parameter, which is either a \texttt{numpy.random.RandomState}
+object or an integer used to generate such an object, to provide fine-grained
+user-level control over the random state of a workflow.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-07-25 scipy-1.0/paper.tex
Add section on API/ABI evolution.

--- 2018-07-25 scipy-1.0/paper.tex
@@
 \subsection*{API and ABI evolution}
-In general, we encourage changes that improve clarity in the API of the library,
-but strongly discourage breaking backwards compatibility, given our position near
-the base of the scientific Python computing stack.
+The application programming interface (API) for SciPy consists of
+approximately 2600 functions and classes. Adding new functionality has to
+meet the following criteria:
+\begin{itemize}
+    \item The algorithm is of relevance to multiple fields of science.
+    \item The implementation is demonstrably important.  E.g., it is classic
+    enough to be included in textbooks, or it is based on a peer-reviewed article
+    which has a significant number of citations.
+\end{itemize}
+
+Our policy for evolving the API over time is that new functionality can be added,
+while removing or changing existing functionality can only be done if the benefits
+of that exceeds the (often significant) costs to users, \textit{and} only
+after giving clear deprecation warnings to those users for at least one year.
+In general, we encourage changes that improve clarity in the API of the
+library but strongly discourage breaking backwards compatibility\footnote{
+    Backwards compatibility means that new versions of SciPy will keep user
+    code written for an older version working.}
+given our position near the base of the scientific Python computing stack.
+
+In addition to the Python API, SciPy has C and Cython interfaces in a number
+of submodules. Therefore, we have to also consider the application binary
+interface (ABI). This ABI has been stable for a long time, and we aim to
+evolve it only in a backwards compatible way.
 

### 2018-07-25 scipy-1.0/paper.tex
Drafted two short paragraphs for sparse tech improvements section.

@@
 The application programming interface (API) for SciPy consists of
-approximately 2600 functions and classes. Adding new functionality has to
-meet the following criteria:
-\begin{itemize}
-    \item The algorithm is of relevance to multiple fields of science.
-    \item The implementation is demonstrably important.  E.g., it is classic
-    enough to be included in textbooks, or it is based on a peer-reviewed article
-    which has a significant number of citations.
-\end{itemize}
-
+approximately 1500 functions and classes.
 Our policy for evolving the API over time is that new functionality can be added,
@@
 
+Over the last three years our sparse matrix handling internals have
+been rewritten and performance has been improved. Iterating over 
+and slicing of CSC and CSR matrices is now faster by up to 35\%,
+and the CSR to COO matrix format conversion is now several times faster.
+COO matrices are now generally more memory efficient. Importantly,
+SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
+the low-level implementations leveraged by a subset of our \texttt{sparse}
+offerings.
+
+From a new features standpoint, \texttt{scipy.sparse} matrices and linear
+operators now support the Python matrix multiplication (@) operator when
+available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
+for computing sparse matrix norms and drawing random variates from arbitrary 
+distributions, respectively. Users can now easily serialize a number of our
+sprase matrix formats with the newly associated \texttt{save\_npz} function.
+We've enhanced the diagonal selection API for sparse matrices to allow
+for more user control.
+
 \subsection*{Unified bindings to compiled code}
@@
 \subsection*{Statistical distributions}
+The \texttt{scipy.stats} module provides provides more than 100 probability
+distributions and a framework to implement additional distributions. Given
+an expression for a probability distribution function (pdf),
+the framework will calculate moments, cumulative distribution function (cdf),
+differential entropy, and more. Those properties are calculated based on
+generic code, e.g., numerical integration of the pdf to obtain the cdf.
+Where possible, the generic code is replaced by explicit formulas or more
+accurate or efficient implementations. For example, most distributions have
+an explicit formula for the cdf, so it is used instead of integrating the
+pdf. The distributions framework also allows the user to draw random
+variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
+parameter, which is either a \texttt{numpy.random.RandomState} object or an
+integer used to generate such an object, to provide fine-grained user-level
+control over the random state of a workflow.
+Currently about 80 continuous univariate distributions, 12 discrete univariate
+distributions and 10 multivariate distributions are implemented.
+Key recent distributions added to \texttt{scipy.stats} include the histogram-based
+distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
+distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
+language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
+far the only multivariate discrete distribution available in SciPy.
 

### 2018-07-25 scipy-1.0/paper.tex
Describe the stats.distributions framework.

--- 2018-07-25 scipy-1.0/paper.tex
@@
 \subsection*{Statistical distributions}
-
-We now provide the cumulative distribution function (using the Fortran code by
-Alan Genz) of the multivariate normal distribution, \\\texttt{scipy.stats.multivariate\_normal},
-via the newly associated \texttt{logcdf} and \texttt{cdf} methods. Other key recent distributions 
-added to \texttt{scipy.stats} include the histogram-based
+The \texttt{scipy.stats} module provides a framework to implement probability
+distributions.  Given an expression for a probability distribution function (pdf),
+the framework will calculate moments, cumulative distribution function (cdf),
+differential entropy, and more. Those properties are calculated based on
+generic code, e.g., numerical integration of the pdf to obtain the cdf. In cases
+where that generic code does not provide sufficient accuracy, it is replaced
+by an implemented specific to the distribution.
+The distributions framework also allows the user to draw random
+variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
+parameter, which is either a \texttt{numpy.random.RandomState} object or an
+integer used to generate such an object, to provide fine-grained user-level
+control over the random state of a workflow.
+Currently about 80 continuous univariate distributions, 12 discrete univariate
+distributions and 10 multivariate distributions are implemented.
+Key recent distributions added to \texttt{scipy.stats} include the histogram-based
 distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
@@
 far the only multivariate discrete distribution available in SciPy.
-
-From a design standpoint, many functions in \texttt{scipy.stats} now support
-a \texttt{nan\_policy} keyword to allow user-level control over the algorithmic
-decision to propagate, omit, or raise an error in the case that \texttt{NaN}
-values are encountered. Much of the functionality offered by
-\texttt{scipy.stats} for drawing random variates now accepts a
-\texttt{random\_state} parameter, which is either a \texttt{numpy.random.RandomState}
-object or an integer used to generate such an object, to provide fine-grained
-user-level control over the random state of a workflow.
 

### 2018-07-25 scipy-1.0/paper.tex

@@
 \subsection*{Statistical distributions}
-The \texttt{scipy.stats} module provides a framework to implement probability
-distributions.  Given an expression for a probability distribution function (pdf),
-the framework will calculate moments, cumulative distribution function (cdf),
-differential entropy, and more. Those properties are calculated based on
-generic code, e.g., numerical integration of the pdf to obtain the cdf. In cases
-where that generic code does not provide sufficient accuracy, it is replaced
-by an implemented specific to the distribution.
-The distributions framework also allows the user to draw random
-variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
-parameter, which is either a \texttt{numpy.random.RandomState} object or an
-integer used to generate such an object, to provide fine-grained user-level
-control over the random state of a workflow.
-Currently about 80 continuous univariate distributions, 12 discrete univariate
-distributions and 10 multivariate distributions are implemented.
-Key recent distributions added to \texttt{scipy.stats} include the histogram-based
-distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
-language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
-far the only multivariate discrete distribution available in SciPy.
 

### 2018-07-25 scipy-1.0/ckdtree.tex
Fuse subsections for kdtree tech improvements; add a few refs.e

--- 2018-07-10 scipy-1.0/ckdtree.tex
@@
-The cKDTree module is rewritten in C++ with templated classes.
-
-The support to a box of periodic boundary condition is added. A periodic
-condition is typically used in computer simulations of physical processes.
-
-\subsubsection{k-NN}
+The cKDTree module was rewritten in C++ with templated classes, and support for
+periodic boundary conditions was added. A periodic boundary condition is typically 
+used in computer simulations of physical processes.
 
@@
 
-\subsubsection{sparse distance matrix}
-
-In 2015 SciPy added support for generating approximated sparse distance matrices 
+Also in 2015, SciPy added support for generating approximated sparse distance matrices 
 between \texttt{KDTree} objects (sets of points organized into a binary space 
@@
 
-\subsubsection{Pair-Counting}
-
-The cKDTree module implements the dual tree counting algorithm in (cite).
-
-An improvement to the pair counting algorithm is added to improve the scale
-with the number of bins.
+The cKDTree module implements a dual tree counting algorithm\cite{pmlr-v28-curtin13},
+with an improvement to the pair counting algorithm to improve the scaling
+with the number of bins. The cKDTree can now be augmented by weights, with 
+weighted paircount essential in many scientific applications, e.g. computing 
+correlation function of galaxies\cite{0004-637X-750-1-38}.
 
 \textbf{Add a Figure to show the scaling, before and after.}
-
-The cKDTree can now be augmented by weights. Weighted paircount is essential
-in many scientific applications, e.g. computing correlation function of galaxies.
-
 \textbf{perhaps give an example or some formula.}
-
 (cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)

### 2018-07-25 scipy-1.0/paper.tex
Add future development section on pydata/sparse

@@
 \textbf{Future development}.
+Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}.
+This is due to the current interface only supporting two dimensional objects.
+Even sub-arrays are two dimensional, and dimensions greater than two aren't supported.
+The new package lacks performance, but it does support dimensions greater than two and a number of other useful features.
+
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}

### 2018-07-25 scipy-1.0/ckdtree.tex
Updated dual tree algorithm reference to match source code; fixed typo in ckdtree tech improvements section.

@@
 
-The cKDTree module implements a dual tree counting algorithm\cite{pmlr-v28-curtin13},
+The cKDTree module implements a dual tree counting algorithm\cite{Moore2000ar},
 with an improvement to the pair counting algorithm to improve the scaling
@@
 weighted paircount essential in many scientific applications, e.g. computing 
-correlation function of galaxies\cite{0004-637X-750-1-38}.
+correlation functions of galaxies\cite{0004-637X-750-1-38}.
 

### 2018-07-25 scipy-1.0/subpackages.tex
Add a bit of context to the submodule organization section.

@@
-The SciPy library is organized as a collection of subpackages.  Here we
-summarize the scope and capabilities of each subpackage.
+The SciPy library is organized as a collection of subpackages.  The 16
+subpackages include mathematical building blocks (e.g. linear algebra, Fourier
+transforms, special functions), data structures (e.g. sparse matrices, k-D trees),
+algorithms (e.g. numerical optimization and integration, clustering, interpolation,
+graph algorithms, computational geometry), and higher-level data analysis
+functionality (e.g. signal and image processing, statistical methods).
+
+Here we summarize the scope and capabilities of each subpackage.
 

### 2018-07-25 scipy-1.0/ckdtree.tex

--- 2018-07-14 scipy-1.0/ckdtree.tex
@@
 
-\subsubsection{k-NN}
+\subsubsection*{k-NN}
 
@@
 
-\subsubsection{sparse distance matrix}
+\subsubsection*{sparse distance matrix}
 
-generating approximated sparse distance matrix
+In 2015 SciPy added support for generating approximated sparse distance matrices 
+between \texttt{KDTree} objects (sets of points organized into a binary space 
+partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
+The user provides a maximum distance value
+above which all distances between points in two provided \texttt{KDTree} objects
+are set to 0. The distance metric used between k-d trees is not constrained
+to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
+bounded by 1 and infinity is valid. By default, a dictionary of keys
+based sparse matrix is the returned data structure. Previously-published
+results had suggested that a hashing approach to sparse matrix assembly
+is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
+The dictionary object maps row and column indices to elements of the sparse matrix, which is
+efficient for matrix construction because all zero values are skipped in
+the dictionary mapping. The C++ level sparse matrix construction releases the Python
+global interpreter lock for increased performance. Once constructed, the
+dictionary of keys sparse matrix has the amortized constant time complexity 
+($O(1)$) for distance value retrieval that one would expect from a 
+dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
+operations on the sparse matrix, SciPy allows the dictionary of keys
+sparse matrix to be directly converted to the common CSR, CSC, and COO
+data structures.
 
-\subsubsection{Pair-Counting}
+\subsubsection*{Pair-Counting}
 

### 2018-07-25 scipy-1.0/paper.tex

--- 2018-07-25 scipy-1.0/paper.tex
@@
 \textbf{Future development}.
+Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}.
+This is due to the current interface only supporting two-dimensional objects.
+Even sub-arrays are two-dimensional, and dimensions greater than two aren't supported.
+The new package supports dimensions greater than two, and it bases itself around arrays rather than matrices (which is strictly more powerful).
+It also features improved inter-operability with NumPy and other community packages.
+
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}

### 2018-07-25 scipy-1.0/paper.tex

@@
 }
+
+\usepackage[dvipsnames]{xcolor}
+\usepackage{todonotes}
+\presetkeys%
+    {todonotes}%
+    {inline,linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum}{}
 

### 2018-07-25 scipy-1.0/supporting_info/measure_scipy_API_surface.py
Add a script to count the number of public objects in SciPy submodules.
This is not perfect (e.g. doesn't count the few things in the main `scipy`
namespace, probably does a bit of double counting of stuff), but it should
be good enough to give an estimate.

@@ -1,0 +1,84 @@
+"""
+This script imports all the public submodules and estimates the number
+of public objects in each of them.  This can be used to give a number
+for the total API surface of SciPy, e.g. "2600 functions and classes".
+
+"""
+
+import scipy
+from scipy import (
+    cluster,
+    constants,
+    fftpack,
+    integrate,
+    interpolate,
+    io,
+    linalg,
+    misc,
+    ndimage,
+    odr,
+    optimize,
+    signal,
+    sparse,
+    spatial,
+    special,
+    stats,
+)
+
+from scipy.cluster import vq, hierarchy
+from scipy.io import arff, harwell_boeing, idl, netcdf, matlab, wavfile
+from scipy.linalg import blas, lapack, interpolative
+from scipy.signal import windows
+from scipy.sparse import linalg as sls
+from scipy.sparse import csgraph
+from scipy.spatial import distance, transform
+from scipy.stats import distributions, mstats
+
+
+modules = [
+    scipy.cluster,
+    scipy.cluster.vq,
+    scipy.cluster.hierarchy,
+    scipy.constants,
+    scipy.fftpack,
+    scipy.integrate,
+    scipy.interpolate,
+    scipy.io,
+    scipy.io.arff,
+    scipy.io.harwell_boeing,
+    scipy.io.idl,
+    scipy.io.matlab,
+    scipy.io.netcdf,
+    scipy.io.wavfile,
+    scipy.linalg,
+    scipy.linalg.blas,
+    scipy.linalg.cython_blas,
+    scipy.linalg.lapack,
+    scipy.linalg.cython_lapack,
+    scipy.linalg.interpolative,
+    scipy.misc,
+    scipy.ndimage,
+    scipy.odr,
+    scipy.optimize,
+    scipy.signal,
+    scipy.signal.windows,
+    scipy.sparse,
+    scipy.sparse.linalg,
+    scipy.sparse.csgraph,
+    scipy.spatial,
+    scipy.spatial.distance,
+    scipy.spatial.transform,
+    scipy.special,
+    scipy.stats,
+    scipy.stats.distributions,
+    scipy.stats.mstats,
+]
+
+api_surface = dict()
+exclude = set(['np', 'numpy', 'division', 'absolute_import', 'print_function'])
+for module in modules:
+    all_objects = set(dir(module))
+    private = set((s for s in all_objects if s.startswith('_')))
+    api_surface[module] = len(all_objects - private - exclude)
+
+print(sum(api_surface.values()))

### 2018-07-25 scipy-1.0/paper.tex
Add a short section on maintainers and contributors.

--- 2018-07-25 scipy-1.0/paper.tex
@@
 \textbf{Maintainers and contributors}
+The SciPy project has approximately 15 active maintainers (people who review
+others' contributions, and in general do everything needed to ensure that
+the software and the project move forward).  It also has an additional ~100
+unique contributors for every 6-monthly release cycle.  Maintainers are
+critical to the health of the project\cite{eghbal2016}; they enable contributions
+from a much larger group, and their skills and efforts largely determine how
+fast the project moves forward.  Anyone with the interest and skills can become
+a maintainer; the SciPy Developer Guide provides guidance on how to do that.
 

### 2018-07-25 scipy-1.0/paper.tex
Improvements to the Impact Now section
* added a more exact count of SciPy PyPI downloads from the year 2017,
along with a citation for the software library we used to obtain
the metric

* removed the previous examples of prominent usages of SciPy and replaced
with more compelling examples including: LIGO work and Apple / Intel shipping
SciPy in distributions

--- 2018-07-25 scipy-1.0/paper.tex
@@
 
-% need some better usage stats sources?
-% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
-From the user side, there have been at least 15 million downloads
-of the SciPy manylinux wheels between the start of 2016 and May of 2018
-(njs blog). The recommended source for citing SciPy is simply a reference
-to the website / library itself\cite{SciPylib}, and has been cited $>3000$
-times. Some of the most prominent citing articles include the IPython
-publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
-article\cite{pmid17287757}, a dark
-matter annihilitation study in the field of
-cosmology\cite{1475-7516-2012-08-007}, and many of the
-other major downstream software libraries in the Python scientific
-computing stack.
+From the user side, there were 13,096,468 PyPI downloads
+of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
+underestimate of total downloads given that we don't have access to the
+popular Anaconda package manager download statistics. The recommended 
+source for citing SciPy is simply a reference to the website / library
+itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
+prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
+scientific collaboration that lead to the observation of gravitational
+waves \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly
+with MacOS and in the Intel Distribution for Python\cite{intel-python}.
 

### 2018-07-25 scipy-1.0/supporting_info/commit_timeline.py
WIP: commit timeline plot

@@ -1,0 +1,28 @@
+import subprocess
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+
+count = subprocess.check_output(
+    'git log | grep -E ^Date | awk \'{print " : "$3" "$6}\' | uniq -c',
+    shell=True
+).decode('utf-8').split('\n')
+
+count_month = [line for line in count if line.strip()]
+for c in count_month:
+    if len(c.split(':')) != 2:
+        print("error:", c)
+
+
+count_month = [line.split(' : ') for line in count_month]
+count_month = [(int(cnt), date) for (cnt, date) in count_month]
+
+cnt, dates = zip(*count_month)
+
+## TODO: use Gaussian filter?
+cnt = np.convolve(cnt, np.ones(3)/3, mode='same')
+
+ts = pd.Series(cnt, index=dates)
+ts.plot(linewidth=2)
+
+plt.show()

### 2018-07-26 scipy-1.0/paper.tex
Add preliminary draft of section on Cython API's for low level libraries.

@@
 
+When writing algorithms or wrappers in Cython, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
+SciPy currently includes Cython wrappers for many BLAS and LAPACK routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls.
+Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.
+Function pointers to wrappers for library routines can also be shared in this way, so that Cython modules using the interface provided by SciPy no longer have to worry about finding and linking against the original library.
+This lets other extensions avoid the complexity of finding and using the correct libraries on the various possible Fortran compiler ABI and OS combinations.
+The low-level Cython wrappers provide an OS and Fortran compiler independent interface for the libraries they wrap.
+
+To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
+The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using f2py, though a small number of routines use hand-written type signatures instead.
+The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
+The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
+
 \subsection*{Numerical optimization}

### 2018-07-26 scipy-1.0/paper.tex
Add an aside about the impact of eliminating Python call overheads.

--- 2018-07-26 scipy-1.0/paper.tex
@@
 These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
-The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls.
+The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls{\textemdash}this can give performance gains of one or two orders of magnitude for many use cases.
 Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.

### 2018-07-26 scipy-1.0/paper.tex
Add a brief description of the compatibility policies used in determining which LAPACK routines can have Cython wrappers. Also mention why there aren't similar concerns for the BLAS and special wrappers.

@@
 
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
+In other words, wrappers are provided for routines that are included in the earliest supported LAPACK version and are not removed or changed in some incompatible way in some subsequent LAPACK release.
+The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers we provide.
+Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
+
 \subsection*{Numerical optimization}

### 2018-07-26 scipy-1.0/paper.tex
Add citations for Cython wrappers section.

@@
 
-When writing algorithms or wrappers in Cython, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
-SciPy currently includes Cython wrappers for many BLAS and LAPACK routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+When writing algorithms or wrappers in Cython\cite{behnel2011cython}, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
+SciPy currently includes Cython wrappers\cite{blas-lapack-wrappers-scipy-2015} for many BLAS and LAPACK\cite{LAPACK} routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
 These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
@@
 To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
-The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using f2py, though a small number of routines use hand-written type signatures instead.
+The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using f2py\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
 The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.

### 2018-07-26 scipy-1.0/paper.tex
Add a section on the scope of SciPy

--- 2018-07-26 scipy-1.0/paper.tex
@@
 \author[23]{Antonio Horta Ribeiro}
-\author[24]{Ian Henriksen}
 
@@
 \affil[23]{Affiliation, department, city, postcode, country}
-\affil[24]{University of Texas at Austin,
-           Instite for Computational Engineering and Sciences,
-	   Austin, TX, 78712, USA}
 
@@
 complete coverage. In other areas it aims to provide fundamental building
-blocks while interacting well with other packages specialized in that area.
-For example, SciPy provides what one expect to find in a
+blocks, while interacting well with other packages specialized in that area.
+For example for statistics, SciPy provides what one expect to find in a
 statistics textbook (probability distributions, hypothesis tests, frequency
@@
     \item The algorithm is of relevance to multiple fields of science.
-    \item The algorithm is demonstrably important.  E.g., it is classic
+    \item The implementation is demonstrably important.  E.g., it is classic
     enough to be included in textbooks, or it is based on a peer-reviewed article
@@
 
-When writing algorithms or wrappers in Cython, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
-SciPy currently includes Cython wrappers for many BLAS and LAPACK routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
-These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
-The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls{\textemdash}this can give performance gains of one or two orders of magnitude for many use cases.
-Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.
-Function pointers to wrappers for library routines can also be shared in this way, so that Cython modules using the interface provided by SciPy no longer have to worry about finding and linking against the original library.
-This lets other extensions avoid the complexity of finding and using the correct libraries on the various possible Fortran compiler ABI and OS combinations.
-The low-level Cython wrappers provide an OS and Fortran compiler independent interface for the libraries they wrap.
-
-To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
-The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using f2py, though a small number of routines use hand-written type signatures instead.
-The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
-The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
-
 \subsection*{Numerical optimization}
@@
 \subsection*{Statistical distributions}
-The \texttt{scipy.stats} module provides provides more than 100 probability
-distributions and a framework to implement additional distributions. Given
-an expression for a probability distribution function (pdf),
-the framework will calculate moments, cumulative distribution function (cdf),
-differential entropy, and more. Those properties are calculated based on
-generic code, e.g., numerical integration of the pdf to obtain the cdf.
-Where possible, the generic code is replaced by explicit formulas or more
-accurate or efficient implementations. For example, most distributions have
-an explicit formula for the cdf, so it is used instead of integrating the
-pdf. The distributions framework also allows the user to draw random
-variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
-parameter, which is either a \texttt{numpy.random.RandomState} object or an
-integer used to generate such an object, to provide fine-grained user-level
-control over the random state of a workflow.
-Currently about 80 continuous univariate distributions, 12 discrete univariate
-distributions and 10 multivariate distributions are implemented.
-Key recent distributions added to \texttt{scipy.stats} include the histogram-based
-distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
-language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
-far the only multivariate discrete distribution available in SciPy.
 
@@
 \textbf{Maintainers and contributors}
-The SciPy project has approximately 15 active maintainers - people who review
+The SciPy project has approximately 15 active maintainers (people who review
 others' contributions, and in general do everything needed to ensure that
-the software and the project move forward. Maintainers are critical to the
-health of the project\cite{eghbal2016}; their skills and efforts largely
-determine how fast the project moves forward, and they enable contributions
-from a much larger group. Accordingly, the project also has an additional
-~100 unique contributors for every 6-monthly release cycle. Anyone with the
-interest and skills can become a contributor or maintainer; the SciPy
-Developer Guide provides guidance on how to do that.
+the software and the project move forward).  It also has an additional ~100
+unique contributors for every 6-monthly release cycle.  Maintainers are
+critical to the health of the project\cite{eghbal2016}; they enable contributions
+from a much larger group, and their skills and efforts largely determine how
+fast the project moves forward.  Anyone with the interest and skills can become
+a maintainer; the SciPy Developer Guide provides guidance on how to do that.
 

### 2018-07-26 scipy-1.0/paper.tex
Minor update to Maintainers section based on comment of @mdhaber.

@@
 \textbf{Maintainers and contributors}
-The SciPy project has approximately 15 active maintainers (people who review
+The SciPy project has approximately 15 active maintainers - people who review
 others' contributions, and in general do everything needed to ensure that
-the software and the project move forward).  It also has an additional ~100
-unique contributors for every 6-monthly release cycle.  Maintainers are
-critical to the health of the project\cite{eghbal2016}; they enable contributions
-from a much larger group, and their skills and efforts largely determine how
-fast the project moves forward.  Anyone with the interest and skills can become
-a maintainer; the SciPy Developer Guide provides guidance on how to do that.
+the software and the project move forward. Maintainers are critical to the
+health of the project\cite{eghbal2016}; their skills and efforts largely
+determine how fast the project moves forward, and they enable contributions
+from a much larger group. Accordingly, the project also has an additional
+~100 unique contributors for every 6-monthly release cycle. Anyone with the
+interest and skills can become a contributor or maintainer; the SciPy
+Developer Guide provides guidance on how to do that.
 

### 2018-07-26 scipy-1.0/paper.tex
Update scipy.stats section for review comments.

--- 2018-07-26 scipy-1.0/paper.tex
@@
 
-\textbf{Project scope}
-SciPy aims to provide fundamental algorithms for scientific computing. The
-breadth of its scope was derived from the Guide to Available Mathematical
-Software classification system (GAMS\cite{boisvert1991guide}). In areas
-that move relatively slowly, e.g. linear algebra, SciPy aims to provide
-complete coverage. In other areas it aims to provide fundamental building
-blocks, while interacting well with other packages specialized in that area.
-For example for statistics, SciPy provides what one expect to find in a
-statistics textbook (probability distributions, hypothesis tests, frequency
-statistics, correlation functions, and more), while Statsmodels provides
-more advanced statistical estimators and inference methods, scikit-learn
-covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
-and probabilistic modelling.
-
-We use the following criteria to determine whether or not to include new
-functionality in SciPy:
-\begin{itemize}
-    \item The algorithm is of relevance to multiple fields of science.
-    \item The implementation is demonstrably important.  E.g., it is classic
-    enough to be included in textbooks, or it is based on a peer-reviewed article
-    which has a significant number of citations.
-\end{itemize}
-
-In terms of software systems and architecture, SciPy's scope matches NumPy's:
-algorithms for in-memory computing on single machines, with support for a wide
-range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) is explicitly out of scope.
+\textbf{Project goals and scope}
 
@@
 \subsection*{API and ABI evolution}
-The application programming interface (API) for SciPy consists of
-approximately 1500 functions and classes.
-Our policy for evolving the API over time is that new functionality can be added,
-while removing or changing existing functionality can only be done if the benefits
-of that exceeds the (often significant) costs to users, \textit{and} only
-after giving clear deprecation warnings to those users for at least one year.
-In general, we encourage changes that improve clarity in the API of the
-library but strongly discourage breaking backwards compatibility\footnote{
-    Backwards compatibility means that new versions of SciPy will keep user
-    code written for an older version working.}
-given our position near the base of the scientific Python computing stack.
-
-In addition to the Python API, SciPy has C and Cython interfaces in a number
-of submodules. Therefore, we have to also consider the application binary
-interface (ABI). This ABI has been stable for a long time, and we aim to
-evolve it only in a backwards compatible way.
+In general, we encourage changes that improve clarity in the API of the library,
+but strongly discourage breaking backwards compatibility, given our position near
+the base of the scientific Python computing stack.
 
@@
 formats and the community continues to move forward here so we will aim
-to take advantage of their improvements\cite{abbasi2018sparse} 
+to take advantage of their improvements\cite{Sparse_Arrays} 
 while preserving backwards compatibility to the extent that it is possible.
@@
 \textbf{Maintainers and contributors}
-The SciPy project has approximately 15 active maintainers (people who review
-others' contributions, and in general do everything needed to ensure that
-the software and the project move forward).  It also has an additional ~100
-unique contributors for every 6-monthly release cycle.  Maintainers are
-critical to the health of the project\cite{eghbal2016}; they enable contributions
-from a much larger group, and their skills and efforts largely determine how
-fast the project moves forward.  Anyone with the interest and skills can become
-a maintainer; the SciPy Developer Guide provides guidance on how to do that.
 
@@
 
-From the user side, there were 13,096,468 PyPI downloads
-of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
-underestimate of total downloads given that we don't have access to the
-popular Anaconda package manager download statistics. The recommended 
-source for citing SciPy is simply a reference to the website / library
-itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
-prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
-scientific collaboration that lead to the observation of gravitational
-waves \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly
-with MacOS and in the Intel Distribution for Python\cite{intel-python}.
+% need some better usage stats sources?
+% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
+From the user side, there have been at least 15 million downloads
+of the SciPy manylinux wheels between the start of 2016 and May of 2018
+(njs blog). The recommended source for citing SciPy is simply a reference
+to the website / library itself\cite{SciPylib}, and has been cited $>3000$
+times. Some of the most prominent citing articles include the IPython
+publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
+article\cite{pmid17287757}, a dark
+matter annihilitation study in the field of
+cosmology\cite{1475-7516-2012-08-007}, and many of the
+other major downstream software libraries in the Python scientific
+computing stack.
 
 \textbf{Future development}.
-Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}.
-This is due to the current interface only supporting two-dimensional objects.
-Even sub-arrays are two-dimensional, and dimensions greater than two aren't supported.
-The new package supports dimensions greater than two, and it bases itself around arrays rather than matrices (which is strictly more powerful).
-It also features improved inter-operability with NumPy and other community packages.
-
 \textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}

### 2018-07-26 scipy-1.0/paper.tex
Address review comments of @mdhaber on scope section.

@@
 complete coverage. In other areas it aims to provide fundamental building
-blocks, while interacting well with other packages specialized in that area.
-For example for statistics, SciPy provides what one expect to find in a
+blocks while interacting well with other packages specialized in that area.
+For example, SciPy provides what one expect to find in a
 statistics textbook (probability distributions, hypothesis tests, frequency
@@
     \item The algorithm is of relevance to multiple fields of science.
-    \item The implementation is demonstrably important.  E.g., it is classic
+    \item The algorithm is demonstrably important.  E.g., it is classic
     enough to be included in textbooks, or it is based on a peer-reviewed article

### 2018-07-26 scipy-1.0/paper.tex

@@
 \subsection*{Cython bindings for BLAS, LAPACK and special}
+
+When writing algorithms or wrappers in Cython\cite{behnel2011cython}, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
+SciPy currently includes Cython wrappers\cite{blas-lapack-wrappers-scipy-2015} for many BLAS and LAPACK\cite{LAPACK} routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls{\textemdash}this can give performance gains of one or two orders of magnitude for many use cases.
+Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.
+Function pointers to wrappers for library routines can also be shared in this way, so that Cython modules using the interface provided by SciPy no longer have to worry about finding and linking against the original library.
+This lets other extensions avoid the complexity of finding and using the correct libraries on the various possible Fortran compiler ABI and OS combinations.
+The low-level Cython wrappers provide an OS and Fortran compiler independent interface for the libraries they wrap.
+
+To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
+The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
+The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
+The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
+
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
+In other words, wrappers are provided for routines that are included in the earliest supported LAPACK version and are not removed or changed in some incompatible way in some subsequent LAPACK release.
+The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers we provide.
+Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
 

### 2018-07-26 scipy-1.0/paper.tex
First round of revisions based on reviewer feedback in manuscript PR 43.

@@
 \textbf{Sparse matrices}
+
+\texttt{scipy.sparse} offers seven sparse matrix data structures. 
+The most important ones are CSR and CSC; these offer fast row or column 
+access and fast matrix-vector multiplication, and are used heavily in 
+scikit-learn.
+
+Over the last three years our sparse matrix handling internals have
+been rewritten and performance has been improved. Iterating over 
+and slicing of CSC and CSR matrices is now faster by up to 35\%,
+and the CSR to COO matrix format conversion is now several times faster.
+COO matrices are now generally more memory efficient. Importantly,
+SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
+the low-level implementations leveraged by a subset of our \texttt{sparse}
+offerings.
+
+From a new features standpoint, \texttt{scipy.sparse} matrices and linear
+operators now support the Python matrix multiplication (@) operator when
+available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
+for computing sparse matrix norms and drawing random variates from arbitrary 
+distributions, respectively.
 

### 2018-07-26 scipy-1.0/paper.tex
Further cleanup of the section on Cython wrappers for low-level libraries. In particular, simplify the discussion of the portability benefits.

--- 2018-07-26 scipy-1.0/paper.tex
@@
 \subsection*{Cython bindings for BLAS, LAPACK and special}
+
+When writing algorithms in Cython\cite{behnel2011cython}, it is typically more efficient to call directly into the libraries SciPy wraps rather than indirectly, using SciPy's Python APIs.
+SciPy includes Cython wrappers for many BLAS and LAPACK\cite{LAPACK} routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+These Cython wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+These low-level interfaces for Cython can also be used outside of the SciPy codebase to gain direct access to the functions in the wrapped libraries while avoiding the overhead of Python function calls.
+This can give performance gains of one or two orders of magnitude for many use cases.
+
+The low-level Cython interfaces can also be used without linking against the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.
+This lets other extensions avoid the complexity of finding and using the correct libraries.
+This is especially important when wrapping libraries written in Fortran since these low-level wrappers can be used without a Fortran compiler and without having to handle all the different Fortran compiler ABI's and name mangling schemes.
+
+To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
+The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
+The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
+The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
+
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
+The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers provided by SciPy.
+Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
 

### 2018-07-27 scipy-1.0/paper.tex
Second round of revisions based on reviewer feedback in manuscript PR 43.

--- 2018-07-26 scipy-1.0/paper.tex
@@
 
-\texttt{scipy.sparse} offers seven sparse matrix data structures. 
-The most important ones are CSR and CSC; these offer fast row or column 
-access and fast matrix-vector multiplication, and are used heavily in 
-scikit-learn.
+\texttt{scipy.sparse} offers seven sparse matrix data structures,
+also known as sparse formats. The most important ones are the row- 
+and column-compressed formats (CSR and CSC, respectively). 
+These offer fast major-axis indexing and fast matrix-vector multiplication,
+and are used heavily in scikit-learn.
 
@@
 for computing sparse matrix norms and drawing random variates from arbitrary 
-distributions, respectively.
+distributions, respectively. Also, we've made a concerted effort to bring the
+\texttt{scipy.sparse} API into line with the equivalent NumPy API where possible,
+typically by adding support for kwargs like axis=None or copy=False to many methods.
 

### 2018-07-27 scipy-1.0/paper.tex
Address comment of @mdhaber on Impact Now section.

--- 2018-07-26 scipy-1.0/paper.tex
@@
 
-\texttt{scipy.sparse} offers seven sparse matrix data structures. 
-The most important ones are CSR and CSC; these offer fast row or column 
-access and fast matrix-vector multiplication, and are used heavily in 
-scikit-learn.
-
-Over the last three years our sparse matrix handling internals have
-been rewritten and performance has been improved. Iterating over 
-and slicing of CSC and CSR matrices is now faster by up to 35\%,
-and the CSR to COO matrix format conversion is now several times faster.
-COO matrices are now generally more memory efficient. Importantly,
-SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
-the low-level implementations leveraged by a subset of our \texttt{sparse}
-offerings.
-
-From a new features standpoint, \texttt{scipy.sparse} matrices and linear
-operators now support the Python matrix multiplication (@) operator when
-available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
-for computing sparse matrix norms and drawing random variates from arbitrary 
-distributions, respectively.
-
 \subsection*{Unified bindings to compiled code}
@@
 
-From the user side, there were 13,096,468 PyPI downloads
-of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
-underestimate of total downloads given that we don't have access to the
-popular Anaconda package manager download statistics. The recommended 
-source for citing SciPy is simply a reference to the website / library
-itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
+From the user side, there were 13,096,468 downloads of SciPy from the
+Python Packaging Index (PyPI) during the year 2017\cite{pypinfo}, which is
+a lower bound on the number of total downloads by users given that downloading
+from PyPI is only one of several popular methods to install SciPy.
+The SciPy website\cite{SciPylib}, which has been the default citation in the
+absence of a peer-reviewed paper, has been cited $>3000$ times. Some of the most
 prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo

### 2018-07-27 scipy-1.0/paper.tex
Third round of revisions based on reviewer feedback in manuscript PR 43.

--- 2018-07-27 scipy-1.0/paper.tex
@@
 and the CSR to COO matrix format conversion is now several times faster.
-COO matrices are now generally more memory efficient. Importantly,
-SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
+Importantly, SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
 the low-level implementations leveraged by a subset of our \texttt{sparse}
@@
 \texttt{scipy.sparse} API into line with the equivalent NumPy API where possible,
-typically by adding support for kwargs like axis=None or copy=False to many methods.
+typically by adding support for kwargs like \texttt{axis=None} or
+\texttt{copy=False} to many methods.
 

### 2018-08-02 scipy-1.0/paper.tex
Fourth round of revisions based on reviewer feedback in manuscript PR 43.

--- 2018-07-27 scipy-1.0/paper.tex
@@
 
-\subsection*{Cython bindings for BLAS, LAPACK and special}
+\subsection*{Cython bindings for BLAS, LAPACK, and special}
 
-When writing algorithms or wrappers in Cython\cite{behnel2011cython}, it is often convenient to call directly into the libraries that SciPy wraps rather than routing the call back through the Python interpreter.
-SciPy currently includes Cython wrappers\cite{blas-lapack-wrappers-scipy-2015} for many BLAS and LAPACK\cite{LAPACK} routines and for many of the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
-These wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
-The low-level interfaces for Cython can also be used by third party libraries to get direct access to the functions in these libraries while avoiding the overhead of Python function calls{\textemdash}this can give performance gains of one or two orders of magnitude for many use cases.
-Cython's cimport machinery allows Cython based modules to share C-level objects with other Cython modules without requiring that the third-party extensions link against any additional libraries.
-Function pointers to wrappers for library routines can also be shared in this way, so that Cython modules using the interface provided by SciPy no longer have to worry about finding and linking against the original library.
-This lets other extensions avoid the complexity of finding and using the correct libraries on the various possible Fortran compiler ABI and OS combinations.
-The low-level Cython wrappers provide an OS and Fortran compiler independent interface for the libraries they wrap.
+SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and LAPACK\cite{LAPACK} routines and for the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
+These Cython wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+When writing algorithms in Cython, it is typically more efficient to call directly into the libraries SciPy wraps rather than indirectly, using SciPy's Python APIs.
+These low-level interfaces for Cython can also be used outside of the SciPy codebase to gain access to the functions in the wrapped libraries while avoiding the overhead of Python function calls.
+This can give performance gains of one or two orders of magnitude for many use cases.
 
-To help with both correctness and ease of maintenance, most of these low-level Cython wrappers are generated automatically.
+Developers can also use the low-level Cython interfaces without linking against the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.
+This lets other extensions avoid the complexity of finding and using the correct libraries.
+Avoiding this complexity is especially important when wrapping libraries written in Fortran.
+Not only can these low-level wrappers be used without a Fortran compiler, they can also be used without having to handle all the different Fortran compiler ABI's and name mangling schemes.
+
+Most of these low-level Cython wrappers are generated automatically to help with both correctness and ease of maintenance. 
 The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
@@
 Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
-In other words, wrappers are provided for routines that are included in the earliest supported LAPACK version and are not removed or changed in some incompatible way in some subsequent LAPACK release.
-The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers we provide.
+The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers provided by SciPy.
 Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
@@
 
-From the user side, there were 13,096,468 downloads of SciPy from the
-Python Packaging Index (PyPI) during the year 2017\cite{pypinfo}, which is
-a lower bound on the number of total downloads by users given that downloading
-from PyPI is only one of several popular methods to install SciPy.
-The SciPy website\cite{SciPylib}, which has been the default citation in the
-absence of a peer-reviewed paper, has been cited $>3000$ times. Some of the most
+From the user side, there were 13,096,468 PyPI downloads
+of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
+underestimate of total downloads given that we don't have access to the
+popular Anaconda package manager download statistics. The recommended 
+source for citing SciPy is simply a reference to the website / library
+itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
 prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo

### 2018-08-11 scipy-1.0/paper.tex

@@
 \textbf{Sparse matrices}
+
+\texttt{scipy.sparse} offers seven sparse matrix data structures,
+also known as sparse formats. The most important ones are the row- 
+and column-compressed formats (CSR and CSC, respectively). 
+These offer fast major-axis indexing and fast matrix-vector multiplication,
+and are used heavily in scikit-learn.
+
+Over the last three years our sparse matrix handling internals have
+been rewritten and performance has been improved. Iterating over 
+and slicing of CSC and CSR matrices is now faster by up to 35\%,
+and the CSR to COO matrix format conversion is now several times faster.
+Importantly, SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
+the low-level implementations leveraged by a subset of our \texttt{sparse}
+offerings.
+
+From a new features standpoint, \texttt{scipy.sparse} matrices and linear
+operators now support the Python matrix multiplication (@) operator when
+available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
+for computing sparse matrix norms and drawing random variates from arbitrary 
+distributions, respectively. Also, we've made a concerted effort to bring the
+\texttt{scipy.sparse} API into line with the equivalent NumPy API where possible.
 

### 2018-08-15 scipy-1.0/paper.tex
Formatting

@@
 
-\textbf{History}
+\subsection*{History}
 
-\textbf{Project scope}
+\subsection*{Project scope}
+
 SciPy aims to provide fundamental algorithms for scientific computing. The
@@
 
-\textbf{Current status (maturity, users)}
+\subsection*{Current status (maturity, users)}
 
@@
 \subsection*{Submodule organisation}
+
 \input{subpackages}
@@
 \subsection*{Language choices}
-Python, Cython, Fortran, C and C++ are the programming languages
-used to implement scientific algorithms in the SciPy library. An analysis
-of our code base using the \texttt{linguist} library provides a detailed
-breakdown as \% composition by programming language in SciPy
-(Figure~\ref{fig:linguist}).
+
+Python, Cython, Fortran, C and C++ are the programming languages used to
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library provides a detailed breakdown as \%
+composition by programming language in SciPy (Figure~\ref{fig:linguist}).
 
@@
 
-For implementing new functionality, we have a clear order of language preference.
-First Python, if performance is not an issue. If it is, then in order of
-preference: Cython, C, C++, Fortran. The main motivation for this is 
-maintainability: Cython has the highest abstraction level and most 
-Python developers will understand it. C is also widely known,
-and easier to deal with for the core development team
-than C++ and especially Fortran.
+For implementing new functionality, we have a clear order of language
+preference.  First Python, if performance is not an issue. If it is, then in
+order of preference: Cython, C, C++, Fortran. The main motivation for this is
+maintainability: Cython has the highest abstraction level and most Python
+developers will understand it. C is also widely known, and easier to deal with
+for the core development team than C++ and especially Fortran.
 
-While it is not surprising that Python is heavily used in SciPy, the
-usage distribution of the other languages warrants some discussion. Fortran
-is an extremely well-established scientific programming language, both
-for historical reasons and because of its continued excellent
-performance\cite{Koelbel:1993:HPF:562354}. We
-wrap the FFTPACK Fortran library for performing Fourier
-transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} since
-this library has been a standard in the field for 33 years and has a license
-that is compatible with our own. Likewise, we wrap the Fortran source
+While it is not surprising that Python is heavily used in SciPy, the usage
+distribution of the other languages warrants some discussion. Fortran is an
+extremely well-established scientific programming language, both for historical
+reasons and because of its continued excellent
+performance\cite{Koelbel:1993:HPF:562354}. We wrap the FFTPACK Fortran library
+for performing Fourier transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}
+since this library has been a standard in the field for 33 years and has a
+license that is compatible with our own. Likewise, we wrap the Fortran source
 for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial
 value problem for ordinary differential equation systems for 30 years. For
-similar reasons we also wrap the Fortran libraries QUADPACK\cite{1983qspa.book.....P} (for numerical
-integration of one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103} (curve-fitting /
-interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal distance regression),
-MINPACK\cite{osti_6997568} (minimization of linear and nonlinear equations),
-ARPACK\cite{leh:sor:yan96} (solving large scale eigenvalue problems), 
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling Bessel Functions), and 
-CDFLIB\cite{CDFLIB_site} (evaluation of cumulative density functions).
+similar reasons we also wrap the Fortran libraries
+QUADPACK\cite{1983qspa.book.....P} (for numerical integration of
+one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103}
+(curve-fitting / interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal
+distance regression), MINPACK\cite{osti_6997568} (minimization of linear and
+nonlinear equations), ARPACK\cite{leh:sor:yan96} (solving large scale
+eigenvalue problems), ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling
+Bessel Functions), and CDFLIB\cite{CDFLIB_site} (evaluation of cumulative
+density functions).
 
@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. Alongside its objected-oriented relative C++, C
-can be leveraged in Python after being exposed using the Cython
-language. Cython has been described as a creole language that mixes
-the best parts of Python and lower-level C / C++
-paradigms\cite{behnel2011cython}. We thus often use Cython as a glue
-between well-established low-level scientific computing libraries and
-the Python interface offered by SciPy. We also use Cython to enable
-performance enhancements in Python code, especially for cases where heavily
-used inner loops benefit from a compiled code with static typing. Some of the C
-libraries that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842} 
-(iterative solving of the trust region problem), SuperLU\cite{li05,superlu_ug99} (solution of
-large, sparse, nonsymmetric systems of linear equations),
-Qhull\cite{Barber:1996:QAC:235815.235821} (computational
-geometry), and Cephes\cite{cephes_netlib} (mathematics algorithms).
+scientific computing. Alongside its objected-oriented relative C++, C can be
+leveraged in Python after being exposed using the Cython language. Cython has
+been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We thus often use Cython
+as a glue between well-established low-level scientific computing libraries and
+the Python interface offered by SciPy. We also use Cython to enable performance
+enhancements in Python code, especially for cases where heavily used inner
+loops benefit from a compiled code with static typing. Some of the C libraries
+that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
+(iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99} (solution of large, sparse, nonsymmetric
+systems of linear equations), Qhull\cite{Barber:1996:QAC:235815.235821}
+(computational geometry), and Cephes\cite{cephes_netlib} (mathematics
+algorithms).
 
-Therefore, the relative abundance of different programming languages
-in the SciPy library results from a combination of the usage of powerful performance
-enhancing languages that interact well with Python (i.e., Cython), and
-the usage of languages (and their libraries) that have proven reliable
-and performant over many decades. The position that SciPy occupies
-near the foundation of the scientific Python ecosystem is such that
-adoption of new languages or major dependencies is generally unlikely--our
-choices are strongly driven by long-term stability. GPU acceleration,
-new transpiling libraries, and the latest JIT compilation approaches
-(i.e., Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall outside the remit
-of the main SciPy library. That said, we have recently increased our
-efforts to support compatibility with some of these options, and having
-our full test suite pass with the PyPy JIT compiler\cite{Bolz:2009:TMP:1565824.1565827}
-is now a requirement in our development workflow.
+Therefore, the relative abundance of different programming languages in the
+SciPy library results from a combination of the usage of powerful performance
+enhancing languages that interact well with Python (i.e., Cython), and the
+usage of languages (and their libraries) that have proven reliable and
+performant over many decades. The position that SciPy occupies near the
+foundation of the scientific Python ecosystem is such that adoption of new
+languages or major dependencies is generally unlikely--our choices are strongly
+driven by long-term stability. GPU acceleration, new transpiling libraries, and
+the latest JIT compilation approaches (i.e.,
+Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall
+outside the remit of the main SciPy library. That said, we have recently
+increased our efforts to support compatibility with some of these options, and
+having our full test suite pass with the PyPy JIT
+compiler\cite{Bolz:2009:TMP:1565824.1565827} is now a requirement in our
+development workflow.
 
 \subsection*{API and ABI evolution}
-The application programming interface (API) for SciPy consists of
-approximately 1500 functions and classes.
-Our policy for evolving the API over time is that new functionality can be added,
-while removing or changing existing functionality can only be done if the benefits
-of that exceeds the (often significant) costs to users, \textit{and} only
-after giving clear deprecation warnings to those users for at least one year.
-In general, we encourage changes that improve clarity in the API of the
-library but strongly discourage breaking backwards compatibility\footnote{
+
+The application programming interface (API) for SciPy consists of approximately
+1500 functions and classes.  Our policy for evolving the API over time is that
+new functionality can be added, while removing or changing existing
+functionality can only be done if the benefits of that exceeds the (often
+significant) costs to users, \textit{and} only after giving clear deprecation
+warnings to those users for at least one year.  In general, we encourage
+changes that improve clarity in the API of the library but strongly discourage
+breaking backwards compatibility\footnote{
     Backwards compatibility means that new versions of SciPy will keep user
@@
 \subsection*{Data structures}
-\textbf{cKDTree}
+
+\subsubsection*{cKDTree}
+
 \input{ckdtree}
-\textbf{Sparse matrices}
+
+\subsubsection*{Sparse matrices}
 
@@
 
-Over the last three years our sparse matrix handling internals have
-been rewritten and performance has been improved. Iterating over 
-and slicing of CSC and CSR matrices is now faster by up to 35\%,
-and the CSR to COO matrix format conversion is now several times faster.
-Importantly, SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing
-the low-level implementations leveraged by a subset of our \texttt{sparse}
+Over the last three years our sparse matrix handling internals have been
+rewritten and performance has been improved. Iterating over and slicing of CSC
+and CSR matrices is now faster by up to 35\%, and the CSR to COO matrix format
+conversion is now several times faster.  Importantly,
+SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
+low-level implementations leveraged by a subset of our \texttt{sparse}
 offerings.
@@
 operators now support the Python matrix multiplication (@) operator when
-available. We've added \texttt{scipy.sparse.norm} and \texttt{scipy.sparse.random} 
-for computing sparse matrix norms and drawing random variates from arbitrary 
-distributions, respectively. Also, we've made a concerted effort to bring the
-\texttt{scipy.sparse} API into line with the equivalent NumPy API where possible.
+available. We've added \texttt{scipy.sparse.norm} and
+\texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
+random variates from arbitrary distributions, respectively. Also, we've made a
+concerted effort to bring the \texttt{scipy.sparse} API into line with the
+equivalent NumPy API where possible.
 
 \subsection*{Unified bindings to compiled code}
+
 \subsubsection*{LowLevelCallable}
-As of SciPy version 0.19, it is possible for users to wrap low-level
-functions in a \texttt{scipy.LowLevelCallable()} object that reduces
-the overhead for calling compiled C functions directly from Python.
-Supported low-level functions include \texttt{PyCapsule} objects,
-ctypes function pointers, and cffi function pointers. The low level
-function signature must be consistent with the expectations of the
-routine it is passed to. For example, the documentation for
-\texttt{scipy.ndimage.generic\_filter} defines two acceptable C callback
-function signatures that may be used to produce functions that operate
-on each element of image data with low overhead. The C code may be
-generated using numba or Cython, for example, as long as the function
-call signature matches the specifications. Furthermore, it is even
-possible to generate a low-level callback function automatically
-from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
+
+As of SciPy version 0.19, it is possible for users to wrap low-level functions
+in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
+calling compiled C functions directly from Python.  Supported low-level
+functions include \texttt{PyCapsule} objects, ctypes function pointers, and
+cffi function pointers. The low level function signature must be consistent
+with the expectations of the routine it is passed to. For example, the
+documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
+C callback function signatures that may be used to produce functions that
+operate on each element of image data with low overhead. The C code may be
+generated using numba or Cython, for example, as long as the function call
+signature matches the specifications. Furthermore, it is even possible to
+generate a low-level callback function automatically from a Cython module using
+\texttt{scipy.LowLevelCallable.from\_cython}.
 
@@
 
-SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and LAPACK\cite{LAPACK} routines and for the special functions provided in the \texttt{scipy.{\allowbreak}special} submodule.
-These Cython wrappers are available in the modules \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas}, \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
-When writing algorithms in Cython, it is typically more efficient to call directly into the libraries SciPy wraps rather than indirectly, using SciPy's Python APIs.
-These low-level interfaces for Cython can also be used outside of the SciPy codebase to gain access to the functions in the wrapped libraries while avoiding the overhead of Python function calls.
-This can give performance gains of one or two orders of magnitude for many use cases.
+SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and
+LAPACK\cite{LAPACK} routines and for the special functions provided in the
+\texttt{scipy.{\allowbreak}special} submodule.  These Cython wrappers are
+available in the modules
+\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},
+\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and
+\texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+When writing algorithms in Cython, it is typically more efficient to call
+directly into the libraries SciPy wraps rather than indirectly, using SciPy's
+Python APIs.  These low-level interfaces for Cython can also be used outside of
+the SciPy codebase to gain access to the functions in the wrapped libraries
+while avoiding the overhead of Python function calls.  This can give
+performance gains of one or two orders of magnitude for many use cases.
 
-Developers can also use the low-level Cython interfaces without linking against the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.
-This lets other extensions avoid the complexity of finding and using the correct libraries.
-Avoiding this complexity is especially important when wrapping libraries written in Fortran.
-Not only can these low-level wrappers be used without a Fortran compiler, they can also be used without having to handle all the different Fortran compiler ABI's and name mangling schemes.
+Developers can also use the low-level Cython interfaces without linking against
+the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.  This lets other
+extensions avoid the complexity of finding and using the correct libraries.
+Avoiding this complexity is especially important when wrapping libraries
+written in Fortran.  Not only can these low-level wrappers be used without a
+Fortran compiler, they can also be used without having to handle all the
+different Fortran compiler ABI's and name mangling schemes.
 
-Most of these low-level Cython wrappers are generated automatically to help with both correctness and ease of maintenance. 
-The wrappers for BLAS and LAPACK are primarily generated using type information that is parsed from the BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small number of routines use hand-written type signatures instead.
-The input and output types of each routine are saved in a data file that is read at build time and used to generate the corresponding Cython wrapper files.
-The wrappers in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also generated from a data file containing type information for the wrapped routines.
+Most of these low-level Cython wrappers are generated automatically to help
+with both correctness and ease of maintenance.  The wrappers for BLAS and
+LAPACK are primarily generated using type information that is parsed from the
+BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small
+number of routines use hand-written type signatures instead.  The input and
+output types of each routine are saved in a data file that is read at build
+time and used to generate the corresponding Cython wrapper files.  The wrappers
+in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also
+generated from a data file containing type information for the wrapped
+routines.
 
-Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only provided for the routines that maintain a consistent interface across all supported LAPACK versions.
-The standard BLAS interface provided by the various existing BLAS libraries is not currently changing, so changes are not generally needed in the wrappers provided by SciPy.
-Changes to the Cython wrappers for the functions in \texttt{scipy.{\allowbreak}special} follow corresponding changes to the interface of that submodule.
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only
+provided for the routines that maintain a consistent interface across all
+supported LAPACK versions.  The standard BLAS interface provided by the various
+existing BLAS libraries is not currently changing, so changes are not generally
+needed in the wrappers provided by SciPy.  Changes to the Cython wrappers for
+the functions in \texttt{scipy.{\allowbreak}special} follow corresponding
+changes to the interface of that submodule.
 
 \subsection*{Numerical optimization}
+
 \input{scipy-optimize}
@@
 \subsection*{Statistical distributions}
+
 The \texttt{scipy.stats} module provides provides more than 100 probability
-distributions and a framework to implement additional distributions. Given
-an expression for a probability distribution function (pdf),
-the framework will calculate moments, cumulative distribution function (cdf),
-differential entropy, and more. Those properties are calculated based on
-generic code, e.g., numerical integration of the pdf to obtain the cdf.
-Where possible, the generic code is replaced by explicit formulas or more
-accurate or efficient implementations. For example, most distributions have
-an explicit formula for the cdf, so it is used instead of integrating the
-pdf. The distributions framework also allows the user to draw random
-variates from the distribution. Drawing random variates now accepts a \texttt{random\_state}
-parameter, which is either a \texttt{numpy.random.RandomState} object or an
-integer used to generate such an object, to provide fine-grained user-level
-control over the random state of a workflow.
-Currently about 80 continuous univariate distributions, 12 discrete univariate
-distributions and 10 multivariate distributions are implemented.
-Key recent distributions added to \texttt{scipy.stats} include the histogram-based
-distribution in \texttt{scipy.stats.rv\_histogram} and the multinomial
-distribution in \texttt{scipy.stats.multinomial} (used, for example, in natural
-language processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so 
-far the only multivariate discrete distribution available in SciPy.
+distributions and a framework to implement additional distributions. Given an
+expression for a probability distribution function (pdf), the framework will
+calculate moments, cumulative distribution function (cdf), differential
+entropy, and more. Those properties are calculated based on generic code, e.g.,
+numerical integration of the pdf to obtain the cdf.  Where possible, the
+generic code is replaced by explicit formulas or more accurate or efficient
+implementations. For example, most distributions have an explicit formula for
+the cdf, so it is used instead of integrating the pdf. The distributions
+framework also allows the user to draw random variates from the distribution.
+Drawing random variates now accepts a \texttt{random\_state} parameter, which
+is either a \texttt{numpy.random.RandomState} object or an integer used to
+generate such an object, to provide fine-grained user-level control over the
+random state of a workflow.  Currently about 80 continuous univariate
+distributions, 12 discrete univariate distributions and 10 multivariate
+distributions are implemented.  Key recent distributions added to
+\texttt{scipy.stats} include the histogram-based distribution in
+\texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
+\texttt{scipy.stats.multinomial} (used, for example, in natural language
+processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
+the only multivariate discrete distribution available in SciPy.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{Polynomial interpolators}
+
 \input{poly}
@@
 
-    The airspeed velocity (asv) library enables benchmarking Python packages over their lifetimes, and the performance of the SciPy
-    code base was monitored with asv starting in February of 2015 (PR \#4501). In addition to ensuring that unit tests are passing (see above),
-    confirming that performance generally remains constant or improves over the commit hash history of the project allows us to objectively
-    measure that our code base is improving, to empower scientific applications.
+The airspeed velocity (asv) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing (see above), confirming that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
 
-    Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning roughly nine years of project history. These demonstrate the gradual performance
-    improvements in a nearest-neighbor search through \texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
+Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
 
-    \texttt{python run.py run -e -s 800 --bench "\textbackslash btime\_query\textbackslash b" "02de46a546..b3ddb2c"}
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"}
 
@@
 \includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-        \caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query() over a roughly nine year commit history time frame.
-The results are based on Python 2.7 performance on the master branch of the project using numpy 1.8.2 and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward compatibility).
-Only the L2 (Euclidean) norm is shown here, and to improve backward compatibility and sampling of the benchmarks there was no application of toroidal topology to the KDTree (boxsize argument was ignored).}
+\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using numpy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the KDTree (boxsize argument was ignored).}
 \label{fig:asvbench}
@@
 
-   Any pull request can be compared against the \texttt{master} branch with the command \texttt{asv continuous master new-feature}.
-This will provide a benchmark against the master branch and the branch a new feature is implemented in.
-More features are available in the documentation, including arguments to select which benchmarks to run.
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the master branch and the branch a new feature is implemented
+in. More features are available in the documentation, including arguments to
+select which benchmarks to run.
 
 \subsubsection*{Test suite}
-    The SciPy test suite is orchestrated by a continuous integration matrix
-    that includes posix and Windows (32/64-bit) platforms managed by Travis CI and appveyor,
-    respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and include
-    code linting with pyflakes and pycodestyle. There are more than $13000$ unit
-    tests in the test suite, which is written for usage with the pytest runner, and
-    with a 76.5 \% coverage of approximately $204000$ lines
-    of code holding steady for the last six months (Figure~\ref{fig:coverage}).
-    Some of the historical components of the code base would still benefit from
-    increased test coverage. Documentation for the code is automatically built and hosted 
-    by the CircleCI service to facilitate evaluation of documentation changes / integrity.
-    Our full test suite also passes with PyPy3, a just-in-time compiled version
-    of the Python language.
+
+The SciPy test suite is orchestrated by a continuous integration matrix that
+includes posix and Windows (32/64-bit) platforms managed by Travis CI and
+appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with pyflakes and pycodestyle. There are more than $13000$
+unit tests in the test suite, which is written for usage with the pytest
+runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
+holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
+the historical components of the code base would still benefit from increased
+test coverage. Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes /
+integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+version of the Python language.
 
@@
 \includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy code base by unit tests over the six
-	 months preceding June 28, 2018 as reported by the \texttt{codecov}
-	 service. The analysis of lines of source code covered is performed
-	 automatically in our continuous integation suite using the
-	 \texttt{pytest-cov} and \texttt{gcov} libraries.}
+\caption{\% coverage of the SciPy code base by unit tests over the six months
+preceding June 28, 2018 as reported by the \texttt{codecov} service. The
+analysis of lines of source code covered is performed automatically in our
+continuous integation suite using the \texttt{pytest-cov} and \texttt{gcov}
+libraries.}
 \label{fig:coverage}
@@
 
-\textbf{Governance}
+\subsection*{Governance}
 
 SciPy adopted an official Governance Document on August 3,
-2017\cite{SciPyProjectGovernance}.
-We have a Benevolant Dictator for Life (BDFL), Pauli Virtanen, leading the
-project, following a detailed discussion of the governance model with the
-community on the mailing list.  In practice, the overruling authority of the BDFL
-is anticipated to be used only when the Steering Council cannot agree on a matter.
-The BDFL is expected to consider any strong indications that they should step down;
-although the BDFL may appoint a successor it is generally expected that the Steering
-Council will be consulted on the matter.
+2017\cite{SciPyProjectGovernance}.  We have a Benevolant Dictator for Life
+(BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
+the governance model with the community on the mailing list.  In practice, the
+overruling authority of the BDFL is anticipated to be used only when the
+Steering Council cannot agree on a matter.  The BDFL is expected to consider
+any strong indications that they should step down; although the BDFL may
+appoint a successor it is generally expected that the Steering Council will be
+consulted on the matter.
 
-The (currently) 18 member Steering Council is chaired by Ralf Gommers and performs a number of
-tasks including nominating and voting on new Council Members, and actively
-contributing to the progress of the project (could be code review but also
-outreach activities, etc.). Generally, 1 year of sustained and substantial
-contributions (not restricted to source code contributions) that improve
-the project are the prerequiste for nomination of new Council Members. Council
-Members do not get special treatment for the vast majority of the day to day
-project activities (code submission / review), though it is anticipated that their
-experience with the project will serve as helpful guidance in many matters.
-Council Members have commit rights to the project, but will typically only
-incorporate changes when there are no substantive community objections. The
-Chair of the Steering Council is responsible for starting more formal
-technical reviews of the direction of the project, ensuring that the
-composition of the Council remains current, and communicating / summarizing
-any Council activities performed privately to the broader community. We also
-describe in some detail how institutional partners may contribute to the
-project, but no financial weight from grants or other sources
+The (currently) 18 member Steering Council is chaired by Ralf Gommers and
+performs a number of tasks including nominating and voting on new Council
+Members, and actively contributing to the progress of the project (could be
+code review but also outreach activities, etc.). Generally, 1 year of sustained
+and substantial contributions (not restricted to source code contributions)
+that improve the project are the prerequiste for nomination of new Council
+Members. Council Members do not get special treatment for the vast majority of
+the day to day project activities (code submission / review), though it is
+anticipated that their experience with the project will serve as helpful
+guidance in many matters.  Council Members have commit rights to the project,
+but will typically only incorporate changes when there are no substantive
+community objections. The Chair of the Steering Council is responsible for
+starting more formal technical reviews of the direction of the project,
+ensuring that the composition of the Council remains current, and communicating
+/ summarizing any Council activities performed privately to the broader
+community. We also describe in some detail how institutional partners may
+contribute to the project, but no financial weight from grants or other sources
 may be leveraged to circumvent or overrule the technical direction of the
-project decided upon by the Steering Committee. The Chair does not have
-a fixed term, but is nominated by the Steering Committee and is expected
-to yield to substantive claims that stepping down is the appropriate action.
+project decided upon by the Steering Committee. The Chair does not have a fixed
+term, but is nominated by the Steering Committee and is expected to yield to
+substantive claims that stepping down is the appropriate action.
 
 SciPy adopted an official Code of Conduct on October 24, 2017
-\cite{SciPyCodeOfConduct}, drawing
-inspiration from the Apache Foundation Code of
-Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
-\cite{ContributorConvenant},
-the Jupyter Code of Conduct \cite{Jupyter_COC}, and the Open Source Guides -
-Code of Conduct\cite{OSG_COC}. In short,
-we stated five specific guidelines: \emph{be open} (to anyone participating in our community),
-\emph{be empathetic and patient} (in resolving conflicts), \emph{be collaborative} (we depend
-on each other to build up the tools in the library), \emph{be inquisitive} (identifying issues
-early on can be helpful to everyone), and \emph{be careful with wording} (do not harass or exclude).
-We outlined a broad diversity inclusion statement, and provide
-instructions for contacting the three members of our Code of Conduct Committee or an
-external representative on the NumFOCUS Board of Directors.
+\cite{SciPyCodeOfConduct}, drawing inspiration from the Apache Foundation Code
+of Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
+\cite{ContributorConvenant}, the Jupyter Code of Conduct \cite{Jupyter_COC},
+and the Open Source Guides - Code of Conduct\cite{OSG_COC}. In short, we stated
+five specific guidelines: \emph{be open} (to anyone participating in our
+community), \emph{be empathetic and patient} (in resolving conflicts), \emph{be
+collaborative} (we depend on each other to build up the tools in the library),
+\emph{be inquisitive} (identifying issues early on can be helpful to everyone),
+and \emph{be careful with wording} (do not harass or exclude).  We outlined a
+broad diversity inclusion statement, and provide instructions for contacting
+the three members of our Code of Conduct Committee or an external
+representative on the NumFOCUS Board of Directors.
 
-\textbf{Roadmap}
+\subsection*{Roadmap}
 
 The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
-maintained by the community that describes some of the major directions
-for improvement for the project, as well as specific limitations and
-matters that require assistance moving forward.
+maintained by the community that describes some of the major directions for
+improvement for the project, as well as specific limitations and matters that
+require assistance moving forward.
 
 We are still striving to increase the number of SciPy usage tutorials beyond
-our current 15 section offering\cite{SciPy_tutorials}, but our
-standard documentation for library features is already in excellent shape.
+our current 15 section offering\cite{SciPy_tutorials}, but our standard
+documentation for library features is already in excellent shape.
 
-The low-level Cython code in our library (which interacts with C-level
-code and exposes it for Python usage) could use some measure of modernization,
-including migration to typed memoryviews to handle NumPy arrays instead of
-the older syntax:
+The low-level Cython code in our library (which interacts with C-level code and
+exposes it for Python usage) could use some measure of modernization, including
+migration to typed memoryviews to handle NumPy arrays instead of the older
+syntax:
 
@@
     pass
-
 \end{lstlisting}
@@
 \texttt{scipy.interpolate} could benefit from a set of new features including
-spline fitting routines with better user control; arithmetic routines for splines;
-transparent tensor-product splines, Non-Uniform Rational B-Splines support;
-and mesh refinement and coarsening of B-splines and corresponding tensor products.
-Some spline functionality in \texttt{scipy.signal} should be migrated to \texttt{interpolate} 
-and Second Order Sections need to be updated to match the capabilities in other routines.
+spline fitting routines with better user control; arithmetic routines for
+splines; transparent tensor-product splines, Non-Uniform Rational B-Splines
+support; and mesh refinement and coarsening of B-splines and corresponding
+tensor products.  Some spline functionality in \texttt{scipy.signal} should be
+migrated to \texttt{interpolate} and Second Order Sections need to be updated
+to match the capabilities in other routines.
 
@@
 
-\textbf{Community beyond the SciPy library}
+\subsection*{Community beyond the SciPy library}
 
-\textbf{Maintainers and contributors}
+TODO
+
+\subsection*{Maintainers and contributors}
+
 The SciPy project has approximately 15 active maintainers - people who review
-others' contributions, and in general do everything needed to ensure that
-the software and the project move forward. Maintainers are critical to the
-health of the project\cite{eghbal2016}; their skills and efforts largely
-determine how fast the project moves forward, and they enable contributions
-from a much larger group. Accordingly, the project also has an additional
-~100 unique contributors for every 6-monthly release cycle. Anyone with the
-interest and skills can become a contributor or maintainer; the SciPy
-Developer Guide provides guidance on how to do that.
+others' contributions, and in general do everything needed to ensure that the
+software and the project move forward. Maintainers are critical to the health
+of the project\cite{eghbal2016}; their skills and efforts largely determine how
+fast the project moves forward, and they enable contributions from a much
+larger group. Accordingly, the project also has an additional ~100 unique
+contributors for every 6-monthly release cycle. Anyone with the interest and
+skills can become a contributor or maintainer; the SciPy Developer Guide
+provides guidance on how to do that.
 
@@
 
-\textbf{Impact now}
-SciPy has a strong developer community and a massive user base. GitHub
-traffic metrics report roughly 20,000 unique visitors to the source
-website between May 14 and May 27, 2018, with 721 unique copies (``clones")
-of the code base over that roughly two-week period. The developer community
-consists of 610 unique contributors of source code, with $>19,000$ commits
-accepted into the code base (GitHub page data).
+\subsection*{Impact now}
 
-From the user side, there were 13,096,468 downloads of SciPy from the
-Python Packaging Index (PyPI) during the year 2017\cite{pypinfo}, which is
-a lower bound on the number of total downloads by users given that downloading
-from PyPI is only one of several popular methods to install SciPy.
-The SciPy website\cite{SciPylib}, which has been the default citation in the
-absence of a peer-reviewed paper, has been cited $>3000$ times. Some of the most
-prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
-scientific collaboration that lead to the observation of gravitational
-waves \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly
-with MacOS and in the Intel Distribution for Python\cite{intel-python}.
+SciPy has a strong developer community and a massive user base. GitHub traffic
+metrics report roughly 20,000 unique visitors to the source website between May
+14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
+that roughly two-week period. The developer community consists of 610 unique
+contributors of source code, with $>19,000$ commits accepted into the code base
+(GitHub page data).
 
-\textbf{Future development}.
-Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}.
-This is due to the current interface only supporting two-dimensional objects.
-Even sub-arrays are two-dimensional, and dimensions greater than two aren't supported.
-The new package supports dimensions greater than two, and it bases itself around arrays rather than matrices (which is strictly more powerful).
-It also features improved inter-operability with NumPy and other community packages.
+From the user side, there were 13,096,468 downloads of SciPy from the Python
+Packaging Index (PyPI) during the year 2017\cite{pypinfo}, which is a lower
+bound on the number of total downloads by users given that downloading from
+PyPI is only one of several popular methods to install SciPy.  The SciPy
+website\cite{SciPylib}, which has been the default citation in the absence of a
+peer-reviewed paper, has been cited $>3000$ times. Some of the most prominent
+usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
+scientific collaboration that lead to the observation of gravitational waves
+\cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
+MacOS and in the Intel Distribution for Python\cite{intel-python}.
 
-\textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}
+\subsection*{Future development}
+
+Future development plans include a major rewrite of the \texttt{scipy.sparse}
+module, which is already underway\cite{abbasi2018sparse}. This is due to the
+current interface only supporting two-dimensional objects. Even sub-arrays are
+two-dimensional, and dimensions greater than two aren't supported. The new
+package supports dimensions greater than two, and it bases itself around arrays
+rather than matrices (which is strictly more powerful). It also features
+improved inter-operability with NumPy and other community packages.
+
+\textit{This section should include key issues: sparse arrays, ndimage pixel vs
+point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained
+submodules.}
 

### 2018-08-15 scipy-1.0/ckdtree.tex
Highlight things that need fixing

--- 2018-07-25 scipy-1.0/ckdtree.tex
@@
 
-\textbf{Add a Figure to show the scaling, before and after.}
-\textbf{perhaps give an example or some formula.}
+\fixme{Add a Figure to show the scaling, before and after.}
+\fixme{perhaps give an example or some formula.}
 (cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)

### 2018-08-15 scipy-1.0/paper.tex
Highlight things that need fixing

@@
 \newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\fixme}[1]{\textcolor{red}{{#1}}}
 
@@
 
-\textit{The Introduction section expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.}
+\fixme{The Introduction section expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.}
 
@@
 
-TODO
+\fixme{TODO}
 
@@
 
-\textit{The Discussion should be succinct and must not contain subheadings.}
+\fixme{The Discussion should be succinct and must not contain subheadings.}
 
@@
 
-\textit{This section should include key issues: sparse arrays, ndimage pixel vs
+\fixme{This section should include key issues: sparse arrays, ndimage pixel vs
 point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained
@@
 
-\noindent Use the cite command for an inline citation, e.g. \cite{behnel2011cython}.
+\section*{Acknowledgements}
 
-\section*{Acknowledgements (not compulsory)}
+\fixme{TODO}
 
-Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
+%Acknowledgements should be brief, and should not include thanks to anonymous
+%referees and editors, or effusive comments. Grant or contribution numbers may
+%be acknowledged.
 
@@
 
-Must include all authors, identified by initials, for example:
-A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
+\fixme{TODO}
+
+%Must include all authors, identified by initials, for example: A.A.
+%conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A.
+%and D.A. analysed the results.  All authors reviewed the manuscript.
 
@@
 
-To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing financial interests} (mandatory statement).
+\fixme{TODO}
 
-The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing financial interests statement} on behalf of all authors of the paper.
-This statement must be included in the submitted article file.
+%To include, in this order: \textbf{Accession codes} (where applicable);
+%\textbf{Competing financial interests} (mandatory statement).
+
+%The corresponding author is responsible for submitting a
+%\href{http://www.nature.com/srep/policies/index.html#competing}{competing
+%financial interests statement} on behalf of all authors of the paper. This
+%statement must be included in the submitted article file.
 

### 2018-08-15 scipy-1.0/paper.tex
Spell check

@@
 \affil[24]{University of Texas at Austin,
-           Instite for Computational Engineering and Sciences,
+           Institute for Computational Engineering and Sciences,
 	   Austin, TX, 78712, USA}
@@
 unique code contributors, millions of downloads per year, 161 dependent
-packages, and 28700 dependent respositories.
+packages, and 28700 dependent repositories.
 The library includes functionality spanning clustering, Fourier transforms,
@@
 covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
-and probabilistic modelling.
+and probabilistic modeling.
 
@@
 
-\subsection*{Submodule organisation}
+\subsection*{Submodule organization}
 
@@
 analysis of lines of source code covered is performed automatically in our
-continuous integation suite using the \texttt{pytest-cov} and \texttt{gcov}
+continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
 libraries.}
@@
 
-\section*{Project organisation and community}
+\section*{Project organization and community}
 
@@
 SciPy adopted an official Governance Document on August 3,
-2017\cite{SciPyProjectGovernance}.  We have a Benevolant Dictator for Life
+2017\cite{SciPyProjectGovernance}.  We have a Benevolent Dictator for Life
 (BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
@@
 and substantial contributions (not restricted to source code contributions)
-that improve the project are the prerequiste for nomination of new Council
+that improve the project are the prerequisites for nomination of new Council
 Members. Council Members do not get special treatment for the vast majority of
@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 

### 2018-09-08 scipy-1.0/paper.tex
Cite astropy in broader community section draft.

@@
 
-\fixme{TODO}
+The scientific Python ecosystem includes many examples
+of domain-specific software libraries building on top
+of SciPy features, and then returning to the base SciPy library
+to suggest and even implement improvements to enable
+more effective science applications. For example, there
+are common contributors to the SciPy and astropy core
+libraries\cite{astropy-2018}, and what works well for 
+one of the code bases, infrastructures, or communities 
+is often transferred in some form to the other.
 

### 2018-09-11 scipy-1.0/paper.tex
Add standard references and timeline from 1.0 ann

@@
 \subsection*{History}
+
+\cite{dubois2007guest}
+\cite{oliphant2007python}
+\cite{perez2007ipython}
+\cite{hunter2007matplotlib}
+\cite{vanderwalt2008scipy}
+\cite{peterson2009f2py}
+\cite{millman2011python}
+\cite{perez2011python}
+\cite{vanderwalt2011numpy}
+\cite{behnel2011cython}
+\cite{pedregosa2011scikit}
+\cite{vanderwalt2014scikit}
+\cite{millman2014developing}
+\cite{meurer2017sympy}
+
+\begin{figure}
+\begin{verbatim}
+    2001: first SciPy release
+    2005: transition to NumPy
+    2007: creation of scikits
+    2008: spatial module and first Cython code added
+    2010: move to a 6-monthly release cycle
+    2011: development moves to GitHub
+    2011: Python 3 support
+    2012: add a sparse graph module and unified optimization interface
+    2012: remove of maxentropy
+    2013: continuous integration with TravisCI
+    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
+    2017: add a unified C API with LowLevelCallable; remove weave
+    2017: SciPy 1.0 release
+\end{verbatim}
+\caption{Timeline}
+\end{figure}
 
@@
 one of the code bases, infrastructures, or communities 
-is often transferred in some form to the other.
+is often transferred in some form to the other. At the code
+base level, the \texttt{binned\_statistic} functionality
+is one such cross-project contribution---it was initially
+developed in an Astropy-affiliated package
+and then placed in SciPy after the fact.
 

### 2018-09-11 scipy-1.0/paper.tex
Outline background section

@@
 
-\subsection*{History}
+\section*{Background}
+
+Python ... late 1980s
+
+An matrix / n-dimensional array object ...
+matrix.py, numeric, numarray, numpy
+
+Paul Dubois
+
+Travis O
+
+Pearu
+
+-
+Gary Strangman
+
+%https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
+%
+
+Eric and Travis V
+
+\subsection*{SciPy Begins}
+
+-  Jun 10, 2001
+Travis N. Vaught announced scipy-dev mailing list
+
+- Aug 20, 2001
+Eric Jones announced
+
+IPython 2001 first release
+
+Matplotlib 2003 first release
+
+\begin{figure}
+\begin{verbatim}
+SciPy is an open source package that builds on the strengths of Python and
+Numeric providing a wide range of fast scientific and numeric functionality.
+SciPy's current module set includes the following:
+
+    Special Functions (Bessel, hanker, Airy, etc.)
+    Signal/Image Processing
+    2D Plotting capabilities
+    Integration
+    ODE solvers
+    Optimization (simplex, BFGS, Netwon-CG, etc.)
+    Genetic Algorithms
+    Numeric -> C++ expression compiler
+    Parallel programming tools
+    Splines and Interpolation
+    And other stuff.
+\end{verbatim}
+\caption{Excerpt from first release announcement.}
+\end{figure}
+
+\subsection*{Conferences}
+
+%https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
+
+SciPy
+
+US, India, Europe
+
+SIAM CSE
+
+PyData
+
+everywhere...
+
+\cite{siamcse09}
+
+%https://archive.siam.org/news/news.php?id=1595
+
+\subsection*{Publications}
+CiSE 2007 special issue
+
+CiSE 2011 special issue
 
@@
 \cite{meurer2017sympy}
+
+\subsection*{SciKits}
+
+% stefan, gael, david, jarrod,
+
+website http://scikits.appspot.com/
+
+statsmodels
+
+scikit-image
+
+scikit-learn
+\subsection*{Release History}
+
+ANN: SciPy 0.10 -- Scientific Computing with Python
 

### 2018-09-12 scipy-1.0/paper.tex
Write background section w/ Jarrod

@@
 \author[24]{Ian Henriksen}
+\author[3,25]{K. Jarrod Millman}
+\author[3]{St\'efan J. van der Walt}
 
@@
 \affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{UC Berkeley Institute for Data Science,
-          Berkeley, CA, 94720, USA}
+\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA, 94720, USA}
 \affil[4]{Los Alamos National Laboratory,
@@
 	   Austin, TX, 78712, USA}
+\affil[25]{Division of Biostatistics, University of California,
+  Berkeley, CA, 94720, USA}
 
@@
 
-Python ... late 1980s
+Python is an interactive computer programming language, designed by
+Guido van Rossum in the late 80s, with a heavy emphasis on readability
+and rapid prototyping.
+% https://legacy.python.org/search/hypermail/python-1992/0000.html
+%
+While it was a general purpose programming language with a ``batteries
+included'' standard library, it had no special support for scientific
+data structures or algorithms, unlike many of the other established computation
+platforms of the time. Yet, scientists soon discovered the language's
+virtues, such as its ability to wrap C and Fortran libraries and to
+then drive those libraries interactively.
+Scientists could thereby gain access to a wide variety of existing
+computational libraries, without concerning themselves with low-level
+programming concepts as as memory management.
 
-An matrix / n-dimensional array object ...
-matrix.py, numeric, numarray, numpy
+In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
+message on the Python Matrix Special Interest Group
+mailing list\footnote{https://mail.python.org/pipermail/matrix-sig/1995-August/000002.html}:
+\begin{quote}
+There seems to be a fair amount of interest in the Python community
+concerning the addition of numeric operations to Python.  My own desire is
+to have as large a library of matrix based functions available as possible
+(linear algebra, eigenfunctions, signal processing, statistics, etc.).  In
+order to ensure that all of these libraries interoperate, there needs to
+be agreement on a basic matrix object that can be used to represent arrays
+of numbers.
+\end{quote}
+Over the next several months, conversations on that mailing
+list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
+Hinsen, and Guido van Rossum, led to the creation of an array object
+with arbitrarily high dimensions.  Jim Hugunin explained the utility
+of Python for numerical computation\footnote{http://hugunin.net/papers/hugunin95numpy.html}:
+\begin{quote}
+I've used almost all of the available numerical languages at one time
+or another over the past 8 years. One thing I've noticed is that over
+time, the designers of these languages are steadily adding more of the
+features that one would expect to find in a general-purpose
+programming language.
+\end{quote}
+This is still one of the distinguishing features of Python for
+science, and why it has been so successful in the realm of data
+science: instead of adding general features to a language designed for
+numerical and scientific computing, here scientific features are added
+to a general purpose language instead.  This broadens the scope of
+problems that can be addressed easily, the sources of data that are
+readily accessible, and increases the size of the community that
+develops general purpose tooling.
 
-Paul Dubois
+The availability of a standard numerical array data structure in
+Python led to a sudden and rapidly growing number of scientific
+packages becoming available for solving various common numeric
+problems.
+
+A number of these packages were written by graduate students and
+postdoctoral researchers to solve the very practical research problems
+that they faced on a daily basis.  While they had access to and were
+familiar with specialized (and often commercial) systems, many found
+it easier to implement the domain-specific functionality they needed
+in a general purpose programming language.  And given Python's innate
+ability to function as a systems language, controlling specialized or
+custom-built hardware was feasible.
+
+\subsection*{SciPy Begins}
 
 Travis O
-
 Pearu
-
 -
 Gary Strangman
-
+%
 %https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
 %
-
+%
 Eric and Travis V
-
-\subsection*{SciPy Begins}
 

### 2018-09-12 scipy-1.0/paper.tex
Minor edits

@@
 computational libraries, without concerning themselves with low-level
-programming concepts as as memory management.
+programming concepts such as memory management.
 
@@
 \end{quote}
-This is still one of the distinguishing features of Python for
-science, and why it has been so successful in the realm of data
-science: instead of adding general features to a language designed for
-numerical and scientific computing, here scientific features are added
-to a general purpose language instead.  This broadens the scope of
-problems that can be addressed easily, the sources of data that are
-readily accessible, and increases the size of the community that
-develops general purpose tooling.
+This is still a distinguishing feature of Python for science, and one of the
+reasons why it has been so successful in the realm of data science: instead of
+adding general features to a language designed for numerical and scientific
+computing, here scientific features are added to a general purpose language
+instead.  This broadens the scope of problems that can be addressed easily, the
+sources of data that are readily accessible, and increases the size of the
+community that develops general purpose tooling.
 
@@
 problems.
-
 A number of these packages were written by graduate students and

### 2018-09-12 scipy-1.0/paper.tex
Start fleshing out 'SciPy begins' section

@@
 
+% Paul Dubois, CDAT
+% https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
+
+% Perry Greenfield, Hubble
+% https://conference.scipy.org/scipy2011/slides/greenfield_keynote_astronomy.pdf
+%
+% Note: plotting was important for them, so they eventually
+% funded development on Matplotlib.
+
+% Numeric, even though written by grad students, became critical for
+% large scale endeavors
+
 \subsection*{SciPy Begins}
 
-Travis O
-Pearu
--
-Gary Strangman
+%With a rich programming environment and a numerical array object in
+%place, the time was ripe for the development of a full scientific
+%software stack.
+
+% PhD Mayo Clinic, started faculty position at BYU in 2001
+By the late 90s, Travis Oliphant, a PhD student at the Mayo Clinic,
+released a number of packages\footnote{https://web.archive.org/web/19990125091242/http://oliphant.netpedia.net:80/}\footnote{https://web.archive.org/web/20001206213500/http://oliphant.netpedia.net:80/} that built on top of the Numeric array
+package, and provided algorithms for signal processing, special
+functions, sparse matrices, quadrature, optimization, and Fast Fourier
+Transforms, and more.  One of these packages, Multipack, was a set of
+extension modules that wrapped Fortran and C libraries such as
+ODEPACK, QUADPACK, and MINPACK, to solve and minimize nonlinear
+equations, integrate differential equations, and fit splines (the
+latter implemented by Pearu Peterson).  Robert Kern, then an
+undergraduate student (and currently a SciPy core developer), provided
+compilation instructions under Windows.
+% 2001 PhD Tallinn Technical University
+Around the same time, Pearu Peterson, a PhD student from Estonia,
+released f2py\footnote{https://mail.python.org/pipermail/numpy-discussion/2000-September/000243.html}, a command line tool for binding Python and Fortran
+codes.
+% Graduated 1999, Duke
+Eric Jones, while a graduate student at Duke, wrote a number of
+packages to support his dissertation, including a parallel job
+scheduler and genetic optimizer.
+% Graduated 1998, PhD from Brown, 97–99 postdoc at Harvard Medical
+% School; at Harvard ever since
 %
-%https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
+% https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
 %
-%
-Eric and Travis V
+Gary Strangman, a postdoctoral fellow at Harvard Medical School,
+published several descriptive and inferential statistical routines\footnote{https://web.archive.org/web/20001022231108/http://www.nmr.mgh.harvard.edu/Neural\_Systems\_Group/gary/python.html}.
+
+In 2001, Eric Jones and Travis Vaught founded Enthought Scientific
+Computing Solutions (now Enthought, Inc.) in Austin, Texas.  In an
+effort to simplify the tool stack, they created the SciPy project,
+centered around the SciPy library which would subsume all the
+above-mentioned packages.
 
@@
 - Aug 20, 2001
-Eric Jones announced
+Eric Jones announced \ref{fig:announce-0.1}
 
@@
 Matplotlib 2003 first release
+
+NumPy 2005
 
@@
 \end{verbatim}
-\caption{Excerpt from first release announcement.}
+\caption{Excerpt from first release announcement.}\label{fig:announce-0.1}
 \end{figure}

### 2018-09-13 scipy-1.0/paper.tex
first round of revisions from manuscript PR 63.

--- 2018-09-12 scipy-1.0/paper.tex
@@
 \author[24]{Ian Henriksen}
-\author[3,25]{K. Jarrod Millman}
-\author[3]{St\'efan J. van der Walt}
 
@@
 \affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA, 94720, USA}
+\affil[3]{UC Berkeley Institute for Data Science,
+          Berkeley, CA, 94720, USA}
 \affil[4]{Los Alamos National Laboratory,
@@
 	   Austin, TX, 78712, USA}
-\affil[25]{Division of Biostatistics, University of California,
-  Berkeley, CA, 94720, USA}
 
@@
 
-\section*{Background}
-
-Python is an interactive computer programming language, designed by
-Guido van Rossum in the late 80s, with a heavy emphasis on readability
-and rapid prototyping.
-% https://legacy.python.org/search/hypermail/python-1992/0000.html
-%
-While it was a general purpose programming language with a ``batteries
-included'' standard library, it had no special support for scientific
-data structures or algorithms, unlike many of the other established computation
-platforms of the time. Yet, scientists soon discovered the language's
-virtues, such as its ability to wrap C and Fortran libraries and to
-then drive those libraries interactively.
-Scientists could thereby gain access to a wide variety of existing
-computational libraries, without concerning themselves with low-level
-programming concepts as as memory management.
-
-In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
-message on the Python Matrix Special Interest Group
-mailing list\footnote{https://mail.python.org/pipermail/matrix-sig/1995-August/000002.html}:
-\begin{quote}
-There seems to be a fair amount of interest in the Python community
-concerning the addition of numeric operations to Python.  My own desire is
-to have as large a library of matrix based functions available as possible
-(linear algebra, eigenfunctions, signal processing, statistics, etc.).  In
-order to ensure that all of these libraries interoperate, there needs to
-be agreement on a basic matrix object that can be used to represent arrays
-of numbers.
-\end{quote}
-Over the next several months, conversations on that mailing
-list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
-Hinsen, and Guido van Rossum, led to the creation of an array object
-with arbitrarily high dimensions.  Jim Hugunin explained the utility
-of Python for numerical computation\footnote{http://hugunin.net/papers/hugunin95numpy.html}:
-\begin{quote}
-I've used almost all of the available numerical languages at one time
-or another over the past 8 years. One thing I've noticed is that over
-time, the designers of these languages are steadily adding more of the
-features that one would expect to find in a general-purpose
-programming language.
-\end{quote}
-This is still one of the distinguishing features of Python for
-science, and why it has been so successful in the realm of data
-science: instead of adding general features to a language designed for
-numerical and scientific computing, here scientific features are added
-to a general purpose language instead.  This broadens the scope of
-problems that can be addressed easily, the sources of data that are
-readily accessible, and increases the size of the community that
-develops general purpose tooling.
-
-The availability of a standard numerical array data structure in
-Python led to a sudden and rapidly growing number of scientific
-packages becoming available for solving various common numeric
-problems.
-
-A number of these packages were written by graduate students and
-postdoctoral researchers to solve the very practical research problems
-that they faced on a daily basis.  While they had access to and were
-familiar with specialized (and often commercial) systems, many found
-it easier to implement the domain-specific functionality they needed
-in a general purpose programming language.  And given Python's innate
-ability to function as a systems language, controlling specialized or
-custom-built hardware was feasible.
-
-\subsection*{SciPy Begins}
-
-Travis O
-Pearu
--
-Gary Strangman
-%
-%https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
-%
-%
-Eric and Travis V
-
--  Jun 10, 2001
-Travis N. Vaught announced scipy-dev mailing list
-
-- Aug 20, 2001
-Eric Jones announced
-
-IPython 2001 first release
-
-Matplotlib 2003 first release
-
-\begin{figure}
-\begin{verbatim}
-SciPy is an open source package that builds on the strengths of Python and
-Numeric providing a wide range of fast scientific and numeric functionality.
-SciPy's current module set includes the following:
-
-    Special Functions (Bessel, hanker, Airy, etc.)
-    Signal/Image Processing
-    2D Plotting capabilities
-    Integration
-    ODE solvers
-    Optimization (simplex, BFGS, Netwon-CG, etc.)
-    Genetic Algorithms
-    Numeric -> C++ expression compiler
-    Parallel programming tools
-    Splines and Interpolation
-    And other stuff.
-\end{verbatim}
-\caption{Excerpt from first release announcement.}
-\end{figure}
-
-\subsection*{Conferences}
-
-%https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
-
-SciPy
-
-US, India, Europe
-
-SIAM CSE
-
-PyData
-
-everywhere...
-
-\cite{siamcse09}
-
-%https://archive.siam.org/news/news.php?id=1595
-
-\subsection*{Publications}
-CiSE 2007 special issue
-
-CiSE 2011 special issue
-
-\cite{dubois2007guest}
-\cite{oliphant2007python}
-\cite{perez2007ipython}
-\cite{hunter2007matplotlib}
-\cite{vanderwalt2008scipy}
-\cite{peterson2009f2py}
-\cite{millman2011python}
-\cite{perez2011python}
-\cite{vanderwalt2011numpy}
-\cite{behnel2011cython}
-\cite{pedregosa2011scikit}
-\cite{vanderwalt2014scikit}
-\cite{millman2014developing}
-\cite{meurer2017sympy}
-
-\subsection*{SciKits}
-
-% stefan, gael, david, jarrod,
-
-website http://scikits.appspot.com/
-
-statsmodels
-
-scikit-image
-
-scikit-learn
-\subsection*{Release History}
-
-ANN: SciPy 0.10 -- Scientific Computing with Python
-
-\begin{figure}
-\begin{verbatim}
-    2001: first SciPy release
-    2005: transition to NumPy
-    2007: creation of scikits
-    2008: spatial module and first Cython code added
-    2010: move to a 6-monthly release cycle
-    2011: development moves to GitHub
-    2011: Python 3 support
-    2012: add a sparse graph module and unified optimization interface
-    2012: remove of maxentropy
-    2013: continuous integration with TravisCI
-    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
-    2017: add a unified C API with LowLevelCallable; remove weave
-    2017: SciPy 1.0 release
-\end{verbatim}
-\caption{Timeline}
-\end{figure}
+\subsection*{History}
 
@@
 is often transferred in some form to the other. At the code
-base level, the \texttt{binned\_statistic} functionality
-is one such cross-project contribution---it was initially
-developed in an Astropy-affiliated package
+base level, the \texttt{binned\_statistics} functionality
+was initially developed in an Astropy-affiliated package
 and then placed in SciPy after the fact.

### 2018-09-13 scipy-1.0/paper.tex
Extend background section to include widening of Numeric ecosystem

--- 2018-09-12 scipy-1.0/paper.tex
@@
 
+With a rich programming environment and a numerical array object in
+place, the time was ripe for the development of a full scientific
+software stack.  Many of the first efforts were driven by graduate
+students and post-doctoral researchers, but soon attracted the
+attention of senior researchers.
+
 % Paul Dubois, CDAT
 % https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
+Paul Dubois in the Program for Climate Model Diagnosis and
+Intercomparison at Lawrence Livermore National Laboratory built their
+Climate Data Analysis Tools (``an open-source set of Python modules
+and tools for accessing, manipulating, and plotting scientific data in
+general and climate data sets in
+particular''\footnote{https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html})
+on Numeric, the newly established Python array object.  Paul later
+became the maintainer for Numeric, and helped fund the writing of the
+first manual [citation needed].
 
@@
 \subsection*{SciPy Begins}
-
-%With a rich programming environment and a numerical array object in
-%place, the time was ripe for the development of a full scientific
-%software stack.
 
@@
 
-%https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
+% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
+% https://str.llnl.gov/str/News896.html
 

### 2018-09-13 scipy-1.0/paper.tex
second round of revisions from manuscript PR 63.

--- 2018-09-13 scipy-1.0/paper.tex
@@
 is often transferred in some form to the other. At the code
-base level, the \texttt{binned\_statistics} functionality
-was initially developed in an Astropy-affiliated package
+base level, the \texttt{binned\_statistic} functionality
+is one such cross-project contribution---it was initially
+developed in an Astropy-affiliated package
 and then placed in SciPy after the fact.

### 2018-09-13 scipy-1.0/paper.tex
Complete outline of SciPy begins section

--- 2018-09-13 scipy-1.0/paper.tex
@@
 \usepackage{pgfplots}
+\usepackage[multiple]{footmisc}
 
@@
 In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
-message on the Python Matrix Special Interest Group
+message on the Python Matrix Special Interest Group (Matrix-SIG)
 mailing list\footnote{https://mail.python.org/pipermail/matrix-sig/1995-August/000002.html}:
@@
 attention of senior researchers.
-
 % Paul Dubois, CDAT
 % https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
-Paul Dubois in the Program for Climate Model Diagnosis and
-Intercomparison at Lawrence Livermore National Laboratory built their
-Climate Data Analysis Tools (``an open-source set of Python modules
-and tools for accessing, manipulating, and plotting scientific data in
-general and climate data sets in
-particular''\footnote{https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html})
-on Numeric, the newly established Python array object.  Paul later
-became the maintainer for Numeric, and helped fund the writing of the
-first manual [citation needed].
+Paul Dubois at Lawrence Livermore National Laboratories took over the
+maintenance of Numeric and funded the writing of its
+manual\footnote{http://people.csail.mit.edu/jrennie/python/numeric/numeric-manual.pdf}.
+\fixme{replace this URL with numpy.org hosted version }% Make sure we host this PDF somewhere in the numpy GitHub org
+%in the Program for Climate Model Diagnosis and Intercomparison at
+%Lawrence Livermore National Laboratory built their Climate Data
+%Analysis Tools (``an open-source set of Python modules and tools for
+%accessing, manipulating, and plotting scientific data in general and
+%climate data sets in
+%particular''\footnote{https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html})
+%on Numeric, the newly established Python array object.  Paul later
+%became the maintainer for Numeric, and helped fund the writing of the
+%first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
+In 1998, the Space Telescope Science Institute, in charge of science
+operations of the Hubble Space Telescope, decided to replace their
+custom scripting language with Python\footnote{https://conference.scipy.org/scipy2011/slides/greenfield\_keynote\_astronomy.pdf}.
 
@@
 % PhD Mayo Clinic, started faculty position at BYU in 2001
-By the late 90s, Travis Oliphant, a PhD student at the Mayo Clinic,
+By the late 90s, discussions started to appear on Matrix-SIG,
+expressing a desire for a complete scientific data analysis environment\footnote{https://conference.scipy.org/scipy2010/slides/travis\_oliphant\_keynote.pdf}.
+Travis Oliphant, a PhD student at the Mayo Clinic,
 released a number of packages\footnote{https://web.archive.org/web/19990125091242/http://oliphant.netpedia.net:80/}\footnote{https://web.archive.org/web/20001206213500/http://oliphant.netpedia.net:80/} that built on top of the Numeric array
@@
 released f2py\footnote{https://mail.python.org/pipermail/numpy-discussion/2000-September/000243.html}, a command line tool for binding Python and Fortran
-codes.
+codes, and wrote modules for linear algebra and interpolation.
 % Graduated 1999, Duke
@@
 packages to support his dissertation, including a parallel job
-scheduler and genetic optimizer.
+scheduler, genetic optimizer, and inline C++ compiler.
 % Graduated 1998, PhD from Brown, 97–99 postdoc at Harvard Medical
@@
 above-mentioned packages.
+The new project quickly gained momentum, with a website and code
+repository appearing in
+February\footnote{https://web.archive.org/web/20010309040805/http://scipy.org:80/}, and
+a mailing list announced in
+June \footnote{https://mail.python.org/pipermail/scipy-dev/2001-June/000000.html}. By
+August, a first release was announced\footnote{https://mail.python.org/pipermail/python-list/2001-August/106419.html}, an excerpt of which is shown in
+Fig.~\ref{fig:announce-0.1}.
+In September, the first documentation was
+published\footnote{https://web.archive.org/web/20021013204556/http://www.scipy.org:80/scipy/site\_content/site\_news/docs\_released1},
+and by the following September the inaugural SciPy workshop was held
+at Caltech\footnote{https://mail.python.org/pipermail/numpy-discussion/2002-June/001511.html}.
 
--  Jun 10, 2001
-Travis N. Vaught announced scipy-dev mailing list
+% Janko Hauser on toolboxes
+% https://mail.python.org/pipermail/scipy-dev/2001-September/000032.html
 
-- Aug 20, 2001
-Eric Jones announced \ref{fig:announce-0.1}
+At this point, scientific Python is starting to attract more serious
+attention; code that started out as side projects by graduate students
+has grown into essential infrastructure at national laboratories and
+research institutes.  But even as SciPy, the algorithms library of the
+ecosystem, started to mature, the next generation of graduate students
+and postdocs were already exploring other aspects of the computational
+environment.
 
-IPython 2001 first release
+In 2000, Prabhu Ramachandran\fixme{also note that he, and the others
+  mentioned below, are also SciPy contributors}, a PhD student at the Indian Institute of
+Technology, started work on a 3D visualization application, based on
+Wingware's C++ Visualization Toolkit, called
+Mayavi\footnote{https://web.archive.org/web/20030731122452/http://pythonology.org/success\&story=mayavi}.
 
-Matplotlib 2003 first release
+SciPy initially contained a plotting module, but it did not have an
+easy and familiar interface. \fixme{There were multiple plotting
+  packages in SciPy, and Chaco, and wxPyPlot(s), IPython + GNUPlot,
+  and ....?}
+John Hunter, a gradute student
+\fixme{check, since John around that time became postdoc, can check
+  with John Milton?} at the University
+of Chicago Medical School, liked the plotting functionality
+available in MATLAB, but had problems accessing their laboratory's
+license, which was governed by a hardware dongle.  In response, he
+wrote a plotting library from scratch, and Matplotlib 0.1 was released
+April 2003\footnote{https://mail.python.org/pipermail/python-list/2003-April/193167.html}.
 
-NumPy 2005
+In 2001, Fernando Pérez, a graduate student at the University of
+Colorado, Boulder, created the IPython interactive shell, built into
+IPP, an earlier attempt by Janko Hauser, and including code from
+Nathan Gray's LazyPython.
+% IPython hosted on SciPy:
+% https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
+
+In 2000, STScI ported their entire IDL-based analysis pipeline, IRAF,
+to Python.  And it was here that a small crisis developed around the
+Python numerical array container.  Numeric, the original array package, was
+suitable for small arrays, but not for the large images processed by
+STScI.  With the Numeric maintainer's blessing, the decision was made
+to write NumArray, a library that could handle data on a different
+scale.  Unfortunately, NumArray proved inefficient for small arrays,
+facing the community with a rather unfortunate choice.  In 2005,
+Travis Oliphant combined the best elements of Numeric and NumArray,
+thereby solving the dilemma and paving the way for Python to become
+a very significant player in the Data Science movement. \fixme{mention
+  first NumPy meeting at Berkeley}
+
+% Future directions for SciPy after meeting at Berkeley
+% https://mail.python.org/pipermail/scipy-user/2005-March/004164.html
 
@@
 
+% While Paul was one of the early adopters at the national laboratories,
+% Python's presence is now ubiquitous.
+
 SciPy has a strong developer community and a massive user base. GitHub traffic

### 2018-09-15 scipy-1.0/paper.tex
Replace Numeric manual URL with new numpy.org hosted one

@@
 maintenance of Numeric and funded the writing of its
-manual\footnote{http://people.csail.mit.edu/jrennie/python/numeric/numeric-manual.pdf}.
-\fixme{replace this URL with numpy.org hosted version }% Make sure we host this PDF somewhere in the numpy GitHub org
+manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}.
 %in the Program for Climate Model Diagnosis and Intercomparison at

### 2018-09-18 scipy-1.0/paper.tex
Fix typos reported by Tyler

@@
 % https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
-Paul Dubois at Lawrence Livermore National Laboratories took over the
+Paul Dubois at Lawrence Livermore National Laboratory took over the
 maintenance of Numeric and funded the writing of its
@@
 scale.  Unfortunately, NumArray proved inefficient for small arrays,
-facing the community with a rather unfortunate choice.  In 2005,
+presenting the community with a rather unfortunate choice.  In 2005,
 Travis Oliphant combined the best elements of Numeric and NumArray,
@@
     2012: add a sparse graph module and unified optimization interface
-    2012: remove of maxentropy
+    2012: removal of maxentropy
     2013: continuous integration with TravisCI

### 2018-09-20 scipy-1.0/paper.tex
Add notes on introduction; transition into conferences section, expand.

@@
 
-\fixme{The Introduction section expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.}
+\begin{verbatim}
+SciPy is not what you think it is
+
+- At risk of failure many times
+- Funding was *hard* to come by; funding agencies didn't easily give
+  money.  Some labs, Enthough committed a bit of money.
+- Java and other enterprize packages were getting support; MATLAB
+  was huge.  Not many wanted to encourage this ``amateur'' effort.
+- Graduate students & post-docs doing the heavy lifting; they found
+  a system they enjoyed using, saw its potential, put energy into it.
+- Driven by eclectic group of people being passionate about it, not
+  because they were paid or told to work on it.
+- And not by EECS people, but by scientists.  Some with fairly little
+  programming background (David), others in varying stages of academic/research careers.
+- Over time, the project became more more professional:
+  - Standardized release formats
+  - Docstrings
+  - Testing & CI infrastructure
+  - APIs
+  - Issue tracking
+  - Revision control (CVS -> SVN -> Git)
+  - Conference
+- Python wasn't perfectly suited in all ways; packaging, compiling,
+  indexing (indexing/array protocols and complex numbers added)
+- Initially, this high quality, stable, widely used project, hinged on
+  three people doing the ``right'' thing to get it established.
+- Bus factor for maintenance still very low.
+
+The background section captures a selective history where we're trying
+to identify some of the more important events.  But really, it relies
+on the collective work of numerous individuals, too many to mention.
+\end{verbatim}
 
@@
 thereby solving the dilemma and paving the way for Python to become
-a very significant player in the Data Science movement. \fixme{mention
-  first NumPy meeting at Berkeley}
+a very significant player in the Data Science movement.
 
@@
 % https://mail.python.org/pipermail/scipy-user/2005-March/004164.html
+
+SciPy now begins a phase of significant growth and adoption.  The
+informal workshops are grown into international conferences with many
+hundreds of attendees, special issues are organized and published in a
+leading scientific magazine, and the scope of the SciPy library
+narrows, while the breadth of the ecosystem is grown and encouraged
+through a new type of auxhilary package, the scikit.  The tooling,
+development and release processes become more professional, grown with
+a patience only affordable in open source development, leading with
+best practices nowadays increasingly common in both industry and other
+open source projects \cite{millman2014developing}.
 
@@
 
-SciPy
-
-US, India, Europe
+In 2008, the SciPy conference went from being a small gathering of
+core developers, to a multi-location event, with increased funding, a
+published proceedings, and scholarships for attending students (some
+of whom are authors on this paper).  In 2010, the conference grew even
+further, with an additional event added in India, and the US
+conference—now with parallel tracks—moving to Austin, Texas.
 
@@
 \cite{vanderwalt2014scikit}
-\cite{millman2014developing}
 \cite{meurer2017sympy}

### 2018-09-20 scipy-1.0/paper.tex
Add note to include NumPy 1.0 release date

@@
 thereby solving the dilemma and paving the way for Python to become
-a very significant player in the Data Science movement.
+a very significant player in the Data Science movement.\fixme{NumPy
+  1.0 was released in October 2006}
 

### 2018-09-20 scipy-1.0/paper.tex
Add some links
Refine conference section

@@
 
+\begin{figure}
+\begin{verbatim}
+SciPy is an open source package that builds on the strengths of Python and
+Numeric providing a wide range of fast scientific and numeric functionality.
+SciPy's current module set includes the following:
+
+    Special Functions (Bessel, hanker, Airy, etc.)
+    Signal/Image Processing
+    2D Plotting capabilities
+    Integration
+    ODE solvers
+    Optimization (simplex, BFGS, Netwon-CG, etc.)
+    Genetic Algorithms
+    Numeric -> C++ expression compiler
+    Parallel programming tools
+    Splines and Interpolation
+    And other stuff.
+\end{verbatim}
+\caption{Excerpt from first release announcement.}\label{fig:announce-0.1}
+\end{figure}
+
 In 2001, Eric Jones and Travis Vaught founded Enthought Scientific
@@
 
-\begin{figure}
-\begin{verbatim}
-SciPy is an open source package that builds on the strengths of Python and
-Numeric providing a wide range of fast scientific and numeric functionality.
-SciPy's current module set includes the following:
-
-    Special Functions (Bessel, hanker, Airy, etc.)
-    Signal/Image Processing
-    2D Plotting capabilities
-    Integration
-    ODE solvers
-    Optimization (simplex, BFGS, Netwon-CG, etc.)
-    Genetic Algorithms
-    Numeric -> C++ expression compiler
-    Parallel programming tools
-    Splines and Interpolation
-    And other stuff.
-\end{verbatim}
-\caption{Excerpt from first release announcement.}\label{fig:announce-0.1}
-\end{figure}
-
 \subsection*{Conferences}
@@
 % https://str.llnl.gov/str/News896.html
+
+http://conference.scipy.org/proceedings/scipy2008/paper\_0/full\_text.pdf
+http://conference.scipy.org/proceedings/scipy2008/paper\_1/full\_text.pdf
+
+The first SciPy workshop was held in September 2002 at CalTech—a
+single track, two day event with 50 participants, many of them
+developers of SciPy and surrounding libraries.  In the early
+workshops, recurrent topics indicate the state of development, with
+emphasis being placed on the underlying array package, plotting,
+parallel processing, acceleration / wrapping, and user interfaces.  By
+2004, a significant shift occurred towards application of SciPy on
+scientific problems.  The event also started to draw in more keynote
+speakers from outside the community, such as Guido van Rossum (creator
+of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
+Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
+
+%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
+%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
+%2004—Jim Hugunin
+https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
+%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
+%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
+%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
+%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
+
+While the US SciPy conference kept growing (by ... it had multiple
+tracks, student sponsorship, proceedings, and funding), satellite
+events were being organized by volunteers elsewhere, such as EuroSciPy
+(2008–) and SciPy India (2009–).
 
@@
 conference—now with parallel tracks—moving to Austin, Texas.
+
+FOSSEE
 
@@
 
-% stefan, gael, david, jarrod,
+% stefan, gael, david, jarrod, robert
 
@@
 scikit-learn
+
+% scikits
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
+
+% roadmap
+%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
+
+% scikit learn
+%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
+
 \subsection*{Release History}
@@
 \end{figure}
+
+% Release management
+
+% Jarrod
+%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
+%Ralf
+%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
+%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
+%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
 

### 2018-09-28 scipy-1.0/paper.tex
Add links
Rearrange background
Write introduction

@@
 
-\begin{verbatim}
-SciPy is not what you think it is
-
-- At risk of failure many times
-- Funding was *hard* to come by; funding agencies didn't easily give
-  money.  Some labs, Enthough committed a bit of money.
-- Java and other enterprize packages were getting support; MATLAB
-  was huge.  Not many wanted to encourage this ``amateur'' effort.
-- Graduate students & post-docs doing the heavy lifting; they found
-  a system they enjoyed using, saw its potential, put energy into it.
-- Driven by eclectic group of people being passionate about it, not
-  because they were paid or told to work on it.
-- And not by EECS people, but by scientists.  Some with fairly little
-  programming background (David), others in varying stages of academic/research careers.
-- Over time, the project became more more professional:
-  - Standardized release formats
-  - Docstrings
-  - Testing & CI infrastructure
-  - APIs
-  - Issue tracking
-  - Revision control (CVS -> SVN -> Git)
-  - Conference
-- Python wasn't perfectly suited in all ways; packaging, compiling,
-  indexing (indexing/array protocols and complex numbers added)
-- Initially, this high quality, stable, widely used project, hinged on
-  three people doing the ``right'' thing to get it established.
-- Bus factor for maintenance still very low.
-
-The background section captures a selective history where we're trying
-to identify some of the more important events.  But really, it relies
-on the collective work of numerous individuals, too many to mention.
-\end{verbatim}
+%% SciPy is not what you think it is
+
+%% - At risk of failure many times
+%% - Funding was *hard* to come by; funding agencies didn't easily give
+%%   money.  Some labs, Enthough committed a bit of money.
+%% - Java and other enterprize packages were getting support; MATLAB
+%%   was huge.  Not many wanted to encourage this ``amateur'' effort.
+%% - Graduate students & post-docs doing the heavy lifting; they found
+%%   a system they enjoyed using, saw its potential, put energy into it.
+%% - Driven by eclectic group of people being passionate about it, not
+%%   because they were paid or told to work on it.
+%% - And not by EECS people, but by scientists.  Some with fairly little
+%%   programming background (David), others in varying stages of academic/research careers.
+%% - Over time, the project became more more professional:
+%%   - Standardized release formats
+%%   - Docstrings
+%%   - Testing & CI infrastructure
+%%   - APIs
+%%   - Issue tracking
+%%   - Revision control (CVS -> SVN -> Git)
+%%   - Conference
+%% - Python wasn't perfectly suited in all ways; packaging, compiling,
+%%   indexing (indexing/array protocols and complex numbers added)
+%% - Initially, this high quality, stable, widely used project, hinged on
+%%   three people doing the ``right'' thing to get it established.
+%% - Bus factor for maintenance still very low.
+
+%% The background section captures a selective history where we're trying
+%% to identify some of the more important events.  But really, it relies
+%% on the collective work of numerous individuals, too many to mention.
+
+SciPy recently released its first stable version: 1.0.  It signals to
+the user community that the package's API has now reached a level of
+maturity and stability that warrants trusting it inside research
+pipelines.  But the version numbering belies the history of a package
+that has already seen extensive adoption in research and industry.
+
+Upon historic introspection, its arrival at this point is surprising,
+and somewhat anomalous.  When first started, the library had very
+little funding, and was written mainly by graduate students---many of
+them without an official computer science education, or even the
+blessing of their advisor.  To even imagine that a small group of
+``rogue'' student programmers could upend the already well-established
+research software ecosystem---backed by millions in funding, and
+thousands of highly qualified engineers---was preposterous.
+
+Yet, the philosophical motivations behind a fully open toolstack, and
+an excited, friendly community with a singular focus, seemed to have
+proven decisive in the long run.  It led not only to the library
+described in this paper, but also to an entire ecosystem of related
+packages, and a variety of social activities around them.  The
+packages in the ecosystem share high standards of implementation,
+documentation, and testing, and a culture eager to learn and adopt
+better practices---both for community management and software
+development.
+
+Despite its success, SciPy still faces challenges: a low ``bus
+factor'' (the minimum number of contributors that needs to leave to
+cause serious difficult) being perhaps the most serious.
+
+In the background section that follows, we capture a selective history
+of some milestones and important events in the growth of SciPy.
+Despite what we might highlight here, it is important to understand
+that a project like SciPy is only possible because of the
+contributions of very many contributors---too many to mention
+individually, but each bringing an important part to the puzzle.
 
@@
 
-\subsection*{Conferences}
-
-% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
-% https://str.llnl.gov/str/News896.html
-
-http://conference.scipy.org/proceedings/scipy2008/paper\_0/full\_text.pdf
-http://conference.scipy.org/proceedings/scipy2008/paper\_1/full\_text.pdf
-
-The first SciPy workshop was held in September 2002 at CalTech—a
-single track, two day event with 50 participants, many of them
-developers of SciPy and surrounding libraries.  In the early
-workshops, recurrent topics indicate the state of development, with
-emphasis being placed on the underlying array package, plotting,
-parallel processing, acceleration / wrapping, and user interfaces.  By
-2004, a significant shift occurred towards application of SciPy on
-scientific problems.  The event also started to draw in more keynote
-speakers from outside the community, such as Guido van Rossum (creator
-of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
-Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
-
-%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
-%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
-%2004—Jim Hugunin
-https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
-%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
-%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
-%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
-%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
-
-While the US SciPy conference kept growing (by ... it had multiple
-tracks, student sponsorship, proceedings, and funding), satellite
-events were being organized by volunteers elsewhere, such as EuroSciPy
-(2008–) and SciPy India (2009–).
-
-In 2008, the SciPy conference went from being a small gathering of
-core developers, to a multi-location event, with increased funding, a
-published proceedings, and scholarships for attending students (some
-of whom are authors on this paper).  In 2010, the conference grew even
-further, with an additional event added in India, and the US
-conference—now with parallel tracks—moving to Austin, Texas.
-
-FOSSEE
-
-SIAM CSE
-
-PyData
-
-everywhere...
-
-\cite{siamcse09}
-
-%https://archive.siam.org/news/news.php?id=1595
-
-\subsection*{Publications}
+\subsection*{SciPy Matures}
+
+%Seeking articles for special issue on Python and Science and Engineering
+%Paul Dubois
+%Wed Apr 12 13:00:04 EDT 2006
+%https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html
+%
+%Python--Batteries Included
+%https://www.computer.org/csdl/mags/cs/2007/03/c3007.pdf
+%
+%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=4160244
+%
+%Python: Batteries Included
+%Paul F. Dubois
+%
+%Python for Scientific Computing
+%Travis E. Oliphant
+%
+%IPython: A System for Interactive Scientific Computing
+%Fernando Perez ; Brian E. Granger
+%
+%Computational Physics Education with Python
+%Arnd Backer
+%
+%Python Unleashed on Systems Biology
+%Christopher R. Myers ; Ryan N. Gutenkunst ; James P. Sethna
+%
+%Reaching for the Stars with Python
+%Perry Greenfield
+%
+%A Python Module for Modeling and Control Design of Flexible Robots
+%Ryan W. Krauss ; Wayne J. Book
+%
+%Python in Nanophotonics Research
+%Peter Bienstman ; Lieven Vanholme ; Wim Bogaerts ; Pieter Dumon ; Peter Vandersteegen
+%
+%Using Python to Solve Partial Differential Equations
+%Kent-Andre Mardal ; Ola Skavhaug ; Glenn T. Lines ; Gunnar A. Staff ; Asmund Odegard
+%
+%Analysis of Functional Magnetic Resonance Imaging in Python
+%K. Jarrod Millman ; Matthew Brett
+%
+%Python for Internet GIS Applications
+%Xuan Shi
+%
+%Python for Education: Computational Methods for Nonlinear Systems
+%Christopher R. Myers ; James P. Sethna
+%
+%Matplotlib: A 2D Graphics Environment
+%John D. Hunter
+%
+%-----
+%
+%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=5725228&punumber=5992
+%
+%Python for Scientists and Engineers
+%K. Jarrod Millman ; Michael Aivazis
+%
+%
+%Python: An Ecosystem for Scientific Computing
+%Fernando Perez ; Brian E. Granger ; John D. Hunter
+%
+%The NumPy Array: A Structure for Efficient Numerical Computation
+%Stefan van der Walt ; S. Chris Colbert ; Gael Varoquaux
+%
+%Cython: The Best of Both Worlds
+%Stefan Behnel ; Robert Bradshaw ; Craig Citro ; Lisandro Dalcin ; Dag Sverre Seljebotn ; Kurt Smith
+%
+%Mayavi: 3D Visualization of Scientific Data
+%Prabhu Ramachandran ; Gael Varoquaux
+%
+%??? these were in the same issue
+%
+%web2py for Scientific Applications
+%Massimo Di Pierro
+%
+%From Equations to Code: Automated Scientific Computing
+%Andy R. Terrel
+%https://www.researchgate.net/publication/260585856_From_Equations_to_Code_Automated_Scientific_Computing
+
 CiSE 2007 special issue
@@
 
-\subsection*{SciKits}
-
-% stefan, gael, david, jarrod, robert
-
-website http://scikits.appspot.com/
-
-statsmodels
-
-scikit-image
-
-scikit-learn
-
-% scikits
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
-
-% roadmap
-%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
-
-% scikit learn
-%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
-
-\subsection*{Release History}
-
-ANN: SciPy 0.10 -- Scientific Computing with Python
+% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
+% https://str.llnl.gov/str/News896.html
+
+http://conference.scipy.org/proceedings/scipy2008/paper\_0/full\_text.pdf
+http://conference.scipy.org/proceedings/scipy2008/paper\_1/full\_text.pdf
+
+The first SciPy workshop was held in September 2002 at CalTech—a
+single track, two day event with 50 participants, many of them
+developers of SciPy and surrounding libraries.  In the early
+workshops, recurrent topics indicate the state of development, with
+emphasis being placed on the underlying array package, plotting,
+parallel processing, acceleration / wrapping, and user interfaces.  By
+2004, a significant shift occurred towards application of SciPy on
+scientific problems.  The event also started to draw in more keynote
+speakers from outside the community, such as Guido van Rossum (creator
+of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
+Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
+
+%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
+%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
+%2004—Jim Hugunin
+https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
+%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
+%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
+%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
+%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
+
+While the US SciPy conference kept growing (by ... it had multiple
+tracks, student sponsorship, proceedings, and funding), satellite
+events were being organized by volunteers elsewhere, such as EuroSciPy
+(2008–) and SciPy India (2009–).
+
+In 2008, the SciPy conference went from being a small gathering of
+core developers, to a multi-location event, with increased funding, a
+published proceedings, and scholarships for attending students (some
+of whom are authors on this paper).  In 2010, the conference grew even
+further, with an additional event added in India, and the US
+conference—now with parallel tracks—moving to Austin, Texas.
+
+FOSSEE
+
+SIAM CSE
+
+PyData
+
+everywhere...
+
+\cite{siamcse09}
+
+%https://archive.siam.org/news/news.php?id=1595
+
+\subsection*{Project scope}
+
+SciPy aims to provide fundamental algorithms for scientific computing. The
+breadth of its scope was derived from the Guide to Available Mathematical
+Software classification system (GAMS\cite{boisvert1991guide}). In areas
+that move relatively slowly, e.g. linear algebra, SciPy aims to provide
+complete coverage. In other areas it aims to provide fundamental building
+blocks while interacting well with other packages specialized in that area.
+For example, SciPy provides what one expect to find in a
+statistics textbook (probability distributions, hypothesis tests, frequency
+statistics, correlation functions, and more), while Statsmodels provides
+more advanced statistical estimators and inference methods, scikit-learn
+covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
+and probabilistic modeling.
+
+We use the following criteria to determine whether or not to include new
+functionality in SciPy:
+\begin{itemize}
+    \item The algorithm is of relevance to multiple fields of science.
+    \item The algorithm is demonstrably important.  E.g., it is classic
+    enough to be included in textbooks, or it is based on a peer-reviewed article
+    which has a significant number of citations.
+\end{itemize}
+
+In terms of software systems and architecture, SciPy's scope matches NumPy's:
+algorithms for in-memory computing on single machines, with support for a wide
+range of data types and process architectures. Distributed computing and support
+for graphics processing units (GPUs) is explicitly out of scope.
+
+\subsection*{Current status (maturity, users)}
+
+SciPy 1.0 was released ... see timeline ...
+
+1.1 ...
+
+1.2
+
+Python 2 / 3?
 
@@
 
-\subsection*{Project scope}
-
-SciPy aims to provide fundamental algorithms for scientific computing. The
-breadth of its scope was derived from the Guide to Available Mathematical
-Software classification system (GAMS\cite{boisvert1991guide}). In areas
-that move relatively slowly, e.g. linear algebra, SciPy aims to provide
-complete coverage. In other areas it aims to provide fundamental building
-blocks while interacting well with other packages specialized in that area.
-For example, SciPy provides what one expect to find in a
-statistics textbook (probability distributions, hypothesis tests, frequency
-statistics, correlation functions, and more), while Statsmodels provides
-more advanced statistical estimators and inference methods, scikit-learn
-covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
-and probabilistic modeling.
-
-We use the following criteria to determine whether or not to include new
-functionality in SciPy:
-\begin{itemize}
-    \item The algorithm is of relevance to multiple fields of science.
-    \item The algorithm is demonstrably important.  E.g., it is classic
-    enough to be included in textbooks, or it is based on a peer-reviewed article
-    which has a significant number of citations.
-\end{itemize}
-
-In terms of software systems and architecture, SciPy's scope matches NumPy's:
-algorithms for in-memory computing on single machines, with support for a wide
-range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) is explicitly out of scope.
-
-\subsection*{Current status (maturity, users)}
+%\subsection*{SciKits}
+
+% stefan, gael, david, jarrod, robert
+
+website http://scikits.appspot.com/
+
+statsmodels
+
+scikit-image
+
+scikit-learn
+
+% scikits
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
+
+% roadmap
+%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
+
+% scikit learn
+%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
 

### 2018-09-28 scipy-1.0/paper.tex
Minor language edits to background section

@@
 SciPy recently released its first stable version: 1.0.  It signals to
-the user community that the package's API has now reached a level of
+the user community that the library's API has now reached a level of
 maturity and stability that warrants trusting it inside research
@@
 little funding, and was written mainly by graduate students---many of
-them without an official computer science education, or even the
-blessing of their advisor.  To even imagine that a small group of
+them without a computer science education, and often without the
+blessing of their advisors.  To even imagine that a small group of
 ``rogue'' student programmers could upend the already well-established
-research software ecosystem---backed by millions in funding, and
+ecosystem of research software---backed by millions in funding, and
 thousands of highly qualified engineers---was preposterous.
@@
 described in this paper, but also to an entire ecosystem of related
-packages, and a variety of social activities around them.  The
-packages in the ecosystem share high standards of implementation,
+packages, and a variety of social activities centered around them.  The
+packages in the SciPy ecosystem share high standards of implementation,
 documentation, and testing, and a culture eager to learn and adopt
@@
 Despite its success, SciPy still faces challenges: a low ``bus
-factor'' (the minimum number of contributors that needs to leave to
-cause serious difficult) being perhaps the most serious.
+factor''---the minimum number of contributors that needs to leave to
+cause serious difficulty---being perhaps the most serious.
 
@@
 %
-While it was a general purpose programming language with a ``batteries
-included'' standard library, it had no special support for scientific
-data structures or algorithms, unlike many of the other established computation
-platforms of the time. Yet, scientists soon discovered the language's
-virtues, such as its ability to wrap C and Fortran libraries and to
-then drive those libraries interactively.
-Scientists could thereby gain access to a wide variety of existing
-computational libraries, without concerning themselves with low-level
-programming concepts such as memory management.
+A general purpose programming language with a ``batteries included''
+standard library, it had no special support for scientific data
+structures or algorithms, unlike many of the other established
+computation platforms of the time. Yet, scientists soon discovered the
+language's virtues, such as its ability to wrap C and Fortran
+libraries and to then drive those libraries interactively.  Scientists
+could thereby gain access to a wide variety of existing computational
+libraries, without concerning themselves with low-level programming
+concepts such as memory management.
 
@@
 Hinsen, and Guido van Rossum, led to the creation of an array object
-with arbitrarily high dimensions.  Jim Hugunin explained the utility
+that supported a high number of dimensions.  Jim Hugunin explained the utility
 of Python for numerical computation\footnote{http://hugunin.net/papers/hugunin95numpy.html}:
@@
 that they faced on a daily basis.  While they had access to and were
-familiar with specialized (and often commercial) systems, many found
+familiar with specialized (often commercial) systems, many found
 it easier to implement the domain-specific functionality they needed
@@
 students and post-doctoral researchers, but soon attracted the
-attention of senior researchers.
+attention of more senior researchers.
 % Paul Dubois, CDAT
 % https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
-Paul Dubois at Lawrence Livermore National Laboratory took over the
+For example, Paul Dubois at Lawrence Livermore National Laboratory took over the
 maintenance of Numeric and funded the writing of its
-manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}.
+manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}, and
 %in the Program for Climate Model Diagnosis and Intercomparison at
@@
 %first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
-In 1998, the Space Telescope Science Institute, in charge of science
+in 1998 the Space Telescope Science Institute, in charge of science
 operations of the Hubble Space Telescope, decided to replace their

### 2018-09-28 scipy-1.0/paper.tex
Revise background

@@
 
-With a rich programming environment and a numerical array object in
-place, the time was ripe for the development of a full scientific
-software stack. Many of the first efforts were driven by graduate
-students and post-doctoral researchers, but soon attracted the
-attention of more senior researchers.
-% Paul Dubois, CDAT
-% https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
+Initial work on numerical computing in Python was driven by graduate
+students, but soon larger research labs became increasingly engaged.
 For example, Paul Dubois at Lawrence Livermore National Laboratory took over the
@@
 
+Many of the first efforts were driven by graduate
+students and post-doctoral researchers, but soon attracted the
+attention of more senior researchers.
+% Paul Dubois, CDAT
+% https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
+For example,
 % Perry Greenfield, Hubble
@@
 
+With a rich programming environment and a numerical array object in
+place, the time was ripe for the development of a full scientific
+software stack.
 In 2001, Eric Jones and Travis Vaught founded Enthought Scientific
@@
 In September, the first documentation was
-published\footnote{https://web.archive.org/web/20021013204556/http://www.scipy.org:80/scipy/site\_content/site\_news/docs\_released1},
-and by the following September the inaugural SciPy workshop was held
-at Caltech\footnote{https://mail.python.org/pipermail/numpy-discussion/2002-June/001511.html}.
+published\footnote{https://web.archive.org/web/20021013204556/http://www.scipy.org:80/scipy/site\_content/site\_news/docs\_released1}.
+The first SciPy
+workshop\footnote{https://mail.python.org/pipermail/numpy-discussion/2002-June/001511.html}
+was held in September 2002 at CalTech---a single track, two day event with 50
+participants, many of them developers of SciPy and surrounding libraries.
+
+%and by the following September the inaugural SciPy workshop was held
+%at Caltech\footnote{https://mail.python.org/pipermail/numpy-discussion/2002-June/001511.html}.
 
@@
 
-At this point, scientific Python is starting to attract more serious
-attention; code that started out as side projects by graduate students
-has grown into essential infrastructure at national laboratories and
-research institutes.  But even as SciPy, the algorithms library of the
-ecosystem, started to mature, the next generation of graduate students
-and postdocs were already exploring other aspects of the computational
-environment.
-
+At this point, scientific Python started attracting more serious attention;
+code that started out as side projects by graduate students has grown into
+essential infrastructure at national laboratories and research institutes.
+For example, starting in 2000, the Space Telescope Science Institute (STScI)
+ported their entire IDL-based analysis pipeline, IRAF, to Python.
+STScI was founded to guide the Hubble Space Telescope including developing
+advanced data, archives, and analysis tools for the most famous observatory
+in history.
+
+As SciPy, the algorithms library of the ecosystem, began attracting the attention
+of large research organizations,
+the next generation of graduate students and postdocs were already exploring
+other aspects of the computational environment.
 In 2000, Prabhu Ramachandran\fixme{also note that he, and the others
@@
 Mayavi\footnote{https://web.archive.org/web/20030731122452/http://pythonology.org/success\&story=mayavi}.
-
-SciPy initially contained a plotting module, but it did not have an
-easy and familiar interface. \fixme{There were multiple plotting
-  packages in SciPy, and Chaco, and wxPyPlot(s), IPython + GNUPlot,
-  and ....?}
+In 2001, Fernando Pérez, a graduate student at the University of
+Colorado, Boulder, created the IPython interactive shell, built into
+IPP, an earlier attempt by Janko Hauser, and including code from
+Nathan Gray's LazyPython.
+% IPython hosted on SciPy:
+% https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
 John Hunter, a gradute student
@@
 
-In 2001, Fernando Pérez, a graduate student at the University of
-Colorado, Boulder, created the IPython interactive shell, built into
-IPP, an earlier attempt by Janko Hauser, and including code from
-Nathan Gray's LazyPython.
-% IPython hosted on SciPy:
-% https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
-
-In 2000, STScI ported their entire IDL-based analysis pipeline, IRAF,
-to Python.  And it was here that a small crisis developed around the
-Python numerical array container.  Numeric, the original array package, was
+% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
+% https://str.llnl.gov/str/News896.html
+
+In the early
+workshops, recurrent topics indicate the state of development, with
+emphasis being placed on the underlying array package, plotting,
+parallel processing, acceleration / wrapping, and user interfaces.  By
+2004, a significant shift occurred towards application of SciPy on
+scientific problems.  The event also started to draw in more keynote
+speakers from outside the community, such as Guido van Rossum (creator
+of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
+Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
+
+%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
+%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
+%2004—Jim Hugunin
+%https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
+%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
+%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
+%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
+%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
+
+As STScI continued to use Python for an increasingly large portion
+of the Hubble Space Telescope data analysis pipeline, they encountered
+problems with the Python numerical array container.
+Numeric, the original array package, was
 suitable for small arrays, but not for the large images processed by
@@
 % https://mail.python.org/pipermail/scipy-user/2005-March/004164.html
+
+CiSE 2007 special issue (steering code)
+\cite{dubois2007guest}
+\cite{oliphant2007python}
+\cite{perez2007ipython}
+\cite{hunter2007matplotlib}
+research applications ...
+
+\subsection*{SciPy Matures}
+
+%Seeking articles for special issue on Python and Science and Engineering
+%Paul Dubois
+%Wed Apr 12 13:00:04 EDT 2006
+%https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html
+%
+%Python--Batteries Included
+%https://www.computer.org/csdl/mags/cs/2007/03/c3007.pdf
+%
+%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=4160244
+%
+%Python: Batteries Included
+%Paul F. Dubois
+%
+%Python for Scientific Computing
+%Travis E. Oliphant
+%
+%IPython: A System for Interactive Scientific Computing
+%Fernando Perez ; Brian E. Granger
+%
+%Computational Physics Education with Python
+%Arnd Backer
+%
+%Python Unleashed on Systems Biology
+%Christopher R. Myers ; Ryan N. Gutenkunst ; James P. Sethna
+%
+%Reaching for the Stars with Python
+%Perry Greenfield
+%
+%A Python Module for Modeling and Control Design of Flexible Robots
+%Ryan W. Krauss ; Wayne J. Book
+%
+%Python in Nanophotonics Research
+%Peter Bienstman ; Lieven Vanholme ; Wim Bogaerts ; Pieter Dumon ; Peter Vandersteegen
+%
+%Using Python to Solve Partial Differential Equations
+%Kent-Andre Mardal ; Ola Skavhaug ; Glenn T. Lines ; Gunnar A. Staff ; Asmund Odegard
+%
+%Analysis of Functional Magnetic Resonance Imaging in Python
+%K. Jarrod Millman ; Matthew Brett
+%
+%Python for Internet GIS Applications
+%Xuan Shi
+%
+%Python for Education: Computational Methods for Nonlinear Systems
+%Christopher R. Myers ; James P. Sethna
+%
+%Matplotlib: A 2D Graphics Environment
+%John D. Hunter
+%
+%-----
+%
+%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=5725228&punumber=5992
+%
+%Python for Scientists and Engineers
+%K. Jarrod Millman ; Michael Aivazis
+%
+%
+%Python: An Ecosystem for Scientific Computing
+%Fernando Perez ; Brian E. Granger ; John D. Hunter
+%
+%The NumPy Array: A Structure for Efficient Numerical Computation
+%Stefan van der Walt ; S. Chris Colbert ; Gael Varoquaux
+%
+%Cython: The Best of Both Worlds
+%Stefan Behnel ; Robert Bradshaw ; Craig Citro ; Lisandro Dalcin ; Dag Sverre Seljebotn ; Kurt Smith
+%
+%Mayavi: 3D Visualization of Scientific Data
+%Prabhu Ramachandran ; Gael Varoquaux
+%
+%??? these were in the same issue
+%
+%web2py for Scientific Applications
+%Massimo Di Pierro
+%
+%From Equations to Code: Automated Scientific Computing
+%Andy R. Terrel
+%https://www.researchgate.net/publication/260585856_From_Equations_to_Code_Automated_Scientific_Computing
 
@@
 
-\subsection*{SciPy Matures}
-
-%Seeking articles for special issue on Python and Science and Engineering
-%Paul Dubois
-%Wed Apr 12 13:00:04 EDT 2006
-%https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html
-%
-%Python--Batteries Included
-%https://www.computer.org/csdl/mags/cs/2007/03/c3007.pdf
-%
-%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=4160244
-%
-%Python: Batteries Included
-%Paul F. Dubois
-%
-%Python for Scientific Computing
-%Travis E. Oliphant
-%
-%IPython: A System for Interactive Scientific Computing
-%Fernando Perez ; Brian E. Granger
-%
-%Computational Physics Education with Python
-%Arnd Backer
-%
-%Python Unleashed on Systems Biology
-%Christopher R. Myers ; Ryan N. Gutenkunst ; James P. Sethna
-%
-%Reaching for the Stars with Python
-%Perry Greenfield
-%
-%A Python Module for Modeling and Control Design of Flexible Robots
-%Ryan W. Krauss ; Wayne J. Book
-%
-%Python in Nanophotonics Research
-%Peter Bienstman ; Lieven Vanholme ; Wim Bogaerts ; Pieter Dumon ; Peter Vandersteegen
-%
-%Using Python to Solve Partial Differential Equations
-%Kent-Andre Mardal ; Ola Skavhaug ; Glenn T. Lines ; Gunnar A. Staff ; Asmund Odegard
-%
-%Analysis of Functional Magnetic Resonance Imaging in Python
-%K. Jarrod Millman ; Matthew Brett
-%
-%Python for Internet GIS Applications
-%Xuan Shi
-%
-%Python for Education: Computational Methods for Nonlinear Systems
-%Christopher R. Myers ; James P. Sethna
-%
-%Matplotlib: A 2D Graphics Environment
-%John D. Hunter
-%
-%-----
-%
-%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=5725228&punumber=5992
-%
-%Python for Scientists and Engineers
-%K. Jarrod Millman ; Michael Aivazis
-%
-%
-%Python: An Ecosystem for Scientific Computing
-%Fernando Perez ; Brian E. Granger ; John D. Hunter
-%
-%The NumPy Array: A Structure for Efficient Numerical Computation
-%Stefan van der Walt ; S. Chris Colbert ; Gael Varoquaux
-%
-%Cython: The Best of Both Worlds
-%Stefan Behnel ; Robert Bradshaw ; Craig Citro ; Lisandro Dalcin ; Dag Sverre Seljebotn ; Kurt Smith
-%
-%Mayavi: 3D Visualization of Scientific Data
-%Prabhu Ramachandran ; Gael Varoquaux
-%
-%??? these were in the same issue
-%
-%web2py for Scientific Applications
-%Massimo Di Pierro
-%
-%From Equations to Code: Automated Scientific Computing
-%Andy R. Terrel
-%https://www.researchgate.net/publication/260585856_From_Equations_to_Code_Automated_Scientific_Computing
-
-CiSE 2007 special issue
+%http://conference.scipy.org/proceedings/scipy2008/paper\_0/full\_text.pdf
+%http://conference.scipy.org/proceedings/scipy2008/paper\_1/full\_text.pdf
+
+In 2008, the SciPy conference went from being a small gathering of
+core developers, to a multi-location event, with increased funding, a
+published proceedings, and scholarships for attending students (some
+of whom are authors on this paper).  In 2010, the conference grew even
+further, with an additional event added in India, and the US
+conference—now with parallel tracks—moving to Austin, Texas.
+While the US SciPy conference kept growing (by ... it had multiple
+tracks, student sponsorship, proceedings, and funding), satellite
+events were being organized by volunteers elsewhere, such as EuroSciPy
+(2008–) and SciPy India (2009–).
 
@@
 
-\cite{dubois2007guest}
-\cite{oliphant2007python}
-\cite{perez2007ipython}
-\cite{hunter2007matplotlib}
 \cite{vanderwalt2008scipy}
@@
 
-% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
-% https://str.llnl.gov/str/News896.html
-
-http://conference.scipy.org/proceedings/scipy2008/paper\_0/full\_text.pdf
-http://conference.scipy.org/proceedings/scipy2008/paper\_1/full\_text.pdf
-
-The first SciPy workshop was held in September 2002 at CalTech—a
-single track, two day event with 50 participants, many of them
-developers of SciPy and surrounding libraries.  In the early
-workshops, recurrent topics indicate the state of development, with
-emphasis being placed on the underlying array package, plotting,
-parallel processing, acceleration / wrapping, and user interfaces.  By
-2004, a significant shift occurred towards application of SciPy on
-scientific problems.  The event also started to draw in more keynote
-speakers from outside the community, such as Guido van Rossum (creator
-of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
-Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
-
-%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
-%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
-%2004—Jim Hugunin
-https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
-%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
-%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
-%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
-%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
-
-While the US SciPy conference kept growing (by ... it had multiple
-tracks, student sponsorship, proceedings, and funding), satellite
-events were being organized by volunteers elsewhere, such as EuroSciPy
-(2008–) and SciPy India (2009–).
-
-In 2008, the SciPy conference went from being a small gathering of
-core developers, to a multi-location event, with increased funding, a
-published proceedings, and scholarships for attending students (some
-of whom are authors on this paper).  In 2010, the conference grew even
-further, with an additional event added in India, and the US
-conference—now with parallel tracks—moving to Austin, Texas.
-
 FOSSEE
@@
 %https://archive.siam.org/news/news.php?id=1595
+
+%\subsection*{SciKits}
+
+% stefan, gael, david, jarrod, robert
+
+website http://scikits.appspot.com/
+
+statsmodels
+
+scikit-image
+
+scikit-learn
+
+% scikits
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
+%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
+%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
+
+% roadmap
+%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
+
+% scikit learn
+%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
 
@@
 %https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
-
-%\subsection*{SciKits}
-
-% stefan, gael, david, jarrod, robert
-
-website http://scikits.appspot.com/
-
-statsmodels
-
-scikit-image
-
-scikit-learn
-
-% scikits
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
-
-% roadmap
-%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
-
-% scikit learn
-%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
 

### 2018-10-05 scipy-1.0/paper.tex
Add missing reference
Proofreading

@@
 
-SciPy recently released its first stable version: 1.0.  It signals to
-the user community that the library's API has now reached a level of
-maturity and stability that warrants trusting it inside research
-pipelines.  But the version numbering belies the history of a package
-that has already seen extensive adoption in research and industry.
-
-Upon historic introspection, its arrival at this point is surprising,
-and somewhat anomalous.  When first started, the library had very
-little funding, and was written mainly by graduate students---many of
-them without a computer science education, and often without the
+Traditionally, a 1.0 release signals that a
+library's API has reached a level of maturity and stability
+that warrants trusting it inside research pipelines.
+The SciPy 1.0 version numbering belies the history of a project
+that has become the standard which others follow
+and has seen extensive adoption in research and industry.
+
+SciPy's arrival at this point is surprising, and somewhat anomalous.
+When started, the library had little funding, and was written mainly
+by graduate students---many of them without a computer science education, and often without the
 blessing of their advisors.  To even imagine that a small group of
@@
 
-Despite its success, SciPy still faces challenges: a low ``bus
-factor''---the minimum number of contributors that needs to leave to
-cause serious difficulty---being perhaps the most serious.
+Despite its success, SciPy still faces challenges:
+a low ``bus factor''---from the phrase ``hit by a bus'' and
+meaning that the project depends on a small number of
+specific contributors---being perhaps the most serious.
 
@@
 Python is an interactive computer programming language, designed by
-Guido van Rossum in the late 80s, with a heavy emphasis on readability
+Guido van Rossum in the late 1980s, with an emphasis on readability
 and rapid prototyping.
@@
 students, but soon larger research labs became increasingly engaged.
-For example, Paul Dubois at Lawrence Livermore National Laboratory took over the
+For example, Paul Dubois at Lawrence Livermore National Laboratory (LLNL) took over the
 maintenance of Numeric and funded the writing of its
@@
 %first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
-in 1998 the Space Telescope Science Institute, in charge of science
+in 1998 the Space Telescope Science Institute (STScI), in charge of science
 operations of the Hubble Space Telescope, decided to replace their
@@
 
-Many of the first efforts were driven by graduate
-students and post-doctoral researchers, but soon attracted the
-attention of more senior researchers.
+%Many of the first efforts were driven by graduate
+%students and post-doctoral researchers, but soon attracted the
+%attention of more senior researchers.
 % Paul Dubois, CDAT
 % https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
-For example,
 % Perry Greenfield, Hubble
@@
 % PhD Mayo Clinic, started faculty position at BYU in 2001
-By the late 90s, discussions started to appear on Matrix-SIG,
+By the late 1990s, discussions appeared on Matrix-SIG,
 expressing a desire for a complete scientific data analysis environment\footnote{https://conference.scipy.org/scipy2010/slides/travis\_oliphant\_keynote.pdf}.
@@
 \end{verbatim}
-\caption{Excerpt from first release announcement.}\label{fig:announce-0.1}
+\caption{Excerpt from SciPy 0.1 release announcement.}\label{fig:announce-0.1}
 \end{figure}
@@
 essential infrastructure at national laboratories and research institutes.
-For example, starting in 2000, the Space Telescope Science Institute (STScI)
-ported their entire IDL-based analysis pipeline, IRAF, to Python.
-STScI was founded to guide the Hubble Space Telescope including developing
-advanced data, archives, and analysis tools for the most famous observatory
-in history.
+For example, starting in 2000, STScI ported their IDL-based analysis pipeline,
+IRAF, to Python.
 
@@
 % https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
-John Hunter, a gradute student
+John Hunter, a graduate student
 \fixme{check, since John around that time became postdoc, can check
@@
 % https://str.llnl.gov/str/News896.html
-
-In the early
-workshops, recurrent topics indicate the state of development, with
-emphasis being placed on the underlying array package, plotting,
-parallel processing, acceleration / wrapping, and user interfaces.  By
-2004, a significant shift occurred towards application of SciPy on
-scientific problems.  The event also started to draw in more keynote
-speakers from outside the community, such as Guido van Rossum (creator
-of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
-Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
 
@@
 
-CiSE 2007 special issue (steering code)
-\cite{dubois2007guest}
+In April 2006, Paul
+Dubois\footnote{https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html} wrote
+to the NumPy mailing list:
+\begin{quote}
+IEEE's magazine, Computing in Science and Engineering (CiSE), has asked me
+to put together a theme issue on the use of Python in Science and
+Engineering. I will write an overview to be accompanied by 3-5 articles [. . .]
+that show a diverse set of applications or
+tools, to give our readers a sense of whether or not Python might be useful
+in their own work.
+\end{quote}
+Python had a strong enough presence in science and engineering so that
+editors were soliciting a special issue, but it was still niche enough
+that the average reader would need additional information in order
+to decide whether it would ``be useful in their own work.''
+In the resulting CiSE May/June 2007 special issue, Paul's overview\cite{dubois2007guest}
+stated:
+\begin{quote}
+LLNL now has many Python-based efforts built from scratch or wrapped around
+legacy codes [. . .] hundreds of thousands
+of lines of C++, Python, and Fortran 95, all working together just as we hoped,
+doing compute-intensive calculations on massively parallel computers.
+\end{quote}
+%https://www.computer.org/csdl/mags/cs/2007/03/c3007.html
+(steering code)
 \cite{oliphant2007python}
@@
 
-SciPy now begins a phase of significant growth and adoption.  The
-informal workshops are grown into international conferences with many
+By the middle to late 2000s, SciPy matured after a long phase of significant
+growth and adoption.
+The informal workshops are grown into international conferences with many
 hundreds of attendees, special issues are organized and published in a
@@
 narrows, while the breadth of the ecosystem is grown and encouraged
-through a new type of auxhilary package, the scikit.  The tooling,
+through a new type of auxiliary package, the scikit.  The tooling,
 development and release processes become more professional, grown with
@@
 
-In 2008, the SciPy conference went from being a small gathering of
+In the early workshops, recurrent topics indicate the state of development, with
+emphasis being placed on the underlying array package, plotting,
+parallel processing, acceleration / wrapping, and user interfaces.  By
+2004, a significant shift occurred towards application of SciPy on
+scientific problems.  The event also started to draw in more keynote
+speakers from outside the community, such as Guido van Rossum (creator
+of Python, 2006), Ivan Krstić (One Laptop per Child, 2007), Alex
+Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
+The SciPy conference went from being a small gathering of
 core developers, to a multi-location event, with increased funding, a
@@
 
-SciPy aims to provide fundamental algorithms for scientific computing. The
+SciPy provides fundamental algorithms for scientific computing. The
 breadth of its scope was derived from the Guide to Available Mathematical
@@
 over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using numpy 1.8.2
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
 and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward

### 2018-10-05 scipy-1.0/paper.tex
Add references to CiSE 2007 issue

@@
 \end{quote}
-Python had a strong enough presence in science and engineering so that
-editors were soliciting a special issue, but it was still niche enough
-that the average reader would need additional information in order
-to decide whether it would ``be useful in their own work.''
 In the resulting CiSE May/June 2007 special issue, Paul's overview\cite{dubois2007guest}
@@
 (steering code)
-\cite{oliphant2007python}
-\cite{perez2007ipython}
-\cite{hunter2007matplotlib}
+SciPy \cite{oliphant2007python}
+IPython \cite{perez2007ipython}
+Matplotlib \cite{hunter2007matplotlib}
 research applications ...
+
+systems biology\cite{myers2007python},
+astronomy\cite{greenfield2007reaching},
+robotics\cite{krauss2007python},
+nanophotonics\cite{bienstman2007python},
+partial differential equations\cite{mardal2007using},
+neuroimaging\cite{millman2007analysis},
+geographic information systems\cite{shi2007python}, and
+education\cite{backer2007computational, myers2007pythona},
 
@@
 
-CiSE 2011 special issue
+When Paul edited the CiSE May/June 2007 special issue, Python already had a
+strong enough presence in science and engineering so that editors solicited
+the special issue, but it was still niche enough that the average reader
+would need additional information in order to decide whether it would
+``be useful in their own work.''
+The follow-up CiSE March/April 2011 Python for Scientists and Engineers
+special issue\cite{millman2011python}, focused more on the core parts
+of the scientific Python ecosystem\cite{perez2011python} including
+NumPy\cite{vanderwalt2011numpy}, Cython\cite{behnel2011cython},
+and Mayavi\cite{ramachandran2011mayavi}.
 
@@
 \cite{peterson2009f2py}
-\cite{millman2011python}
-\cite{perez2011python}
-\cite{vanderwalt2011numpy}
-\cite{behnel2011cython}
 \cite{pedregosa2011scikit}

### 2018-10-05 scipy-1.0/paper.tex
Sync w/ Stefan

@@
 (2008–) and SciPy India (2009–).
+Special sessions and minisymposia dedicated to scientific Python began
+appearing at many events.
+For example, a three-part minisymposium organized for CSE 2009 was
+featured in SIAM News\cite{siamcse09}.
+
+In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
+and Fernando Perez founded NumFOCUS\footnote{https://numfocus.org}.
+
+over 25 sponsored projects
+
+PyData
+
+everywhere...
 
@@
 
+Python has become so pervasive, there are now domain-specific special issues.
+For example, Frontiers in Neuroinformatics has a collection of 25 articles---covering
+topics including modeling and simulation, data collection, electrophysiology, visualization,
+as well as stimulus generation and presentation---called Python in
+neuroscience\footnote{\url{https://www.frontiersin.org/research-topics/8/python-in-neuroscience}}
+
 \cite{vanderwalt2008scipy}
@@
 FOSSEE
-
-SIAM CSE
-
-PyData
-
-everywhere...
-
-\cite{siamcse09}
-
-%https://archive.siam.org/news/news.php?id=1595
+https://fossee.in/about-us
 

### 2018-10-05 scipy-1.0/paper.tex
Add scikit links
Draft CiSE 2007 and NumFOCUS paragraphs

@@
 \end{quote}
+While it is now more commonly accepted, Paul took the time to explain to
+the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
+He also shares his own interest in Python for ``computational steering''
+where ``Python serves as the input language to a scientific application, and the actual
+computations are performed both in Python itself and in compiled extensions.''
 %https://www.computer.org/csdl/mags/cs/2007/03/c3007.html
-(steering code)
-SciPy \cite{oliphant2007python}
-IPython \cite{perez2007ipython}
-Matplotlib \cite{hunter2007matplotlib}
-research applications ...
-
+Using Python for computational steering was one of the earliest uses of Python
+in large scientific processing pipelines.
+In addition to Paul's overview there were also articles introducing
+Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
+and  Matplotlib \cite{hunter2007matplotlib}.
+In addition to these longer introductory articles, there was a diverse set of
+research applications including
 systems biology\cite{myers2007python},
@@
 geographic information systems\cite{shi2007python}, and
-education\cite{backer2007computational, myers2007pythona},
+education\cite{backer2007computational, myers2007pythona}.
 
@@
 
-In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
-and Fernando Perez founded NumFOCUS\footnote{https://numfocus.org}.
-
-over 25 sponsored projects
-
-PyData
-
-everywhere...
-
 When Paul edited the CiSE May/June 2007 special issue, Python already had a
@@
 neuroscience\footnote{\url{https://www.frontiersin.org/research-topics/8/python-in-neuroscience}}
+
+In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
+and Fernando Pérez founded NumFOCUS\footnote{https://numfocus.org}.
+NumFOCUS is a 501(c)3 public charity whose mission is to
+``promote sustainable high-level programming languages, open code development,
+and reproducible scientific research.''
+While NumFOCUS is language agnostic, many of the early sponsored projects
+came from the scientific Python stack.
+Today it has over 25 sponsored projects including NumPy, SciPy, IPython, and
+Matplotlib.
+Among its other projects, NumFOCUS organizes a global network of 
+community driven educational program called PyData.
 
@@
     \item The algorithm is of relevance to multiple fields of science.
-    \item The algorithm is demonstrably important.  E.g., it is classic
+    \item The algorithm is demonstrably important.  For example, it is classic
     enough to be included in textbooks, or it is based on a peer-reviewed article

### 2018-10-05 scipy-1.0/paper.tex
Add a 'what is SciPy' section to the introduction

@@
 
-Traditionally, a 1.0 release signals that a
-library's API has reached a level of maturity and stability
-that warrants trusting it inside research pipelines.
-The SciPy 1.0 version numbering belies the history of a project
-that has become the standard which others follow
-and has seen extensive adoption in research and industry.
+SciPy is a library of numerical routines for the Python programming
+language, providing the fundamental building blocks for modeling and
+solving scientific problems.  For example, when LIGO, the Laser
+Interferometer Gravitational-Wave Observatory, recently announced the
+discovery of gravitational waves, their published Python analysis
+scripts imported three different SciPy submodules: signal processing,
+interpolation, and input/output.  SciPy has routines for solving
+classes of generic problems such as optimization, integration,
+interpolation, eigenvalues, and linear systems.  It also provides
+specialized data structures, such as sparse arrays and k-dimensional
+trees.  It is the foundation upon which higher level scientific
+libraries, including scikit-learn and scikit-image, are built.
+
+Recently, SciPy released version 1.0, a milestone that traditionally
+signals a library's API (Applications Programming Interface) being
+mature enough to be be trusted in production pipelines.  The SciPy 1.0
+version numbering, however, belies the history of a project that has
+become the standard which others follow and has seen extensive
+adoption in research and industry.
 
@@
 % scikits
+
+% First scikit (mail moving over from nipy list?)
+% https://mail.python.org/pipermail/scipy-dev/2007-March/006805.html
+
+% First scikit development:
+% https://mail.python.org/pipermail/scipy-dev/2007-April/006987.html
+
 %https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html

### 2018-10-05 scipy-1.0/paper.tex
Spellcheck

--- 2018-10-05 scipy-1.0/paper.tex
@@
 foundation of the scientific Python ecosystem is such that adoption of new
-languages or major dependencies is generally unlikely--our choices are strongly
+languages or major dependencies is generally unlikely---our choices are strongly
 driven by long-term stability. GPU acceleration, new transpiling libraries, and
@@
 Members, and actively contributing to the progress of the project (could be
-code review but also outreach activities, etc.). Generally, 1 year of sustained
+code review but also outreach activities, etc.). Generally, one year of sustained
 and substantial contributions (not restricted to source code contributions)
@@
 
-The SciPy project has approximately 15 active maintainers - people who review
+The SciPy project has approximately 15 active maintainers---people who review
 others' contributions, and in general do everything needed to ensure that the

### 2018-10-05 scipy-1.0/scipy-optimize.tex
Spellcheck

@@
 Implementing the core algorithm of the commercial solver MOSEK \cite{andersen2000mosek}, it solves all of the 90+ NETLIB LP benchmark problems \cite{netlib} tested.
-Unlike some interior point methods, this homogenous self-dual formulation provides certificates of infeasibility or unboundedness as appropriate.
+Unlike some interior point methods, this homogeneous self-dual formulation provides certificates of infeasibility or unboundedness as appropriate.
 
 A presolve routine \cite{andersen1995presolving} solves trivial problems and otherwise performs problem simplifications, such as bound tightening and removal of fixed variables, and one of several routines for eliminating redundant equality constraints is automatically chosen to reduce the chance of numerical difficulties caused by singular matrices.
-Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers --- often for the same system with multiple right hand sides due to the predictor-corrector approach --- provide speed sufficient for problems with tens of thousands of variables and constraints.
+Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers---often for the same system with multiple right hand sides due to the predictor-corrector approach---provide speed sufficient for problems with tens of thousands of variables and constraints.
 

### 2018-10-05 scipy-1.0/paper.tex
Add NumPy -> SciPy -> scikits description
Weave probably happened mostly as part of Enthought

@@
 specialized data structures, such as sparse arrays and k-dimensional
-trees.  It is the foundation upon which higher level scientific
-libraries, including scikit-learn and scikit-image, are built.
+trees.  SciPy is built on top of NumPy, and is itself the foundation
+upon which higher level scientific libraries, including scikit-learn
+and scikit-image, are built.
 
@@
 packages to support his dissertation, including a parallel job
-scheduler, genetic optimizer, and inline C++ compiler.
+scheduler and genetic optimizer.
 % Graduated 1998, PhD from Brown, 97–99 postdoc at Harvard Medical

### 2018-10-11 scipy-1.0/paper.tex
Cleanup outline

@@
 trees.  SciPy is built on top of NumPy, and is itself the foundation
-upon which higher level scientific libraries, including scikit-learn
-and scikit-image, are built.
+upon which higher level scientific libraries, including scikit-learn\cite{pedregosa2011scikit}
+and scikit-image\cite{vanderwalt2014scikit}, are built.
 
@@
 
-\cite{vanderwalt2008scipy}
-\cite{peterson2009f2py}
-\cite{pedregosa2011scikit}
-\cite{vanderwalt2014scikit}
-\cite{meurer2017sympy}
-
-FOSSEE
-https://fossee.in/about-us
-
-%\subsection*{SciKits}
-
-% stefan, gael, david, jarrod, robert
-
-website http://scikits.appspot.com/
-
-statsmodels
-
-scikit-image
-
-scikit-learn
-
-% scikits
-
-% First scikit (mail moving over from nipy list?)
-% https://mail.python.org/pipermail/scipy-dev/2007-March/006805.html
-
-% First scikit development:
-% https://mail.python.org/pipermail/scipy-dev/2007-April/006987.html
-
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006315.html
-%https://mail.python.org/pipermail/scipy-dev/2006-November/006316.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006972.html
-%https://mail.python.org/pipermail/scipy-dev/2007-April/006835.html
-
-% roadmap
-%https://mail.python.org/pipermail/scipy-dev/2007-March/006808.html
-
-% scikit learn
-%https://mail.python.org/pipermail/scipy-dev/2007-July/007451.html
-
 \subsection*{Project scope}
@@
 and probabilistic modeling.
+
+scikit-image
+
+sympy\cite{meurer2017sympy}
+
+NetworkX\cite{hagberg2008networkx}
 

### 2018-12-14 scipy-1.0/paper.tex
Minor revisions

@@
 Despite its success, SciPy still faces challenges:
-a low ``bus factor''---from the phrase ``hit by a bus'' and
-meaning that the project depends on a small number of
-specific contributors---being perhaps the most serious.
+a low ``bus factor''---meaning that the entire project depends on a small number of
+active contributors---being perhaps the most serious.
 
@@
 ability to function as a systems language, controlling specialized or
-custom-built hardware was feasible.
+custom-built hardware was also possible.
 
@@
 package, and provided algorithms for signal processing, special
-functions, sparse matrices, quadrature, optimization, and Fast Fourier
+functions, sparse matrices, quadrature, optimization, Fast Fourier
 Transforms, and more.  One of these packages, Multipack, was a set of
@@
 other aspects of the computational environment.
-In 2000, Prabhu Ramachandran\fixme{also note that he, and the others
-  mentioned below, are also SciPy contributors}, a PhD student at the Indian Institute of
+In 2000, Prabhu Ramachandran, a PhD student at the Indian Institute of
 Technology, started work on a 3D visualization application, based on
@@
 In 2001, Fernando Pérez, a graduate student at the University of
-Colorado, Boulder, created the IPython interactive shell, built into
+Colorado, Boulder, created the IPython interactive shell (that would
+eventually blossom into Project Jupyter), built into
 IPP, an earlier attempt by Janko Hauser, and including code from
@@
 % https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
-John Hunter, a graduate student
-\fixme{check, since John around that time became postdoc, can check
-  with John Milton?} at the University
-of Chicago Medical School, liked the plotting functionality
+John Hunter, a postdoc
+%John got his PhD in 2001, then a post-doc with John Milton (dept of
+%Neurology, UChicago?) a couple of years; then research professor
+%under Kurt Hecox, whereafter at Tradelink from 2006
+at the University
+of Chicago, liked the plotting functionality
 available in MATLAB, but had problems accessing their laboratory's
@@
 
-In April 2006, Paul
-Dubois\footnote{https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html} wrote
-to the NumPy mailing list:
-\begin{quote}
-IEEE's magazine, Computing in Science and Engineering (CiSE), has asked me
-to put together a theme issue on the use of Python in Science and
-Engineering. I will write an overview to be accompanied by 3-5 articles [. . .]
-that show a diverse set of applications or
-tools, to give our readers a sense of whether or not Python might be useful
-in their own work.
-\end{quote}
-In the resulting CiSE May/June 2007 special issue, Paul's overview\cite{dubois2007guest}
-stated:
+%In April 2006, Paul
+%Dubois\footnote{https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html} wrote
+%to the NumPy mailing list:
+%\begin{quote}
+%IEEE's magazine, Computing in Science and Engineering (CiSE), has asked me
+%to put together a theme issue on the use of Python in Science and
+%Engineering. I will write an overview to be accompanied by 3-5 articles [. . .]
+%that show a diverse set of applications or
+%tools, to give our readers a sense of whether or not Pyth%on might be useful
+%in their own work.
+%\end{quote}
+In a May/June 2007 special issue of IEEE Computing in Science and
+Engineering, Paul Dubois wrote\cite{dubois2007guest}:
 \begin{quote}
@@
 
-By the middle to late 2000s, SciPy matured after a long phase of significant
+By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
 growth and adoption.
@@
 through a new type of auxiliary package, the scikit.  The tooling,
-development and release processes become more professional, grown with
-a patience only affordable in open source development, leading with
+development, and release processes become more professional, grown with
+the patience affordable in open source development, leading with
 best practices nowadays increasingly common in both industry and other
@@
 parallel processing, acceleration / wrapping, and user interfaces.  By
-2004, a significant shift occurred towards application of SciPy on
+2004, a significant shift occurred towards application of SciPy to
 scientific problems.  The event also started to draw in more keynote
@@
 as well as stimulus generation and presentation---called Python in
-neuroscience\footnote{\url{https://www.frontiersin.org/research-topics/8/python-in-neuroscience}}
+neuroscience\footnote{\url{https://www.frontiersin.org/research-topics/8/python-in-neuroscience}}.
 
 In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
-and Fernando Pérez founded NumFOCUS\footnote{https://numfocus.org}.
-NumFOCUS is a 501(c)3 public charity whose mission is to
+and Fernando Pérez founded NumFOCUS\footnote{https://numfocus.org},
+ a 501(c)3 public charity whose mission is to
 ``promote sustainable high-level programming languages, open code development,

### 2019-01-13 scipy-1.0/paper.tex
Add linguist citation

@@
 implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library provides a detailed breakdown as \%
-composition by programming language in SciPy (Figure~\ref{fig:linguist}).
+base using the \texttt{linguist} library\cite{linguistref} provides a 
+detailed breakdown as \% composition by programming language in 
+SciPy (Figure~\ref{fig:linguist}).
 

### 2019-01-14 scipy-1.0/paper.tex
Add Jupyter citation / reference

@@
 Colorado, Boulder, created the IPython interactive shell (that would
-eventually blossom into Project Jupyter), built into
+eventually blossom into Project Jupyter\cite{Kluyver:2016aa}), built into
 IPP, an earlier attempt by Janko Hauser, and including code from

### 2019-01-14 scipy-1.0/paper.tex
Cite asv repo

@@
 
-The airspeed velocity (asv) library enables benchmarking Python packages
+The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
 over their lifetimes, and the performance of the SciPy code base was monitored

### 2019-01-14 scipy-1.0/paper.tex
Add SI Figures for Sparse Performance
* add / cite SI Figures for our sparse
performance improvements

* currently, it appears that the CSR to COO
performance improvement is not substantiated
by the asv results

@@
 ecosystem of research software---backed by millions in funding, and
-thousands of highly qualified engineers---was preposterous.
+many hundreds of highly qualified engineers\cite{mathworks-globe-97,
+esri-revenue,bloom-wolfram}---was preposterous.
 
@@
 message on the Python Matrix Special Interest Group (Matrix-SIG)
-mailing list\footnote{https://mail.python.org/pipermail/matrix-sig/1995-August/000002.html}:
+mailing list\cite{Hugunin-first}:
 \begin{quote}
@@
 that supported a high number of dimensions.  Jim Hugunin explained the utility
-of Python for numerical computation\footnote{http://hugunin.net/papers/hugunin95numpy.html}:
+of Python for numerical computation\cite{Hugunin-whitepaper}:
 \begin{quote}
@@
 maintenance of Numeric and funded the writing of its
-manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}, and
+manual\cite{Numeric-manual}, and
 %in the Program for Climate Model Diagnosis and Intercomparison at
@@
 operations of the Hubble Space Telescope, decided to replace their
-custom scripting language with Python\footnote{https://conference.scipy.org/scipy2011/slides/greenfield\_keynote\_astronomy.pdf}.
+custom scripting language with Python\cite{STScI-slither}.
 
@@
 By the late 1990s, discussions appeared on Matrix-SIG,
-expressing a desire for a complete scientific data analysis environment\footnote{https://conference.scipy.org/scipy2010/slides/travis\_oliphant\_keynote.pdf}.
+expressing a desire for a complete scientific data analysis environment\cite{Travis-Keynote-2010}.
 Travis Oliphant, a PhD student at the Mayo Clinic,
-released a number of packages\footnote{https://web.archive.org/web/19990125091242/http://oliphant.netpedia.net:80/}\footnote{https://web.archive.org/web/20001206213500/http://oliphant.netpedia.net:80/} that built on top of the Numeric array
+released a number of packages\cite{Travis-some-modules,Travis-enhance} 
+that built on top of the Numeric array
 package, and provided algorithms for signal processing, special
@@
 Around the same time, Pearu Peterson, a PhD student from Estonia,
-released f2py\footnote{https://mail.python.org/pipermail/numpy-discussion/2000-September/000243.html}, a command line tool for binding Python and Fortran
+released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
 codes, and wrote modules for linear algebra and interpolation.
@@
 Gary Strangman, a postdoctoral fellow at Harvard Medical School,
-published several descriptive and inferential statistical routines\footnote{https://web.archive.org/web/20001022231108/http://www.nmr.mgh.harvard.edu/Neural\_Systems\_Group/gary/python.html}.
+published several descriptive and inferential statistical 
+routines\cite{Strangman-modules}.
 
@@
 repository appearing in
-February\footnote{https://web.archive.org/web/20010309040805/http://scipy.org:80/}, and
+February\cite{archived-scipyorg}, and
 a mailing list announced in
-June \footnote{https://mail.python.org/pipermail/scipy-dev/2001-June/000000.html}. By
-August, a first release was announced\footnote{https://mail.python.org/pipermail/python-list/2001-August/106419.html}, an excerpt of which is shown in
+June \cite{new-scipy-list}. By
+August, a first release was announced\cite{first-scipy-rel}, an excerpt of which is shown in
 Fig.~\ref{fig:announce-0.1}.
 In September, the first documentation was
-published\footnote{https://web.archive.org/web/20021013204556/http://www.scipy.org:80/scipy/site\_content/site\_news/docs\_released1}.
+published\cite{first-scipy-docs}.
 The first SciPy
-workshop\footnote{https://mail.python.org/pipermail/numpy-discussion/2002-June/001511.html}
+workshop\cite{first-scipy-workshop}
 was held in September 2002 at CalTech---a single track, two day event with 50
@@
 Technology, started work on a 3D visualization application, based on
-Wingware's C++ Visualization Toolkit, called
-Mayavi\footnote{https://web.archive.org/web/20030731122452/http://pythonology.org/success\&story=mayavi}.
+Kitware's C++ Visualization Toolkit\cite{schroeder:2006:VTK}, called
+Mayavi\cite{mayavi-intro}.
 In 2001, Fernando Pérez, a graduate student at the University of
@@
 wrote a plotting library from scratch, and Matplotlib 0.1 was released
-April 2003\footnote{https://mail.python.org/pipermail/python-list/2003-April/193167.html}.
+April 2003\cite{matplotlib-rel}.
 
@@
 \end{quote}
-While it is now more commonly accepted, Paul took the time to explain to
+While it is now more commonly accepted, Dubois took the time to explain to
 the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
@@
 in large scientific processing pipelines.
-In addition to Paul's overview there were also articles introducing
+In addition to Dubois' overview there were also articles introducing
 Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
 and  Matplotlib \cite{hunter2007matplotlib}.
-In addition to these longer introductory articles, there was a diverse set of
+There was also a diverse set of
 research applications including
@@
 growth and adoption.
-The informal workshops are grown into international conferences with many
-hundreds of attendees, special issues are organized and published in a
-leading scientific magazine, and the scope of the SciPy library
-narrows, while the breadth of the ecosystem is grown and encouraged
-through a new type of auxiliary package, the scikit.  The tooling,
-development, and release processes become more professional, grown with
-the patience affordable in open source development, leading with
-best practices nowadays increasingly common in both industry and other
-open source projects \cite{millman2014developing}.
+The informal workshops grew into international conferences with many
+hundreds of attendees. Special issues were organized and published in a
+leading scientific journal\cite{dubois2007guest}, and the scope of the SciPy library
+narrowed, while the breadth of the ecosystem grew
+through a new type of auxiliary package, the scikit\cite{scikits-general}. 
+The tooling, development, and release processes became more professional.
+SciPy was expanded carefully, with the patience affordable in open source 
+projects and via best practices common in industry \cite{millman2014developing}.
 
@@
 
-In the early workshops, recurrent topics indicate the state of development, with
+In the early workshops, recurrent topics reflected the state of development, with
 emphasis being placed on the underlying array package, plotting,
@@
 conference—now with parallel tracks—moving to Austin, Texas.
-While the US SciPy conference kept growing (by ... it had multiple
+While the US SciPy conference kept growing (by 2010 it had multiple
 tracks, student sponsorship, proceedings, and funding), satellite
@@
 
-When Paul edited the CiSE May/June 2007 special issue, Python already had a
-strong enough presence in science and engineering so that editors solicited
-the special issue, but it was still niche enough that the average reader
+In 2007, Python had a strong enough presence in science and engineering
+that the editors of IEEE Computing in Science and Engineering
+(CiSE) solicited a special issue\cite{dubois2007guest}, edited by Paul
+Dubois. However, Python was still sufficiently niche that the average reader
 would need additional information in order to decide whether it would
@@
 as well as stimulus generation and presentation---called Python in
-neuroscience\footnote{\url{https://www.frontiersin.org/research-topics/8/python-in-neuroscience}}.
+neuroscience\cite{python-FIN}.
 
 In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
-and Fernando Pérez founded NumFOCUS\footnote{https://numfocus.org},
+and Fernando Pérez founded NumFOCUS\cite{numfocus},
  a 501(c)3 public charity whose mission is to
@@
 Among its other projects, NumFOCUS organizes a global network of 
-community driven educational program called PyData.
+community driven educational programs, meetups, and conferences
+called PyData.
 
@@
 blocks while interacting well with other packages specialized in that area.
-For example, SciPy provides what one expect to find in a
+For example, SciPy provides what one expects to find in a
 statistics textbook (probability distributions, hypothesis tests, frequency
-statistics, correlation functions, and more), while Statsmodels provides
-more advanced statistical estimators and inference methods, scikit-learn
-covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
-and probabilistic modeling.
-
-scikit-image
-
-sympy\cite{meurer2017sympy}
-
-NetworkX\cite{hagberg2008networkx}
+statistics, correlation functions, and more), while 
+Statsmodels\cite{statsmodels2010} provides
+more advanced statistical estimators and inference methods;
+scikit-learn\cite{pedregosa2011scikit} covers machine learning; and
+PyMC3\cite{10.7717/peerj-cs.55}, emcee\cite{2013PASP-emcee} and 
+PyStan\cite{pystan-ref} cover Bayesian statistics and probabilistic modeling.
+scikit-image\cite{vanderwalt2014scikit} provides image processing
+capabilities beyond \texttt{scipy.ndimage}, sympy\cite{meurer2017sympy}
+provides a Python interface for symbolic computation, and
+NetworkX\cite{hagberg2008networkx} is a Python library for probing
+graphs and networks in a more specialized way than SciPy components
+like \texttt{sparse.csgraph} or \texttt{spatial}.
 
@@
 
-\subsection*{Current status (maturity, users)}
-
-SciPy 1.0 was released ... see timeline ...
-
-1.1 ...
-
-1.2
-
-Python 2 / 3?
-
-\begin{figure}
+\subsection*{Current status}
+
+At the time of writing, SciPy produces a new release
+approximately every six months, and some of the major
+feature highlights from the three years preceding
+SciPy 1.0 are discussed in a technical improvements
+section below. Some longer-term milestone events are
+highlighted in Figure~\ref{fig:timeline}.
+
+\begin{figure}[H]
 \begin{verbatim}
@@
 \caption{Timeline}
+\label{fig:timeline}
 \end{figure}
@@
 changes that improve clarity in the API of the library but strongly discourage
-breaking backwards compatibility\footnote{
-    Backwards compatibility means that new versions of SciPy will keep user
-    code written for an older version working.}
-given our position near the base of the scientific Python computing stack.
+breaking backwards compatibility given our position near the base of the 
+scientific Python computing stack.
 
@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\%, and the CSR to COO matrix format
-conversion is now several times faster.  Importantly,
+and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
+ and the CSR to COO matrix format
+conversion is now several times faster (Figure~\ref{fig:sparse-conv}).  Importantly,
 SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
@@
 with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see above), confirming that performance generally
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
 remains constant or improves over the commit hash history of the project allows
@@
 performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
 
 \texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"}
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
 
@@
 benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the documentation, including arguments to
-select which benchmarks to run.
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
 
@@
 
+\section*{Supporting Information}
+
+% separate Figure numbering for SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
+\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-iter}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_csr_to_coo_bench}
+\caption{Airspeed velocity benchmarks for SciPy CSR to COO matrix format conversion over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-conv}
+\end{figure}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-01-14 scipy-1.0/paper.tex
Add citation for LIGO SciPy code usage.

@@
 discovery of gravitational waves, their published Python analysis
-scripts imported three different SciPy submodules: signal processing,
+scripts\cite{LIGO-open,GW150914-Python} imported three different
+ SciPy submodules: signal processing,
 interpolation, and input/output.  SciPy has routines for solving
@@
 
-Python is an interactive computer programming language, designed by
-Guido van Rossum in the late 1980s, with an emphasis on readability
-and rapid prototyping.
+Python is an interpreted, high-level, general-purpose computer programming
+language, designed by Guido van Rossum in the late 1980s,
+with a dynamic type system and an emphasis on readability and rapid prototyping.
+The reference and most popular implementation of Python is 
+CPython\cite{cpython-refman,cpython-source}, which is written
+in the C and Python languages and assumed throughout this paper.
 % https://legacy.python.org/search/hypermail/python-1992/0000.html
 %
-A general purpose programming language with a ``batteries included''
+As a general purpose programming language with a ``batteries included''
 standard library, it had no special support for scientific data
@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the CSR to COO matrix format
-conversion is now several times faster (Figure~\ref{fig:sparse-conv}).  Importantly,
+and CSR matrices is now faster by up to 35\%, and the CSR to COO matrix format
+conversion is now several times faster.  Importantly,
 SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
@@
 
-\section*{Supporting Information}
-
-% separate Figure numbering for SI
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
-\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
-Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
-The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
-\label{fig:sparse-iter}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_csr_to_coo_bench}
-\caption{Airspeed velocity benchmarks for SciPy CSR to COO matrix format conversion over roughly three years of commits prior to the release of version 1.0.
-Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
-The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
-\label{fig:sparse-conv}
-\end{figure}
-
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-01-14 scipy-1.0/paper.tex
Associate VTK with correct company; cite preferred source for VTK

--- 2019-01-14 scipy-1.0/paper.tex
@@
 Technology, started work on a 3D visualization application, based on
-Wingware's C++ Visualization Toolkit, called
+Kitware's C++ Visualization Toolkit\cite{schroeder:2006:VTK}, called
 Mayavi\footnote{https://web.archive.org/web/20030731122452/http://pythonology.org/success\&story=mayavi}.

### 2019-01-15 scipy-1.0/paper.tex
Paul Dubois CiSE edit cleanup

@@
 Kitware's C++ Visualization Toolkit\cite{schroeder:2006:VTK}, called
-Mayavi\footnote{https://web.archive.org/web/20030731122452/http://pythonology.org/success\&story=mayavi}.
+Mayavi\cite{mayavi-intro}.
 In 2001, Fernando Pérez, a graduate student at the University of
@@
 \end{quote}
-While it is now more commonly accepted, Paul took the time to explain to
+While it is now more commonly accepted, Paul Dubois took the time to explain to
 the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
@@
 in large scientific processing pipelines.
-In addition to Paul's overview there were also articles introducing
+In addition to Paul Dubois' overview there were also articles introducing
 Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
 and  Matplotlib \cite{hunter2007matplotlib}.
-In addition to these longer introductory articles, there was a diverse set of
+There was also a diverse set of
 research applications including
@@
 
-When Paul edited the CiSE May/June 2007 special issue, Python already had a
-strong enough presence in science and engineering so that editors solicited
-the special issue, but it was still niche enough that the average reader
+When Paul Dubois edited the May/June 2007 special issue of the 
+IEEE publication Computing in Science and Engineering (CiSE), 
+Python already had a strong enough presence in science and engineering
+that the editors solicited the endeavor. However, Python was still sufficiently
+niche that the average reader
 would need additional information in order to decide whether it would

### 2019-01-16 scipy-1.0/paper.tex
Convert footnotes to proper references; footnotes not allowed in Scientific Reports

@@
 ecosystem of research software---backed by millions in funding, and
-many hundreds of highly qualified engineers\cite{mathworks-globe-97,
-esri-revenue,bloom-wolfram}---was preposterous.
+thousands of highly qualified engineers---was preposterous.
 
@@
 \end{quote}
-While it is now more commonly accepted, Paul Dubois took the time to explain to
+While it is now more commonly accepted, Paul took the time to explain to
 the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
@@
 in large scientific processing pipelines.
-In addition to Paul Dubois' overview there were also articles introducing
+In addition to Paul's overview there were also articles introducing
 Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
@@
 
-When Paul Dubois edited the May/June 2007 special issue of the 
-IEEE publication Computing in Science and Engineering (CiSE), 
-Python already had a strong enough presence in science and engineering
-that the editors solicited the endeavor. However, Python was still sufficiently
-niche that the average reader
+When Paul edited the CiSE May/June 2007 special issue, Python already had a
+strong enough presence in science and engineering so that editors solicited
+the special issue, but it was still niche enough that the average reader
 would need additional information in order to decide whether it would
@@
 changes that improve clarity in the API of the library but strongly discourage
-breaking backwards compatibility given our position near the base of the 
-scientific Python computing stack.
+breaking backwards compatibility (Backwards compatibility means that new 
+versions of SciPy will keep user code written for an older version working.)
+given our position near the base of the scientific Python computing stack.
 
@@
 with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
-that performance generally
+unit tests are passing (see above), confirming that performance generally
 remains constant or improves over the commit hash history of the project allows
@@
 performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
 
 \texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"}
 
@@
 benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
+in. More features are available in the documentation, including arguments to
+select which benchmarks to run.
 

### 2019-01-16 scipy-1.0/paper.tex

--- 2019-01-15 scipy-1.0/paper.tex
@@
 
-\subsection*{Current status}
-
-At the time of writing, SciPy produces a new release
-approximately every six months, and some of the major
-feature highlights from the three years preceding
-SciPy 1.0 are discussed in a technical improvements
-section below. Some longer-term milestone events are
-highlighted in Figure~\ref{fig:timeline}.
-
-\begin{figure}[H]
+\subsection*{Current status (maturity, users)}
+
+SciPy 1.0 was released ... see timeline ...
+
+1.1 ...
+
+1.2
+
+Python 2 / 3?
+
+\begin{figure}
 \begin{verbatim}
@@
 \caption{Timeline}
-\label{fig:timeline}
 \end{figure}

### 2019-01-16 scipy-1.0/paper.tex
Fix tense of first paragraph in SciPy Matures subsection; fix a repeated 'In addition.'

@@
 and  Matplotlib \cite{hunter2007matplotlib}.
-In addition to these longer introductory articles, there was a diverse set of
+There was also a diverse set of
 research applications including
@@
 growth and adoption.
-The informal workshops are grown into international conferences with many
-hundreds of attendees, special issues are organized and published in a
-leading scientific magazine, and the scope of the SciPy library
-narrows, while the breadth of the ecosystem is grown and encouraged
+The informal workshops grew into international conferences with many
+hundreds of attendees. Special issues were organized and published in a
+leading scientific journal, and the scope of the SciPy library
+narrowed, while the breadth of the ecosystem grew
 through a new type of auxiliary package, the scikit.  The tooling,
-development, and release processes become more professional, grown with
-the patience affordable in open source development, leading with
-best practices nowadays increasingly common in both industry and other
+development, and release processes became more professional, grown with
+the patience affordable in open source development. Leading with
+best practices nowadays is increasingly common in both industry and other
 open source projects \cite{millman2014developing}.

### 2019-01-16 scipy-1.0/paper.tex
Benchmark suite section revisions
* the "see above" comment re: unit tests is now "below"

* add clarification about the `run.py` script that ships
with SciPy for benchmarking purposes

* more precise reference to asv docs

--- 2019-01-16 scipy-1.0/paper.tex
@@
 changes that improve clarity in the API of the library but strongly discourage
-breaking backwards compatibility (Backwards compatibility means that new 
-versions of SciPy will keep user code written for an older version working.)
-given our position near the base of the scientific Python computing stack.
+breaking backwards compatibility given our position near the base of the 
+scientific Python computing stack.
 
@@
 with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see above), confirming that performance generally
+unit tests are passing (see below), confirming that performance generally
 remains constant or improves over the commit hash history of the project allows
@@
 performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using the command:
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using the
+following command, where \texttt{run.py} is a light wrapper
+script around \texttt{asv}, packaged with SciPy:
 
@@
 benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the documentation, including arguments to
-select which benchmarks to run.
+in. More features are available in the \texttt{asv} documentation, including 
+arguments to select which benchmarks to run.
 

### 2019-01-18 scipy-1.0/paper.tex
Substantiate competition resources in intro

--- 2019-01-16 scipy-1.0/paper.tex
@@
 ecosystem of research software---backed by millions in funding, and
-thousands of highly qualified engineers---was preposterous.
+thousands of highly qualified engineers\cite{mathworks-globe-97,
+esri-revenue}---was preposterous.
 
@@
 hundreds of attendees. Special issues were organized and published in a
-leading scientific journal, and the scope of the SciPy library
+leading scientific journal\cite{dubois2007guest}, and the scope of the SciPy library
 narrowed, while the breadth of the ecosystem grew
-through a new type of auxiliary package, the scikit.  The tooling,
-development, and release processes became more professional, grown with
-the patience affordable in open source development. Leading with
-best practices nowadays is increasingly common in both industry and other
-open source projects \cite{millman2014developing}.
+through a new type of auxiliary package, the scikit\cite{scikits-general}. 
+The tooling, development, and release processes became more professional.
+SciPy was expanded carefully, with the patience affordable in open source 
+projects and via best practices common in industry \cite{millman2014developing}.
 

### 2019-01-22 scipy-1.0/paper.tex

@@
 ecosystem of research software---backed by millions in funding, and
-thousands of highly qualified engineers\cite{mathworks-globe-97,
-esri-revenue}---was preposterous.
+thousands of highly qualified engineers---was preposterous.
 

### 2019-01-22 scipy-1.0/paper.tex
MAINT: address first round of revisions in PR 76.

@@
 with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see below), confirming that performance generally
+unit tests are passing (see above), confirming that performance generally
 remains constant or improves over the commit hash history of the project allows

### 2019-01-22 scipy-1.0/paper.tex
MAINT: first round of revisions in PR 75.
Signed-off-by: Tyler Reddy <tyler.je.reddy@gmail.com>

--- 2019-01-22 scipy-1.0/paper.tex
@@
 in large scientific processing pipelines.
-In addition to Paul's overview there were also articles introducing
+In addition to Dubois' overview there were also articles introducing
 Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
@@
 
-When Paul edited the CiSE May/June 2007 special issue, Python already had a
-strong enough presence in science and engineering so that editors solicited
-the special issue, but it was still niche enough that the average reader
+When Paul Dubois edited the May/June 2007 special issue of the 
+IEEE publication Computing in Science and Engineering (CiSE)\cite{dubois2007guest}, 
+Python already had a strong enough presence in science and engineering
+that the editors solicited the endeavor. However, Python was still sufficiently
+niche that the average reader
 would need additional information in order to decide whether it would
@@
 with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see below), confirming that performance generally
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
 remains constant or improves over the commit hash history of the project allows

### 2019-01-23 scipy-1.0/paper.tex

--- 2019-01-22 scipy-1.0/paper.tex
@@
 
-When Paul Dubois edited the May/June 2007 special issue of the 
-IEEE publication Computing in Science and Engineering (CiSE)\cite{dubois2007guest}, 
-Python already had a strong enough presence in science and engineering
-that the editors solicited the endeavor. However, Python was still sufficiently
-niche that the average reader
+In 2007, Python had a strong enough presence in science and engineering
+that the editors of IEEE Computing in Science and Engineering
+(CiSE) solicited a special issue\cite{dubois2007guest}, edited by Paul
+Dubois. However, Python was still sufficiently niche that the average reader
 would need additional information in order to decide whether it would

### 2019-01-23 scipy-1.0/paper.tex
MAINT: first round of revisions in PR 69.

@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\%, and the CSR to COO matrix format
-conversion is now several times faster.  Importantly,
+and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
+ and the COO / DIA to CSR / CSC matrix format
+conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
 SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
@@
 
+\section*{Supporting Information}
+
+% separate Figure numbering for SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
+\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-iter}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
+\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-conv}
+\end{figure}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-01-23 scipy-1.0/paper.tex
MAINT: provide likely explanation for performance regression in Fig. S1.
MAINT: cite PyCBC usage for LIGO project.

--- 2019-01-23 scipy-1.0/paper.tex
@@
 discovery of gravitational waves, their published Python analysis
-scripts imported three different SciPy submodules: signal processing,
-interpolation, and input/output.  SciPy has routines for solving
+scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python} 
+imported many different SciPy submodules for: signal processing,
+interpolation, input / output, optimization, statistics, integration,
+and more. SciPy has routines for solving
 classes of generic problems such as optimization, integration,

### 2019-01-24 scipy-1.0/paper.tex
More accurate description of CPython & citations

@@
 discovery of gravitational waves, their published Python analysis
-scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python} 
-imported many different SciPy submodules for: signal processing,
-interpolation, input / output, optimization, statistics, integration,
-and more. SciPy has routines for solving
+scripts imported three different SciPy submodules: signal processing,
+interpolation, and input/output.  SciPy has routines for solving
 classes of generic problems such as optimization, integration,

### 2019-01-24 scipy-1.0/paper.tex
additional LIGO citaton based on feedback in PR 70.

@@
 discovery of gravitational waves, their published Python analysis
-scripts imported three different SciPy submodules: signal processing,
-interpolation, and input/output.  SciPy has routines for solving
+scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
+PhysRevD.93.122003} imported many different SciPy submodules for:
+signal processing, interpolation, input / output, optimization, statistics,
+integration, and more. SciPy has routines for solving
 classes of generic problems such as optimization, integration,

### 2019-01-24 scipy-1.0/paper.tex
Add brief description + citation of NumPy in first paragraph of intro.

--- 2019-01-24 scipy-1.0/paper.tex
@@
 specialized data structures, such as sparse arrays and k-dimensional
-trees.  SciPy is built on top of NumPy, and is itself the foundation
-upon which higher level scientific libraries, including scikit-learn\cite{pedregosa2011scikit}
+trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
+which provides array data structures and related fast numerical routines.
+SciPy is itself the foundation upon which higher level scientific libraries,
+including scikit-learn\cite{pedregosa2011scikit}
 and scikit-image\cite{vanderwalt2014scikit}, are built.

### 2019-01-24 scipy-1.0/paper.tex
Citation for 'entire ecosystem' of related packages.

--- 2019-01-24 scipy-1.0/paper.tex
@@
 described in this paper, but also to an entire ecosystem of related
-packages, and a variety of social activities centered around them.  The
-packages in the SciPy ecosystem share high standards of implementation,
+packages\cite{scipy-ecosystem}, and a variety of social activities 
+centered around them. The packages in the SciPy ecosystem share
+high standards of implementation,
 documentation, and testing, and a culture eager to learn and adopt

### 2019-01-24 scipy-1.0/paper.tex
Condense list-like roadmap mini-paragraphs into a Table.

--- 2019-01-24 scipy-1.0/paper.tex
@@
 unique code contributors, millions of downloads per year, 161 dependent
-packages, and 28700 dependent repositories.
+packages, and 28700 dependent repositories. This includes usage of SciPy
+in almost half of all machine learning projects on GitHub.
 The library includes functionality spanning clustering, Fourier transforms,
@@
 described in this paper, but also to an entire ecosystem of related
-packages, and a variety of social activities centered around them.  The
-packages in the SciPy ecosystem share high standards of implementation,
+packages \cite{scipy-ecosystem}, and a variety of social activities centered around
+them\cite{social-python}. The packages in the SciPy ecosystem 
+share high standards of implementation,
 documentation, and testing, and a culture eager to learn and adopt
@@
 thereby solving the dilemma and paving the way for Python to become
-a very significant player in the Data Science movement.\fixme{NumPy
-  1.0 was released in October 2006}
+a very significant player in the Data Science movement. NumPy
+1.0 was released in October 2006\cite{numpy-1.0-tag}, with about 30
+authors recognized for major contributions in its release notes.
 
@@
 
-SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and
-LAPACK\cite{LAPACK} routines and for the special functions provided in the
-\texttt{scipy.{\allowbreak}special} submodule.  These Cython wrappers are
-available in the modules
+SciPy has provided special functions and leveraged BLAS and
+LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
+includes Cython\cite{behnel2011cython} wrappers for:
+many BLAS and LAPACK routines (added in 2015), and the special functions 
+provided in the \texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
+These Cython wrappers are available in the modules
 \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},
@@
 
-\subsection*{Roadmap}
-
-The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
-maintained by the community that describes some of the major directions for
-improvement for the project, as well as specific limitations and matters that
-require assistance moving forward.
-
-We are still striving to increase the number of SciPy usage tutorials beyond
-our current 15 section offering\cite{SciPy_tutorials}, but our standard
-documentation for library features is already in excellent shape.
-
-The low-level Cython code in our library (which interacts with C-level code and
-exposes it for Python usage) could use some measure of modernization, including
-migration to typed memoryviews to handle NumPy arrays instead of the older
-syntax:
-
-\begin{lstlisting}[language=Python]
-# example of a function that takes a single
-# 3 dimensional typed memoryview as an argument
-# this allows Cython-level handling of a NumPy
-# array object
-cpdef int func_new(int[:, :, :] arr):
-    pass
-
-# example of the old syntax for handling
-# the same scenario
-cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
-    pass
-\end{lstlisting}
-
-\texttt{fftpack} and {linalg} have too much functionality overlap with
-\texttt{numpy} equivalents.
-
-\texttt{scipy.interpolate} could benefit from a set of new features including
-spline fitting routines with better user control; arithmetic routines for
-splines; transparent tensor-product splines, Non-Uniform Rational B-Splines
-support; and mesh refinement and coarsening of B-splines and corresponding
-tensor products.  Some spline functionality in \texttt{scipy.signal} should be
-migrated to \texttt{interpolate} and Second Order Sections need to be updated
-to match the capabilities in other routines.
-
-For \texttt{scipy.linalg}, the major change needed is to support a more recent
-version of LAPACK. Users should be able access an increasing repertoire of
-LAPACK functions using either \texttt{get\_lapack\_funcs} or
-\texttt{sp.linalg.lapack.name\_of\_function}.
-
-\texttt{scipy.optimize} would benefit from a few more good global optimizers 
-as well as large-scale optimizers.
-
-\texttt{scipy.ndimage} should be completely migrated to a data point
-model (values on a grid), instead of a pixel model (elements with centers),
-both of which have been used in the past, because the data point model is now
-proving more effective in practice.
-
-While \texttt{scipy.sparse} is mature, there are many data
-formats and the community continues to move forward here so we will aim
-to take advantage of their improvements\cite{abbasi2018sparse} 
-while preserving backwards compatibility to the extent that it is possible.
-\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
-for faster sparse SVD computation.
-
-\texttt{scipy.spatial} will benefit from \texttt{spatial.transforms} with
-support for rotation matrices.
-
-\texttt{scipy.special} needs precision improvements for hypergeometric 
-functions, parabolic cylinder functions, and spheroidal wave functions.
-
 \subsection*{Community beyond the SciPy library}
@@
 \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}.
-
-\subsection*{Future development}
-
-Future development plans include a major rewrite of the \texttt{scipy.sparse}
-module, which is already underway\cite{abbasi2018sparse}. This is due to the
-current interface only supporting two-dimensional objects. Even sub-arrays are
-two-dimensional, and dimensions greater than two aren't supported. The new
-package supports dimensions greater than two, and it bases itself around arrays
-rather than matrices (which is strictly more powerful). It also features
-improved inter-operability with NumPy and other community packages.
-
-\fixme{This section should include key issues: sparse arrays, ndimage pixel vs
-point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained
-submodules.}
+MacOS and in the Intel Distribution for Python\cite{intel-python}. SciPy is used
+by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
+
+\subsection*{Future Work}
+
+The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
+maintained by the community that describes some of the major directions for
+improvement for the project, as well as specific limitations and matters that
+require assistance moving forward.
+
+We are still striving to increase the number of SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}, but our standard
+documentation for library features is already in excellent shape.
+
+The low-level Cython code in our library (which interacts with C-level code and
+exposes it for Python usage) could use some measure of modernization, including
+migration to typed memoryviews to handle NumPy arrays instead of the older
+syntax:
+
+\begin{lstlisting}[language=Python]
+# example of a function that takes a single
+# 3 dimensional typed memoryview as an argument
+# this allows Cython-level handling of a NumPy
+# array object
+cpdef int func_new(int[:, :, :] arr):
+    pass
+
+# example of the old syntax for handling
+# the same scenario
+cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    pass
+\end{lstlisting}
+
+Some additional roadmap items are summarized in Table~\ref{table:roadmap}.
+
+\begin{table}[H]
+\caption{Summary of SciPy Roadmap items following 1.0 release}
+\begin{tabular}{ c|c }
+    SciPy submodule & Summary of Change \\
+    \hline
+    \texttt{optimize} & a few more high quality global optimizers\\
+    \texttt{fftpack} & reduce overlap with NumPy equivalent \\
+    \texttt{linalg} & reduce overlap with NumPy equivalent \\
+    \texttt{interpolate} & new spline fitting and arithmetic routines \\
+    \texttt{interpolate} & new transparent tensor-product splines\\
+    \texttt{interpolate} & new Non-Uniform Rational B-Splines\\
+    \texttt{interpolate} & mesh refinement \& coarsening of B-splines and tensor products\\
+    \texttt{signal} & migrate spline functionality to \texttt{interpolate}\\
+    \texttt{signal} & Second Order Sections update to match capabilities in other routines\\
+    \texttt{linalg} & support a more recent version of LAPACK\\
+    \texttt{ndimage} & migrate to data point model from pixel model\\
+    \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\
+    \texttt{sparse.linalg} & add PROPACK wrappers for faster SVD\\
+    \texttt{spatial} & add support for (quaternion) rotation matrices\\
+    \texttt{special} & precision improvements for hypergeometric, parabolic cylinder, and spheroidal wave functions\\
+\end{tabular}\label{table:roadmap}
+\end{table}
 
@@
 \fixme{TODO}
+
+\section*{Supporting Information}
+
+% separate Figure numbering for SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
+\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
+The partial performance regression observed shortly after the 1.0 release is likely related to a correctness fix\cite{sparse-regress}.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-iter}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
+\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-conv}
+\end{figure}
 

### 2019-01-25 scipy-1.0/paper.tex
Clarify that Cython bindings are recent

--- 2019-01-24 scipy-1.0/paper.tex
@@
 
-SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and
-LAPACK\cite{LAPACK} routines and for the special functions provided in the
-\texttt{scipy.{\allowbreak}special} submodule.  These Cython wrappers are
-available in the modules
+SciPy now includes Cython\cite{behnel2011cython} wrappers for many BLAS and
+LAPACK\cite{LAPACK} routines (added in 2015) and for the special functions
+provided in the
+\texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
+These Cython wrappers are available in the modules
 \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},

### 2019-01-25 scipy-1.0/paper.tex

--- 2019-01-24 scipy-1.0/paper.tex
@@
 described in this paper, but also to an entire ecosystem of related
-packages\cite{scipy-ecosystem}, and a variety of social activities 
-centered around them. The packages in the SciPy ecosystem share
-high standards of implementation,
+packages, and a variety of social activities centered around
+them\cite{social-python}. The packages in the SciPy ecosystem 
+share high standards of implementation,
 documentation, and testing, and a culture eager to learn and adopt

### 2019-01-25 scipy-1.0/paper.tex

@@
 described in this paper, but also to an entire ecosystem of related
-packages, and a variety of social activities centered around
+packages \cite{scipy-ecosystem}, and a variety of social activities centered around
 them\cite{social-python}. The packages in the SciPy ecosystem 

### 2019-01-25 scipy-1.0/paper.tex

--- 2019-01-25 scipy-1.0/paper.tex
@@
 
-SciPy now includes Cython\cite{behnel2011cython} wrappers for many BLAS and
-LAPACK\cite{LAPACK} routines (added in 2015) and for the special functions
-provided in the
-\texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
-These Cython wrappers are available in the modules
+SciPy includes Cython\cite{behnel2011cython} wrappers for many BLAS and
+LAPACK\cite{LAPACK} routines and for the special functions provided in the
+\texttt{scipy.{\allowbreak}special} submodule.  These Cython wrappers are
+available in the modules
 \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},

### 2019-01-25 scipy-1.0/paper.tex
Move roadmap into Future Work subsection

@@
 
-\subsection*{Future development}
-
-Future development plans include a major rewrite of the \texttt{scipy.sparse}
-module, which is already underway\cite{abbasi2018sparse}. This is due to the
-current interface only supporting two-dimensional objects. Even sub-arrays are
-two-dimensional, and dimensions greater than two aren't supported. The new
-package supports dimensions greater than two, and it bases itself around arrays
-rather than matrices (which is strictly more powerful). It also features
-improved inter-operability with NumPy and other community packages.
-
-\fixme{This section should include key issues: sparse arrays, ndimage pixel vs
-point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained
-submodules.}
+\subsection*{Future Work}
+
+The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
+maintained by the community that describes some of the major directions for
+improvement for the project, as well as specific limitations and matters that
+require assistance moving forward.
+
+We are still striving to increase the number of SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}, but our standard
+documentation for library features is already in excellent shape.
+
+The low-level Cython code in our library (which interacts with C-level code and
+exposes it for Python usage) could use some measure of modernization, including
+migration to typed memoryviews to handle NumPy arrays instead of the older
+syntax:
+
+\begin{lstlisting}[language=Python]
+# example of a function that takes a single
+# 3 dimensional typed memoryview as an argument
+# this allows Cython-level handling of a NumPy
+# array object
+cpdef int func_new(int[:, :, :] arr):
+    pass
+
+# example of the old syntax for handling
+# the same scenario
+cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    pass
+\end{lstlisting}
+
+\texttt{fftpack} and {linalg} have too much functionality overlap with
+\texttt{numpy} equivalents.
+
+\texttt{scipy.interpolate} could benefit from a set of new features including
+spline fitting routines with better user control; arithmetic routines for
+splines; transparent tensor-product splines, Non-Uniform Rational B-Splines
+support; and mesh refinement and coarsening of B-splines and corresponding
+tensor products.  Some spline functionality in \texttt{scipy.signal} should be
+migrated to \texttt{interpolate} and Second Order Sections need to be updated
+to match the capabilities in other routines.
+
+For \texttt{scipy.linalg}, the major change needed is to support a more recent
+version of LAPACK. Users should be able access an increasing repertoire of
+LAPACK functions using either \texttt{get\_lapack\_funcs} or
+\texttt{sp.linalg.lapack.name\_of\_function}.
+
+\texttt{scipy.optimize} would benefit from a few more good global optimizers 
+as well as large-scale optimizers.
+
+\texttt{scipy.ndimage} should be completely migrated to a data point
+model (values on a grid), instead of a pixel model (elements with centers),
+both of which have been used in the past, because the data point model is now
+proving more effective in practice.
+
+While \texttt{scipy.sparse} is mature, there are many data
+formats and the community continues to move forward here so we will aim
+to take advantage of their improvements\cite{abbasi2018sparse} 
+while preserving backwards compatibility to the extent that it is possible.
+\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
+for faster sparse SVD computation.
+
+\texttt{scipy.spatial} will benefit from \texttt{spatial.transforms} with
+support for rotation matrices.
+
+\texttt{scipy.special} needs precision improvements for hypergeometric 
+functions, parabolic cylinder functions, and spheroidal wave functions.
 

### 2019-01-25 scipy-1.0/paper.tex
Reviewer feedback: add 1.0 and dev roadmap citations

@@
 \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}.
+MacOS and in the Intel Distribution for Python\cite{intel-python}. SciPy is used
+by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
 
@@
 
-The SciPy Roadmap\cite{SciPy_roadmap} is a continuously-updated document
+The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev} is a
+continuously-updated document
 maintained by the community that describes some of the major directions for
@@
 
-\texttt{fftpack} and {linalg} have too much functionality overlap with
-\texttt{numpy} equivalents.
-
-\texttt{scipy.interpolate} could benefit from a set of new features including
-spline fitting routines with better user control; arithmetic routines for
-splines; transparent tensor-product splines, Non-Uniform Rational B-Splines
-support; and mesh refinement and coarsening of B-splines and corresponding
-tensor products.  Some spline functionality in \texttt{scipy.signal} should be
-migrated to \texttt{interpolate} and Second Order Sections need to be updated
-to match the capabilities in other routines.
-
-For \texttt{scipy.linalg}, the major change needed is to support a more recent
-version of LAPACK. Users should be able access an increasing repertoire of
-LAPACK functions using either \texttt{get\_lapack\_funcs} or
-\texttt{sp.linalg.lapack.name\_of\_function}.
-
-\texttt{scipy.optimize} would benefit from a few more good global optimizers 
-as well as large-scale optimizers.
-
-\texttt{scipy.ndimage} should be completely migrated to a data point
-model (values on a grid), instead of a pixel model (elements with centers),
-both of which have been used in the past, because the data point model is now
-proving more effective in practice.
-
-While \texttt{scipy.sparse} is mature, there are many data
-formats and the community continues to move forward here so we will aim
-to take advantage of their improvements\cite{abbasi2018sparse} 
-while preserving backwards compatibility to the extent that it is possible.
-\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
-for faster sparse SVD computation.
-
-\texttt{scipy.spatial} will benefit from \texttt{spatial.transforms} with
-support for rotation matrices.
-
-\texttt{scipy.special} needs precision improvements for hypergeometric 
-functions, parabolic cylinder functions, and spheroidal wave functions.
+Some additional roadmap items are summarized in Table~\ref{table:roadmap}.
+
+\begin{table}[H]
+\caption{Summary of SciPy Roadmap items following 1.0 release}
+\begin{tabular}{ c|c }
+    SciPy submodule & Summary of Change \\
+    \hline
+    \texttt{optimize} & a few more high quality global optimizers\\
+    \texttt{fftpack} & reduce overlap with NumPy equivalent \\
+    \texttt{linalg} & reduce overlap with NumPy equivalent \\
+    \texttt{interpolate} & new spline fitting and arithmetic routines \\
+    \texttt{interpolate} & new transparent tensor-product splines\\
+    \texttt{interpolate} & new Non-Uniform Rational B-Splines\\
+    \texttt{interpolate} & mesh refinement \& coarsening of B-splines and tensor products\\
+    \texttt{signal} & migrate spline functionality to \texttt{interpolate}\\
+    \texttt{signal} & Second Order Sections update to match capabilities in other routines\\
+    \texttt{linalg} & support a more recent version of LAPACK\\
+    \texttt{ndimage} & migrate to data point model from pixel model\\
+    \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\
+    \texttt{sparse.linalg} & add PROPACK wrappers for faster SVD\\
+    \texttt{spatial} & add support for (quaternion) rotation matrices\\
+    \texttt{special} & precision improvements for hypergeometric, parabolic cylinder, and spheroidal wave functions\\
+\end{tabular}\label{table:roadmap}
+\end{table}
 

### 2019-01-28 scipy-1.0/ckdtree.tex
Start cleaning up cKDTree citations / accuracy.

@@
 
+Specifically, in 2015, we enhanced \texttt{cKDTree.query} by reimplementing in C++,
+removing some of the algorithm memory leaks, and allowing 
+release of the GIL so that multiple threads may be used\cite{gh-4374}. In 2013,
+the time complexity of the k nearest neighbor search from \texttt{cKDTree.query}
+was approximately loglinear \cite{knn-jake}, and we've confirmed that this time
+complexity has been preserved in SciPy \texttt{1.0.0} (Figure~\ref{fig:knn-complexity}),
+but with generally improved efficiency (Figure~\ref{fig:asvbench}). The
+discussion in the originally-cited nearest neighbor search
+algorithm publication\cite{kdtree-search-algo}, suggests that there are
+at least some scenarios where $O(log n)$ search time complexity may
+be achieved, but the influence of dimensionality appears to be consistent
+with our empirical observations of approximately $0(n log n)$ performance.
+
 % TODO: please check / improve literature citations used here
-% what about the time complexity of the kNN implementation?
-In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
-parameter. This is an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
-because it generates and returns a data structure that may be only a fraction 
-of the size of the full neighbor list. A C++ implementation is provided that releases
-the Python global interpreter lock for the neighbor search process, uses
-a memory pool to allocate and automatically reclaim \texttt{structs}, and
-heapsort to sort the $k$ nearest neighbors distances, even in the case
-of periodic boundary conditions. When $k$ is a non-contiguous list of nearest
-neighbors to account for (i.e., \texttt{[1, 3, 4]}), the intervening distances
-between the nearest neighbor ($k = 1$) and the farthest neighbor requested
-are all interrogated, but only those neighbors that are requested are retained
-in memory. Thus, the maximum heap size for this algorithm is conceptually determined
-as \texttt{np.arange(max(k))}.
-
 Also in 2015, SciPy added support for generating approximated sparse distance matrices 

### 2019-01-28 scipy-1.0/paper.tex
Start cleaning up cKDTree citations / accuracy.

--- 2019-01-25 scipy-1.0/paper.tex
@@
 
-\subsection*{Common infrastructure}
-
 \subsection*{Language choices}
@@
 contributors for every 6-monthly release cycle. Anyone with the interest and
-skills can become a contributor or maintainer; the SciPy Developer Guide
-provides guidance on how to do that.
+skills can become a contributor or maintainer; the SciPy Developer
+Guide\cite{scipy-dev-guide} provides guidance on how to do that.
 
@@
 
+\section*{Supporting Information}
+% separate Figure numbering for SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
+\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
+\label{fig:knn-complexity}
+\end{figure}
+
 %Must include all authors, identified by initials, for example: A.A.

### 2019-01-29 scipy-1.0/paper.tex

--- 2019-01-25 scipy-1.0/paper.tex
@@
 thereby solving the dilemma and paving the way for Python to become
-a very significant player in the Data Science movement. NumPy
-1.0 was released in October 2006\cite{numpy-1.0-tag}, with about 30
-authors recognized for major contributions in its release notes.
+a very significant player in the Data Science movement.\fixme{NumPy
+  1.0 was released in October 2006}
 

### 2019-01-31 scipy-1.0/ckdtree.tex
Remove extraneous content near end of KDTree section

--- 2018-08-15 scipy-1.0/ckdtree.tex
@@
 correlation functions of galaxies\cite{0004-637X-750-1-38}.
-
-\fixme{Add a Figure to show the scaling, before and after.}
-\fixme{perhaps give an example or some formula.}
-(cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)
-
-The main purpose of the paircounting algorithm in cKDTree is to provide a readily
-available tool. The KDTree based algorithm is not necessarily the fastest algorithm
-depending on the application. E.g. for low number densities a chaining mesh based algorithm
-can be easily many times faster than a tree based algorithm.

### 2019-01-31 scipy-1.0/paper.tex

@@
 
+\section*{Supporting Information}
+
+% separate Figure numbering for SI
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
+\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
+The partial performance regression observed shortly after the 1.0 release is likely related to a correctness fix\cite{sparse-regress}.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-iter}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
+\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0.
+Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
+The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
+\label{fig:sparse-conv}
+\end{figure}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-01-31 scipy-1.0/paper.tex
Cite SciPy Developer Guide.

@@
 
+\subsection*{Common infrastructure}
+
 \subsection*{Language choices}

### 2019-02-01 scipy-1.0/ckdtree.tex
Update scipy-1.0/ckdtree.tex
Co-Authored-By: tylerjereddy <tyler.je.reddy@gmail.com>

--- 2019-01-28 scipy-1.0/ckdtree.tex
@@
 algorithm publication\cite{kdtree-search-algo}, suggests that there are
-at least some scenarios where $O(log n)$ search time complexity may
+at least some scenarios where $\mathcal{O}(\log{n})$ search time complexity may
 be achieved, but the influence of dimensionality appears to be consistent

### 2019-02-01 scipy-1.0/paper.tex
Update scipy-1.0/ckdtree.tex
Co-Authored-By: tylerjereddy <tyler.je.reddy@gmail.com>

@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
+\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
+\label{fig:knn-complexity}
+\end{figure}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-01 scipy-1.0/ckdtree.tex
MAINT: revisions based on PR 91 feeback
* Fix placement of Figure S1 / knn complexity
after rebase / conflict resolution

* revisions to cKDTree knn algorithm paragaph
based on feedback in PR

--- 2019-01-28 scipy-1.0/ckdtree.tex
@@
 
-Specifically, in 2015, we enhanced \texttt{cKDTree.query} by reimplementing in C++,
-removing some of the algorithm memory leaks, and allowing 
-release of the GIL so that multiple threads may be used\cite{gh-4374}. In 2013,
-the time complexity of the k nearest neighbor search from \texttt{cKDTree.query}
-was approximately loglinear \cite{knn-jake}, and we've confirmed that this time
-complexity has been preserved in SciPy \texttt{1.0.0} (Figure~\ref{fig:knn-complexity}),
-but with generally improved efficiency (Figure~\ref{fig:asvbench}). The
-discussion in the originally-cited nearest neighbor search
-algorithm publication\cite{kdtree-search-algo}, suggests that there are
-at least some scenarios where $O(log n)$ search time complexity may
-be achieved, but the influence of dimensionality appears to be consistent
-with our empirical observations of approximately $0(n log n)$ performance.
+In 2013, the time complexity of the k nearest neighbor search from
+\texttt{cKDTree.query} was approximately loglinear \cite{knn-jake},
+consistent with its formal description \cite{kdtree-search-algo}.
+Since then, we've enhanced \texttt{cKDTree.query} by reimplementing in
+C++, removing memory leaks, and allowing release of the GIL so that
+multiple threads may be used\cite{gh-4374}. This generally improved
+performance on any given problem (Figure~\ref{fig:asvbench}) while
+preserving the asymptotic complexity (Figure~\ref{fig:knn-complexity}).
 

### 2019-02-01 scipy-1.0/ckdtree.tex
Improvements to cKDTree paragraph 3
I have tried to shrink a bit paragraph 3.
I also removed the reference to `(sets of points organized into a binary space 
partitioning data structure\cite{Bentley:1975:MBS:361002.361007})` ... if we think we need a definition of what a kdtree is, maybe it is better to put it at the beginning of the section.

--- 2019-01-31 scipy-1.0/ckdtree.tex
@@
 
-Also in 2015, SciPy added support for generating approximated sparse distance matrices 
-between \texttt{KDTree} objects (sets of points organized into a binary space 
-partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
-The user provides a maximum distance value
-above which all distances between points in two provided \texttt{KDTree} objects
-are set to 0. The distance metric used between k-d trees is not constrained
-to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
-bounded by 1 and infinity is valid. By default, a dictionary of keys
-based sparse matrix is the returned data structure. Previously-published
-results had suggested that a hashing approach to sparse matrix assembly
-is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
-The dictionary object maps row and column indices to elements of the sparse matrix, which is
-efficient for matrix construction because all zero values are skipped in
-the dictionary mapping. The C++ level sparse matrix construction releases the Python
-global interpreter lock for increased performance. Once constructed, the
-dictionary of keys sparse matrix has the amortized constant time complexity 
-($O(1)$) for distance value retrieval that one would expect from a 
-dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
-operations on the sparse matrix, SciPy allows the dictionary of keys
-sparse matrix to be directly converted to the common CSR, CSC, and COO
-data structures.
+In 2015, SciPy added support for generating approximated sparse distance matrices 
+between \texttt{KDTree} objects by zeroing all distances that exceed a user-provided value. 
+The distance metric used between k-d trees is not constrained to the 
+conventional L2 (Euclidean) norm---any Minkowski p-norm value bounded by 1 and 
+infinity is valid. By default, the returned data structure is a dictionary of 
+keys based sparse matrix, which is very efficient for matrix construction. 
+Previously-published results had suggested that this hashing approach to sparse 
+matrix assembly is 7 times faster than constructing with compressed row format (CSR)
+\cite{10.1007/978-3-540-75755-9_107} and the C++ level sparse matrix construction 
+releases the Python GIL for increased performance. Also, once the matrix is 
+contructed, distance value retrieval has an amortized constant time complexity 
+\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic operations 
+on it, SciPy allows the sparse matrix to be directly converted to the common 
+CSR, CSC and COO data structures.
 
@@
 correlation functions of galaxies\cite{0004-637X-750-1-38}.
+
+\fixme{Add a Figure to show the scaling, before and after.}
+\fixme{perhaps give an example or some formula.}
+(cite / mention faster implementions of paircounting algorithms / treecorr, corrfunc)
+
+The main purpose of the paircounting algorithm in cKDTree is to provide a readily
+available tool. The KDTree based algorithm is not necessarily the fastest algorithm
+depending on the application. E.g. for low number densities a chaining mesh based algorithm
+can be easily many times faster than a tree based algorithm.

### 2019-02-02 scipy-1.0/ckdtree.tex
Update scipy-1.0/ckdtree.tex
Co-Authored-By: tylerjereddy <tyler.je.reddy@gmail.com>

--- 2019-02-01 scipy-1.0/ckdtree.tex
@@
 consistent with its formal description \cite{kdtree-search-algo}.
-Since then, we've enhanced \texttt{cKDTree.query} by reimplementing in
+Since then, we've enhanced \texttt{cKDTree.query} by reimplementing it in
 C++, removing memory leaks, and allowing release of the GIL so that

### 2019-02-02 scipy-1.0/ckdtree.tex
Update scipy-1.0/ckdtree.tex
Co-Authored-By: tylerjereddy <tyler.je.reddy@gmail.com>

--- 2019-02-01 scipy-1.0/ckdtree.tex
@@
 consistent with its formal description \cite{kdtree-search-algo}.
-Since then, we've enhanced \texttt{cKDTree.query} by reimplementing in
-C++, removing memory leaks, and allowing release of the GIL so that
+Since then, we've enhanced \texttt{cKDTree.query} by reimplementing it in
+C++, removing memory leaks, and allowing release of the global interpreter lock (GIL) so that
 multiple threads may be used\cite{gh-4374}. This generally improved

### 2019-02-06 scipy-1.0/ckdtree.tex
review

--- 2019-02-02 scipy-1.0/ckdtree.tex
@@
 
-In 2013, the time complexity of the k nearest neighbor search from
-\texttt{cKDTree.query} was approximately loglinear \cite{knn-jake},
-consistent with its formal description \cite{kdtree-search-algo}.
-Since then, we've enhanced \texttt{cKDTree.query} by reimplementing it in
-C++, removing memory leaks, and allowing release of the GIL so that
-multiple threads may be used\cite{gh-4374}. This generally improved
-performance on any given problem (Figure~\ref{fig:asvbench}) while
-preserving the asymptotic complexity (Figure~\ref{fig:knn-complexity}).
+% TODO: please check / improve literature citations used here
+% what about the time complexity of the kNN implementation?
+In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
+parameter. This is an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
+because it generates and returns a data structure that may be only a fraction 
+of the size of the full neighbor list. A C++ implementation is provided that releases
+the Python global interpreter lock for the neighbor search process, uses
+a memory pool to allocate and automatically reclaim \texttt{structs}, and
+heapsort to sort the $k$ nearest neighbors distances, even in the case
+of periodic boundary conditions. When $k$ is a non-contiguous list of nearest
+neighbors to account for (i.e., \texttt{[1, 3, 4]}), the intervening distances
+between the nearest neighbor ($k = 1$) and the farthest neighbor requested
+are all interrogated, but only those neighbors that are requested are retained
+in memory. Thus, the maximum heap size for this algorithm is conceptually determined
+as \texttt{np.arange(max(k))}.
 
-% TODO: please check / improve literature citations used here
-Also in 2015, SciPy added support for generating approximated sparse distance matrices 
-between \texttt{KDTree} objects (sets of points organized into a binary space 
-partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
-The user provides a maximum distance value
-above which all distances between points in two provided \texttt{KDTree} objects
-are set to 0. The distance metric used between k-d trees is not constrained
-to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
-bounded by 1 and infinity is valid. By default, a dictionary of keys
-based sparse matrix is the returned data structure. Previously-published
-results had suggested that a hashing approach to sparse matrix assembly
-is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
-The dictionary object maps row and column indices to elements of the sparse matrix, which is
-efficient for matrix construction because all zero values are skipped in
-the dictionary mapping. The C++ level sparse matrix construction releases the Python
-global interpreter lock for increased performance. Once constructed, the
-dictionary of keys sparse matrix has the amortized constant time complexity 
-($O(1)$) for distance value retrieval that one would expect from a 
-dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
-operations on the sparse matrix, SciPy allows the dictionary of keys
-sparse matrix to be directly converted to the common CSR, CSC, and COO
-data structures.
+In 2015, SciPy added the \texttt{sparse_distance_matrix} routine for generating approximate 
+sparse distance matrices between \texttt{KDTree} objects by ignoring all distances that 
+exceed a user-provided value. Also, this routine allows any Minkowski p-norm between 1 and 
+infinity and it is not limited to the conventional L2 (Euclidean) norm. 
+By default, the returned data structure is a dictionary of keys (DOK) based sparse matrix, 
+which is very efficient for matrix construction. 
+This hashing approach to sparse matrix assembly can be 7 times faster than constructing 
+with compressed row format (CSR) \cite{10.1007/978-3-540-75755-9_107} and the C++ level 
+sparse matrix construction releases the Python GIL for increased performance. 
+Once the matrix is constructed, distance value retrieval has an amortized constant time 
+complexity \cite{Cormen:2001:IA:580470}, and the DOK structure can be efficiently converted 
+to a CSR, CSC or COO matrix to allow for speedy arithmetic operations.
 

### 2019-02-06 scipy-1.0/ckdtree.tex
fix texttt

@@
 
-In 2015, SciPy added the \texttt{sparse_distance_matrix} routine for generating approximate 
+In 2015, SciPy added the \texttt{sparse\_distance\_matrix} routine for generating approximate 
 sparse distance matrices between \texttt{KDTree} objects by ignoring all distances that 

### 2019-02-06 scipy-1.0/ckdtree.tex

@@
 
-% TODO: please check / improve literature citations used here
-% what about the time complexity of the kNN implementation?
-In 2015, we enhanced \texttt{cKDTree.query} with a $k$ nearest neighbors search
-parameter. This is an efficient operation\cite{Sproull:1991:RNS:3118219.3118331} 
-because it generates and returns a data structure that may be only a fraction 
-of the size of the full neighbor list. A C++ implementation is provided that releases
-the Python global interpreter lock for the neighbor search process, uses
-a memory pool to allocate and automatically reclaim \texttt{structs}, and
-heapsort to sort the $k$ nearest neighbors distances, even in the case
-of periodic boundary conditions. When $k$ is a non-contiguous list of nearest
-neighbors to account for (i.e., \texttt{[1, 3, 4]}), the intervening distances
-between the nearest neighbor ($k = 1$) and the farthest neighbor requested
-are all interrogated, but only those neighbors that are requested are retained
-in memory. Thus, the maximum heap size for this algorithm is conceptually determined
-as \texttt{np.arange(max(k))}.
+In 2013, the time complexity of the k nearest neighbor search from
+\texttt{cKDTree.query} was approximately loglinear \cite{knn-jake},
+consistent with its formal description \cite{kdtree-search-algo}.
+Since then, we've enhanced \texttt{cKDTree.query} by reimplementing it in
+C++, removing memory leaks, and allowing release of the global interpreter lock (GIL) so that
+multiple threads may be used\cite{gh-4374}. This generally improved
+performance on any given problem (Figure~\ref{fig:asvbench}) while
+preserving the asymptotic complexity (Figure~\ref{fig:knn-complexity}).
 

### 2019-02-06 scipy-1.0/ckdtree.tex

@@
 
-In 2015, SciPy added the \texttt{sparse\_distance\_matrix} routine for generating approximate 
-sparse distance matrices between \texttt{KDTree} objects by ignoring all distances that 
-exceed a user-provided value. Also, this routine allows any Minkowski p-norm between 1 and 
-infinity and it is not limited to the conventional L2 (Euclidean) norm. 
-By default, the returned data structure is a dictionary of keys (DOK) based sparse matrix, 
-which is very efficient for matrix construction. 
-This hashing approach to sparse matrix assembly can be 7 times faster than constructing 
-with compressed row format (CSR) \cite{10.1007/978-3-540-75755-9_107} and the C++ level 
-sparse matrix construction releases the Python GIL for increased performance. 
-Once the matrix is constructed, distance value retrieval has an amortized constant time 
-complexity \cite{Cormen:2001:IA:580470}, and the DOK structure can be efficiently converted 
-to a CSR, CSC or COO matrix to allow for speedy arithmetic operations.
+% TODO: please check / improve literature citations used here
+Also in 2015, SciPy added support for generating approximated sparse distance matrices 
+between \texttt{KDTree} objects (sets of points organized into a binary space 
+partitioning data structure\cite{Bentley:1975:MBS:361002.361007}). 
+The user provides a maximum distance value
+above which all distances between points in two provided \texttt{KDTree} objects
+are set to 0. The distance metric used between k-d trees is not constrained
+to the conventional L2 (Euclidean) norm---any Minkowski p-norm value
+bounded by 1 and infinity is valid. By default, a dictionary of keys
+based sparse matrix is the returned data structure. Previously-published
+results had suggested that a hashing approach to sparse matrix assembly
+is 7 times faster than constructing with compressed row format (CSR)\cite{10.1007/978-3-540-75755-9_107}.
+The dictionary object maps row and column indices to elements of the sparse matrix, which is
+efficient for matrix construction because all zero values are skipped in
+the dictionary mapping. The C++ level sparse matrix construction releases the Python
+global interpreter lock for increased performance. Once constructed, the
+dictionary of keys sparse matrix has the amortized constant time complexity 
+($O(1)$) for distance value retrieval that one would expect from a 
+dictionary\cite{Cormen:2001:IA:580470}. For efficiently performing arithmetic
+operations on the sparse matrix, SciPy allows the dictionary of keys
+sparse matrix to be directly converted to the common CSR, CSC, and COO
+data structures.
 
-The cKDTree module implements a dual tree counting algorithm\cite{Moore2000ar},
-with an improvement to the pair counting algorithm to improve the scaling
-with the number of bins. The cKDTree can now be augmented by weights, with 
-weighted paircount essential in many scientific applications, e.g. computing 
-correlation functions of galaxies\cite{0004-637X-750-1-38}.
+In 2015 the cKDTree dual tree counting algorithm\cite{Moore2000ar}
+was enhanced to support weights\cite{ckdtree-weights}, which are
+essential in many scientific applications, e.g. computing correlation
+functions of galaxies\cite{0004-637X-750-1-38}.

### 2019-02-06 scipy-1.0/ckdtree.tex
Update ckdtree.tex

--- 2019-02-06 scipy-1.0/ckdtree.tex
@@
 
-In 2015, SciPy added the \texttt{sparse\_distance\_matrix} routine for generating approximate 
-sparse distance matrices between \texttt{KDTree} objects by ignoring all distances that 
-exceed a user-provided value. Also, this routine allows any Minkowski p-norm between 1 and 
-infinity and it is not limited to the conventional L2 (Euclidean) norm. 
-By default, the returned data structure is a dictionary of keys (DOK) based sparse matrix, 
-which is very efficient for matrix construction. 
-This hashing approach to sparse matrix assembly can be 7 times faster than constructing 
-with compressed row format (CSR) \cite{10.1007/978-3-540-75755-9_107} and the C++ level 
-sparse matrix construction releases the Python GIL for increased performance. 
-Once the matrix is constructed, distance value retrieval has an amortized constant time 
-complexity \cite{Cormen:2001:IA:580470}, and the DOK structure can be efficiently converted 
-to a CSR, CSC or COO matrix to allow for speedy arithmetic operations.
+In 2015, SciPy added the \texttt{sparse\_distance\_matrix} routine for generating 
+approximate sparse distance matrices between \texttt{KDTree} objects by ignoring 
+all distances that exceed a user-provided value. Also, this routine it is not 
+limited to the conventional L2 (Euclidean) norm but supports any Minkowski 
+p-norm between 1 and infinity. By default, the returned data structure is a 
+Dictionary Of Keys based sparse matrix (DOK), which is very efficient for matrix 
+construction. This hashing approach to sparse matrix assembly can be 7 times 
+faster than constructing with Compressed Sparse Row format (CSR) 
+\cite{10.1007/978-3-540-75755-9_107} and the C++ level sparse matrix construction 
+releases the Python GIL for increased performance. Once the matrix is constructed, 
+distance value retrieval has an amortized constant time complexity 
+\cite{Cormen:2001:IA:580470}, and the DOK structure can be efficiently converted 
+to a CSR, Compressed Sparse Column (CSC) or COOrdinate (COO) matrix to allow for 
+speedy arithmetic operations.
 

### 2019-02-07 scipy-1.0/subpackages.tex
fix two typos

@@
     \emph{statistical tests}, including Pearson's correlation, Spearman's rank-order
-    correlations, Kendal's tau, chi-squared test and its generalization as the
+    correlations, Kendall's tau, chi-squared test and its generalization as the
     Cressie-Read power divergence, contingency table tests including Fisher's

### 2019-02-14 scipy-1.0/paper.tex
MAINT: revisions based on PR 91 feeback
* Fix placement of Figure S1 / knn complexity
after rebase / conflict resolution

* revisions to cKDTree knn algorithm paragaph
based on feedback in PR
Fix cov plot

@@
 
-The \texttt{scipy.stats} module provides provides more than 100 probability
+The \texttt{scipy.stats} module provides more than 100 probability
 distributions and a framework to implement additional distributions. Given an
@@
 unit tests in the test suite, which is written for usage with the pytest
-runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
-holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the code base would still benefit from increased
+runner, and with 87 \% and 45 \% line coverages for Python and compiled
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
+Some of the historical components of the code base would still benefit from increased
 test coverage. Documentation for the code is automatically built and hosted by
@@
 \centering
-\includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy code base by unit tests over the six months
-preceding June 28, 2018 as reported by the \texttt{codecov} service. The
-analysis of lines of source code covered is performed automatically in our
-continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
-libraries.}
+\includegraphics[width=\textwidth]{static/cov}
+\caption{\% coverage of the SciPy code base by unit tests over a five year
+time span predeceding the 1.0 release point. The reconstitution of historical
+build and test environments was performed using a docker-based approach
+\cite{scipy-cov}. Analysis of lines of source code covered is performed 
+automatically in our continuous integration suite using the \texttt{pytest-cov}
+and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
+inclusion of additional vendored code.}
 \label{fig:coverage}
@@
 \centering
+\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
+\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
+\label{fig:knn-complexity}
+\end{figure}
+
+\begin{figure}[H]
+\centering
 \includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
-The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
-\label{fig:knn-complexity}
-\end{figure}
-
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-15 scipy-1.0/ckdtree.tex
Update scipy-1.0/ckdtree.tex
Co-Authored-By: apbard <apbard@users.noreply.github.com>

@@
 approximate sparse distance matrices between \texttt{KDTree} objects by ignoring 
-all distances that exceed a user-provided value. Also, this routine it is not 
+all distances that exceed a user-provided value. Also, this routine is not 
 limited to the conventional L2 (Euclidean) norm but supports any Minkowski 
@@
 
-The cKDTree module implements a dual tree counting algorithm\cite{Moore2000ar},
-with an improvement to the pair counting algorithm to improve the scaling
-with the number of bins. The cKDTree can now be augmented by weights, with 
-weighted paircount essential in many scientific applications, e.g. computing 
-correlation functions of galaxies\cite{0004-637X-750-1-38}.
+In 2015 the cKDTree dual tree counting algorithm\cite{Moore2000ar}
+was enhanced to support weights\cite{ckdtree-weights}, which are
+essential in many scientific applications, e.g. computing correlation
+functions of galaxies\cite{0004-637X-750-1-38}.

### 2019-02-15 scipy-1.0/paper.tex
Simplify polynomial interpolation content

--- 2019-02-01 scipy-1.0/paper.tex
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
-The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
-\label{fig:knn-complexity}
-\end{figure}
+\section{Polynomial Interpolators: Additional Details}
+FITPACK provides a collection of routines for calculating
+smoothing splines and automatic knot selection. One issue 
+with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
+and the \code{splrep}/\code{splev} combination, is the monolithic design:
+specific algorithms for construction of interpolating and 
+smoothing splines and knot selection are tightly coupled 
+at the Fortran level. The library offers a very
+limited degree of user control over the details 
+of the spline fitting it performs. 
+%
+Routines for manipulating piecewise polynomials and b-splines were either
+not available or lacking in ease of use and/or performance characteristics.
+
+An alternative set of spline interpolation routines was implemented in the
+(now deprecated) \code{splmake}/\code{spleval} combination, and used as a
+backend of the \code{interp1d} routine. To further complicate matters,
+the spline format used by \code{splmake}/\code{spleval} was incompatible 
+with \code{splrep}/\code{splev}.
+
+Using \code{UnivariateSpline} in new code is not recommended. Given data points,
+the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
+instance, representing a spline which interpolates the data. Interpolated values
+at intermediate points can be computed either via its \code{\_\_call\_\_} method,
+or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
+accepts both \code{tck} tuples and \code{Bspline} instances.
+
+Fitting \emph{smoothing} splines to data can currently only be done with 
+FITPACK. Building alternative algorithms is certainly possible,
+but is still to be implemented.
+
+Interpolation with known or estimated derivatives, i.e., cubic or
+higher-order Hermite splines, can be constructed by either
+subclassing \code{PPoly} or \code{BPoly}, or using the factory
+method \code{BPoly.from\_derivatives}.
+
+The spline representation used by \code{BSpline} is directly
+compatible with FITPACK. We recommend that users work with
+\code{BSpline} objects instead of directly manipulating \code{tck}
+tuples of knots, coefficients and the spline degree.
+%
 

### 2019-02-15 scipy-1.0/poly.tex
Simplify polynomial interpolation content

@@
 Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for
-univariate interpolation and approximation of data. FITPACK provides a
-collection of routines for calculating smoothing splines and automatic knot
-selection. One issue with FITPACK and its Python wrappers, \code{UnivariateSpline} and \code{splrep}/\code{splev} combo, is the monolithic design: specific
-algorithms for construction of interpolating and smoothing splines and knot
-selection are tightly coupled at the Fortran level. The library offers very
-limited degree of user control over the details of the spline fitting it performs. 
-%
-An alternative set of spline interpolation routines was implemented in the
-(now deprecated) \code{splmake}/\code{spleval} combo, and used as a backend of
-the \code{interp1d} routine. To further complicate matters, the spline format
-used by \code{splmake}/\code{spleval} was incompatible with
-\code{splrep}/\code{splev}.
-%
-Routines for manipulating piecewise polynomials and b-splines were either not available or lacking in ease of use and/or performance characteristics.
+univariate interpolation and approximation of data. The original monolithic
+design and API for interaction between SciPy and FITPACK was limiting for both
+users and developers, as detailed in the Supporting Information.
 
@@
 represents piecewise polynomials in the Bernstein basis (which is suitable
-for e.g., constructing Bezier curves).
+for e.g., constructing Bezier curves). \code{BSpline} represents spline
+curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}.
 
-In the next layer, these polynomial classes are used for constructing several common ways of interpolating data: \code{CubicSpline} constructs a twice differentiable piecewise cubic function, \code{Akima1DInterpolator} and \code{PCHIPInterpolator} implement two classic prescriptions for constructing a $C^1$ continuous monotone shape-preserving interpolator.
+In the next layer, these polynomial classes are used for constructing several
+common ways of interpolating data: \code{CubicSpline} constructs a twice 
+differentiable piecewise cubic function, \code{Akima1DInterpolator} 
+and \code{PCHIPInterpolator} implement two classic prescriptions for
+constructing a $C^1$ continuous monotone shape-preserving interpolator.
 \cite{FritschCarlson1980, Akima1970}
-%
-Interpolation with known or estimated derivatives, i.e., cubic or higher-order Hermite splines, can be constructed by either subclassing \code{PPoly} or \code{BPoly}, or using the factory method \code{BPoly.from\_derivatives}.
-
-For working with b-splines, there are several options.
-The basic object is \code{BSpline}, which represents spline curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}.
-The spline
-representation used by \code{BSpline} is directly compatible with FITPACK. We
-recommend that users work with \code{BSpline} objects instead of directly
-manipulating \code{tck} tuples of knots, coefficients and the spline degree.
-%
-Using \code{UnivariateSpline} in new code is not recommended. Given data points,
-the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
-instance, representing a spline which interpolates the data. Interpolated values
-at intermediate points can be computed either via its \code{\_\_call\_\_} method,
-or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
-accepts both \code{tck} tuples and \code{Bspline} instances.
-
-Fitting \emph{smoothing} splines to data can currently only be done with FITPACK. Building alternative algorithms is certainly possible, but is still to be implemented.

### 2019-02-15 scipy-1.0/paper.tex
Shift optimize background content for concision of main text

--- 2019-02-01 scipy-1.0/paper.tex
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
-The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
-\label{fig:knn-complexity}
-\end{figure}
+\section*{Numerical Optimization: Additional Details}
+The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
+\begin{enumerate}
+\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
+\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
+\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
+\item general, linear optimization problems (\texttt{linprog}),
+\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
+\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
+\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
+\end{enumerate}
 

### 2019-02-15 scipy-1.0/scipy-optimize.tex
Shift optimize background content for concision of main text

@@
 \newcommand{\RR}{\ensuremath{\mathbb{R}}}
-The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
-\begin{enumerate}
-\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
-\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
-\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
-\item general, linear optimization problems (\texttt{linprog}),
-\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
-\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
-\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
-\end{enumerate}
-Documentation of SciPy's functionality in each these areas can be found in (cite SciPy documentation), but here we highlight recent additions through SciPy 1.0.
+The \texttt{scipy.optimize} module provides functions for the numerical
+solution of several classes of root finding and optimization problems.
+Here we highlight recent additions through SciPy 1.0.
 

### 2019-02-15 scipy-1.0/paper.tex
Table 1 concision / focus-related adjustments for optimization.

@@
 
-% separate Figure numbering for SI
+% separate Figure / Table numbering for SI
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \setcounter{figure}{0}
+\setcounter{table}{0}
 
@@
 
-\section*{Numerical Optimization: Additional Details}
-The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
-\begin{enumerate}
-\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
-\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
-\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
-\item general, linear optimization problems (\texttt{linprog}),
-\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
-\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
-\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
-\end{enumerate}
-
+\begin{table}[H]
+  \centering
+  \caption{Optimization methods from \texttt{minimize} added prior to SciPy 0.13.
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}}\\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark \\
+    \hline
+    Equality constraints &&&&&&&\cmark& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize-si}
+\end{table}
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-15 scipy-1.0/scipy-optimize.tex
Table 1 concision / focus-related adjustments for optimization.

--- 2018-10-05 scipy-1.0/scipy-optimize.tex
@@
 
-The \texttt{minimize} function provides methods for finding local minima of nonlinear optimization
-problems. It unifies several methods with a common interface in order to facilitate
-interchanging between solvers with different characteristics, summarized in Table~\ref{tab:minimize}.
+The \texttt{minimize} function provides methods for finding local minima of
+nonlinear optimization problems. It unifies several methods with a common 
+interface in order to facilitate interchanging between solvers with different
+characteristics. Some recent improvements are summarized in Table~\ref{tab:minimize}.
 %% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).
@@
   \centering
-  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+  \caption{Recently-added optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
 The field \textit{version added} specifies the algorithm's first appearance in SciPy.
-Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
     The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
-    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
-    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
-    with $\sim$ the algorithm does not requires second-order derivatives from
-    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    (e.g. C or FORTRAN). First and second order derivates are required for all methods in the table. 
     \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
@@
     has a superlinear rate for the parameters used in SciPy, but can achieve a quadratic convergence rate with other parameter choices.
-    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
-    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
-  \begin{tabular}{cccccccccccccc}
-      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    These algorithms all guarantee convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. None of the listed algorithms 
+    can deal with constraints on the variables. A similar analysis for optimization algorithms added in older versions of SciPy is available in Table~\ref{tab:minimize-si}.}
+
+  \begin{tabular}{ccccc}
+      & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
     \hline
-    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    Version added & 0.13 & 0.13 & 0.19 & 1.0 \\
     \hline
-    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
-    \hline
-    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
-    \hline
-    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    Wrapper & & & & \cmark \\
     \hline
     \makecell{Iterative Hessian \\
-    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    factorization} &  & \cmark &  & \cmark \\
     \hline
-    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    Local convergence& Q & S$^*$ & Q & S$^*$  \\
     \hline
-    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & TR & TR & TR & TR \\
     \hline
-    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
-    \hline
-    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
-    \hline
-    Equality constraints &&&&&&&\cmark&&&&& \\
-    \hline
-    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
-    \hline
-    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
-      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
-      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
-      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
-      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+    References &
       \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &

### 2019-02-16 scipy-1.0/poly.tex
Add citations from reviewer comments in PR 100.

@@
 At the lowest level of the new design are classes which represent univariate
-piecewise polynomials: \code{PPoly}, \code{BPoly} and \code{BSpline}, which allow
+piecewise polynomials: \code{PPoly} (SciPy 0.13)\cite{scipy-gh5827},
+\code{BPoly} (SciPy 0.13) and \code{BSpline} (SciPy 0.19)\cite{scipy-gh3174},
+which allow
 efficient vectorized evaluations, differentiation, integration and root-finding.
@@
 for e.g., constructing Bezier curves). \code{BSpline} represents spline
-curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}.
+curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}
 
 In the next layer, these polynomial classes are used for constructing several
-common ways of interpolating data: \code{CubicSpline} constructs a twice 
+common ways of interpolating data: \code{CubicSpline} (SciPy 0.18)
+\cite{scipy-gh5653} constructs a twice 
 differentiable piecewise cubic function, \code{Akima1DInterpolator} 

### 2019-02-16 scipy-1.0/paper.tex
added author affiliation

--- 2019-02-15 scipy-1.0/paper.tex
@@
 \affil[19]{Affiliation, department, city, postcode, country}
-\affil[20]{Affiliation, department, city, postcode, country}
+\affil[20]{BioResource and Agricultural Engineering, California Polytechnic State University, San Luis Obispo, CA, 93407, USA}
 \affil[21]{Affiliation, department, city, postcode, country}
@@
 
-\section*{Numerical Optimization: Additional Details}
-The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
-\begin{enumerate}
-\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
-\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
-\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
-\item general, linear optimization problems (\texttt{linprog}),
-\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
-\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
-\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
-\end{enumerate}
-
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: removed comma to state more simply what SciPy is Also, "the fundamental building blocks" suggest _all_ the building blocks. We're not done yet.

@@
 SciPy is a library of numerical routines for the Python programming
-language, providing the fundamental building blocks for modeling and
+language that provides fundamental building blocks for modeling and
 solving scientific problems.  For example, when LIGO, the Laser

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: commented out LIGO sentence in first paragraph "For example" is not precisely appropriate here because the LIGO is not an example of SciPy _providing_ building blocks but of a project making use of them. It is a strong example and belongs somewhere around here... just not quite here.

@@
 language that provides fundamental building blocks for modeling and
-solving scientific problems.  For example, when LIGO, the Laser
+solving scientific problems. 
+// For example, when LIGO, the Laser
 Interferometer Gravitational-Wave Observatory, recently announced the
@@
 signal processing, interpolation, input / output, optimization, statistics,
-integration, and more. SciPy has routines for solving
+integration, and more.
+SciPy has routines for solving
 classes of generic problems such as optimization, integration,

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: nitpicks to improve language precision SciPy routines solve problems of certain classes, but SciPy routines don't solve classes of problems. The routines solve problems of a general form, but the problems they solve are not generic.

@@
 integration, and more.
-SciPy has routines for solving
-classes of generic problems such as optimization, integration,
-interpolation, eigenvalues, and linear systems.  It also provides
+SciPy has routines for the numerical solution of problems from classes
+including linear systems, integration, differential equations, eigenvalue 
+problems, and interpolation.  It also provides
 specialized data structures, such as sparse arrays and k-dimensional

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: style suggestion

--- 2019-02-16 scipy-1.0/paper.tex
@@
 integration, and more.
-SciPy has routines for solving
-classes of generic problems such as optimization, integration,
-interpolation, eigenvalues, and linear systems.  It also provides
+SciPy has routines for the numerical solution of problems from classes
+including linear systems, integration, differential equations, eigenvalue 
+problems, and interpolation; it also provides
 specialized data structures, such as sparse arrays and k-dimensional
 trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
-which provides array data structures and related fast numerical routines.
+which provides array data structures and related fast numerical routines, and
 SciPy is itself the foundation upon which higher level scientific libraries,

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: adding back LIGO example GW150914-Python seemed like a weak reference, and the link between SciPy and the published LIGO results is already established, so I removed it. Added another example of LIGO results that used SciPy via PyCBC.

@@
 solving scientific problems. 
-// For example, when LIGO, the Laser
-Interferometer Gravitational-Wave Observatory, recently announced the
-discovery of gravitational waves, their published Python analysis
-scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
-PhysRevD.93.122003} imported many different SciPy submodules for:
-signal processing, interpolation, input / output, optimization, statistics,
-integration, and more.
 SciPy has routines for the numerical solution of problems from classes
@@
 and scikit-image\cite{vanderwalt2014scikit}, are built. 
+SciPy is relied upon by scientists, engineers, and others 
+around the world. For example, the 
+scripts\cite{alex_nitz_2018_1596771, LIGO-open}
+used in the analysis of gravitational waves
+\cite{PhysRevD.93.122003, abbott2017gw170817} 
+import several submodules of SciPy.
 

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: tweaks

@@
 solving scientific problems. 
-SciPy has routines for the numerical solution of problems from classes
-including linear systems, integration, differential equations, eigenvalue 
-problems, and interpolation; it also provides
+SciPy includes algorithms for optimization, integration, interpolation,
+eigenvalue problems, algebraic equations, differential equations, and many other
+classes of problems; it also provides
 specialized data structures, such as sparse arrays and k-dimensional
@@
 SciPy is relied upon by scientists, engineers, and others 
-around the world. For example, the 
+around the world. For example, published 
 scripts\cite{alex_nitz_2018_1596771, LIGO-open}

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: SciPy 1.0 -> convention belies "The SciPy 1.0 version numbering" doesn't fail to accurately represent the history of the project, but the version numbering convention does.

@@
 signals a library's API (Applications Programming Interface) being
-mature enough to be be trusted in production pipelines.  The SciPy 1.0
-version numbering, however, belies the history of a project that has
+mature enough to be be trusted in production pipelines. This version 
+numbering convention, however, belies the history of a project that has
 become the standard which others follow and has seen extensive

### 2019-02-16 scipy-1.0/paper.tex
decisive doesn't seem like the right word -> auspicious

@@
 an excited, friendly community with a singular focus, seemed to have
-proven decisive in the long run.  It led not only to the library
+proven auspicious in the long run.  It led not only to the library
 described in this paper, but also to an entire ecosystem of related

### 2019-02-16 scipy-1.0/paper.tex
ENH: intro: "Despite its success" doesn't belong here This is sort of negative, so I don't think it really belongs in the introduction. Perhaps it would fit somewhere that describes challenges ahead.
ENH: intro: part -> piece of the puzzle

@@
 
-Despite its success, SciPy still faces challenges:
-a low ``bus factor''---meaning that the entire project depends on a small number of
-active contributors---being perhaps the most serious.
+% Despite its success, SciPy still faces challenges:
+% a low ``bus factor''---meaning that the entire project depends on a small number of
+% active contributors---being perhaps the most serious.
 
@@
 contributions of very many contributors---too many to mention
-individually, but each bringing an important part to the puzzle.
+individually, but each bringing an important piece to the puzzle.
 

### 2019-02-16 scipy-1.0/paper.tex
ENH: background: removed "batteries included standard library" Having "batteries included" suggests that perhaps it should have some special support for scientific computing, then we say it doesn't. I don't think this bit supports what we are really trying to say with this sentence - there were no special features for scientific computing.

@@
 %
-As a general purpose programming language with a ``batteries included''
-standard library, it had no special support for scientific data
-structures or algorithms, unlike many of the other established
+As a general purpose programming language it had no special support for 
+scientific data structures or algorithms, unlike many of the other established
 computation platforms of the time. Yet, scientists soon discovered the

### 2019-02-16 scipy-1.0/paper.tex
ENH: background: removed unneeded comma

@@
 could thereby gain access to a wide variety of existing computational
-libraries, without concerning themselves with low-level programming
+libraries without concerning themselves with low-level programming
 concepts such as memory management.

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: acknowledge typos in original announcement

@@
 
-In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
-message on the Python Matrix Special Interest Group (Matrix-SIG)
+In 1995, Jim Hugunin, a graduate student from MIT, wrote the first
+message in a new Python Matrix Special Interest Group (Matrix-SIG)
 mailing list\cite{Hugunin-first}:
@@
 \end{verbatim}
-\caption{Excerpt from SciPy 0.1 release announcement.}\label{fig:announce-0.1}
+\caption{Excerpt from SciPy 0.1 release announcement (typos included).}\label{fig:announce-0.1}
 \end{figure}

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: removed redundant instead

@@
 adding general features to a language designed for numerical and scientific
-computing, here scientific features are added to a general purpose language
-instead.  This broadens the scope of problems that can be addressed easily, the
+computing, here scientific features are added to a general purpose language.  
+This broadens the scope of problems that can be addressed easily, the
 sources of data that are readily accessible, and increases the size of the

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: sentence needed a verb here; added one

@@
 computing, here scientific features are added to a general purpose language.  
-This broadens the scope of problems that can be addressed easily, the
+This broadens the scope of problems that can be addressed easily, expands the
 sources of data that are readily accessible, and increases the size of the

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: general purpose tooling-> code for the platform

--- 2019-02-18 scipy-1.0/paper.tex
@@
 computing, here scientific features are added to a general purpose language.  
-This broadens the scope of problems that can be addressed easily, the
+This broadens the scope of problems that can be addressed easily, expands the
 sources of data that are readily accessible, and increases the size of the
-community that develops general purpose tooling.
+community that develops code for the platform.
 

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: 'sudden' should describe 'growth', not 'number'

@@
 The availability of a standard numerical array data structure in
-Python led to a sudden and rapidly growing number of scientific
-packages becoming available for solving various common numeric
-problems.
+Python led to a sudden, rapid growth in the number of scientific
+packages available for solving common numeric problems.
 A number of these packages were written by graduate students and

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: ', in charge' -> ', which was in charge'

@@
 %first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
-in 1998 the Space Telescope Science Institute (STScI), in charge of science
-operations of the Hubble Space Telescope, decided to replace their
+in 1998 the Space Telescope Science Institute (STScI), which was in charge of 
+Hubble Space Telescope science operations, decided to replace their
 custom scripting language with Python\cite{STScI-slither}.

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: introduced 'Numeric' (originally referred to without introduction)

@@
 list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
-Hinsen, and Guido van Rossum, led to the creation of an array object
+Hinsen, and Guido van Rossum, led to the creation of a package called Numeric with an array object
 that supported a high number of dimensions.  Jim Hugunin explained the utility

### 2019-02-18 scipy-1.0/paper.tex
ENH: background: added back inadvertently removed comma

@@
 %
-As a general purpose programming language it had no special support for 
+As a general purpose programming language, it had no special support for 
 scientific data structures or algorithms, unlike many of the other established

### 2019-02-19 scipy-1.0/paper.tex
Address reviewer comments in PR 102.

--- 2019-02-18 scipy-1.0/paper.tex
@@
 \affil[19]{Affiliation, department, city, postcode, country}
-\affil[20]{BioResource and Agricultural Engineering, California Polytechnic State University, San Luis Obispo, CA, 93407, USA}
+\affil[20]{Affiliation, department, city, postcode, country}
 \affil[21]{Affiliation, department, city, postcode, country}
@@
 SciPy is a library of numerical routines for the Python programming
-language that provides fundamental building blocks for modeling and
-solving scientific problems. 
-SciPy includes algorithms for optimization, integration, interpolation,
-eigenvalue problems, algebraic equations, differential equations, and many other
-classes of problems; it also provides
+language, providing the fundamental building blocks for modeling and
+solving scientific problems.  For example, when LIGO, the Laser
+Interferometer Gravitational-Wave Observatory, recently announced the
+discovery of gravitational waves, their published Python analysis
+scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
+PhysRevD.93.122003} imported many different SciPy submodules for:
+signal processing, interpolation, input / output, optimization, statistics,
+integration, and more. SciPy has routines for solving
+classes of generic problems such as optimization, integration,
+interpolation, eigenvalues, and linear systems.  It also provides
 specialized data structures, such as sparse arrays and k-dimensional
 trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
-which provides array data structures and related fast numerical routines, and
+which provides array data structures and related fast numerical routines.
 SciPy is itself the foundation upon which higher level scientific libraries,
@@
 and scikit-image\cite{vanderwalt2014scikit}, are built.
-SciPy is relied upon by scientists, engineers, and others 
-around the world. For example, published 
-scripts\cite{alex_nitz_2018_1596771, LIGO-open}
-used in the analysis of gravitational waves
-\cite{PhysRevD.93.122003, abbott2017gw170817} 
-import several submodules of SciPy.
 
@@
 signals a library's API (Applications Programming Interface) being
-mature enough to be be trusted in production pipelines. This version 
-numbering convention, however, belies the history of a project that has
+mature enough to be be trusted in production pipelines.  The SciPy 1.0
+version numbering, however, belies the history of a project that has
 become the standard which others follow and has seen extensive
@@
 an excited, friendly community with a singular focus, seemed to have
-proven auspicious in the long run.  It led not only to the library
+proven decisive in the long run.  It led not only to the library
 described in this paper, but also to an entire ecosystem of related
@@
 
-% Despite its success, SciPy still faces challenges:
-% a low ``bus factor''---meaning that the entire project depends on a small number of
-% active contributors---being perhaps the most serious.
+Despite its success, SciPy still faces challenges:
+a low ``bus factor''---meaning that the entire project depends on a small number of
+active contributors---being perhaps the most serious.
 
@@
 contributions of very many contributors---too many to mention
-individually, but each bringing an important piece to the puzzle.
+individually, but each bringing an important part to the puzzle.
 
@@
 %
-As a general purpose programming language it had no special support for 
-scientific data structures or algorithms, unlike many of the other established
+As a general purpose programming language with a ``batteries included''
+standard library, it had no special support for scientific data
+structures or algorithms, unlike many of the other established
 computation platforms of the time. Yet, scientists soon discovered the
@@
 could thereby gain access to a wide variety of existing computational
-libraries without concerning themselves with low-level programming
+libraries, without concerning themselves with low-level programming
 concepts such as memory management.
 
-In 1995, Jim Hugunin, a graduate student from MIT, wrote the first
-message in a new Python Matrix Special Interest Group (Matrix-SIG)
+In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
+message on the Python Matrix Special Interest Group (Matrix-SIG)
 mailing list\cite{Hugunin-first}:
@@
 adding general features to a language designed for numerical and scientific
-computing, here scientific features are added to a general purpose language.  
-This broadens the scope of problems that can be addressed easily, expands the
+computing, here scientific features are added to a general purpose language
+instead.  This broadens the scope of problems that can be addressed easily, the
 sources of data that are readily accessible, and increases the size of the
-community that develops code for the platform.
+community that develops general purpose tooling.
 
 The availability of a standard numerical array data structure in
-Python led to a sudden, rapid growth in the number of scientific
-packages available for solving common numeric problems.
+Python led to a sudden and rapidly growing number of scientific
+packages becoming available for solving various common numeric
+problems.
 A number of these packages were written by graduate students and
@@
 %first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
-in 1998 the Space Telescope Science Institute (STScI), which was in charge of 
-Hubble Space Telescope science operations, decided to replace their
+in 1998 the Space Telescope Science Institute (STScI), in charge of science
+operations of the Hubble Space Telescope, decided to replace their
 custom scripting language with Python\cite{STScI-slither}.
@@
 \end{verbatim}
-\caption{Excerpt from SciPy 0.1 release announcement (typos included).}\label{fig:announce-0.1}
+\caption{Excerpt from SciPy 0.1 release announcement.}\label{fig:announce-0.1}
 \end{figure}
@@
 
+\begin{table}[H]
+  \centering
+  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
+    \hline
+    Equality constraints &&&&&&&\cmark&&&&& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize-si}
+\end{table}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-19 scipy-1.0/scipy-optimize.tex
Address reviewer comments in PR 102.

@@
 interface in order to facilitate interchanging between solvers with different
-characteristics. Some recent improvements are summarized in Table~\ref{tab:minimize}.
+characteristics. Some recent improvements are summarized in Table~\ref{tab:minimize-si}.
 %% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).
@@
 
-\begin{table}[H]
-  \centering
-  \caption{Recently-added optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
-The field \textit{version added} specifies the algorithm's first appearance in SciPy.
-    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
-    (e.g. C or FORTRAN). First and second order derivates are required for all methods in the table. 
-    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
-    which does not require explicit matrix factorization or storage of the Hessian.
-    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
-    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
-    has a superlinear rate for the parameters used in SciPy, but can achieve a quadratic convergence rate with other parameter choices.
-    These algorithms all guarantee convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. None of the listed algorithms 
-    can deal with constraints on the variables. A similar analysis for optimization algorithms added in older versions of SciPy is available in Table~\ref{tab:minimize-si}.}
-
-  \begin{tabular}{ccccc}
-      & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
-    \hline
-    Version added & 0.13 & 0.13 & 0.19 & 1.0 \\
-    \hline
-    Wrapper & & & & \cmark \\
-    \hline
-    \makecell{Iterative Hessian \\
-    factorization} &  & \cmark &  & \cmark \\
-    \hline
-    Local convergence& Q & S$^*$ & Q & S$^*$  \\
-    \hline
-    \makecell{Line-search (LS) or\\ trust-region (TR)} & TR & TR & TR & TR \\
-    \hline
-    References &
-      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
-      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
-    \hline
-  \end{tabular}
-  \label{tab:minimize}
-\end{table}
-
 \paragraph{Global Minimization}

### 2019-02-19 scipy-1.0/paper.tex

@@
 \end{figure}
+
+\section*{Numerical Optimization: Additional Details}
+The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
+\begin{enumerate}
+\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
+\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
+\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
+\item general, linear optimization problems (\texttt{linprog}),
+\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
+\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
+\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
+\end{enumerate}
 

### 2019-02-19 scipy-1.0/scipy-optimize.tex
Updated table reference (now includes all methods)

@@
 interface in order to facilitate interchanging between solvers with different
-characteristics. Some recent improvements are summarized in Table~\ref{tab:minimize-si}.
+characteristics. Characteristics of the different methods are summarized in 
+Table~\ref{tab:minimize-si}.
 %% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).

### 2019-02-22 scipy-1.0/paper.tex
Change roadmap summary of ndimage
removed unnecessary comma

--- 2019-02-19 scipy-1.0/paper.tex
@@
 % PhD Mayo Clinic, started faculty position at BYU in 2001
-By the late 1990s, discussions appeared on Matrix-SIG,
+By the late 1990s, discussions appeared on Matrix-SIG
 expressing a desire for a complete scientific data analysis environment\cite{Travis-Keynote-2010}.
@@
 
-% separate Figure / Table numbering for SI
+% separate Figure numbering for SI
 \renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
 \setcounter{figure}{0}
-\setcounter{table}{0}
 
@@
 
-\begin{table}[H]
-  \centering
-  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
-The field \textit{version added} specifies the algorithm's first appearance in SciPy.
-Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
-    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
-    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
-    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
-    with $\sim$ the algorithm does not requires second-order derivatives from
-    the user; it computes an approximation internally and uses it to accelerate method convergence.
-    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
-    which does not require explicit matrix factorization or storage of the Hessian.
-    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
-    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
-    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
-    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
-    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
-  \begin{tabular}{cccccccccccccc}
-      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
-    \hline
-    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
-    \hline
-    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
-    \hline
-    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
-    \hline
-    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
-    \hline
-    \makecell{Iterative Hessian \\
-    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
-    \hline
-    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
-    \hline
-    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
-    \hline
-    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
-    \hline
-    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
-    \hline
-    Equality constraints &&&&&&&\cmark&&&&& \\
-    \hline
-    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
-    \hline
-    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
-      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
-      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
-      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
-      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
-      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
-      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
-    \hline
-  \end{tabular}
-  \label{tab:minimize-si}
-\end{table}
-
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-22 scipy-1.0/paper.tex
f2py appears to be stylized F2PY online

@@
 Around the same time, Pearu Peterson, a PhD student from Estonia,
-released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
+released F2PY\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
 codes, and wrote modules for linear algebra and interpolation.

### 2019-02-22 scipy-1.0/paper.tex
fixed F2PY citation

@@
 Around the same time, Pearu Peterson, a PhD student from Estonia,
-released F2PY\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
+released F2PY\cite{peterson2009f2py}, a command line tool for binding Python and Fortran
 codes, and wrote modules for linear algebra and interpolation.

### 2019-02-22 scipy-1.0/paper.tex
CalTech should be Caltech

@@
 workshop\cite{first-scipy-workshop}
-was held in September 2002 at CalTech---a single track, two day event with 50
+was held in September 2002 at Caltech---a single track, two day event with 50
 participants, many of them developers of SciPy and surrounding libraries.

### 2019-02-22 scipy-1.0/paper.tex
removed details about IPython

@@
 In 2001, Fernando Pérez, a graduate student at the University of
-Colorado, Boulder, created the IPython interactive shell (that would
-eventually blossom into Project Jupyter\cite{Kluyver:2016aa}), built into
-IPP, an earlier attempt by Janko Hauser, and including code from
-Nathan Gray's LazyPython.
+Colorado, Boulder, created the IPython interactive shell, which 
+eventually blossomed into Project Jupyter\cite{Kluyver:2016aa}.
 % IPython hosted on SciPy:

### 2019-02-22 scipy-1.0/paper.tex
different scale -> larger scale
He also shares should be past tense

@@
 STScI.  With the Numeric maintainer's blessing, the decision was made
-to write NumArray, a library that could handle data on a different
+to write NumArray, a library that could handle data on a larger
 scale.  Unfortunately, NumArray proved inefficient for small arrays,
@@
 the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
-He also shares his own interest in Python for ``computational steering''
+He also shared his own interest in Python for ``computational steering''
 where ``Python serves as the input language to a scientific application, and the actual

### 2019-02-22 scipy-1.0/paper.tex
removed unnecessary commas

--- 2019-02-22 scipy-1.0/paper.tex
@@
 % PhD Mayo Clinic, started faculty position at BYU in 2001
-By the late 1990s, discussions appeared on Matrix-SIG
+By the late 1990s, discussions appeared on Matrix-SIG,
 expressing a desire for a complete scientific data analysis environment\cite{Travis-Keynote-2010}.
@@
 Around the same time, Pearu Peterson, a PhD student from Estonia,
-released F2PY\cite{peterson2009f2py}, a command line tool for binding Python and Fortran
+released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
 codes, and wrote modules for linear algebra and interpolation.
@@
 workshop\cite{first-scipy-workshop}
-was held in September 2002 at Caltech---a single track, two day event with 50
+was held in September 2002 at CalTech---a single track, two day event with 50
 participants, many of them developers of SciPy and surrounding libraries.
@@
 The SciPy conference went from being a small gathering of
-core developers, to a multi-location event, with increased funding, a
+core developers to a multi-location event with increased funding, a
 published proceedings, and scholarships for attending students (some
@@
 
+\section*{Numerical Optimization: Additional Details}
+The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
+\begin{enumerate}
+\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
+\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
+\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
+\item general, linear optimization problems (\texttt{linprog}),
+\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
+\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
+\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
+\end{enumerate}
+
+\begin{table}[H]
+  \centering
+  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
+    \hline
+    Equality constraints &&&&&&&\cmark&&&&& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize-si}
+\end{table}
+
 %To include, in this order: \textbf{Accession codes} (where applicable);

### 2019-02-22 scipy-1.0/paper.tex
removed '(some of whom are authors)'

@@
 core developers to a multi-location event with increased funding, a
-published proceedings, and scholarships for attending students (some
-of whom are authors on this paper).  In 2010, the conference grew even
+published proceedings, and scholarships for attending students.  
+In 2010, the conference grew even
 further, with an additional event added in India, and the US

### 2019-02-22 scipy-1.0/paper.tex
moving (present continuous) -> moved (past participle)

@@
 further, with an additional event added in India, and the US
-conference—now with parallel tracks—moving to Austin, Texas.
+conference—now with parallel tracks—moved to Austin, Texas.
 While the US SciPy conference kept growing (by 2010 it had multiple

### 2019-02-22 scipy-1.0/paper.tex
revised redundant text about SciPy conference growth

@@
 published proceedings, and scholarships for attending students.
-In 2010, the conference grew even
-further, with an additional event added in India, and the US
-conference—now with parallel tracks—moved to Austin, Texas.
-While the US SciPy conference kept growing (by 2010 it had multiple
-tracks, student sponsorship, proceedings, and funding), satellite
-events were being organized by volunteers elsewhere, such as EuroSciPy
+By 2010, the US SciPy conference had multiple tracks, and satellite
+conferences were being organized by volunteers elsewhere, such as EuroSciPy
 (2008–) and SciPy India (2009–).
 Special sessions and minisymposia dedicated to scientific Python began
-appearing at many events.
+appearing at many other events.
 For example, a three-part minisymposium organized for CSE 2009 was

### 2019-02-22 scipy-1.0/paper.tex
replaced inappropriate comma with 'that'
Capitalized Python in Neuroscience The article title is not capitalized online, but it is referred to with capitalization within the article. In any case, it makes it more clear that it's an article title.

@@
 
-Python has become so pervasive, there are now domain-specific special issues.
+Python has become so pervasive that there are now domain-specific special issues.
 For example, Frontiers in Neuroinformatics has a collection of 25 articles---covering
@@
 as well as stimulus generation and presentation---called Python in
-neuroscience\cite{python-FIN}.
+Neuroscience\cite{python-FIN}.
 

### 2019-02-22 scipy-1.0/paper.tex
another change to past tense

--- 2019-02-22 scipy-1.0/paper.tex
@@
 NumPy\cite{vanderwalt2011numpy}, Cython\cite{behnel2011cython},
-and Mayavi\cite{ramachandran2011mayavi}.
-
-Python has become so pervasive, there are now domain-specific special issues.
-For example, Frontiers in Neuroinformatics has a collection of 25 articles---covering
+and Mayavi\cite{ramachandran2011mayavi}. Python became so pervasive that 
+journals began publishing domain-specific special issues.
+For example, in 2015 Frontiers in Neuroinformatics published a collection of 25 articles---covering
 topics including modeling and simulation, data collection, electrophysiology, visualization,
 as well as stimulus generation and presentation---called Python in
-neuroscience\cite{python-FIN}.
+Neuroscience\cite{python-FIN}.
 

### 2019-02-22 scipy-1.0/paper.tex
avoiding present tense again

@@
 and Fernando Pérez founded NumFOCUS\cite{numfocus},
- a 501(c)3 public charity whose mission is to
-``promote sustainable high-level programming languages, open code development,
+ a 501(c)3 public charity with a mission
+``to promote sustainable high-level programming languages, open code development,
 and reproducible scientific research.''

### 2019-02-22 scipy-1.0/paper.tex
SciPy is not a NumFOCUS sponsored project at this time
reworked Current Status -> SciPy Today

--- 2019-02-22 scipy-1.0/paper.tex
@@
 The SciPy conference went from being a small gathering of
-core developers to a multi-location event with increased funding, a
-published proceedings, and scholarships for attending students.
-By 2010, the US SciPy conference had multiple tracks, and satellite
-conferences were being organized by volunteers elsewhere, such as EuroSciPy
+core developers, to a multi-location event, with increased funding, a
+published proceedings, and scholarships for attending students (some
+of whom are authors on this paper).  In 2010, the conference grew even
+further, with an additional event added in India, and the US
+conference—now with parallel tracks—moving to Austin, Texas.
+While the US SciPy conference kept growing (by 2010 it had multiple
+tracks, student sponsorship, proceedings, and funding), satellite
+events were being organized by volunteers elsewhere, such as EuroSciPy
 (2008–) and SciPy India (2009–).
 Special sessions and minisymposia dedicated to scientific Python began
-appearing at many other events.
+appearing at many events.
 For example, a three-part minisymposium organized for CSE 2009 was
@@
 
-Python has become so pervasive that there are now domain-specific special issues.
+Python has become so pervasive, there are now domain-specific special issues.
 For example, Frontiers in Neuroinformatics has a collection of 25 articles---covering
@@
 as well as stimulus generation and presentation---called Python in
-Neuroscience\cite{python-FIN}.
+neuroscience\cite{python-FIN}.
 
@@
 called PyData.
+
+\subsection*{SciPy Today}
+At the time of writing, the SciPy library consists of nearly 
+600,000 lines of code organized in 16 subpackages. 
+Over 85,000 GitHub repositories and almost 5000 packages depend
+on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
+visitors every two weeks, and more than 100 people contribute to a 
+new release every six months. Some of the major
+feature highlights from the three years preceding
+SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
+and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.7\textwidth]{static/scipy_timeline}
+\caption{Major milestones from SciPy's initial release in 2001 to
+the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
+\label{fig:timeline}
+\end{figure}
 
@@
 for graphics processing units (GPUs) is explicitly out of scope.
-
-\subsection*{Current status}
-
-At the time of writing, SciPy produces a new release
-approximately every six months, and some of the major
-feature highlights from the three years preceding
-SciPy 1.0 are discussed in a technical improvements
-section below. Some longer-term milestone events are
-highlighted in Figure~\ref{fig:timeline}.
-
-\begin{figure}[H]
-\begin{verbatim}
-    2001: first SciPy release
-    2005: transition to NumPy
-    2007: creation of scikits
-    2008: spatial module and first Cython code added
-    2010: move to a 6-monthly release cycle
-    2011: development moves to GitHub
-    2011: Python 3 support
-    2012: add a sparse graph module and unified optimization interface
-    2012: removal of maxentropy
-    2013: continuous integration with TravisCI
-    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
-    2017: add a unified C API with LowLevelCallable; remove weave
-    2017: SciPy 1.0 release
-\end{verbatim}
-\caption{Timeline}
-\label{fig:timeline}
-\end{figure}
 
@@
 \section*{Key technical improvements}
-
+\label{sec:technical_improvements}
 Here we describe key technical improvements made in the last three years.

### 2019-02-22 scipy-1.0/paper.tex
Move project scope section I'm not sure if it belongs in Architecture and implementation choices, but it makes more sense there than where it was.

@@
 
-\subsection*{SciPy Today}
-At the time of writing, the SciPy library consists of nearly 
-600,000 lines of code organized in 16 subpackages. 
-Over 85,000 GitHub repositories and almost 5000 packages depend
-on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
-visitors every two weeks, and more than 100 people contribute to a 
-new release every six months. Some of the major
+\subsection*{Current status}
+
+At the time of writing, SciPy produces a new release
+approximately every six months, and some of the major
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
-and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
+SciPy 1.0 are discussed in a technical improvements
+section below. Some longer-term milestone events are
+highlighted in Figure~\ref{fig:timeline}.
 
 \begin{figure}[H]
-\centering
-\includegraphics[width=0.7\textwidth]{static/scipy_timeline}
-\caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
+\begin{verbatim}
+    2001: first SciPy release
+    2005: transition to NumPy
+    2007: creation of scikits
+    2008: spatial module and first Cython code added
+    2010: move to a 6-monthly release cycle
+    2011: development moves to GitHub
+    2011: Python 3 support
+    2012: add a sparse graph module and unified optimization interface
+    2012: removal of maxentropy
+    2013: continuous integration with TravisCI
+    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
+    2017: add a unified C API with LowLevelCallable; remove weave
+    2017: SciPy 1.0 release
+\end{verbatim}
+\caption{Timeline}
 \label{fig:timeline}
 \end{figure}
+
+% Release management
+
+% Jarrod
+%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
+%Ralf
+%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
+%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
+%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
+
+\section*{Architecture and implementation choices}
 
@@
 for graphics processing units (GPUs) is explicitly out of scope.
-
-% Release management
-
-% Jarrod
-%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
-%Ralf
-%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
-%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
-%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
-
-\section*{Architecture and implementation choices}
 
@@
 \section*{Key technical improvements}
-\label{sec:technical_improvements}
+
 Here we describe key technical improvements made in the last three years.

### 2019-02-24 scipy-1.0/subpackages.tex
Adjusted formatting of package organization to save space

@@
 
-\begin{description}
-
-\item[\texttt{cluster}] ~ \newline
+\begin{description}[leftmargin=!, labelwidth=\widthof{\bfseries \texttt{interpolate}}]
+\itemsep0em
+\item[\texttt{cluster}]
     The \texttt{cluster} subpackage contains two submodules:
@@
     agglomerative clustering.
-\item[\texttt{constants}] ~ \newline
+\item[\texttt{constants}]
     Physical and mathematical constants, including the CODATA recommended
     values of the fundamental physical constants\cite{CODATA2014}.
-\item[\texttt{fftpack}] ~ \newline
+\item[\texttt{fftpack}]
     Fast Fourier Transform routines.  In addition to the FFT, the subpackage
@@
     pseudo-differential operators.
-\item[\texttt{integrate}] ~ \newline
+\item[\texttt{integrate}]
     The \texttt{integrate} subpackage provides tools for the numerical
@@
     problems and two-point boundary value problems.
-\item[\texttt{interpolate}] ~ \newline
+\item[\texttt{interpolate}]
     The \texttt{interpolate} subpackage contains spline functions and
@@
     interpolators, and wrappers for FITPACK and DFITPACK functions.
-\item[\texttt{io}] ~ \newline
+\item[\texttt{io}]
     A collection of functions and classes for reading and writing MATLAB, IDL,
     Matrix Market, Fortran, NetCDF, Harwell-Boeing, WAV and ARFF data files. 
-\item[\texttt{linalg}] ~ \newline
+\item[\texttt{linalg}]
     Linear algebra functions, including:
@@
     Hankel, companion, Hilbert, and more.
-\item[\texttt{ndimage}] ~ \newline
+\item[\texttt{ndimage}]
     This package contains various functions for multi-dimensional image
@@
     functions.
-\item[\texttt{misc}] ~ \newline
+\item[\texttt{misc}]
     A collection of functions that did not fit into the other subpackages.
@@
     to deprecate the contents of this subpackage and eventually remove it.
-\item[\texttt{odr}] ~ \newline
+\item[\texttt{odr}]
     Orthogonal distance regression, including Python wrappers for the Fortran
     library ODRPACK.
-\item[\texttt{optimize}] ~ \newline
-    The package includes:
+\item[\texttt{optimize}]
+    The package includes the following, with additional details in the SI:
     implementations of many minimization algorithms; a general linear
@@
     collection of general nonlinear solvers for root-finding.
-\item[\texttt{signal}] ~ \newline
+\item[\texttt{signal}]
     The \texttt{signal} subpackage focuses on signal processing (plus some
@@
     analysis.  
-\item[\texttt{sparse}] ~ \newline
+\item[\texttt{sparse}]
     This package includes implementations of several representations of
@@
     and more.
-\item[\texttt{spatial}] ~ \newline
+\item[\texttt{spatial}]
     This subpackage provides spatial data structures and algorithms,
@@
     of points or between all pairs from two collections of points.
-\item[\texttt{special}] ~ \newline
+\item[\texttt{special}]
     The name comes the class of functions traditionally known as \emph{special
@@
     and more.
-\item[\texttt{stats}] ~ \newline
+\item[\texttt{stats}]
     The \texttt{stats} subpackage provides: a large collection of continuous

### 2019-02-24 scipy-1.0/subpackages.tex
Adjustments for consistent use of package, subpackage, and module

@@
 \item[\texttt{cluster}]
-    The \texttt{cluster} subpackage contains two submodules:
+    The \texttt{cluster} subpackage contains two modules:
     \texttt{cluster.vq} provides vector quantization and k-means algorithms,
@@
     This package includes implementations of several representations of
-    sparse matrices.  It contains two subpackages, 
+    sparse matrices.  It contains two modules, 
     \texttt{scipy.sparse.linalg} and \texttt{scipy.sparse.csgraph}.
@@
     The name comes the class of functions traditionally known as \emph{special
-    functions}, but over time, the module has grown to include functions
+    functions}, but over time, the subpackage has grown to include functions
     beyond the classical special functions.  A more appropriate characterization

### 2019-02-24 scipy-1.0/paper.tex
Submodule organization -> Package organization Part of an attempt to make the use of words package, subpackage, and module consistent.
attempt to stylize k-... consistently

@@
 interpolation, eigenvalues, and linear systems.  It also provides
-specialized data structures, such as sparse arrays and k-dimensional
+specialized data structures, such as sparse arrays and $k$-dimensional
 trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
@@
 
-\subsection*{Current status}
-
-At the time of writing, SciPy produces a new release
-approximately every six months, and some of the major
-feature highlights from the three years preceding
-SciPy 1.0 are discussed in a technical improvements
-section below. Some longer-term milestone events are
-highlighted in Figure~\ref{fig:timeline}.
-
-\begin{figure}[H]
-\begin{verbatim}
-    2001: first SciPy release
-    2005: transition to NumPy
-    2007: creation of scikits
-    2008: spatial module and first Cython code added
-    2010: move to a 6-monthly release cycle
-    2011: development moves to GitHub
-    2011: Python 3 support
-    2012: add a sparse graph module and unified optimization interface
-    2012: removal of maxentropy
-    2013: continuous integration with TravisCI
-    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
-    2017: add a unified C API with LowLevelCallable; remove weave
-    2017: SciPy 1.0 release
-\end{verbatim}
-\caption{Timeline}
-\label{fig:timeline}
-\end{figure}
-
-% Release management
-
-% Jarrod
-%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
-%Ralf
-%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
-%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
-%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
-
-\section*{Architecture and implementation choices}
-
 \subsection*{Project scope}
@@
 
-\subsection*{Submodule organization}
+\subsection*{Current status}
+
+At the time of writing, SciPy produces a new release
+approximately every six months, and some of the major
+feature highlights from the three years preceding
+SciPy 1.0 are discussed in a technical improvements
+section below. Some longer-term milestone events are
+highlighted in Figure~\ref{fig:timeline}.
+
+\begin{figure}[H]
+\begin{verbatim}
+    2001: first SciPy release
+    2005: transition to NumPy
+    2007: creation of scikits
+    2008: spatial module and first Cython code added
+    2010: move to a 6-monthly release cycle
+    2011: development moves to GitHub
+    2011: Python 3 support
+    2012: add a sparse graph module and unified optimization interface
+    2012: removal of maxentropy
+    2013: continuous integration with TravisCI
+    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
+    2017: add a unified C API with LowLevelCallable; remove weave
+    2017: SciPy 1.0 release
+\end{verbatim}
+\caption{Timeline}
+\label{fig:timeline}
+\end{figure}
+
+% Release management
+
+% Jarrod
+%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
+%Ralf
+%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
+%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
+%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
+
+\section*{Architecture and implementation choices}
+
+\subsection*{Package organization}
 
@@
 
-\subsubsection*{cKDTree}
+\subsubsection*{\texttt{cKDTree}}
 
@@
 \includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query()}
 over a roughly nine year commit history time frame. The results are based on
@@
 backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the KDTree (boxsize argument was ignored).}
+of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
 \label{fig:asvbench}
@@
 \includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
 The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}

### 2019-02-24 scipy-1.0/subpackages.tex
attempt to stylize k-... consistently

@@
 subpackages include mathematical building blocks (e.g. linear algebra, Fourier
-transforms, special functions), data structures (e.g. sparse matrices, k-D trees),
+transforms, special functions), data structures (e.g. sparse matrices, $k$-D trees),
 algorithms (e.g. numerical optimization and integration, clustering, interpolation,
@@
     The \texttt{cluster} subpackage contains two modules:
-    \texttt{cluster.vq} provides vector quantization and k-means algorithms,
+    \texttt{cluster.vq} provides vector quantization and $k$-means algorithms,
     and \texttt{cluster.hierarchy} provides functions for hierarchical and
@@
     This subpackage provides spatial data structures and algorithms,
-    including the k-d tree, Delaunay triangulation, convex hulls and Voronoi
+    including the $k$-d tree, Delaunay triangulation, convex hulls and Voronoi
     diagrams.  The subpackage \texttt{scipy.spatial.distance} provides

### 2019-02-24 scipy-1.0/subpackages.tex
eliminated inappropriate comma Doesn't match any of the rules for use here: https://owl.purdue.edu/owl/general_writing/punctuation/commas/extended_rules_for_commas.html Similar to example of comma abuse in #13

@@
 \item[\texttt{fftpack}]
-    Fast Fourier Transform routines.  In addition to the FFT, the subpackage
+    Fast Fourier Transform routines.  In addition to the FFT itself, the subpackage
     includes functions for the discrete sine and cosine transforms and for
@@
     The \texttt{integrate} subpackage provides tools for the numerical
-    computation of single and multiple definite integrals, and for the
+    computation of single and multiple definite integrals and for the
     solution of ordinary differential equations, including initial value

### 2019-02-24 scipy-1.0/subpackages.tex
removed inappropriate colons http://www.kentlaw.edu/academics/lrw/grinker/LwtaColons.htm

@@
     This package contains various functions for multi-dimensional image
-    processing, including: convolution and assorted linear and nonlinear
+    processing, including convolution and assorted linear and nonlinear
     filters (Gaussian filter, median filter, Sobel filter, etc.);
@@
     of this subpackage is simply \emph{useful functions}.
-    It includes: a large collection of the classical special functions
+    It includes a large collection of the classical special functions
     such as Airy, Bessel, etc.; orthogonal families of polynomials;
@@
 \item[\texttt{stats}]
-    The \texttt{stats} subpackage provides: a large collection of continuous
+    The \texttt{stats} subpackage provides a large collection of continuous
     and discrete \emph{probability distributions}, each with methods to compute

### 2019-02-24 scipy-1.0/subpackages.tex
to deprecate -> to deprecate or relocate
distinguished between linear and nonline minimization

@@
     A collection of functions that did not fit into the other subpackages.
-    While this subpackage still exists in version 1.0.0, effort is continuing
-    to deprecate the contents of this subpackage and eventually remove it.
+    While this subpackage still exists in SciPy 1.0, effort is underway
+    to deprecate or relocate the contents of this subpackage and eventually remove it.
 \item[\texttt{odr}]
@@
     The package includes the following, with additional details in the SI:
-    implementations of many minimization algorithms; a general linear
-    programming solver; a routine for least-squares curve fitting; and a
-    collection of general nonlinear solvers for root-finding.
+    simplex and interior-point linear programming solvers,
+		implementations of many nonlinear minimization algorithms,  
+		a routine for least-squares curve fitting, and 
+		a collection of general nonlinear solvers for root-finding.
 \item[\texttt{signal}]

### 2019-02-24 scipy-1.0/subpackages.tex
'(plus some...' -> 'and'. More professional?
found another inappropriate colon

@@
 \item[\texttt{signal}]
-    The \texttt{signal} subpackage focuses on signal processing (plus some
-    basic linear systems theory).  Functionality includes:
+    The \texttt{signal} subpackage focuses on signal processing and
+    basic linear systems theory.  Functionality includes
     convolution and correlation; splines; filtering and filter design;

### 2019-02-24 scipy-1.0/paper.tex
removed inappropriate colons http://www.kentlaw.edu/academics/lrw/grinker/LwtaColons.htm
fixed incomplete sentence... and got carried away

--- 2019-02-22 scipy-1.0/paper.tex
@@
 
-For implementing new functionality, we have a clear order of language
-preference.  First Python, if performance is not an issue. If it is, then in
-order of preference: Cython, C, C++, Fortran. The main motivation for this is
+For implementing new functionality, Python is the first choice. If Python performance is an issue, then 
+we use Cython, C, C++, or Fortran (listed in order of preference). The main motivation for this is
 maintainability: Cython has the highest abstraction level and most Python

### 2019-02-24 scipy-1.0/paper.tex
gave Cython its own paragraph
Modified for clarity/precision "Cython has been described as a creole language..." does not suggest that Cython should be used "as a glue between well-established low-level scientific computing libraries", so "Thus" was inappropriate here. Removed unecessary bit about Cython in previous paragraph as the use of Cython to expose C/C++ is covered here, and I don't think it was needed as a transition into this paragraph.

@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. Alongside its objected-oriented relative C++, C can be
-leveraged in Python after being exposed using the Cython language. Cython has
-been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We thus often use Cython
-as a glue between well-established low-level scientific computing libraries and
-the Python interface offered by SciPy. We also use Cython to enable performance
-enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing. Some of the C libraries
+scientific computing. Some of the C libraries
 that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
@@
 algorithms).
+
+Cython has been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
+as a glue between well-established low-level scientific computing libraries written
+in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
+enhancements in Python code, especially for cases where heavily used inner
+loops benefit from a compiled code with static typing.
 

### 2019-02-24 scipy-1.0/paper.tex
"core development team" -> "current core development team" Trying to explain the disconnect between the substantial use of Fortran and C in existing SciPy code and our order of language preference.
Additional modifications to explain difference between language preferences and historical use

@@
 
-For implementing new functionality, Python is the first choice. If Python performance is an issue, then 
-we use Cython, C, C++, or Fortran (listed in order of preference). The main motivation for this is
-maintainability: Cython has the highest abstraction level and most Python
-developers will understand it. C is also widely known, and easier to deal with
-for the core development team than C++ and especially Fortran.
-
-While it is not surprising that Python is heavily used in SciPy, the usage
-distribution of the other languages warrants some discussion. Fortran is an
+For implementing \emph{new} functionality, Python is the language of choice. If Python 
+performance is an issue, then we use Cython, C, C++, or Fortran (listed in order 
+of preference). The main motivation for this is maintainability: Cython has the 
+highest abstraction level and most Python developers will understand it. C is 
+also widely known, and easier to deal with for the current core development team 
+than C++ and especially Fortran. While it is not surprising that Python is heavily 
+used in SciPy, the usage distribution of the other languages \textemdash despite 
+our preferences for implementing new functionality \textemdash warrants discussion.
+
+Fortran is an
 extremely well-established scientific programming language, both for historical

### 2019-02-24 scipy-1.0/paper.tex
dictionaries agree that performance-enhancing is hyphenated

@@
 Therefore, the relative abundance of different programming languages in the
-SciPy library results from a combination of the usage of powerful performance
+SciPy library results from a combination of the usage of powerful performance-
 enhancing languages that interact well with Python (i.e., Cython), and the

### 2019-02-24 scipy-1.0/paper.tex
many more suggestions

--- 2019-02-24 scipy-1.0/paper.tex
@@
 
-Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library\cite{linguistref} provides a 
-detailed breakdown as \% composition by programming language in 
-SciPy (Figure~\ref{fig:linguist}).
+Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
+in SciPy. While it is not surprising that Python is predominant, the usage distribution 
+of the other languages warrants discussion.
 
@@
 \centering
-
-    \begin{tikzpicture}[font=\sffamily]
-    \begin{axis}[height=6cm, width=10cm,
-    ybar, ymax=60, ymin=0,bar width=1cm,
-    ylabel={\% of Scipy code base},
-    ytick={0,20,...,60},
-    axis lines*=left,
-    axis y line shift=0pt,
-    enlarge x limits=0.15,
-    symbolic x coords={py,f,c,cy,cpp},
-    xticklabels={Python,FORTRAN,C,Cython,C++},
-    ticklabel style={align=center,font=\sffamily\small,
-    /pgf/number format/assume math mode=true},
-    xtick=data,
-    nodes near coords,
-    nodes near coords style={/pgf/number format/assume math mode=true}
-    ]
-    \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
-    \end{axis}
-    \end{tikzpicture}
-    \caption{The breakdown of programming languages used in the
-             SciPy library determined using the linguist library.
+\includegraphics[width=0.5\textwidth]{static/scipy_languages}
+\caption{Composition of SciPy: lines of code in each programming language as determined by the \texttt{linguist} package.
 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
-    	 and Makefile are excluded for clarity and mostly
-    	 provide supporting roles in tests, building, and
-    	 documentation.}
+and Makefile (not visible) provide supporting roles in tests,
+building, and documentation.}
 \label{fig:linguist}
@@
 
-For implementing new functionality, Python is the first choice. If Python performance is an issue, then 
-we use Cython, C, C++, or Fortran (listed in order of preference). The main motivation for this is
-maintainability: Cython has the highest abstraction level and most Python
-developers will understand it. C is also widely known, and easier to deal with
-for the core development team than C++ and especially Fortran.
-
-While it is not surprising that Python is heavily used in SciPy, the usage
-distribution of the other languages warrants some discussion. Fortran is an
+%\begin{figure}[H]
+    %\centering
+%
+    %\begin{tikzpicture}[font=\sffamily]
+    %\begin{axis}[height=6cm, width=10cm,
+    %ybar, ymax=60, ymin=0,bar width=1cm,
+    %ylabel={\% of Scipy code base},
+    %ytick={0,20,...,60},
+    %axis lines*=left,
+    %axis y line shift=0pt,
+    %enlarge x limits=0.15,
+    %symbolic x coords={py,f,c,cy,cpp},
+    %xticklabels={Python,FORTRAN,C,Cython,C++},
+    %ticklabel style={align=center,font=\sffamily\small,
+    %/pgf/number format/assume math mode=true},
+    %xtick=data,
+    %nodes near coords,
+    %nodes near coords style={/pgf/number format/assume math mode=true}
+    %]
+    %\addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
+    %\end{axis}
+    %\end{tikzpicture}
+    %\caption{The breakdown of programming languages used in the
+             %SciPy library determined using the \texttt{linguist} library.
+    	 %Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
+    	 %and Makefile are excluded for clarity and mostly
+    	 %provide supporting roles in tests, building, and
+    	 %documentation.}
+    %\label{fig:linguist}
+%\end{figure}
+
+Fortran is an
 extremely well-established scientific programming language, both for historical
@@
 
-Therefore, the relative abundance of different programming languages in the
-SciPy library results from a combination of the usage of powerful performance
-enhancing languages that interact well with Python (i.e., Cython), and the
+For implementing new functionality, Python is the still the language of choice. If Python 
+performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
+order). The main motivation for this is maintainability: Cython has the 
+highest abstraction level and most Python developers will understand it. C is 
+also widely known, and easier for the current core development team to manage
+than C++ and especially Fortran.
+
+The distribution of secondary programming languages in SciPy is a compromise between
+a powerful, performance-
+enhancing language that interacts well with Python (i.e. Cython) and the
 usage of languages (and their libraries) that have proven reliable and

### 2019-02-25 scipy-1.0/paper.tex
revised figure

--- 2019-02-24 scipy-1.0/paper.tex
@@
 
-\subsection*{Current status}
-
-At the time of writing, SciPy produces a new release
-approximately every six months, and some of the major
+\subsection*{SciPy Today}
+At the time of writing, the SciPy library consists of nearly 
+600,000 lines of code organized in 16 subpackages. 
+Over 85,000 GitHub repositories and almost 5000 packages depend
+on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
+visitors every two weeks, and more than 100 people contribute to a 
+new release every six months. Some of the major
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in a technical improvements
-section below. Some longer-term milestone events are
-highlighted in Figure~\ref{fig:timeline}.
+SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
+and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
 
 \begin{figure}[H]
-\begin{verbatim}
-    2001: first SciPy release
-    2005: transition to NumPy
-    2007: creation of scikits
-    2008: spatial module and first Cython code added
-    2010: move to a 6-monthly release cycle
-    2011: development moves to GitHub
-    2011: Python 3 support
-    2012: add a sparse graph module and unified optimization interface
-    2012: removal of maxentropy
-    2013: continuous integration with TravisCI
-    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
-    2017: add a unified C API with LowLevelCallable; remove weave
-    2017: SciPy 1.0 release
-\end{verbatim}
-\caption{Timeline}
+\centering
+\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
+\caption{Major milestones from SciPy's initial release in 2001 to
+the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
 \label{fig:timeline}
 \end{figure}
-
-% Release management
-
-% Jarrod
-%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
-%Ralf
-%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
-%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
-%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
-
-\section*{Architecture and implementation choices}
 
@@
 for graphics processing units (GPUs) is explicitly out of scope.
+
+% Release management
+
+% Jarrod
+%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
+%Ralf
+%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
+%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
+%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
+
+\section*{Architecture and implementation choices}
 
@@
 
-For implementing \emph{new} functionality, Python is the language of choice. If Python 
-performance is an issue, then we use Cython, C, C++, or Fortran (listed in order 
-of preference). The main motivation for this is maintainability: Cython has the 
-highest abstraction level and most Python developers will understand it. C is 
-also widely known, and easier to deal with for the current core development team 
-than C++ and especially Fortran. While it is not surprising that Python is heavily 
-used in SciPy, the usage distribution of the other languages \textemdash despite 
-our preferences for implementing new functionality \textemdash warrants discussion.
-
-Fortran is an
+For implementing new functionality, we have a clear order of language
+preference.  First Python, if performance is not an issue. If it is, then in
+order of preference: Cython, C, C++, Fortran. The main motivation for this is
+maintainability: Cython has the highest abstraction level and most Python
+developers will understand it. C is also widely known, and easier to deal with
+for the core development team than C++ and especially Fortran.
+
+While it is not surprising that Python is heavily used in SciPy, the usage
+distribution of the other languages warrants some discussion. Fortran is an
 extremely well-established scientific programming language, both for historical
@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. Some of the C libraries
+scientific computing. Alongside its objected-oriented relative C++, C can be
+leveraged in Python after being exposed using the Cython language. Cython has
+been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We thus often use Cython
+as a glue between well-established low-level scientific computing libraries and
+the Python interface offered by SciPy. We also use Cython to enable performance
+enhancements in Python code, especially for cases where heavily used inner
+loops benefit from a compiled code with static typing. Some of the C libraries
 that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
@@
 algorithms).
-
-Cython has been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
-as a glue between well-established low-level scientific computing libraries written
-in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
-enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing.
 

### 2019-02-25 scipy-1.0/paper.tex
Scipy matures revisions (#106)
* removed unnecessary commas

* removed '(some of whom are authors)'

* moving (present continuous) -> moved (past participle)

* revised redundant text about SciPy conference growth

* replaced inappropriate comma with 'that'

* Capitalized Python in Neuroscience
The article title is not capitalized online, but it is referred to with
capitalization within the article. In any case, it makes it more clear that
it's an article title.

* another change to past tense

* avoiding present tense again

* SciPy is not a NumFOCUS sponsored project at this time

@@
 The SciPy conference went from being a small gathering of
-core developers, to a multi-location event, with increased funding, a
-published proceedings, and scholarships for attending students (some
-of whom are authors on this paper).  In 2010, the conference grew even
-further, with an additional event added in India, and the US
-conference—now with parallel tracks—moving to Austin, Texas.
-While the US SciPy conference kept growing (by 2010 it had multiple
-tracks, student sponsorship, proceedings, and funding), satellite
-events were being organized by volunteers elsewhere, such as EuroSciPy
+core developers to a multi-location event with increased funding, a
+published proceedings, and scholarships for attending students.
+By 2010, the US SciPy conference had multiple tracks, and satellite
+conferences were being organized by volunteers elsewhere, such as EuroSciPy
 (2008–) and SciPy India (2009–).
 Special sessions and minisymposia dedicated to scientific Python began
-appearing at many events.
+appearing at many other events.
 For example, a three-part minisymposium organized for CSE 2009 was
@@
 came from the scientific Python stack.
-Today it has over 25 sponsored projects including NumPy, SciPy, IPython, and
+Today it has 50 affiliated and sponsored projects including NumPy, SciPy, IPython, and
 Matplotlib.
@@
 
-\subsection*{SciPy Today}
-At the time of writing, the SciPy library consists of nearly 
-600,000 lines of code organized in 16 subpackages. 
-Over 85,000 GitHub repositories and almost 5000 packages depend
-on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
-visitors every two weeks, and more than 100 people contribute to a 
-new release every six months. Some of the major
+\subsection*{Current status}
+
+At the time of writing, SciPy produces a new release
+approximately every six months, and some of the major
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
-and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
+SciPy 1.0 are discussed in a technical improvements
+section below. Some longer-term milestone events are
+highlighted in Figure~\ref{fig:timeline}.
 
 \begin{figure}[H]
-\centering
-\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
-\caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
+\begin{verbatim}
+    2001: first SciPy release
+    2005: transition to NumPy
+    2007: creation of scikits
+    2008: spatial module and first Cython code added
+    2010: move to a 6-monthly release cycle
+    2011: development moves to GitHub
+    2011: Python 3 support
+    2012: add a sparse graph module and unified optimization interface
+    2012: removal of maxentropy
+    2013: continuous integration with TravisCI
+    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
+    2017: add a unified C API with LowLevelCallable; remove weave
+    2017: SciPy 1.0 release
+\end{verbatim}
+\caption{Timeline}
 \label{fig:timeline}
 \end{figure}
+
+% Release management
+
+% Jarrod
+%https://mail.python.org/pipermail/scipy-dev/2007-August/007613.html
+%Ralf
+%https://mail.python.org/pipermail/numpy-discussion/2010-March/049057.html
+%https://mail.python.org/pipermail/scipy-dev/2010-March/014034.html
+%https://mail.python.org/pipermail/scipy-dev/2010-April/014091.html
+
+\section*{Architecture and implementation choices}
 

### 2019-02-27 scipy-1.0/paper.tex

@@
 Around the same time, Pearu Peterson, a PhD student from Estonia,
-released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
+released F2PY\cite{peterson2009f2py}, a command line tool for binding Python and Fortran
 codes, and wrote modules for linear algebra and interpolation.

### 2019-02-27 scipy-1.0/paper.tex

@@
 \section*{Architecture and implementation choices}
-
 \subsection*{Project scope}
@@
 
-\subsection*{Submodule organization}
-
+\subsection*{Package organization}
 \input{subpackages}
@@
     point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
     \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}

### 2019-03-01 scipy-1.0/paper.tex
not sure what I changed

--- 2019-02-27 scipy-1.0/paper.tex
@@
 
-\subsubsection*{cKDTree}
-
-\input{ckdtree}
-
 \subsubsection*{Sparse matrices}
@@
 These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily in scikit-learn.
-
-Over the last three years our sparse matrix handling internals have been
+and are used heavily throughout SciPy and dependent packages.
+
+Over the last three years, our sparse matrix handling internals have been
 rewritten and performance has been improved. Iterating over and slicing of CSC
 and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the COO / DIA to CSR / CSC matrix format
+ and the coordinate (COO)/ diagonal (DIA) to CSR / CSC matrix format
 conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
@@
 equivalent NumPy API where possible.
+
+\subsubsection*{cKDTree}
+
+\input{ckdtree}
 

### 2019-03-01 scipy-1.0/ckdtree.tex
introduced kdtrees

@@
-The cKDTree module was rewritten in C++ with templated classes, and support for
-periodic boundary conditions was added. A periodic boundary condition is typically 
-used in computer simulations of physical processes.
+The `cKDTree` module, which implements a space-partitioning data structure that
+organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 
+Support was added for periodic boundary conditions, which are often used 
+in simulations of physical processes.
 

### 2019-03-01 scipy-1.0/paper.tex
LowLevelCallable() -> LowLevelCallable

@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
+in a \texttt{scipy.LowLevelCallable} object that reduces the overhead for
 calling compiled C functions directly from Python.  Supported low-level

### 2019-03-01 scipy-1.0/paper.tex
reduces the overhead for -> overhead of

@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable} object that reduces the overhead for
+in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
 calling compiled C functions directly from Python.  Supported low-level

### 2019-03-01 scipy-1.0/paper.tex
ctypes/cffi should be teletype
Tightened LowLevelCallable section I removed details that, in my opinion, weakened the section. If these details are very important and can potentially strengthen the section, please let me know. Perhaps I just didn't understand. If that's the case, we can work together on how they are presented.

@@
 in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions directly from Python.  Supported low-level
-functions include \texttt{PyCapsule} objects, ctypes function pointers, and
-cffi function pointers. The low level function signature must be consistent
-with the expectations of the routine it is passed to. For example, the
-documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
-C callback function signatures that may be used to produce functions that
-operate on each element of image data with low overhead. The C code may be
-generated using numba or Cython, for example, as long as the function call
-signature matches the specifications. Furthermore, it is even possible to
-generate a low-level callback function automatically from a Cython module using
-\texttt{scipy.LowLevelCallable.from\_cython}.
+calling compiled C functions directly from Python.  
+Supported low-level functions include \texttt{PyCapsule} 
+objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers. 
+Furthermore, it is even possible to generate a low-level callback function 
+automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
 

### 2019-03-01 scipy-1.0/paper.tex
Cython bindings... should be a subsubsection
removed inappropriate colon

@@
 
-\subsection*{Cython bindings for BLAS, LAPACK, and special}
+\subsubsection*{Cython bindings for BLAS, LAPACK, and special}
 
@@
 LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
-includes Cython\cite{behnel2011cython} wrappers for:
+includes Cython\cite{behnel2011cython} wrappers for
 many BLAS and LAPACK routines (added in 2015), and the special functions 

### 2019-03-01 scipy-1.0/paper.tex
removed inappropriate comma

@@
 includes Cython\cite{behnel2011cython} wrappers for
-many BLAS and LAPACK routines (added in 2015), and the special functions 
+many BLAS and LAPACK routines (added in 2015) and the special functions 
 provided in the \texttt{scipy.{\allowbreak}special} submodule (added in 2016).  

### 2019-03-01 scipy-1.0/paper.tex
ABIs less objectionable than ABI's

@@
 Fortran compiler, they can also be used without having to handle all the
-different Fortran compiler ABI's and name mangling schemes.
+different Fortran compiler ABIs and name mangling schemes.
 

### 2019-03-01 scipy-1.0/scipy-optimize.tex
minimized minimize

@@
 
-The \texttt{minimize} function provides methods for finding local minima of
-nonlinear optimization problems. It unifies several methods with a common 
-interface in order to facilitate interchanging between solvers with different
-characteristics. Characteristics of the different methods are summarized in 
-Table~\ref{tab:minimize-si}.
-%% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).
-The reader can easily find the standard of meaning of `nonlinear programming', so I believe it should be removed.
-If we decide that it is appropriate to include, then we should add the entire problem statement - including constraints - and we should also include the entire linear programming problem statement in the previous section.
-% summarizes all the methods available for solving minimization problems of the type:
-%\begin{equation}
-%  \label{eq:minimization-prob}
-%  \min_x f(x)
-%\end{equation}
-%where $f$ is a multivariate function $f: \mathbb{R}^n \rightarrow \mathbb{R}$ and $x \in \mathbb{R}^n$.
-%%  Without the problem statement this sentence isn't really appropriate.
-%\texttt{COBYLA} and \texttt{SLSQP} methods support nonlinear constraints, \texttt{L-BFGS-B} and \texttt{TNC} allow box constraints $x^l \le x \le x^u$,
-%and the remaining methods are for unconstrained problems.
-
-% The available methods implement a variety of solution strategies.
-Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
-noisy objective functions. When derivatives are available, however, other methods offer faster convergence rates
-and have proven convergence properties. 
-The derivative-based methods start with an initial guess and refine it iteratively, seeking
-a local solution to the optimization problem. Each of these algorithms implements one of two complementary strategies:
-line-search or trust-region. At each iteration, the line-search (LS) methods choose a direction
-and approximate the optimal point in the given direction. Trust-region (TR) methods, on the other
-hand, choose a ``trust-region'' region for which a local model is valid and approximate the best point
-inside.
-
-One recently-added trust-region method is \texttt{trust-exact},
-which achieves fast convergence by solving the trust-region subproblem almost exactly.
-Unfortunately, it requires the Hessian to be stored and factored every iteration, which may preclude
-the solution of large problems (e.g.~problems with more than 1000 variables).
-Other methods are appropriate for large-scale problems, including
-\texttt{CG}, which does not use second derivatives and thus has low memory requirements, and \texttt{L-BFGS-B}, which maintains a simple and compact representation of the Hessian.
-\texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use second order derivatives to achieve faster
-convergence, but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way that does not require the Hessian to be stored.
-The method \texttt{trust-krylov}, introduced in the latest release, is a good compromise between
-\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
-subproblem than \texttt{trust-ncg}, but it does not require the Hessian to be stored and factored as in \texttt{trust-exact}.
-
-While most \texttt{minimize} development prior to SciPy 1.0 has focused on unconstrained minimization, 
-% methods \texttt{SLSQP} and \texttt{COBYLA}\footnote{We note that any equality constraint $c_{\text{eq}}(x) = 0$ can be expressed as two inequality constraints $c_{\text{eq}}(x) \leq 0$ and $-c_{\text{eq}}(x) \leq 0$}
-\texttt{SLSQP} supports fully general nonlinear programming.
-However, it does not provide special treatment for sparsity, limiting its utility to small problems.
-Accordingly, ongoing development efforts for \texttt{minimize} are focused on the addition of nonlinear programming algorithms for large, sparse problems.
+The \texttt{minimize} function unifies several methods of finding local minima of nonlinear optimization problems in a common interface.
+Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
+All are trust-region methods that build a local model of the objective function based on first- and second-derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
+For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second-derivative Hessian matrix to be stored and factored every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
+On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second-derivative information in a faster, approximate way.
+A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}.
 

### 2019-03-01 scipy-1.0/scipy-optimize.tex
eliminated old global minimizers brute and basinhopping

@@
 As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values, especially if the objective function possesses many local minima that other solvers could get stuck in.
-% e.g.
-if there is concern that `minimize` could get stuck in an unacceptable local minimum?
-The \texttt{brute} function performs a grid search over the entire domain, and the best grid point is ``polished'' by a local minimizer.
-However, this approach becomes prohibitively expensive as the problem dimensionality increases.
-% As far as I understand how brute works, the run time does not depend on the number of extrema, as it only polishes the best one
-% , or if the objective function has many extrema. 
-To address this, the
-\texttt{basinhopping} \cite{Wales1997} function uses random perturbation of the decision variables in conjunction with local minimization.
-The acceptance of the perturbed coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
-\texttt{differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic solver that improves by gradual evolution of a population of candidate solutions.
-In each iteration, trial candidates are generated by combination of candidates from the existing population; if the trial candidates represent an improvement then the population is updated.
-% Is the following a unique feature of differential_evolution?
-If so, state it (although it seems that neither `brute`or `basinhopping` inherently require the a differentiable problem).
-If not, why mention it?
-% \texttt{differential\textunderscore evolution} does not require that the objective function be differentiable.
-%  basinhopping may need differentiable objective for the local minimization.
-Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems
-% They're ``comprehensive''
-% of varying difficulties, 
-% that can be used to evaluate the performance of the solvers. 
-%This suite is used to decide
-for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits inclusion in the package. 
-% have good enough performance to be added to the package.
+The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by ``evolving'' a population of candidate solutions.
+In each iteration, trial candidates are generated by combination of candidates from the existing population.
+If the trial candidates represent an improvement, then the population is updated.
+Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits inclusion in the package.

### 2019-03-01 scipy-1.0/scipy-optimize.tex
removed the unnecessary quotes I added

@@
 As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values, especially if the objective function possesses many local minima that other solvers could get stuck in.
-The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by ``evolving'' a population of candidate solutions.
+The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by evolving a population of candidate solutions.
 In each iteration, trial candidates are generated by combination of candidates from the existing population.

### 2019-03-01 scipy-1.0/paper.tex
attempt at clarification

@@
 calculate moments, cumulative distribution function (cdf), differential
-entropy, and more. Those properties are calculated based on generic code, e.g.,
-numerical integration of the pdf to obtain the cdf.  Where possible, the
-generic code is replaced by explicit formulas or more accurate or efficient
-implementations. For example, most distributions have an explicit formula for
+entropy, and more. By default, these properties are calculated by generic code, e.g.
+numerical integration of the pdf to obtain the cdf, but the
+generic code can be replaced by explicit formulas or more accurate/efficient
+implementations when available. For example, most distributions have an explicit formula for
 the cdf, so it is used instead of integrating the pdf. The distributions

### 2019-03-01 scipy-1.0/paper.tex
tweaked bit about multinomial distribution

@@
 \texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
-\texttt{scipy.stats.multinomial} (used, for example, in natural language
-processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
-the only multivariate discrete distribution available in SciPy.
+\texttt{scipy.stats.multinomial}, the first discrete multivariate distribution 
+available in SciPy.
 

### 2019-03-01 scipy-1.0/poly.tex
b-spline -> B-spline

--- 2018-07-17 scipy-1.0/poly.tex
@@
 %
-Routines for manipulating piecewise polynomials and b-splines were either not available or lacking in ease of use and/or performance characteristics.
+Routines for manipulating piecewise polynomials and B-splines were either not available or lacking in ease of use and/or performance characteristics.
 
@@
 
-For working with b-splines, there are several options.
-The basic object is \code{BSpline}, which represents spline curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}.
+For working with B-splines, there are several options.
+The basic object is \code{BSpline}, which represents spline curves, i.e., linear combinations of B-spline basis elements.\cite{deBoor1978}.
 The spline

### 2019-03-01 scipy-1.0/paper.tex
consistently stylized asv in teletype

@@
 
-The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
+The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
 over their lifetimes, and the performance of the SciPy code base was monitored
-with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
+with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
 unit tests are passing (see Test suite section below), confirming 
@@
 
-Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+Consider the \texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
 roughly nine years of project history. These demonstrate the gradual

### 2019-03-01 scipy-1.0/paper.tex
removed unnecessary `to empower scientific applications'

@@
 remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
+us to objectively measure that our code base is improving.
 

### 2019-03-01 scipy-1.0/paper.tex
attempted to make consistent code base / codebase

@@
 Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library\cite{linguistref} provides a 
+implement scientific algorithms in the SciPy library. An analysis of our 
+codebase using the \texttt{linguist} library\cite{linguistref} provides a 
 detailed breakdown as \% composition by programming language in 
@@
     ybar, ymax=60, ymin=0,bar width=1cm,
-    ylabel={\% of Scipy code base},
+    ylabel={\% of Scipy codebase},
     ytick={0,20,...,60},
@@
 The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
+over their lifetimes, and the performance of the SciPy codebase was monitored
 with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
@@
 remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving.
+us to objectively measure that our codebase is improving.
 
@@
 holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the code base would still benefit from increased
+the historical components of the codebase would still benefit from increased
 test coverage. Documentation for the code is automatically built and hosted by
@@
 \includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy code base by unit tests over the six months
+\caption{\% coverage of the SciPy codebase by unit tests over the six months
 preceding June 28, 2018 as reported by the \texttt{codecov} service. The
@@
 libraries\cite{astropy-2018}, and what works well for 
-one of the code bases, infrastructures, or communities 
-is often transferred in some form to the other. At the code
-base level, the \texttt{binned\_statistic} functionality
+one of the codebases, infrastructures, or communities 
+is often transferred in some form to the other. At the
+codebase level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution---it was initially
@@
 metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
+14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
 that roughly two-week period. The developer community consists of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the code base
+contributors of source code, with $>19,000$ commits accepted into the codebase
 (GitHub page data).

### 2019-03-01 scipy-1.0/paper.tex
benchmark suite refers to test suite -> put test suite first.

@@
 
+\subsubsection*{Test suite}
+\label{sec:testsuite}
+
+The SciPy test suite is orchestrated by a continuous integration matrix that
+includes posix and Windows (32/64-bit) platforms managed by Travis CI and
+appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with pyflakes and pycodestyle. There are more than $13000$
+unit tests in the test suite, which is written for usage with the pytest
+runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
+holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
+the historical components of the codebase would still benefit from increased
+test coverage. Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes /
+integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+version of the Python language.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/coverage-chart}
+\caption{\% coverage of the SciPy codebase by unit tests over the six months
+preceding June 28, 2018 as reported by the \texttt{codecov} service. The
+analysis of lines of source code covered is performed automatically in our
+continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
+libraries.}
+\label{fig:coverage}
+\end{figure}
+
 \subsubsection*{Benchmark suite}
@@
 with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
+unit tests are passing (see section \ref{sec:testsuite}), confirming 
 that performance generally
@@
 arguments to select which benchmarks to run.
-
-\subsubsection*{Test suite}
-
-The SciPy test suite is orchestrated by a continuous integration matrix that
-includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
-holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the codebase would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
-version of the Python language.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy codebase by unit tests over the six months
-preceding June 28, 2018 as reported by the \texttt{codecov} service. The
-analysis of lines of source code covered is performed automatically in our
-continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
-libraries.}
-\label{fig:coverage}
-\end{figure}
 

### 2019-03-01 scipy-1.0/paper.tex
relocated command to generate asv benchmark to caption. consider removal.

@@
 with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see section \ref{sec:testsuite}), confirming 
-that performance generally
+unit tests are passing, confirming that performance generally
 remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our codebase is improving.
-
-Consider the \texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
+us to objectively measure that our codebase is improving. Figure~\ref{fig:asvbench}
+shows an example benchmark results for \texttt{scipy.spatial.cKDTree.query} over 
+roughly nine years of project history.
 
@@
 \includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
-over a roughly nine year commit history time frame. The results are based on
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}. 
+The results are based on
 Python 2.7 performance on the master branch of the project using NumPy 1.8.2
@@
 backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the KDTree (boxsize argument was ignored).}
+of toroidal topology to the \texttt{KDTree} (\texttt{boxsize} argument was ignored). Such results
+can be generated using can be run using the command:
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.}
 \label{fig:asvbench}

### 2019-03-01 scipy-1.0/paper.tex
tightened asv section

@@
 
-The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy codebase was monitored
-with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing, confirming that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our codebase is improving. Figure~\ref{fig:asvbench}
-shows an example benchmark results for \texttt{scipy.spatial.cKDTree.query} over 
+In addition to ensuring that unit tests are passing, it is important to ensure that the
+the performance of the SciPy codebase improves over time. Since February 2015, the 
+performance of the SciPy codebase has been monitored with Airspeed Velocity
+(\texttt{asv}\cite{asvref}) library. Figure~\ref{fig:asvbench}
+shows example benchmark results for \texttt{scipy.spatial.cKDTree.query} over 
 roughly nine years of project history.
@@
 backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (\texttt{boxsize} argument was ignored). Such results
-can be generated using can be run using the command:
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.}
+of toroidal topology to the \texttt{KDTree} (i.e. the \texttt{boxsize} argument 
+was ignored).}
 \label{fig:asvbench}
 \end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
 

### 2019-03-01 scipy-1.0/paper.tex
further asv section revisions

@@
 and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
+compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
 backward compatibility and sampling of the benchmarks there was no application
 of toroidal topology to the \texttt{KDTree} (i.e. the \texttt{boxsize} argument 
-was ignored).}
+is $None$).}
 \label{fig:asvbench}

### 2019-03-01 scipy-1.0/paper.tex
stylized pyflakes and pycodestyle
stylized AppVeyor

@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13000$
 unit tests in the test suite, which is written for usage with the pytest

### 2019-03-01 scipy-1.0/paper.tex
further revisions to Test Suite

@@
 AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
-holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the codebase would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which uses the \texttt{pytest} framework, and coverage of SciPy's approximately $204000$ lines of code
+has held steady at 76.5 \% for the last six months (Figure~\ref{fig:coverage}). 
+Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes/integrity.  
+Our full test suite also passes with PyPy3, a just-in-time compiled
 version of the Python language.

### 2019-03-01 scipy-1.0/paper.tex
last tweaks to Test Suite for now

@@
 include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which uses the \texttt{pytest} framework, and coverage of SciPy's approximately $204000$ lines of code
-has held steady at 76.5 \% for the last six months (Figure~\ref{fig:coverage}). 
+unit tests in the test suite, which uses the \texttt{pytest} framework, and coverage of SciPy's approximately $204,000$ lines of code
+has held steady at 76.5\% for the last six months (Figure~\ref{fig:coverage}). 
 Documentation for the code is automatically built and hosted by

### 2019-03-01 scipy-1.0/ckdtree.tex
`cKDTree` -> \texttt{cKDTree}

@@
-The `cKDTree` module, which implements a space-partitioning data structure that
+The \texttt{cKDTree} module, which implements a space-partitioning data structure that
 organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 
@@
 
-In 2015 the cKDTree dual tree counting algorithm\cite{Moore2000ar}
+In 2015 the \texttt{cKDTree} dual tree counting algorithm\cite{Moore2000ar}
 was enhanced to support weights\cite{ckdtree-weights}, which are

### 2019-03-01 scipy-1.0/paper.tex

--- 2019-03-01 scipy-1.0/paper.tex
@@
 
-\subsection*{Current status}
-
-At the time of writing, SciPy produces a new release
-approximately every six months, and some of the major
+\subsection*{SciPy Today}
+At the time of writing, the SciPy library consists of nearly 
+600,000 lines of code organized in 16 subpackages. 
+Over 85,000 GitHub repositories and almost 5000 packages depend
+on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
+visitors every two weeks, and more than 100 people contribute to a 
+new release every six months. Some of the major
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in a technical improvements
-section below. Some longer-term milestone events are
-highlighted in Figure~\ref{fig:timeline}.
+SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
+and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
+% At the time of writing, SciPy produces a new release
+% approximately every six months, and some of the major
+% feature highlights from the three years preceding
+% SciPy 1.0 are discussed in a technical improvements
+% section below. Some longer-term milestone events are
+% highlighted in Figure~\ref{fig:timeline}.
 
 \begin{figure}[H]
-\begin{verbatim}
-    2001: first SciPy release
-    2005: transition to NumPy
-    2007: creation of scikits
-    2008: spatial module and first Cython code added
-    2010: move to a 6-monthly release cycle
-    2011: development moves to GitHub
-    2011: Python 3 support
-    2012: add a sparse graph module and unified optimization interface
-    2012: removal of maxentropy
-    2013: continuous integration with TravisCI
-    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
-    2017: add a unified C API with LowLevelCallable; remove weave
-    2017: SciPy 1.0 release
-\end{verbatim}
-\caption{Timeline}
+\centering
+\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
+\caption{Major milestones from SciPy's initial release in 2001 to
+the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
 \label{fig:timeline}
@@
 Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our 
-codebase using the \texttt{linguist} library\cite{linguistref} provides a 
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library\cite{linguistref} provides a 
 detailed breakdown as \% composition by programming language in 
@@
     ybar, ymax=60, ymin=0,bar width=1cm,
-    ylabel={\% of Scipy codebase},
+    ylabel={\% of Scipy code base},
     ytick={0,20,...,60},
@@
 \subsection*{Data structures}
+
+\subsubsection*{cKDTree}
+
+\input{ckdtree}
 
@@
 These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily throughout SciPy and dependent packages.
-
-Over the last three years, our sparse matrix handling internals have been
+and are used heavily in scikit-learn.
+
+Over the last three years our sparse matrix handling internals have been
 rewritten and performance has been improved. Iterating over and slicing of CSC
 and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the coordinate (COO)/ diagonal (DIA) to CSR / CSC matrix format
+ and the COO / DIA to CSR / CSC matrix format
 conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
@@
 
-\subsubsection*{cKDTree}
-
-\input{ckdtree}
-
 \subsection*{Unified bindings to compiled code}
@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions directly from Python.  
-Supported low-level functions include \texttt{PyCapsule} 
-objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers. 
-Furthermore, it is even possible to generate a low-level callback function 
-automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
-
-\subsubsection*{Cython bindings for BLAS, LAPACK, and special}
+in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
+calling compiled C functions directly from Python.  Supported low-level
+functions include \texttt{PyCapsule} objects, ctypes function pointers, and
+cffi function pointers. The low level function signature must be consistent
+with the expectations of the routine it is passed to. For example, the
+documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
+C callback function signatures that may be used to produce functions that
+operate on each element of image data with low overhead. The C code may be
+generated using numba or Cython, for example, as long as the function call
+signature matches the specifications. Furthermore, it is even possible to
+generate a low-level callback function automatically from a Cython module using
+\texttt{scipy.LowLevelCallable.from\_cython}.
+
+\subsection*{Cython bindings for BLAS, LAPACK, and special}
 
@@
 LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
-includes Cython\cite{behnel2011cython} wrappers for
-many BLAS and LAPACK routines (added in 2015) and the special functions 
+includes Cython\cite{behnel2011cython} wrappers for:
+many BLAS and LAPACK routines (added in 2015), and the special functions 
 provided in the \texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
@@
 Fortran compiler, they can also be used without having to handle all the
-different Fortran compiler ABIs and name mangling schemes.
+different Fortran compiler ABI's and name mangling schemes.
 
@@
 calculate moments, cumulative distribution function (cdf), differential
-entropy, and more. By default, these properties are calculated by generic code, e.g.
-numerical integration of the pdf to obtain the cdf, but the
-generic code can be replaced by explicit formulas or more accurate/efficient
-implementations when available. For example, most distributions have an explicit formula for
+entropy, and more. Those properties are calculated based on generic code, e.g.,
+numerical integration of the pdf to obtain the cdf.  Where possible, the
+generic code is replaced by explicit formulas or more accurate or efficient
+implementations. For example, most distributions have an explicit formula for
 the cdf, so it is used instead of integrating the pdf. The distributions
@@
 \texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
-\texttt{scipy.stats.multinomial}, the first discrete multivariate distribution 
-available in SciPy.
+\texttt{scipy.stats.multinomial} (used, for example, in natural language
+processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
+the only multivariate discrete distribution available in SciPy.
 
@@
 
+\subsubsection*{Benchmark suite}
+
+The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
+
+Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the KDTree (boxsize argument was ignored).}
+\label{fig:asvbench}
+\end{figure}
+
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the master branch and the branch a new feature is implemented
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
+
 \subsubsection*{Test suite}
-\label{sec:testsuite}
 
@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13000$
+appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with pyflakes and pycodestyle. There are more than $13000$
 unit tests in the test suite, which is written for usage with the pytest
@@
 holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the codebase would still benefit from increased
+the historical components of the code base would still benefit from increased
 test coverage. Documentation for the code is automatically built and hosted by
@@
 \includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy codebase by unit tests over the six months
+\caption{\% coverage of the SciPy code base by unit tests over the six months
 preceding June 28, 2018 as reported by the \texttt{codecov} service. The
@@
 \label{fig:coverage}
-\end{figure}
-
-\subsubsection*{Benchmark suite}
-
-In addition to ensuring that unit tests are passing, it is important to ensure that the
-the performance of the SciPy codebase improves over time. Since February 2015, the 
-performance of the SciPy codebase has been monitored with Airspeed Velocity
-(\texttt{asv}\cite{asvref}) library. Figure~\ref{fig:asvbench}
-shows example benchmark results for \texttt{scipy.spatial.cKDTree.query} over 
-roughly nine years of project history.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}. 
-The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (i.e. the \texttt{boxsize} argument 
-is $None$).}
-\label{fig:asvbench}
 \end{figure}
@@
 libraries\cite{astropy-2018}, and what works well for 
-one of the codebases, infrastructures, or communities 
-is often transferred in some form to the other. At the
-codebase level, the \texttt{binned\_statistic} functionality
+one of the code bases, infrastructures, or communities 
+is often transferred in some form to the other. At the code
+base level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution---it was initially
@@
 metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
+14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
 that roughly two-week period. The developer community consists of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the codebase
+contributors of source code, with $>19,000$ commits accepted into the code base
 (GitHub page data).

### 2019-03-01 scipy-1.0/paper.tex

@@
 \end{figure}
+
+\section{Polynomial Interpolators: Additional Details}
+FITPACK provides a collection of routines for calculating
+smoothing splines and automatic knot selection. One issue 
+with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
+and the \code{splrep}/\code{splev} combination, is the monolithic design:
+specific algorithms for construction of interpolating and 
+smoothing splines and knot selection are tightly coupled 
+at the Fortran level. The library offers a very
+limited degree of user control over the details 
+of the spline fitting it performs. 
+%
+Routines for manipulating piecewise polynomials and b-splines were either
+not available or lacking in ease of use and/or performance characteristics.
+
+An alternative set of spline interpolation routines was implemented in the
+(now deprecated) \code{splmake}/\code{spleval} combination, and used as a
+backend of the \code{interp1d} routine. To further complicate matters,
+the spline format used by \code{splmake}/\code{spleval} was incompatible 
+with \code{splrep}/\code{splev}.
+
+Using \code{UnivariateSpline} in new code is not recommended. Given data points,
+the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
+instance, representing a spline which interpolates the data. Interpolated values
+at intermediate points can be computed either via its \code{\_\_call\_\_} method,
+or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
+accepts both \code{tck} tuples and \code{Bspline} instances.
+
+Fitting \emph{smoothing} splines to data can currently only be done with 
+FITPACK. Building alternative algorithms is certainly possible,
+but is still to be implemented.
+
+Interpolation with known or estimated derivatives, i.e., cubic or
+higher-order Hermite splines, can be constructed by either
+subclassing \code{PPoly} or \code{BPoly}, or using the factory
+method \code{BPoly.from\_derivatives}.
+
+The spline representation used by \code{BSpline} is directly
+compatible with FITPACK. We recommend that users work with
+\code{BSpline} objects instead of directly manipulating \code{tck}
+tuples of knots, coefficients and the spline degree.
 

### 2019-03-01 scipy-1.0/subpackages.tex

--- 2019-02-24 scipy-1.0/subpackages.tex
@@
 subpackages include mathematical building blocks (e.g. linear algebra, Fourier
-transforms, special functions), data structures (e.g. sparse matrices, $k$-D trees),
+transforms, special functions), data structures (e.g. sparse matrices, k-D trees),
 algorithms (e.g. numerical optimization and integration, clustering, interpolation,
@@
 
-\begin{description}[leftmargin=!, labelwidth=\widthof{\bfseries \texttt{interpolate}}]
-\itemsep0em
-\item[\texttt{cluster}]
-    The \texttt{cluster} subpackage contains two modules:
-    \texttt{cluster.vq} provides vector quantization and $k$-means algorithms,
+\begin{description}
+
+\item[\texttt{cluster}] ~ \newline
+    The \texttt{cluster} subpackage contains two submodules:
+    \texttt{cluster.vq} provides vector quantization and k-means algorithms,
     and \texttt{cluster.hierarchy} provides functions for hierarchical and
     agglomerative clustering.
-\item[\texttt{constants}]
+\item[\texttt{constants}] ~ \newline
     Physical and mathematical constants, including the CODATA recommended
     values of the fundamental physical constants\cite{CODATA2014}.
-\item[\texttt{fftpack}]
-    Fast Fourier Transform routines.  In addition to the FFT itself, the subpackage
+\item[\texttt{fftpack}] ~ \newline
+    Fast Fourier Transform routines.  In addition to the FFT, the subpackage
     includes functions for the discrete sine and cosine transforms and for
     pseudo-differential operators.
-\item[\texttt{integrate}]
+\item[\texttt{integrate}] ~ \newline
     The \texttt{integrate} subpackage provides tools for the numerical
-    computation of single and multiple definite integrals and for the
+    computation of single and multiple definite integrals, and for the
     solution of ordinary differential equations, including initial value
     problems and two-point boundary value problems.
-\item[\texttt{interpolate}]
+\item[\texttt{interpolate}] ~ \newline
     The \texttt{interpolate} subpackage contains spline functions and
@@
     interpolators, and wrappers for FITPACK and DFITPACK functions.
-\item[\texttt{io}]
+\item[\texttt{io}] ~ \newline
     A collection of functions and classes for reading and writing MATLAB, IDL,
     Matrix Market, Fortran, NetCDF, Harwell-Boeing, WAV and ARFF data files. 
-\item[\texttt{linalg}]
+\item[\texttt{linalg}] ~ \newline
     Linear algebra functions, including:
@@
     Hankel, companion, Hilbert, and more.
-\item[\texttt{ndimage}]
+\item[\texttt{ndimage}] ~ \newline
     This package contains various functions for multi-dimensional image
-    processing, including convolution and assorted linear and nonlinear
+    processing, including: convolution and assorted linear and nonlinear
     filters (Gaussian filter, median filter, Sobel filter, etc.);
@@
     functions.
-\item[\texttt{misc}]
+\item[\texttt{misc}] ~ \newline
     A collection of functions that did not fit into the other subpackages.
@@
     to deprecate the contents of this subpackage and eventually remove it.
-\item[\texttt{odr}]
+\item[\texttt{odr}] ~ \newline
     Orthogonal distance regression, including Python wrappers for the Fortran
     library ODRPACK.
-\item[\texttt{optimize}]
+\item[\texttt{optimize}] ~ \newline
     The package includes the following, with additional details in the SI:
@@
     collection of general nonlinear solvers for root-finding.
-\item[\texttt{signal}]
+\item[\texttt{signal}] ~ \newline
     The \texttt{signal} subpackage focuses on signal processing (plus some
@@
     analysis.  
-\item[\texttt{sparse}]
+\item[\texttt{sparse}] ~ \newline
     This package includes implementations of several representations of
-    sparse matrices.  It contains two modules, 
+    sparse matrices.  It contains two subpackages, 
     \texttt{scipy.sparse.linalg} and \texttt{scipy.sparse.csgraph}.
@@
     and more.
-\item[\texttt{spatial}]
+\item[\texttt{spatial}] ~ \newline
     This subpackage provides spatial data structures and algorithms,
-    including the $k$-d tree, Delaunay triangulation, convex hulls and Voronoi
+    including the k-d tree, Delaunay triangulation, convex hulls and Voronoi
     diagrams.  The subpackage \texttt{scipy.spatial.distance} provides
@@
     of points or between all pairs from two collections of points.
-\item[\texttt{special}]
+\item[\texttt{special}] ~ \newline
     The name comes the class of functions traditionally known as \emph{special
-    functions}, but over time, the subpackage has grown to include functions
+    functions}, but over time, the module has grown to include functions
     beyond the classical special functions.  A more appropriate characterization
     of this subpackage is simply \emph{useful functions}.
-    It includes a large collection of the classical special functions
+    It includes: a large collection of the classical special functions
     such as Airy, Bessel, etc.; orthogonal families of polynomials;
@@
     and more.
-\item[\texttt{stats}]
-    The \texttt{stats} subpackage provides a large collection of continuous
+\item[\texttt{stats}] ~ \newline
+    The \texttt{stats} subpackage provides: a large collection of continuous
     and discrete \emph{probability distributions}, each with methods to compute

### 2019-03-01 scipy-1.0/paper.tex

--- 2019-03-01 scipy-1.0/paper.tex
@@
 
+\subsubsection*{cKDTree}
+
+\input{ckdtree}
+
 \subsection*{Unified bindings to compiled code}
@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
-calling compiled C functions directly from Python.  Supported low-level
-functions include \texttt{PyCapsule} objects, ctypes function pointers, and
-cffi function pointers. The low level function signature must be consistent
-with the expectations of the routine it is passed to. For example, the
-documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
-C callback function signatures that may be used to produce functions that
-operate on each element of image data with low overhead. The C code may be
-generated using numba or Cython, for example, as long as the function call
-signature matches the specifications. Furthermore, it is even possible to
-generate a low-level callback function automatically from a Cython module using
-\texttt{scipy.LowLevelCallable.from\_cython}.
-
-\subsection*{Cython bindings for BLAS, LAPACK, and special}
+in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
+calling compiled C functions directly from Python.  
+Supported low-level functions include \texttt{PyCapsule} 
+objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers. 
+Furthermore, it is even possible to generate a low-level callback function 
+automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
+
+\subsubsection*{Cython bindings for BLAS, LAPACK, and special}
 
@@
 LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
-includes Cython\cite{behnel2011cython} wrappers for:
-many BLAS and LAPACK routines (added in 2015), and the special functions 
+includes Cython\cite{behnel2011cython} wrappers for
+many BLAS and LAPACK routines (added in 2015) and the special functions 
 provided in the \texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
@@
 
-\subsubsection*{Benchmark suite}
-
-The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
-with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
-that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
-
-Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the KDTree (boxsize argument was ignored).}
-\label{fig:asvbench}
-\end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
-
 \subsubsection*{Test suite}
+\label{sec:testsuite}
 
@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
-holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
-the historical components of the code base would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which uses the \texttt{pytest} framework, and coverage of SciPy's approximately $204,000$ lines of code
+has held steady at 76.5\% for the last six months (Figure~\ref{fig:coverage}). 
+Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes/integrity.  
+Our full test suite also passes with PyPy3, a just-in-time compiled
 version of the Python language.
@@
 \label{fig:coverage}
+\end{figure}
+
+\subsubsection*{Benchmark suite}
+
+In addition to ensuring that unit tests are passing, it is important to ensure that the
+the performance of the SciPy codebase improves over time. Since February 2015, the 
+performance of the SciPy codebase has been monitored with Airspeed Velocity
+(\texttt{asv}\cite{asvref}) library. Figure~\ref{fig:asvbench}
+shows example benchmark results for \texttt{scipy.spatial.cKDTree.query} over 
+roughly nine years of project history.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}. 
+The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the \texttt{KDTree} (i.e. the \texttt{boxsize} argument 
+is $None$).}
+\label{fig:asvbench}
 \end{figure}

### 2019-03-01 scipy-1.0/paper.tex
Update paper.tex

--- 2019-03-01 scipy-1.0/paper.tex
@@
 \usepackage{float}
-\usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \usepackage[multiple]{footmisc}
@@
 \centering
-\includegraphics[width=\textwidth]{static/cov}
-\caption{\% coverage of the SciPy code base by unit tests over a five year
-time span predeceding the 1.0 release point. The reconstitution of historical
-build and test environments was performed using a docker-based approach
+\begin{tikzpicture}[]
+\pgfplotstableread[column/ver/.style=string type]{
+ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
+v0.12.1 33844 25721 8123 273634  142290 131344
+v0.13.3 36638 28944 7694 338926  172852 166074
+v0.14.1 37301 30587 6714 277643  147151 130492
+v0.15.1 38339 30288 8051 304898  158547 146351
+v0.16.1 40094 32075 8019 322398  167647 154751
+v0.17.1 42566 34478 8088 340903  170452 170451
+v0.18.1 44711 36216 8495 353417  173174 180243
+v0.19.1 43823 36373 7450 436900  200974 235926
+v1.0.0 106878 92984 13894 462574 208158 254416
+}\covtable
+    \begin{axis}[
+    width=\textwidth,
+    height=7cm,
+    enlargelimits=false,
+    ymin=0,
+    ymax=6e5,
+    stack plots=y,%   
+    area style,
+    xtick=data,
+    ytick={2e5,4e5,6e5},
+    xticklabels from table={\covtable}{ver},
+    scaled y ticks=false,
+    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
+    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
+    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
+    legend columns=-1,
+    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
+    ]
+  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
+  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
+  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
+  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
+\end{axis}
+\end{tikzpicture}
+\caption{Lines of Python (red) and compiled (blue) code in SciPy over time.
+Deep-shaded area represents lines of code covered by unit tests; light-shaded area represents lines not covered.
+With the exception of the removal of $\approx$ 61,000 lines of compiled code for SciPy v0.14, the volume of both compiled and Python code has increased between releases, as has the number of lines covered by unit tests.
+SciPy v1.0.0 was composed of 462,574 lines of compiled code with test coverage near 45\% and 106,878 lines of Python code with test coverage near 87\%.
+
+The reconstitution of historical build and test environments was performed using a docker-based approach
 \cite{scipy-cov}. Analysis of lines of source code covered is performed 
@@
 and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.}
+inclusion of additional vendored code.
+}
 \label{fig:coverage}

### 2019-03-01 scipy-1.0/paper.tex
Incorporating new plot

@@
 \end{tikzpicture}
-\caption{Lines of Python (red) and compiled (blue) code in SciPy over time.
-Deep-shaded area represents lines of code covered by unit tests; light-shaded area represents lines not covered.
-With the exception of the removal of $\approx$ 61,000 lines of compiled code for SciPy v0.14, the volume of both compiled and Python code has increased between releases, as has the number of lines covered by unit tests.
-SciPy v1.0.0 was composed of 462,574 lines of compiled code with test coverage near 45\% and 106,878 lines of Python code with test coverage near 87\%.
-
-The reconstitution of historical build and test environments was performed using a docker-based approach
+\caption{\% coverage of the SciPy code base by unit tests over a five year
+time span predeceding the 1.0 release point. The reconstitution of historical
+build and test environments was performed using a docker-based approach
 \cite{scipy-cov}. Analysis of lines of source code covered is performed 
@@
 and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.
-}
+inclusion of additional vendored code.}
 \label{fig:coverage}

### 2019-03-01 scipy-1.0/paper.tex
Update caption

@@
 \end{tikzpicture}
-\caption{\% coverage of the SciPy code base by unit tests over a five year
-time span predeceding the 1.0 release point. The reconstitution of historical
+\caption{
+Python (red) and compiled (blue) code volume in SciPy over time. 
+Deep-shaded area represents lines of code covered by units tests; 
+light-shaded area represents lines not covered. With the exception
+of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
+v0.14 due to (insert reason), the volume of both compiled and Python
+code has increased between releases, as has the number of lines
+covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
+compiled code with test coverage near 45\% and 106,878 lines of Python 
+code with test coverage near 87\%.
+
+The reconstitution of historical
 build and test environments was performed using a docker-based approach

### 2019-03-01 scipy-1.0/paper.tex

--- 2019-03-01 scipy-1.0/paper.tex
@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the code base would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which uses the \texttt{pytest} framework, and coverage of SciPy's approximately $204,000$ lines of code
+has held steady at 76.5\% for the last six months (Figure~\ref{fig:coverage}). 
+Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes/integrity.  
+Our full test suite also passes with PyPy3, a just-in-time compiled
 version of the Python language.
@@
 \end{tikzpicture}
-\caption{\% coverage of the SciPy code base by unit tests over a five year
-time span predeceding the 1.0 release point. The reconstitution of historical
+\caption{
+% This paragraph of caption won't compile
+%Python (red) and compiled (blue) code volume in SciPy over time. 
+%Deep-shaded area represents lines of code covered by units tests; 
+%light-shaded area represents lines not covered. With the exception
+%of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
+%v0.14 due to (insert reason), the volume of both compiled and Python
+%code has increased between releases, as has the number of lines
+%covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
+%compiled code with test coverage near 45\% and 106,878 lines of Python 
+%code with test coverage near 87\%.
+The reconstitution of historical
 build and test environments was performed using a docker-based approach

### 2019-03-01 scipy-1.0/ckdtree.tex

@@
-The \texttt{cKDTree} module, which implements a space-partitioning data structure that
-organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 
-Support was added for periodic boundary conditions, which are often used 
-in simulations of physical processes.
+The \texttt{cKDTree} module was rewritten in C++ with templated classes, and support for
+periodic boundary conditions was added. A periodic boundary condition is typically 
+used in computer simulations of physical processes.
 
-In 2013, the time complexity of the k nearest neighbor search from
+In 2013, the time complexity of the $k$-nearest neighbor search from
 \texttt{cKDTree.query} was approximately loglinear \cite{knn-jake},

### 2019-03-01 scipy-1.0/subpackages.tex

--- 2019-02-24 scipy-1.0/subpackages.tex
@@
 \item[\texttt{linalg}]
-    Linear algebra functions, including:
+    Linear algebra functions, including
     elementary functions of a matrix, such as the trace, determinant, norm and
@@
 \item[\texttt{special}]
-    The name comes the class of functions traditionally known as \emph{special
-    functions}, but over time, the subpackage has grown to include functions
+    The name comes the class of functions traditionally known as ``special
+    functions'', but over time, the subpackage has grown to include functions
     beyond the classical special functions.  A more appropriate characterization
-    of this subpackage is simply \emph{useful functions}.
+    of this subpackage is simply ``useful functions''.
     It includes a large collection of the classical special functions
@@
     The \texttt{stats} subpackage provides a large collection of continuous
-    and discrete \emph{probability distributions}, each with methods to compute
+    and discrete probability distributions, each with methods to compute
     the PDF or PMF, CDF, moments and other statistics, generation of random
     variates, and more;
-    \emph{statistical tests}, including Pearson's correlation, Spearman's rank-order
+    statistical tests, including Pearson's correlation, Spearman's rank-order
     correlations, Kendall's tau, chi-squared test and its generalization as the
@@
     exact test and Mood's median test, and many more;
-    and assorted \emph{transformations and statistics} of data.
+    and assorted transformations and statistics of data.
 \end{description}

### 2019-03-01 scipy-1.0/paper.tex

@@
 SciPy is a library of numerical routines for the Python programming
-language, providing the fundamental building blocks for modeling and
-solving scientific problems.  For example, when LIGO, the Laser
-Interferometer Gravitational-Wave Observatory, recently announced the
-discovery of gravitational waves, their published Python analysis
-scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
-PhysRevD.93.122003} imported many different SciPy submodules for:
-signal processing, interpolation, input / output, optimization, statistics,
-integration, and more. SciPy has routines for solving
-classes of generic problems such as optimization, integration,
-interpolation, eigenvalues, and linear systems.  It also provides
-specialized data structures, such as sparse arrays and $k$-dimensional
-trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
-which provides array data structures and related fast numerical routines.
+language that provides fundamental building blocks for modeling and
+solving scientific problems. 
+SciPy includes algorithms for optimization, integration, interpolation,
+eigenvalue problems, algebraic equations, differential equations, and many other
+classes of problems; it also provides
+specialized data structures, such as sparse arrays
+and $k$-dimensional trees. SciPy is built on top of 
+NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
+which provides array data structures and related fast numerical routines, and
 SciPy is itself the foundation upon which higher level scientific libraries,
@@
 and scikit-image\cite{vanderwalt2014scikit}, are built. 
+SciPy is relied upon by scientists, engineers, and others 
+around the world. For example, published 
+scripts\cite{alex_nitz_2018_1596771, LIGO-open}
+used in the analysis of gravitational waves
+\cite{PhysRevD.93.122003, abbott2017gw170817} 
+import several submodules of SciPy.
 
@@
 %
-As a general purpose programming language with a ``batteries included''
-standard library, it had no special support for scientific data
-structures or algorithms, unlike many of the other established
+As a general purpose programming language, it had no special support for 
+scientific data structures or algorithms, unlike many of the other established
 computation platforms of the time. Yet, scientists soon discovered the
@@
 could thereby gain access to a wide variety of existing computational
-libraries, without concerning themselves with low-level programming
+libraries without concerning themselves with low-level programming
 concepts such as memory management.
 
-In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
-message on the Python Matrix Special Interest Group (Matrix-SIG)
+In 1995, Jim Hugunin, a graduate student from MIT, wrote the first
+message in a new Python Matrix Special Interest Group (Matrix-SIG)
 mailing list\cite{Hugunin-first}:
@@
 adding general features to a language designed for numerical and scientific
-computing, here scientific features are added to a general purpose language
-instead.  This broadens the scope of problems that can be addressed easily, the
+computing, here scientific features are added to a general purpose language.  
+This broadens the scope of problems that can be addressed easily, expands the
 sources of data that are readily accessible, and increases the size of the
-community that develops general purpose tooling.
+community that develops code for the platform.
 
 The availability of a standard numerical array data structure in
-Python led to a sudden and rapidly growing number of scientific
-packages becoming available for solving various common numeric
-problems.
+Python led to a sudden, rapid growth in the number of scientific
+packages available for solving common numeric problems.
 A number of these packages were written by graduate students and

### 2019-03-03 scipy-1.0/paper.tex

@@
 
-\subsubsection*{\texttt{cKDTree}}
-
-\input{ckdtree}
-
 \subsubsection*{Sparse matrices}
@@
 \subsubsection*{Test suite}
+\label{sec:testsuite}
 
@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the code base would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which uses the \texttt{pytest} framework, 
+and with 87 \% and 45 \% line coverages for Python and compiled
+code, respectively, at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
+Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes/integrity.  
+Our full test suite also passes with PyPy3, a just-in-time compiled
 version of the Python language.
@@
     \texttt{linalg} & support a more recent version of LAPACK\\
-    \texttt{ndimage} & migrate to data point model from pixel model\\
+    \texttt{ndimage} & clarify usage of the ``data point'' coordinate
+    model, and add additional wrapping modes\\
     \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\

### 2019-03-03 scipy-1.0/paper.tex
Update paper.tex

@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which uses the \texttt{pytest} framework, 
-and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively, at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Documentation for the code is automatically built and hosted by
-the CircleCI service to facilitate evaluation of documentation changes/integrity.  
-Our full test suite also passes with PyPy3, a just-in-time compiled
+appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with pyflakes and pycodestyle. There are more than $13000$
+unit tests in the test suite, which is written for usage with the pytest
+runner, and with 87 \% and 45 \% line coverages for Python and compiled
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
+Some of the historical components of the code base would still benefit from increased
+test coverage. Documentation for the code is automatically built and hosted by
+the CircleCI service to facilitate evaluation of documentation changes /
+integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
 version of the Python language.

### 2019-03-03 scipy-1.0/paper.tex
reverted to old figure

--- 2019-03-01 scipy-1.0/paper.tex
@@
 \usepackage{float}
-\usepackage{pgfplotstable}
+\usepackage{pgfplots}
 \usepackage[multiple]{footmisc}
@@
 SciPy is a library of numerical routines for the Python programming
-language that provides fundamental building blocks for modeling and
-solving scientific problems. 
-SciPy includes algorithms for optimization, integration, interpolation,
-eigenvalue problems, algebraic equations, differential equations, and many other
-classes of problems; it also provides
-specialized data structures, such as sparse arrays
-and $k$-dimensional trees. SciPy is built on top of 
-NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
-which provides array data structures and related fast numerical routines, and
+language, providing the fundamental building blocks for modeling and
+solving scientific problems.  For example, when LIGO, the Laser
+Interferometer Gravitational-Wave Observatory, recently announced the
+discovery of gravitational waves, their published Python analysis
+scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
+PhysRevD.93.122003} imported many different SciPy submodules for:
+signal processing, interpolation, input / output, optimization, statistics,
+integration, and more. SciPy has routines for solving
+classes of generic problems such as optimization, integration,
+interpolation, eigenvalues, and linear systems.  It also provides
+specialized data structures, such as sparse arrays and k-dimensional
+trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
+which provides array data structures and related fast numerical routines.
 SciPy is itself the foundation upon which higher level scientific libraries,
@@
 %
-As a general purpose programming language, it had no special support for 
-scientific data structures or algorithms, unlike many of the other established
+As a general purpose programming language with a ``batteries included''
+standard library, it had no special support for scientific data
+structures or algorithms, unlike many of the other established
 computation platforms of the time. Yet, scientists soon discovered the
@@
 list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
-Hinsen, and Guido van Rossum, led to the creation of a package called Numeric with an array object
+Hinsen, and Guido van Rossum, led to the creation of an array object
 that supported a high number of dimensions.  Jim Hugunin explained the utility
@@
 In 2001, Fernando Pérez, a graduate student at the University of
-Colorado, Boulder, created the IPython interactive shell, which 
-eventually blossomed into Project Jupyter\cite{Kluyver:2016aa}.
+Colorado, Boulder, created the IPython interactive shell (that would
+eventually blossom into Project Jupyter\cite{Kluyver:2016aa}), built into
+IPP, an earlier attempt by Janko Hauser, and including code from
+Nathan Gray's LazyPython.
 % IPython hosted on SciPy:
@@
 STScI.  With the Numeric maintainer's blessing, the decision was made
-to write NumArray, a library that could handle data on a larger
+to write NumArray, a library that could handle data on a different
 scale.  Unfortunately, NumArray proved inefficient for small arrays,
@@
 the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
-He also shared his own interest in Python for ``computational steering''
+He also shares his own interest in Python for ``computational steering''
 where ``Python serves as the input language to a scientific application, and the actual
@@
 NumPy\cite{vanderwalt2011numpy}, Cython\cite{behnel2011cython},
-and Mayavi\cite{ramachandran2011mayavi}. Python became so pervasive that 
-journals began publishing domain-specific special issues.
-For example, in 2015 Frontiers in Neuroinformatics published a collection of 25 articles---covering
+and Mayavi\cite{ramachandran2011mayavi}.
+
+Python has become so pervasive, there are now domain-specific special issues.
+For example, Frontiers in Neuroinformatics has a collection of 25 articles---covering
 topics including modeling and simulation, data collection, electrophysiology, visualization,
 as well as stimulus generation and presentation---called Python in
-Neuroscience\cite{python-FIN}.
+neuroscience\cite{python-FIN}.
 
@@
 and Fernando Pérez founded NumFOCUS\cite{numfocus},
- a 501(c)3 public charity with a mission
-``to promote sustainable high-level programming languages, open code development,
+ a 501(c)3 public charity whose mission is to
+``promote sustainable high-level programming languages, open code development,
 and reproducible scientific research.''
@@
 came from the scientific Python stack.
-Today it has 50 affiliated and sponsored projects including NumPy, SciPy, IPython, and
+Today it has over 25 sponsored projects including NumPy, SciPy, IPython, and
 Matplotlib.
@@
 
-\subsection*{SciPy Today}
-At the time of writing, the SciPy library consists of nearly 
-600,000 lines of code organized in 16 subpackages. 
-Over 85,000 GitHub repositories and almost 5000 packages depend
-on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
-visitors every two weeks, and more than 100 people contribute to a 
-new release every six months. Some of the major
+\subsection*{Current status}
+
+At the time of writing, SciPy produces a new release
+approximately every six months, and some of the major
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
-and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
-% At the time of writing, SciPy produces a new release
-% approximately every six months, and some of the major
-% feature highlights from the three years preceding
-% SciPy 1.0 are discussed in a technical improvements
-% section below. Some longer-term milestone events are
-% highlighted in Figure~\ref{fig:timeline}.
+SciPy 1.0 are discussed in a technical improvements
+section below. Some longer-term milestone events are
+highlighted in Figure~\ref{fig:timeline}.
 
 \begin{figure}[H]
-\centering
-\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
-\caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
+\begin{verbatim}
+    2001: first SciPy release
+    2005: transition to NumPy
+    2007: creation of scikits
+    2008: spatial module and first Cython code added
+    2010: move to a 6-monthly release cycle
+    2011: development moves to GitHub
+    2011: Python 3 support
+    2012: add a sparse graph module and unified optimization interface
+    2012: removal of maxentropy
+    2013: continuous integration with TravisCI
+    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
+    2017: add a unified C API with LowLevelCallable; remove weave
+    2017: SciPy 1.0 release
+\end{verbatim}
+\caption{Timeline}
 \label{fig:timeline}
@@
 \section*{Architecture and implementation choices}
+
 \subsection*{Project scope}
@@
 
-\subsection*{Package organization}
+\subsection*{Submodule organization}
+
 \input{subpackages}
@@
 
-Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library\cite{linguistref} provides a 
-detailed breakdown as \% composition by programming language in 
-SciPy (Figure~\ref{fig:linguist}).
+Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
+in SciPy. While it is not surprising that Python is predominant, the usage distribution 
+of the other languages warrants discussion.
+
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=0.5\textwidth]{static/scipy_languages}
+%\caption{Composition of SciPy: lines of code in each programming language as determined by the \texttt{linguist} package.
+%Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
+%and Makefile (not visible) provide supporting roles in tests,
+%building, and documentation.}
+%\label{fig:linguist}
+%\end{figure}
 
@@
     \caption{The breakdown of programming languages used in the
-             SciPy library determined using the linguist library.
+             SciPy library determined using the \texttt{linguist} library.
     	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
@@
 
-For implementing new functionality, we have a clear order of language
-preference.  First Python, if performance is not an issue. If it is, then in
-order of preference: Cython, C, C++, Fortran. The main motivation for this is
-maintainability: Cython has the highest abstraction level and most Python
-developers will understand it. C is also widely known, and easier to deal with
-for the core development team than C++ and especially Fortran.
-
-While it is not surprising that Python is heavily used in SciPy, the usage
-distribution of the other languages warrants some discussion. Fortran is an
+Fortran is an
 extremely well-established scientific programming language, both for historical
@@
 
-Therefore, the relative abundance of different programming languages in the
-SciPy library results from a combination of the usage of powerful performance
-enhancing languages that interact well with Python (i.e., Cython), and the
+Cython has been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
+as a glue between well-established low-level scientific computing libraries written
+in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
+enhancements in Python code, especially for cases where heavily used inner
+loops benefit from a compiled code with static typing.
+
+For implementing new functionality, Python is the still the language of choice. If Python 
+performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
+order). The main motivation for this is maintainability: Cython has the 
+highest abstraction level and most Python developers will understand it. C is 
+also widely known, and easier for the current core development team to manage
+than C++ and especially Fortran.
+
+The distribution of secondary programming languages in SciPy is a compromise between
+a powerful, performance-
+enhancing language that interacts well with Python (i.e. Cython) and the
 usage of languages (and their libraries) that have proven reliable and
@@
 
-\subsubsection*{\texttt{cKDTree}}
+\subsubsection*{cKDTree}
 
@@
 \includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query()}
+\caption{Airspeed velocity benchmarks for scipy.spatial.cKDTree.query()
 over a roughly nine year commit history time frame. The results are based on
@@
 backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+of toroidal topology to the KDTree (boxsize argument was ignored).}
 \label{fig:asvbench}
@@
 unit tests in the test suite, which is written for usage with the pytest
-runner, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the code base would still benefit from increased
+runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
+holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
+the historical components of the code base would still benefit from increased
 test coverage. Documentation for the code is automatically built and hosted by
@@
 \centering
-\begin{tikzpicture}[]
-\pgfplotstableread[column/ver/.style=string type]{
-ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
-v0.12.1 33844 25721 8123 273634  142290 131344
-v0.13.3 36638 28944 7694 338926  172852 166074
-v0.14.1 37301 30587 6714 277643  147151 130492
-v0.15.1 38339 30288 8051 304898  158547 146351
-v0.16.1 40094 32075 8019 322398  167647 154751
-v0.17.1 42566 34478 8088 340903  170452 170451
-v0.18.1 44711 36216 8495 353417  173174 180243
-v0.19.1 43823 36373 7450 436900  200974 235926
-v1.0.0 106878 92984 13894 462574 208158 254416
-}\covtable
-    \begin{axis}[
-    width=\textwidth,
-    height=7cm,
-    enlargelimits=false,
-    ymin=0,
-    ymax=6e5,
-    stack plots=y,%   
-    area style,
-    xtick=data,
-    ytick={2e5,4e5,6e5},
-    xticklabels from table={\covtable}{ver},
-    scaled y ticks=false,
-    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
-    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
-    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
-    legend columns=-1,
-    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
-    ]
-  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
-  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
-  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
-  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
-\end{axis}
-\end{tikzpicture}
-\caption{
-% This paragraph of caption won't compile
-%Python (red) and compiled (blue) code volume in SciPy over time. 
-%Deep-shaded area represents lines of code covered by units tests; 
-%light-shaded area represents lines not covered. With the exception
-%of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
-%v0.14 due to (insert reason), the volume of both compiled and Python
-%code has increased between releases, as has the number of lines
-%covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
-%compiled code with test coverage near 45\% and 106,878 lines of Python 
-%code with test coverage near 87\%.
-The reconstitution of historical
-build and test environments was performed using a docker-based approach
-\cite{scipy-cov}. Analysis of lines of source code covered is performed 
-automatically in our continuous integration suite using the \texttt{pytest-cov}
-and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.}
+\includegraphics[width=\textwidth]{static/coverage-chart}
+\caption{\% coverage of the SciPy code base by unit tests over the six months
+preceding June 28, 2018 as reported by the \texttt{codecov} service. The
+analysis of lines of source code covered is performed automatically in our
+continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
+libraries.}
 \label{fig:coverage}
@@
 \includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+\caption{Assessment of cKDTree nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
 The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
@@
 \end{figure}
-
-\section{Polynomial Interpolators: Additional Details}
-FITPACK provides a collection of routines for calculating
-smoothing splines and automatic knot selection. One issue 
-with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
-and the \code{splrep}/\code{splev} combination, is the monolithic design:
-specific algorithms for construction of interpolating and 
-smoothing splines and knot selection are tightly coupled 
-at the Fortran level. The library offers a very
-limited degree of user control over the details 
-of the spline fitting it performs. 
-%
-Routines for manipulating piecewise polynomials and b-splines were either
-not available or lacking in ease of use and/or performance characteristics.
-
-An alternative set of spline interpolation routines was implemented in the
-(now deprecated) \code{splmake}/\code{spleval} combination, and used as a
-backend of the \code{interp1d} routine. To further complicate matters,
-the spline format used by \code{splmake}/\code{spleval} was incompatible 
-with \code{splrep}/\code{splev}.
-
-Using \code{UnivariateSpline} in new code is not recommended. Given data points,
-the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
-instance, representing a spline which interpolates the data. Interpolated values
-at intermediate points can be computed either via its \code{\_\_call\_\_} method,
-or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
-accepts both \code{tck} tuples and \code{Bspline} instances.
-
-Fitting \emph{smoothing} splines to data can currently only be done with 
-FITPACK. Building alternative algorithms is certainly possible,
-but is still to be implemented.
-
-Interpolation with known or estimated derivatives, i.e., cubic or
-higher-order Hermite splines, can be constructed by either
-subclassing \code{PPoly} or \code{BPoly}, or using the factory
-method \code{BPoly.from\_derivatives}.
-
-The spline representation used by \code{BSpline} is directly
-compatible with FITPACK. We recommend that users work with
-\code{BSpline} objects instead of directly manipulating \code{tck}
-tuples of knots, coefficients and the spline degree.
 
@@
     point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
     \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}

### 2019-03-03 scipy-1.0/paper.tex
alternative language breakdown figure

--- 2019-03-03 scipy-1.0/paper.tex
@@
     \centering
-
-    \begin{tikzpicture}[font=\sffamily]
-    \begin{axis}[height=6cm, width=10cm,
-    ybar, ymax=60, ymin=0,bar width=1cm,
-    ylabel={\% of Scipy codebase},
-    ytick={0,20,...,60},
-    axis lines*=left,
-    axis y line shift=0pt,
-    enlarge x limits=0.15,
-    symbolic x coords={py,f,c,cy,cpp},
-    xticklabels={Python,FORTRAN,C,Cython,C++},
-    ticklabel style={align=center,font=\sffamily\small,
-    /pgf/number format/assume math mode=true},
-    xtick=data,
-    nodes near coords,
-    nodes near coords style={/pgf/number format/assume math mode=true}
-    ]
-    \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
-    \end{axis}
-    \end{tikzpicture}
+		\includegraphics[width=\textwidth]{static/composition}
     \caption{The breakdown of programming languages used in the
@@
 
+\subsubsection*{\texttt{cKDTree}}
+
+\input{ckdtree}
+
 \subsubsection*{Sparse matrices}
@@
 \subsubsection*{Test suite}
-\label{sec:testsuite}
 

### 2019-03-05 scipy-1.0/paper.tex
test suite stylization, figure caption

@@
 includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which is written for usage with the \texttt{pytest}
 runner, and with 87 \% and 45 \% line coverages for Python and compiled
@@
 \caption{
-% This paragraph of caption won't compile
-%Python (red) and compiled (blue) code volume in SciPy over time. 
-%Deep-shaded area represents lines of code covered by units tests; 
-%light-shaded area represents lines not covered. With the exception
-%of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
-%v0.14 due to (insert reason), the volume of both compiled and Python
-%code has increased between releases, as has the number of lines
-%covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
-%compiled code with test coverage near 45\% and 106,878 lines of Python 
-%code with test coverage near 87\%.
+Python (red) and compiled (blue) code volume in SciPy over time.
+Deep-shaded area represents lines of code covered by units tests;
+light-shaded area represents lines not covered. With the exception
+of the removal of $\approx 61,000$ lines of compiled code for SciPy
+v0.14, the volume of both compiled and Python
+code has increased between releases, as has the number of lines
+covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
+compiled code with test coverage near $45 \%$ and $106,878$ lines of 
+Python code with test coverage near $87 \%$.\\
 The reconstitution of historical

### 2019-03-05 scipy-1.0/paper.tex
POSIX acronym capitalization

@@
 The SciPy test suite is orchestrated by a continuous integration matrix that
-includes posix and Windows (32/64-bit) platforms managed by Travis CI and
+includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
 AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and

### 2019-03-05 scipy-1.0/paper.tex
code base -> codebase
moved benchmark suite after test suite
stylization fixes

--- 2019-03-03 scipy-1.0/paper.tex
@@
 \subsection*{Test and benchmark suite}
-
-\subsubsection*{Benchmark suite}
-
-The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
-with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
-that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
-
-Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query()}
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
-\label{fig:asvbench}
-\end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the master branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
 
@@
 
+\subsubsection*{Benchmark suite}
+
+The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
+
+Consider the (\texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+\label{fig:asvbench}
+\end{figure}
+
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the \texttt{master} branch and the branch a new feature is implemented
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
+
 \section*{Project organization and community}

### 2019-03-05 scipy-1.0/paper.tex
removed '(see Test suite section below)'

@@
 with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
+unit tests are passing, confirming 
 that performance generally

### 2019-03-05 scipy-1.0/paper.tex
capitalized Airspeed Velocity

@@
 
-The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
+The Airspeed Velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
 over their lifetimes, and the performance of the SciPy code base was monitored

### 2019-03-05 scipy-1.0/paper.tex
pytest runner -> pytest framework

--- 2019-03-05 scipy-1.0/paper.tex
@@
 
+\subsubsection*{Benchmark suite}
+
+The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
+
+Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query()}
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+\label{fig:asvbench}
+\end{figure}
+
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the master branch and the branch a new feature is implemented
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
+
 \subsubsection*{Test suite}
@@
 The SciPy test suite is orchestrated by a continuous integration matrix that
-includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with 87 \% and 45 \% line coverages for Python and compiled
+includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which is written for usage with the \texttt{pytest}
+framework, and with 87 \% and 45 \% line coverages for Python and compiled
 code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the code base would still benefit from increased
+Some of the historical components of the codebase would still benefit from increased
 test coverage. Documentation for the code is automatically built and hosted by
@@
 % NOTE: is there a citation for PyPy?
-
-\subsubsection*{Benchmark suite}
-
-The airspeed velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
-with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing (see Test suite section below), confirming 
-that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
-
-Consider the (\texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
-\label{fig:asvbench}
-\end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the \texttt{master} branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
 

### 2019-03-06 scipy-1.0/paper.tex
save 'Some of the historical components...' for discussion
comma between dependent and independent clauses

@@
 framework, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the codebase would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 any strong indications that they should step down; although the BDFL may
-appoint a successor it is generally expected that the Steering Council will be
+appoint a successor, it is generally expected that the Steering Council will be
 consulted on the matter.

### 2019-03-06 scipy-1.0/paper.tex
removing comma; clause on right is not independent

@@
 project decided upon by the Steering Committee. The Chair does not have a fixed
-term, but is nominated by the Steering Committee and is expected to yield to
+term but is nominated by the Steering Committee and is expected to yield to
 substantive claims that stepping down is the appropriate action.

### 2019-03-06 scipy-1.0/paper.tex
removing comma; not a list and no independent clause on right

@@
 performs a number of tasks including nominating and voting on new Council
-Members, and actively contributing to the progress of the project (could be
+Members and actively contributing to the progress of the project (could be
 code review but also outreach activities, etc.). Generally, one year of sustained

### 2019-03-06 scipy-1.0/paper.tex
removed commented out figure

--- 2019-03-06 scipy-1.0/paper.tex
@@
     \centering
-		\includegraphics[width=\textwidth]{static/composition}
+
+    \begin{tikzpicture}[font=\sffamily]
+    \begin{axis}[height=6cm, width=10cm,
+    ybar, ymax=60, ymin=0,bar width=1cm,
+    ylabel={\% of Scipy code base},
+    ytick={0,20,...,60},
+    axis lines*=left,
+    axis y line shift=0pt,
+    enlarge x limits=0.15,
+    symbolic x coords={py,f,c,cy,cpp},
+    xticklabels={Python,FORTRAN,C,Cython,C++},
+    ticklabel style={align=center,font=\sffamily\small,
+    /pgf/number format/assume math mode=true},
+    xtick=data,
+    nodes near coords,
+    nodes near coords style={/pgf/number format/assume math mode=true}
+    ]
+    \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
+    \end{axis}
+    \end{tikzpicture}
     \caption{The breakdown of programming languages used in the
-             SciPy library determined using the linguist library.
+             SciPy library determined using the \texttt{linguist} library.
     	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
@@
 The SciPy test suite is orchestrated by a continuous integration matrix that
-includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
-AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which is written for usage with the \texttt{pytest}
-framework, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
+includes posix and Windows (32/64-bit) platforms managed by Travis CI and
+appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with pyflakes and pycodestyle. There are more than $13000$
+unit tests in the test suite, which is written for usage with the pytest
+runner, and with a 76.5 \% coverage of approximately $204000$ lines of code
+holding steady for the last six months (Figure~\ref{fig:coverage}).  Some of
+the historical components of the code base would still benefit from increased
+test coverage. Documentation for the code is automatically built and hosted by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 \centering
-\begin{tikzpicture}[]
-\pgfplotstableread[column/ver/.style=string type]{
-ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
-v0.12.1 33844 25721 8123 273634  142290 131344
-v0.13.3 36638 28944 7694 338926  172852 166074
-v0.14.1 37301 30587 6714 277643  147151 130492
-v0.15.1 38339 30288 8051 304898  158547 146351
-v0.16.1 40094 32075 8019 322398  167647 154751
-v0.17.1 42566 34478 8088 340903  170452 170451
-v0.18.1 44711 36216 8495 353417  173174 180243
-v0.19.1 43823 36373 7450 436900  200974 235926
-v1.0.0 106878 92984 13894 462574 208158 254416
-}\covtable
-    \begin{axis}[
-    width=\textwidth,
-    height=7cm,
-    enlargelimits=false,
-    ymin=0,
-    ymax=6e5,
-    stack plots=y,%   
-    area style,
-    xtick=data,
-    ytick={2e5,4e5,6e5},
-    xticklabels from table={\covtable}{ver},
-    scaled y ticks=false,
-    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
-    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
-    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
-    legend columns=-1,
-    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
-    ]
-  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
-  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
-  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
-  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
-\end{axis}
-\end{tikzpicture}
-\caption{
-Python (red) and compiled (blue) code volume in SciPy over time.
-Deep-shaded area represents lines of code covered by units tests;
-light-shaded area represents lines not covered. With the exception
-of the removal of $\approx 61,000$ lines of compiled code for SciPy
-v0.14, the volume of both compiled and Python
-code has increased between releases, as has the number of lines
-covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
-compiled code with test coverage near $45 \%$ and $106,878$ lines of 
-Python code with test coverage near $87 \%$.\\
-The reconstitution of historical
-build and test environments was performed using a docker-based approach
-\cite{scipy-cov}. Analysis of lines of source code covered is performed 
-automatically in our continuous integration suite using the \texttt{pytest-cov}
-and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.}
+\includegraphics[width=\textwidth]{static/coverage-chart}
+\caption{\% coverage of the SciPy code base by unit tests over the six months
+preceding June 28, 2018 as reported by the \texttt{codecov} service. The
+analysis of lines of source code covered is performed automatically in our
+continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
+libraries.}
 \label{fig:coverage}
@@
 any strong indications that they should step down; although the BDFL may
-appoint a successor, it is generally expected that the Steering Council will be
+appoint a successor it is generally expected that the Steering Council will be
 consulted on the matter.
@@
     \texttt{linalg} & support a more recent version of LAPACK\\
-    \texttt{ndimage} & clarify usage of the ``data point'' coordinate
-    model, and add additional wrapping modes\\
+    \texttt{ndimage} & migrate to data point model from pixel model\\
     \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\

### 2019-03-06 scipy-1.0/paper.tex
streamlined Fortran paragraph per @ilayn's comment

@@
 
-Fortran is an
-extremely well-established scientific programming language, both for historical
-reasons and because of its continued excellent
-performance\cite{Koelbel:1993:HPF:562354}. We wrap the FFTPACK Fortran library
-for performing Fourier transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}
-since this library has been a standard in the field for 33 years and has a
-license that is compatible with our own. Likewise, we wrap the Fortran source
-for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial
-value problem for ordinary differential equation systems for 30 years. For
-similar reasons we also wrap the Fortran libraries
-QUADPACK\cite{1983qspa.book.....P} (for numerical integration of
-one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103}
-(curve-fitting / interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal
-distance regression), MINPACK\cite{osti_6997568} (minimization of linear and
-nonlinear equations), ARPACK\cite{leh:sor:yan96} (solving large scale
-eigenvalue problems), ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling
-Bessel Functions), and CDFLIB\cite{CDFLIB_site} (evaluation of cumulative
-density functions).
+Fortran, despite its age, is still a high-performance scientific programming language with 
+continued contemporary usage. Thus, we wrap the following excellent field-tested Fortran
+libraries in order to provide the Python convenience while still benefiting from their respective
+performance: 
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
+ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
+QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
+FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
+ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
+MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
+ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
+CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
 

### 2019-03-06 scipy-1.0/paper.tex
streamlined C paragraph per @ilayn's comment

@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. Some of the C libraries
-that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
-(iterative solving of the trust region problem),
-SuperLU\cite{li05,superlu_ug99} (solution of large, sparse, nonsymmetric
-systems of linear equations), Qhull\cite{Barber:1996:QAC:235815.235821}
-(computational geometry), and Cephes\cite{cephes_netlib} (mathematics
-algorithms).
+scientific computing. The C libraries that we wrap in SciPy include 
+trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry) 
+Cephes\cite{cephes_netlib}. % (mathematics algorithms).
 

### 2019-03-06 scipy-1.0/paper.tex
tweaks to Fortran and C paragraphs

@@
 Fortran, despite its age, is still a high-performance scientific programming language with 
-continued contemporary usage. Thus, we wrap the following excellent field-tested Fortran
-libraries in order to provide the Python convenience while still benefiting from their respective
-performance: 
+continued contemporary usage. Thus, we wrap the following excellent, field-tested Fortran
+libraries in order to provide Python convenience while benefiting from their performance: 
 FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 

### 2019-03-06 scipy-1.0/paper.tex
added Fortran reference back
replaced language choices figure with table

--- 2019-03-06 scipy-1.0/paper.tex
@@
 \usepackage{float}
-\usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \usepackage[multiple]{footmisc}
+\usepackage{booktabs,siunitx}
 
@@
 SciPy is a library of numerical routines for the Python programming
-language, providing the fundamental building blocks for modeling and
-solving scientific problems.  For example, when LIGO, the Laser
-Interferometer Gravitational-Wave Observatory, recently announced the
-discovery of gravitational waves, their published Python analysis
-scripts\cite{LIGO-open,alex_nitz_2018_1596771,GW150914-Python,
-PhysRevD.93.122003} imported many different SciPy submodules for:
-signal processing, interpolation, input / output, optimization, statistics,
-integration, and more. SciPy has routines for solving
-classes of generic problems such as optimization, integration,
-interpolation, eigenvalues, and linear systems.  It also provides
-specialized data structures, such as sparse arrays and k-dimensional
-trees. SciPy is built on top of NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
-which provides array data structures and related fast numerical routines.
+language that provides fundamental building blocks for modeling and
+solving scientific problems. 
+SciPy includes algorithms for optimization, integration, interpolation,
+eigenvalue problems, algebraic equations, differential equations, and many other
+classes of problems; it also provides
+specialized data structures, such as sparse arrays
+and $k$-dimensional trees. SciPy is built on top of 
+NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
+which provides array data structures and related fast numerical routines, and
 SciPy is itself the foundation upon which higher level scientific libraries,
@@
 
-Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
-in SciPy. While it is not surprising that Python is predominant, the usage distribution 
-of the other languages warrants discussion.
-
-\begin{figure}[H]
+Python, Cython, Fortran, C and C++ are the programming languages used to
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library\cite{linguistref} provides a 
+detailed breakdown as \% composition by programming language in 
+SciPy (Table~\ref{tab:linguist}).
+
+\begin{table}
 \centering
-
-    \begin{tikzpicture}[font=\sffamily]
-    \begin{axis}[height=6cm, width=10cm,
-    ybar, ymax=60, ymin=0,bar width=1cm,
-    ylabel={\% of Scipy code base},
-    ytick={0,20,...,60},
-    axis lines*=left,
-    axis y line shift=0pt,
-    enlarge x limits=0.15,
-    symbolic x coords={py,f,c,cy,cpp},
-    xticklabels={Python,FORTRAN,C,Cython,C++},
-    ticklabel style={align=center,font=\sffamily\small,
-    /pgf/number format/assume math mode=true},
-    xtick=data,
-    nodes near coords,
-    nodes near coords style={/pgf/number format/assume math mode=true}
-    ]
-    \addplot coordinates {(py,49.5) (f,25.6) (c,19.5) (cy, 3.0) (cpp,2.3)};
-    \end{axis}
-    \end{tikzpicture}
-    \caption{The breakdown of programming languages used in the
-             SciPy library determined using the \texttt{linguist} library.
-    	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
-    	 and Makefile are excluded for clarity and mostly
-    	 provide supporting roles in tests, building, and
-    	 documentation.}
-    \label{fig:linguist}
-\end{figure}
-
-Fortran, despite its age, is still a high-performance scientific programming language with 
-continued contemporary usage. Thus, we wrap the following excellent field-tested Fortran
-libraries in order to provide the Python convenience while still benefiting from their respective
-performance: 
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
-ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
-MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
-ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
-CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
+\begin{tabular}{p{3cm}S[table-format = <0.1]}
+\toprule
+Language & {Percent}\\
+\midrule
+Python  & 49.5 \\
+Fortran & 25.6 \\
+C       & 19.5 \\
+Cython  & 3.0  \\
+C++     & 2.3  \\[\defaultaddspace]
+{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
+\bottomrule
+\end{tabular}
+\caption{Language composition of SciPy codebase: lines of code in each
+programming language as determined by the \texttt{linguist} package.
+The last row denotes tools used in supporting roles in tests,
+building, and documentation.}
+\label{tab:linguist}
+\end{table}
+
+For implementing new functionality, we have a clear order of language
+preference.  First Python, if performance is not an issue. If it is, then in
+order of preference: Cython, C, C++, Fortran. The main motivation for this is
+maintainability: Cython has the highest abstraction level and most Python
+developers will understand it. C is also widely known, and easier to deal with
+for the core development team than C++ and especially Fortran.
+
+While it is not surprising that Python is heavily used in SciPy, the usage
+distribution of the other languages warrants some discussion. Fortran is an
+extremely well-established scientific programming language, both for historical
+reasons and because of its continued excellent
+performance\cite{Koelbel:1993:HPF:562354}. We wrap the FFTPACK Fortran library
+for performing Fourier transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}
+since this library has been a standard in the field for 33 years and has a
+license that is compatible with our own. Likewise, we wrap the Fortran source
+for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial
+value problem for ordinary differential equation systems for 30 years. For
+similar reasons we also wrap the Fortran libraries
+QUADPACK\cite{1983qspa.book.....P} (for numerical integration of
+one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103}
+(curve-fitting / interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal
+distance regression), MINPACK\cite{osti_6997568} (minimization of linear and
+nonlinear equations), ARPACK\cite{leh:sor:yan96} (solving large scale
+eigenvalue problems), ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling
+Bessel Functions), and CDFLIB\cite{CDFLIB_site} (evaluation of cumulative
+density functions).
 
@@
 
-Cython has been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
-as a glue between well-established low-level scientific computing libraries written
-in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
-enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing.
-
-For implementing new functionality, Python is the still the language of choice. If Python 
-performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
-order). The main motivation for this is maintainability: Cython has the 
-highest abstraction level and most Python developers will understand it. C is 
-also widely known, and easier for the current core development team to manage
-than C++ and especially Fortran.
-
-The distribution of secondary programming languages in SciPy is a compromise between
-a powerful, performance-
-enhancing language that interacts well with Python (i.e. Cython) and the
+Therefore, the relative abundance of different programming languages in the
+SciPy library results from a combination of the usage of powerful performance
+enhancing languages that interact well with Python (i.e., Cython), and the
 usage of languages (and their libraries) that have proven reliable and
@@
 \centering
-\includegraphics[width=\textwidth]{static/coverage-chart}
-\caption{\% coverage of the SciPy code base by unit tests over the six months
-preceding June 28, 2018 as reported by the \texttt{codecov} service. The
-analysis of lines of source code covered is performed automatically in our
-continuous integration suite using the \texttt{pytest-cov} and \texttt{gcov}
-libraries.}
+\begin{tikzpicture}[]
+\pgfplotstableread[column/ver/.style=string type]{
+ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
+v0.12.1 33844 25721 8123 273634  142290 131344
+v0.13.3 36638 28944 7694 338926  172852 166074
+v0.14.1 37301 30587 6714 277643  147151 130492
+v0.15.1 38339 30288 8051 304898  158547 146351
+v0.16.1 40094 32075 8019 322398  167647 154751
+v0.17.1 42566 34478 8088 340903  170452 170451
+v0.18.1 44711 36216 8495 353417  173174 180243
+v0.19.1 43823 36373 7450 436900  200974 235926
+v1.0.0 106878 92984 13894 462574 208158 254416
+}\covtable
+    \begin{axis}[
+    width=\textwidth,
+    height=7cm,
+    enlargelimits=false,
+    ymin=0,
+    ymax=6e5,
+    stack plots=y,%   
+    area style,
+    xtick=data,
+    ytick={2e5,4e5,6e5},
+    xticklabels from table={\covtable}{ver},
+    scaled y ticks=false,
+    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
+    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
+    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
+    legend columns=-1,
+    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
+    ]
+  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
+  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
+  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
+  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
+\end{axis}
+\end{tikzpicture}
+\caption{
+% This paragraph of caption won't compile
+%Python (red) and compiled (blue) code volume in SciPy over time. 
+%Deep-shaded area represents lines of code covered by units tests; 
+%light-shaded area represents lines not covered. With the exception
+%of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
+%v0.14 due to (insert reason), the volume of both compiled and Python
+%code has increased between releases, as has the number of lines
+%covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
+%compiled code with test coverage near 45\% and 106,878 lines of Python 
+%code with test coverage near 87\%.
+The reconstitution of historical
+build and test environments was performed using a docker-based approach
+\cite{scipy-cov}. Analysis of lines of source code covered is performed 
+automatically in our continuous integration suite using the \texttt{pytest-cov}
+and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
+inclusion of additional vendored code.}
 \label{fig:coverage}

### 2019-03-06 scipy-1.0/paper.tex

@@
 \usepackage[multiple]{footmisc}
-\usepackage{booktabs,siunitx}
 
@@
 detailed breakdown as \% composition by programming language in 
-SciPy (Table~\ref{tab:linguist}).
-
-\begin{table}
+SciPy (Figure~\ref{fig:linguist}).
+
+\begin{figure}[H]
     \centering
-\begin{tabular}{p{3cm}S[table-format = <0.1]}
-\toprule
-Language & {Percent}\\
-\midrule
-Python  & 49.5 \\
-Fortran & 25.6 \\
-C       & 19.5 \\
-Cython  & 3.0  \\
-C++     & 2.3  \\[\defaultaddspace]
-{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
-\bottomrule
-\end{tabular}
-\caption{Language composition of SciPy codebase: lines of code in each
-programming language as determined by the \texttt{linguist} package.
-The last row denotes tools used in supporting roles in tests,
-building, and documentation.}
-\label{tab:linguist}
-\end{table}
+		\includegraphics[width=\textwidth]{static/composition}
+    \caption{The breakdown of programming languages used in the
+             SciPy library determined using the linguist library.
+    	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
+    	 and Makefile are excluded for clarity and mostly
+    	 provide supporting roles in tests, building, and
+    	 documentation.}
+    \label{fig:linguist}
+\end{figure}
 
@@
 The SciPy test suite is orchestrated by a continuous integration matrix that
-includes posix and Windows (32/64-bit) platforms managed by Travis CI and
-appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with pyflakes and pycodestyle. There are more than $13000$
-unit tests in the test suite, which is written for usage with the pytest
-runner, and with 87 \% and 45 \% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
-Some of the historical components of the code base would still benefit from increased
-test coverage. Documentation for the code is automatically built and hosted by
+includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which is written for usage with the \texttt{pytest}
+framework, and with 87 \% and 45 \% line coverages for Python and compiled
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 
-SciPy adopted an official Governance Document on August 3,
-2017\cite{SciPyProjectGovernance}.  We have a Benevolent Dictator for Life
-(BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
-the governance model with the community on the mailing list.  In practice, the
-overruling authority of the BDFL is anticipated to be used only when the
-Steering Council cannot agree on a matter.  The BDFL is expected to consider
-any strong indications that they should step down; although the BDFL may
-appoint a successor it is generally expected that the Steering Council will be
-consulted on the matter.
-
-The (currently) 18 member Steering Council is chaired by Ralf Gommers and
-performs a number of tasks including nominating and voting on new Council
-Members, and actively contributing to the progress of the project (could be
-code review but also outreach activities, etc.). Generally, one year of sustained
-and substantial contributions (not restricted to source code contributions)
-that improve the project are the prerequisites for nomination of new Council
-Members. Council Members do not get special treatment for the vast majority of
-the day to day project activities (code submission / review), though it is
-anticipated that their experience with the project will serve as helpful
-guidance in many matters.  Council Members have commit rights to the project,
-but will typically only incorporate changes when there are no substantive
-community objections. The Chair of the Steering Council is responsible for
-starting more formal technical reviews of the direction of the project,
-ensuring that the composition of the Council remains current, and communicating
-/ summarizing any Council activities performed privately to the broader
-community. We also describe in some detail how institutional partners may
-contribute to the project, but no financial weight from grants or other sources
-may be leveraged to circumvent or overrule the technical direction of the
-project decided upon by the Steering Committee. The Chair does not have a fixed
-term, but is nominated by the Steering Committee and is expected to yield to
-substantive claims that stepping down is the appropriate action.
-
-SciPy adopted an official Code of Conduct on October 24, 2017
-\cite{SciPyCodeOfConduct}, drawing inspiration from the Apache Foundation Code
-of Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
-\cite{ContributorConvenant}, the Jupyter Code of Conduct \cite{Jupyter_COC},
-and the Open Source Guides - Code of Conduct\cite{OSG_COC}. In short, we stated
-five specific guidelines: \emph{be open} (to anyone participating in our
-community), \emph{be empathetic and patient} (in resolving conflicts), \emph{be
-collaborative} (we depend on each other to build up the tools in the library),
-\emph{be inquisitive} (identifying issues early on can be helpful to everyone),
-and \emph{be careful with wording} (do not harass or exclude).  We outlined a
-broad diversity inclusion statement, and provide instructions for contacting
-the three members of our Code of Conduct Committee or an external
-representative on the NumFOCUS Board of Directors.
+SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
+A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
+Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
+The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community.
+The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.
+
+SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
+\emph{be open} to everyone participating in our community;
+\emph{be empathetic and patient} in resolving conflicts;
+\emph{be collaborative}, as we depend on each other to build the library;
+\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
+\emph{be careful with wording}.
+The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response.
+Our Diversity Statement ``welcomes and encourages participation by everyone''.
 

### 2019-03-06 scipy-1.0/paper.tex
capitalized Astropy

@@
 
-SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
-A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
-Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
-The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community.
-The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.
-
-SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
-\emph{be open} to everyone participating in our community;
-\emph{be empathetic and patient} in resolving conflicts;
-\emph{be collaborative}, as we depend on each other to build the library;
-\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
-\emph{be careful with wording}.
-The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response.
-Our Diversity Statement ``welcomes and encourages participation by everyone''.
+SciPy adopted an official Governance Document on August 3,
+2017\cite{SciPyProjectGovernance}.  We have a Benevolent Dictator for Life
+(BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
+the governance model with the community on the mailing list.  In practice, the
+overruling authority of the BDFL is anticipated to be used only when the
+Steering Council cannot agree on a matter.  The BDFL is expected to consider
+any strong indications that they should step down; although the BDFL may
+appoint a successor, it is generally expected that the Steering Council will be
+consulted on the matter.
+
+The (currently) 18 member Steering Council is chaired by Ralf Gommers and
+performs a number of tasks including nominating and voting on new Council
+Members and actively contributing to the progress of the project (could be
+code review but also outreach activities, etc.). Generally, one year of sustained
+and substantial contributions (not restricted to source code contributions)
+that improve the project are the prerequisites for nomination of new Council
+Members. Council Members do not get special treatment for the vast majority of
+the day to day project activities (code submission / review), though it is
+anticipated that their experience with the project will serve as helpful
+guidance in many matters.  Council Members have commit rights to the project,
+but will typically only incorporate changes when there are no substantive
+community objections. The Chair of the Steering Council is responsible for
+starting more formal technical reviews of the direction of the project,
+ensuring that the composition of the Council remains current, and communicating
+/ summarizing any Council activities performed privately to the broader
+community. We also describe in some detail how institutional partners may
+contribute to the project, but no financial weight from grants or other sources
+may be leveraged to circumvent or overrule the technical direction of the
+project decided upon by the Steering Committee. The Chair does not have a fixed
+term but is nominated by the Steering Committee and is expected to yield to
+substantive claims that stepping down is the appropriate action.
+
+SciPy adopted an official Code of Conduct on October 24, 2017
+\cite{SciPyCodeOfConduct}, drawing inspiration from the Apache Foundation Code
+of Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
+\cite{ContributorConvenant}, the Jupyter Code of Conduct \cite{Jupyter_COC},
+and the Open Source Guides - Code of Conduct\cite{OSG_COC}. In short, we stated
+five specific guidelines: \emph{be open} (to anyone participating in our
+community), \emph{be empathetic and patient} (in resolving conflicts), \emph{be
+collaborative} (we depend on each other to build up the tools in the library),
+\emph{be inquisitive} (identifying issues early on can be helpful to everyone),
+and \emph{be careful with wording} (do not harass or exclude).  We outlined a
+broad diversity inclusion statement, and provide instructions for contacting
+the three members of our Code of Conduct Committee or an external
+representative on the NumFOCUS Board of Directors.
 
@@
 more effective science applications. For example, there
-are common contributors to the SciPy and astropy core
+are common contributors to the SciPy and Astropy core
 libraries\cite{astropy-2018}, and what works well for 

### 2019-03-06 scipy-1.0/paper.tex
making codebase consistent

@@
 libraries\cite{astropy-2018}, and what works well for 
-one of the code bases, infrastructures, or communities 
-is often transferred in some form to the other. At the code
-base level, the \texttt{binned\_statistic} functionality
+one of the codebases, infrastructures, or communities 
+is often transferred in some form to the other. At the codebase 
+level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution---it was initially

### 2019-03-06 scipy-1.0/paper.tex
removed idiom 'after the fact'

@@
 developed in an Astropy-affiliated package
-and then placed in SciPy after the fact.
+and then placed in SciPy afterwards.
 

### 2019-03-06 scipy-1.0/paper.tex
removed subheadings from discussion

--- 2019-03-06 scipy-1.0/paper.tex
@@
 more effective science applications. For example, there
-are common contributors to the SciPy and Astropy core
+are common contributors to the SciPy and astropy core
 libraries\cite{astropy-2018}, and what works well for 
@@
 
-\fixme{The Discussion should be succinct and must not contain subheadings.}
-
-\subsection*{Impact now}
-
-% While Paul was one of the early adopters at the national laboratories,
-% Python's presence is now ubiquitous.
+% The Discussion should be succinct and must not contain subheadings.
 
@@
 by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
-
-\subsection*{Future Work}
 

### 2019-03-06 scipy-1.0/paper.tex
removed mention of 20,000 visitors from SciPy Today

@@
 Over 85,000 GitHub repositories and almost 5000 packages depend
-on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
-visitors every two weeks, and more than 100 people contribute to a 
-new release every six months. Some of the major
+on SciPy. Some of the major
 feature highlights from the three years preceding
@@
 14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
-that roughly two-week period. The developer community consists of 610 unique
+that roughly two-week period. The developer community at that time consisted of 610 unique
 contributors of source code, with $>19,000$ commits accepted into the code base

### 2019-03-06 scipy-1.0/paper.tex
making codebase consistent

@@
 metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
+14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
 that roughly two-week period. The developer community at that time consisted of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the code base
+contributors of source code, with $>19,000$ commits accepted into the codebase
 (GitHub page data).
@@
 scientific collaboration that lead to the observation of gravitational waves
-\cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}. SciPy is used
+\cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
+MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
 by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.

### 2019-03-28 scipy-1.0/paper.tex
revised description of stats module

--- 2019-03-06 scipy-1.0/paper.tex
@@
 
-The \texttt{scipy.stats} module provides more than 100 probability
-distributions and a framework to implement additional distributions. Given an
-expression for a probability distribution function (pdf), the framework will
-calculate moments, cumulative distribution function (cdf), differential
-entropy, and more. Those properties are calculated based on generic code, e.g.,
-numerical integration of the pdf to obtain the cdf.  Where possible, the
-generic code is replaced by explicit formulas or more accurate or efficient
-implementations. For example, most distributions have an explicit formula for
-the cdf, so it is used instead of integrating the pdf. The distributions
-framework also allows the user to draw random variates from the distribution.
-Drawing random variates now accepts a \texttt{random\_state} parameter, which
-is either a \texttt{numpy.random.RandomState} object or an integer used to
-generate such an object, to provide fine-grained user-level control over the
-random state of a workflow.  Currently about 80 continuous univariate
-distributions, 12 discrete univariate distributions and 10 multivariate
-distributions are implemented.  Key recent distributions added to
-\texttt{scipy.stats} include the histogram-based distribution in
-\texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
-\texttt{scipy.stats.multinomial} (used, for example, in natural language
-processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
-the only multivariate discrete distribution available in SciPy.
+The \texttt{scipy.stats} module contains more than 100 probability
+distributions (96 continuous and 13 dicrete univariate distributions
+and 10 multivariate distributions). The implementation relies on a
+consistent framework that provides methods to sample random variates,
+to evaluate the cumulative distribution function (cdf), the probability
+density function (pdf) and to fit parameters for every distribution. 
+Generally, the methods rely on specific implementations for each
+distribution if available (such as a closed form expression of the
+cdf or a sampling algorithm). Otherwise, default methods are used
+based on generic code, e.g., numerical integration of the pdf to
+obtain the cdf.
 
@@
 
-SciPy adopted an official Governance Document on August 3,
-2017\cite{SciPyProjectGovernance}.  We have a Benevolent Dictator for Life
-(BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
-the governance model with the community on the mailing list.  In practice, the
-overruling authority of the BDFL is anticipated to be used only when the
-Steering Council cannot agree on a matter.  The BDFL is expected to consider
-any strong indications that they should step down; although the BDFL may
-appoint a successor, it is generally expected that the Steering Council will be
-consulted on the matter.
-
-The (currently) 18 member Steering Council is chaired by Ralf Gommers and
-performs a number of tasks including nominating and voting on new Council
-Members and actively contributing to the progress of the project (could be
-code review but also outreach activities, etc.). Generally, one year of sustained
-and substantial contributions (not restricted to source code contributions)
-that improve the project are the prerequisites for nomination of new Council
-Members. Council Members do not get special treatment for the vast majority of
-the day to day project activities (code submission / review), though it is
-anticipated that their experience with the project will serve as helpful
-guidance in many matters.  Council Members have commit rights to the project,
-but will typically only incorporate changes when there are no substantive
-community objections. The Chair of the Steering Council is responsible for
-starting more formal technical reviews of the direction of the project,
-ensuring that the composition of the Council remains current, and communicating
-/ summarizing any Council activities performed privately to the broader
-community. We also describe in some detail how institutional partners may
-contribute to the project, but no financial weight from grants or other sources
-may be leveraged to circumvent or overrule the technical direction of the
-project decided upon by the Steering Committee. The Chair does not have a fixed
-term but is nominated by the Steering Committee and is expected to yield to
-substantive claims that stepping down is the appropriate action.
-
-SciPy adopted an official Code of Conduct on October 24, 2017
-\cite{SciPyCodeOfConduct}, drawing inspiration from the Apache Foundation Code
-of Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
-\cite{ContributorConvenant}, the Jupyter Code of Conduct \cite{Jupyter_COC},
-and the Open Source Guides - Code of Conduct\cite{OSG_COC}. In short, we stated
-five specific guidelines: \emph{be open} (to anyone participating in our
-community), \emph{be empathetic and patient} (in resolving conflicts), \emph{be
-collaborative} (we depend on each other to build up the tools in the library),
-\emph{be inquisitive} (identifying issues early on can be helpful to everyone),
-and \emph{be careful with wording} (do not harass or exclude).  We outlined a
-broad diversity inclusion statement, and provide instructions for contacting
-the three members of our Code of Conduct Committee or an external
-representative on the NumFOCUS Board of Directors.
+SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
+A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
+Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
+The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community.
+The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.
+
+SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
+\emph{be open} to everyone participating in our community;
+\emph{be empathetic and patient} in resolving conflicts;
+\emph{be collaborative}, as we depend on each other to build the library;
+\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
+\emph{be careful with wording}.
+The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response.
+Our Diversity Statement ``welcomes and encourages participation by everyone''.
 
@@
 
-% The Discussion should be succinct and must not contain subheadings.
+\fixme{The Discussion should be succinct and must not contain subheadings.}
+
+\subsection*{Impact now}
+
+% While Paul was one of the early adopters at the national laboratories,
+% Python's presence is now ubiquitous.
 
@@
 by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
+
+\subsection*{Future Work}
 

### 2019-03-28 scipy-1.0/subpackages.tex
revised description of stats module

@@
 \item[\texttt{stats}]
-    The \texttt{stats} subpackage provides a large collection of continuous
-    and discrete probability distributions, each with methods to compute
+    The \texttt{stats} subpackage provides: a large collection of continuous
+    and discrete \emph{probability distributions}, each with methods to compute
     the PDF or PMF, CDF, moments and other statistics, generation of random
     variates, and more;
-    statistical tests, including Pearson's correlation, Spearman's rank-order
-    correlations, Kendall's tau, chi-squared test and its generalization as the
-    Cressie-Read power divergence, contingency table tests including Fisher's
-    exact test and Mood's median test, and many more;
-    and assorted transformations and statistics of data.
+    statistical tests, including tests on equality of means / medians /
+    variance (such as the t-test) and tests whether a sample is drawn from a
+    certain distribution (such as the Kolmogorov-Smirnov test);
+    measures of correlation, including Pearson's, Kendall's and Spearman's
+    coefficients;
+    descriptive statistics including trimmed values;
+    kernel density estimation;
+    and transformations of data such as the Box-Cox power transformation.
 \end{description}

### 2019-04-30 scipy-1.0/paper.tex
add comment on features recently added
Cite M87 black hole project
* mention in abstract & introduction,
along with appropriate citation for this
high-profile black hole imaging project

@@
 packages, and 28700 dependent repositories. This includes usage of SciPy
-in almost half of all machine learning projects on GitHub.
+in almost half of all machine learning projects on GitHub, and usage by
+high profile projects including M87 black hole imaging and gravitational
+wave analysis.
 The library includes functionality spanning clustering, Fourier transforms,
@@
 \cite{PhysRevD.93.122003, abbott2017gw170817} 
-import several submodules of SciPy.
+import several submodules of SciPy, and the M87 black
+hole imaging project directly cites SciPy\cite{2019ApJ875L3E}.
 
@@
 
-The \texttt{scipy.stats} module contains more than 100 probability
-distributions (96 continuous and 13 dicrete univariate distributions
-and 10 multivariate distributions). The implementation relies on a
-consistent framework that provides methods to sample random variates,
-to evaluate the cumulative distribution function (cdf), the probability
-density function (pdf) and to fit parameters for every distribution. 
-Generally, the methods rely on specific implementations for each
-distribution if available (such as a closed form expression of the
-cdf or a sampling algorithm). Otherwise, default methods are used
-based on generic code, e.g., numerical integration of the pdf to
-obtain the cdf.
+The \texttt{scipy.stats} module provides more than 100 probability
+distributions and a framework to implement additional distributions. Given an
+expression for a probability distribution function (pdf), the framework will
+calculate moments, cumulative distribution function (cdf), differential
+entropy, and more. Those properties are calculated based on generic code, e.g.,
+numerical integration of the pdf to obtain the cdf.  Where possible, the
+generic code is replaced by explicit formulas or more accurate or efficient
+implementations. For example, most distributions have an explicit formula for
+the cdf, so it is used instead of integrating the pdf. The distributions
+framework also allows the user to draw random variates from the distribution.
+Drawing random variates now accepts a \texttt{random\_state} parameter, which
+is either a \texttt{numpy.random.RandomState} object or an integer used to
+generate such an object, to provide fine-grained user-level control over the
+random state of a workflow.  Currently about 80 continuous univariate
+distributions, 12 discrete univariate distributions and 10 multivariate
+distributions are implemented.  Key recent distributions added to
+\texttt{scipy.stats} include the histogram-based distribution in
+\texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
+\texttt{scipy.stats.multinomial} (used, for example, in natural language
+processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
+the only multivariate discrete distribution available in SciPy.
 

### 2019-05-01 scipy-1.0/paper.tex

@@
 in almost half of all machine learning projects on GitHub, and usage by
-high profile projects including M87 black hole imaging and gravitational
-wave analysis.
+high profile projects including LIGO gravitational wave analysis and creation
+of the first-ever image of a black hole (M87).
 The library includes functionality spanning clustering, Fourier transforms,

### 2019-05-03 scipy-1.0/paper.tex

@@
 more effective science applications. For example, there
-are common contributors to the SciPy and astropy core
+are common contributors to the SciPy and Astropy core
 libraries\cite{astropy-2018}, and what works well for 
-one of the code bases, infrastructures, or communities 
-is often transferred in some form to the other. At the code
-base level, the \texttt{binned\_statistic} functionality
+one of the codebases, infrastructures, or communities 
+is often transferred in some form to the other. At the codebase 
+level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution---it was initially
 developed in an Astropy-affiliated package
-and then placed in SciPy after the fact.
+and then placed in SciPy afterwards.
 

### 2019-05-03 scipy-1.0/paper.tex
Swap subsections in Project Organization The "Governance" and "Maintainers and contributors" subsections are focused on SciPy itself (internal). "Commmunity beyond the SciPy library" is about relationships with other projects (external), but its sandwiched between the other two internal sections. I changed the order of subsections to "Governance", "Maintainers and contributors", and "Community beyond the SciPy library"; they are now in order from closest to the core of the project to further away.

@@
 metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
-that roughly two-week period. The developer community consists of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the code base
+14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
+that roughly two-week period. The developer community at that time consisted of 610 unique
+contributors of source code, with $>19,000$ commits accepted into the codebase
 (GitHub page data).
@@
 scientific collaboration that lead to the observation of gravitational waves
-\cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}. SciPy is used
+\cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
+MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
 by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
-
-\subsection*{Future Work}
 

### 2019-05-03 scipy-1.0/paper.tex
Brought back numba by request

@@
 in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions directly from Python.  
+calling compiled C functions, such as those generated using \texttt{numba}
+or Cython, directly from Python.  
 Supported low-level functions include \texttt{PyCapsule} 

### 2019-05-03 scipy-1.0/scipy-optimize.tex
Reductions to paragraphs on local and global optimizers As discussed in #113, #102

@@
 \paragraph{Local Minimization}
-
 The \texttt{minimize} function unifies several methods of finding local minima of nonlinear optimization problems in a common interface.

### 2019-05-03 scipy-1.0/paper.tex
Trim LowLevelCallable As discussed in #113

--- 2019-05-03 scipy-1.0/paper.tex
@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
-calling compiled C functions directly from Python.  Supported low-level
-functions include \texttt{PyCapsule} objects, ctypes function pointers, and
-cffi function pointers. The low level function signature must be consistent
-with the expectations of the routine it is passed to. For example, the
-documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
-C callback function signatures that may be used to produce functions that
-operate on each element of image data with low overhead. The C code may be
-generated using numba or Cython, for example, as long as the function call
-signature matches the specifications. Furthermore, it is even possible to
-generate a low-level callback function automatically from a Cython module using
-\texttt{scipy.LowLevelCallable.from\_cython}.
+in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
+calling compiled C functions, such as those generated using \texttt{numba}
+or Cython, directly from Python.
+Supported low-level functions include \texttt{PyCapsule}
+objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
+Furthermore, it is even possible to generate a low-level callback function
+automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
 

### 2019-05-03 scipy-1.0/paper.tex
Reorder cKDTree / sparse The sparse section is a more natural place to define sparse matrix format acronyms than cKDTree. By swapping the order of cKDTree and sparse, we define the acronyms in sparse then use them in cKDTree.

@@
 
+\subsubsection*{\texttt{cKDTree}}
+
+\input{ckdtree}
+
 \subsection*{Unified bindings to compiled code}
@@
 As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions, such as those generated using \texttt{numba}
-or Cython, directly from Python.
-Supported low-level functions include \texttt{PyCapsule}
-objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
-Furthermore, it is even possible to generate a low-level callback function
-automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
+in a \texttt{scipy.LowLevelCallable()} object that reduces the overhead for
+calling compiled C functions directly from Python.  Supported low-level
+functions include \texttt{PyCapsule} objects, ctypes function pointers, and
+cffi function pointers. The low level function signature must be consistent
+with the expectations of the routine it is passed to. For example, the
+documentation for \texttt{scipy.ndimage.generic\_filter} defines two acceptable
+C callback function signatures that may be used to produce functions that
+operate on each element of image data with low overhead. The C code may be
+generated using numba or Cython, for example, as long as the function call
+signature matches the specifications. Furthermore, it is even possible to
+generate a low-level callback function automatically from a Cython module using
+\texttt{scipy.LowLevelCallable.from\_cython}.
 

### 2019-05-03 scipy-1.0/paper.tex
Mention of scikit-learn seemed out of place As discussed in #113
update Community beyond the SciPy library section

@@
 equivalent NumPy API where possible.
-
-\subsubsection*{\texttt{cKDTree}}
-
-\input{ckdtree}
 
@@
 The scientific Python ecosystem includes many examples
-of domain-specific software libraries building on top
-of SciPy features, and then returning to the base SciPy library
-to suggest and even implement improvements to enable
-more effective science applications. For example, there
-are common contributors to the SciPy and Astropy core
-libraries\cite{astropy-2018}, and what works well for 
-one of the codebases, infrastructures, or communities 
+of domain-specific software libraries that are built 
+on top of SciPy features. The work done on these libraries
+often turns out to be beneficial to the whole ecosystem
+being Scipy the catalyst of this cross-fertilisation process. 
+For example, there are common contributors to the SciPy and 
+Astropy core libraries\cite{astropy-2018}, and what works 
+well for one of the codebases, infrastructures, or communities 
 is often transferred in some form to the other. At the codebase 

### 2019-05-03 scipy-1.0/paper.tex

--- 2019-05-03 scipy-1.0/paper.tex
@@
 Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library\cite{linguistref} provides a 
+implement scientific algorithms in the SciPy library. An analysis of our
+codebase using the \texttt{linguist} library\cite{linguistref} provides a 
 detailed breakdown as \% composition by programming language in 
@@
 
-Over the last three years our sparse matrix handling internals have been
+Over the last three years, our sparse matrix handling internals have been
 rewritten and performance has been improved. Iterating over and slicing of CSC
 and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the COO / DIA to CSR / CSC matrix format
+ and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
 conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,

### 2019-05-03 scipy-1.0/ckdtree.tex

@@
 p-norm between 1 and infinity. By default, the returned data structure is a 
-Dictionary Of Keys based sparse matrix (DOK), which is very efficient for matrix 
+Dictionary Of Keys (DOK) based sparse matrix, which is very efficient for matrix 
 construction. This hashing approach to sparse matrix assembly can be 7 times 
-faster than constructing with Compressed Sparse Row format (CSR) 
-\cite{10.1007/978-3-540-75755-9_107} and the C++ level sparse matrix construction 
+faster than constructing with CSR format
+\cite{10.1007/978-3-540-75755-9_107}, and the C++ level sparse matrix construction 
 releases the Python GIL for increased performance. Once the matrix is constructed, 
@@
 \cite{Cormen:2001:IA:580470}, and the DOK structure can be efficiently converted 
-to a CSR, Compressed Sparse Column (CSC) or COOrdinate (COO) matrix to allow for 
+to a CSR, CSC, or COO matrix to allow for 
 speedy arithmetic operations.

### 2019-05-03 scipy-1.0/paper.tex

@@
 These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily in scikit-learn.
+and are used heavily throughout SciPy and dependent packages.
 

### 2019-05-04 scipy-1.0/paper.tex
change section title to Relation with downstream projects

@@
 
-Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our
-codebase using the \texttt{linguist} library\cite{linguistref} provides a 
-detailed breakdown as \% composition by programming language in 
-SciPy (Figure~\ref{fig:linguist}).
+Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
+in SciPy. While it is not surprising that Python is predominant, the usage distribution 
+of the other languages warrants discussion.
 
@@
 
-For implementing new functionality, we have a clear order of language
-preference.  First Python, if performance is not an issue. If it is, then in
-order of preference: Cython, C, C++, Fortran. The main motivation for this is
-maintainability: Cython has the highest abstraction level and most Python
-developers will understand it. C is also widely known, and easier to deal with
-for the core development team than C++ and especially Fortran.
-
-While it is not surprising that Python is heavily used in SciPy, the usage
-distribution of the other languages warrants some discussion. Fortran is an
-extremely well-established scientific programming language, both for historical
-reasons and because of its continued excellent
-performance\cite{Koelbel:1993:HPF:562354}. We wrap the FFTPACK Fortran library
-for performing Fourier transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}
-since this library has been a standard in the field for 33 years and has a
-license that is compatible with our own. Likewise, we wrap the Fortran source
-for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial
-value problem for ordinary differential equation systems for 30 years. For
-similar reasons we also wrap the Fortran libraries
-QUADPACK\cite{1983qspa.book.....P} (for numerical integration of
-one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103}
-(curve-fitting / interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal
-distance regression), MINPACK\cite{osti_6997568} (minimization of linear and
-nonlinear equations), ARPACK\cite{leh:sor:yan96} (solving large scale
-eigenvalue problems), ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling
-Bessel Functions), and CDFLIB\cite{CDFLIB_site} (evaluation of cumulative
-density functions).
+Fortran, despite its age, is still a high-performance scientific programming language with 
+continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
+libraries in order to provide Python convenience while benefiting from their performance: 
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
+ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
+QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
+FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
+ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
+MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
+ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
+CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
 
@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. Alongside its objected-oriented relative C++, C can be
-leveraged in Python after being exposed using the Cython language. Cython has
-been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We thus often use Cython
-as a glue between well-established low-level scientific computing libraries and
-the Python interface offered by SciPy. We also use Cython to enable performance
+scientific computing. The C libraries that we wrap in SciPy include 
+trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry) 
+Cephes\cite{cephes_netlib}. % (mathematics algorithms).
+
+Cython has been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
+as a glue between well-established low-level scientific computing libraries written
+in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
 enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing. Some of the C libraries
-that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
-(iterative solving of the trust region problem),
-SuperLU\cite{li05,superlu_ug99} (solution of large, sparse, nonsymmetric
-systems of linear equations), Qhull\cite{Barber:1996:QAC:235815.235821}
-(computational geometry), and Cephes\cite{cephes_netlib} (mathematics
-algorithms).
-
-Therefore, the relative abundance of different programming languages in the
-SciPy library results from a combination of the usage of powerful performance
-enhancing languages that interact well with Python (i.e., Cython), and the
+loops benefit from a compiled code with static typing.
+
+For implementing new functionality, Python is the still the language of choice. If Python 
+performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
+order). The main motivation for this is maintainability: Cython has the 
+highest abstraction level and most Python developers will understand it. C is 
+also widely known, and easier for the current core development team to manage
+than C++ and especially Fortran.
+
+The distribution of secondary programming languages in SciPy is a compromise between
+a powerful, performance-
+enhancing language that interacts well with Python (i.e. Cython) and the
 usage of languages (and their libraries) that have proven reliable and

### 2019-05-04 scipy-1.0/poly.tex

--- 2019-02-16 scipy-1.0/poly.tex
@@
 At the lowest level of the new design are classes which represent univariate
-piecewise polynomials: \code{PPoly} (SciPy 0.13)\cite{scipy-gh5827},
+piecewise polynomials: \code{PPoly} (SciPy 0.13)\cite{scipy-gh2885},
 \code{BPoly} (SciPy 0.13) and \code{BSpline} (SciPy 0.19)\cite{scipy-gh3174},
@@
 for e.g., constructing Bezier curves). \code{BSpline} represents spline
-curves, i.e., linear combinations of b-spline basis elements.\cite{deBoor1978}
+curves, i.e., linear combinations of B-spline basis elements.\cite{deBoor1978}
 

### 2019-05-04 scipy-1.0/paper.tex

--- 2019-05-03 scipy-1.0/paper.tex
@@
 
+\subsubsection*{Benchmark suite}
+
+The Airspeed Velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing, confirming 
+that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
+
+Consider the (\texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+\label{fig:asvbench}
+\end{figure}
+
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the \texttt{master} branch and the branch a new feature is implemented
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
+
 \section*{Project organization and community}

### 2019-05-04 scipy-1.0/paper.tex
Tightened Benchmark suite section Removed the command to generate the figure. As #117 records, it doesn't "just work", so I don't think it's worth including. Removed the material after the figure about how to use asv. To me it seemed like a short tutorial on how to use asv with SciPy, which seemed out of place. Removed the last part of "allows us to objectively measure that our code base is improving~~, to empower scientific applications.~~" Reworded "generally remains constant or improves over the commit hash history of the project" -> "improves over time." Reworded "the performance of the Sc

@@
 
-The Airspeed Velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
-with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing, confirming 
-that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
-
-Consider the (\texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
+In addition to ensuring that unit tests are passing, it is important to confirm that the
+the performance of the SciPy codebase improves over time. Since February 2015, the
+performance of SciPy has been monitored with Airspeed Velocity (\texttt{asv} \cite{asvref}).
+SciPy's \texttt{run.py} script conveniently wraps \texttt{asv} features such that benchmark 
+results over time can be generated with a single console command. For example,
+Figure~\ref{fig:asvbench} illustrates the improvement of\texttt{scipy.spatial.cKDTree.query}
+over roughly nine years of project history.
 
@@
 \end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the \texttt{master} branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
 

### 2019-05-06 scipy-1.0/paper.tex
update title of Downstream projects section

--- 2019-05-04 scipy-1.0/paper.tex
@@
 Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our
-codebase using the \texttt{linguist} library\cite{linguistref} provides a 
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library\cite{linguistref} provides a 
 detailed breakdown as \% composition by programming language in 
@@
 These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily throughout SciPy and dependent packages.
-
-Over the last three years, our sparse matrix handling internals have been
+and are used heavily in scikit-learn.
+
+Over the last three years our sparse matrix handling internals have been
 rewritten and performance has been improved. Iterating over and slicing of CSC
 and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
+ and the COO / DIA to CSR / CSC matrix format
 conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
@@
 
-\subsubsection*{\texttt{cKDTree}}
-
-\input{ckdtree}
-
 \subsection*{Unified bindings to compiled code}
@@
 \subsection*{Test and benchmark suite}
+
+\subsubsection*{Benchmark suite}
+
+The airspeed velocity (asv\cite{asvref}) library enables benchmarking Python packages
+over their lifetimes, and the performance of the SciPy code base was monitored
+with asv starting in February of 2015 (PR \#4501). In addition to ensuring that
+unit tests are passing (see Test suite section below), confirming 
+that performance generally
+remains constant or improves over the commit hash history of the project allows
+us to objectively measure that our code base is improving, to empower
+scientific applications.
+
+Consider the asv benchmark results shown in Figure~\ref{fig:asvbench}, spanning
+roughly nine years of project history. These demonstrate the gradual
+performance improvements in a nearest-neighbor search through
+\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
+
+\texttt{python run.py run -e -s 800 --bench "\textbackslash
+btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
+
+where \texttt{run.py} is a light wrapper script around \texttt{asv}
+that is packaged with SciPy.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
+\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query()}
+over a roughly nine year commit history time frame. The results are based on
+Python 2.7 performance on the master branch of the project using NumPy 1.8.2
+and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
+compatibility). Only the L2 (Euclidean) norm is shown here, and to improve
+backward compatibility and sampling of the benchmarks there was no application
+of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+\label{fig:asvbench}
+\end{figure}
+
+Any pull request can be compared against the \texttt{master} branch with the
+command \texttt{asv continuous master new-feature}. This will provide a
+benchmark against the master branch and the branch a new feature is implemented
+in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
+arguments to select which benchmarks to run.
 
@@
 
-\subsubsection*{Benchmark suite}
-
-The Airspeed Velocity (\texttt{asv}\cite{asvref}) library enables benchmarking Python packages
-over their lifetimes, and the performance of the SciPy code base was monitored
-with \texttt{asv} starting in February of 2015 (PR \#4501). In addition to ensuring that
-unit tests are passing, confirming 
-that performance generally
-remains constant or improves over the commit hash history of the project allows
-us to objectively measure that our code base is improving, to empower
-scientific applications.
-
-Consider the (\texttt{asv} benchmark results shown in Figure~\ref{fig:asvbench}, spanning
-roughly nine years of project history. These demonstrate the gradual
-performance improvements in a nearest-neighbor search through
-\texttt{scipy.spatial.cKDTree.query()}, and can be run using:
-
-\texttt{python run.py run -e -s 800 --bench "\textbackslash
-btime\_query\textbackslash b" "02de46a546..b3ddb2c"},
-
-where \texttt{run.py} is a light wrapper script around \texttt{asv}
-that is packaged with SciPy.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
-\label{fig:asvbench}
-\end{figure}
-
-Any pull request can be compared against the \texttt{master} branch with the
-command \texttt{asv continuous master new-feature}. This will provide a
-benchmark against the \texttt{master} branch and the branch a new feature is implemented
-in. More features are available in the \texttt{asv} documentation\cite{asvdocs}, including 
-arguments to select which benchmarks to run.
-
 \section*{Project organization and community}
@@
 
-\subsection*{Community beyond the SciPy library}
+\subsection*{Downstream projects}
 
 The scientific Python ecosystem includes many examples
-of domain-specific software libraries building on top
-of SciPy features, and then returning to the base SciPy library
-to suggest and even implement improvements to enable
-more effective science applications. For example, there
-are common contributors to the SciPy and Astropy core
-libraries\cite{astropy-2018}, and what works well for 
-one of the codebases, infrastructures, or communities 
+of domain-specific software libraries that are built 
+on top of SciPy features. The work done on these libraries
+often turns out to be beneficial to the whole ecosystem
+being Scipy the catalyst of this cross-fertilisation process. 
+For example, there are common contributors to the SciPy and 
+Astropy core libraries\cite{astropy-2018}, and what works 
+well for one of the codebases, infrastructures, or communities 
 is often transferred in some form to the other. At the codebase 

### 2019-05-06 scipy-1.0/paper.tex
typo

@@
 often turns out to be beneficial to the whole ecosystem
-being Scipy the catalyst of this cross-fertilisation process. 
+being SciPy the catalyst of this cross-fertilisation process. 
 For example, there are common contributors to the SciPy and 

### 2019-05-13 scipy-1.0/paper.tex
minor rewording / style fixes

@@
 packages, and 28700 dependent repositories. This includes usage of SciPy
-in almost half of all machine learning projects on GitHub, and usage by
-high profile projects including LIGO gravitational wave analysis and creation
-of the first-ever image of a black hole (M87).
+in almost half of all machine learning projects on GitHub.
 The library includes functionality spanning clustering, Fourier transforms,
@@
 \cite{PhysRevD.93.122003, abbott2017gw170817} 
-import several submodules of SciPy, and the M87 black
-hole imaging project directly cites SciPy\cite{2019ApJ875L3E}.
+import several submodules of SciPy.
 
@@
 Over 85,000 GitHub repositories and almost 5000 packages depend
-on SciPy. Some of the major
+on SciPy. SciPy's GitHub repository attracts approximately 20,000 unique 
+visitors every two weeks, and more than 100 people contribute to a 
+new release every six months. Some of the major
 feature highlights from the three years preceding
@@
 
-The \texttt{scipy.stats} module provides more than 100 probability
-distributions and a framework to implement additional distributions. Given an
-expression for a probability distribution function (pdf), the framework will
-calculate moments, cumulative distribution function (cdf), differential
-entropy, and more. Those properties are calculated based on generic code, e.g.,
-numerical integration of the pdf to obtain the cdf.  Where possible, the
-generic code is replaced by explicit formulas or more accurate or efficient
-implementations. For example, most distributions have an explicit formula for
-the cdf, so it is used instead of integrating the pdf. The distributions
-framework also allows the user to draw random variates from the distribution.
-Drawing random variates now accepts a \texttt{random\_state} parameter, which
-is either a \texttt{numpy.random.RandomState} object or an integer used to
-generate such an object, to provide fine-grained user-level control over the
-random state of a workflow.  Currently about 80 continuous univariate
-distributions, 12 discrete univariate distributions and 10 multivariate
-distributions are implemented.  Key recent distributions added to
-\texttt{scipy.stats} include the histogram-based distribution in
-\texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
-\texttt{scipy.stats.multinomial} (used, for example, in natural language
-processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
-the only multivariate discrete distribution available in SciPy.
+The \texttt{scipy.stats} module contains more than 100 probability
+distributions: 96 continuous and 13 discrete univariate distributions
+and 10 multivariate distributions. The implementation relies on a
+consistent framework that provides methods to sample random variates,
+to evaluate the cumulative distribution function (cdf) and the probability
+density function (pdf) and to fit parameters for every distribution. 
+Generally, the methods rely on specific implementations for each
+distribution such as a closed form expression of the cdf or a sampling 
+algorithm, if available. Otherwise, default methods are used
+based on generic code, e.g., numerical integration of the pdf to
+obtain the cdf. 
+Key recent distributions added to \texttt{scipy.stats} include the
+histogram-based distribution in \texttt{scipy.stats.rv\_histogram}
+and the multinomial distribution in \texttt{scipy.stats.multinomial}
+(used, for example, in natural language processing, see
+\cite{Griffiths5228}).
 
@@
 
-\subsection*{Downstream projects}
+\subsection*{Community beyond the SciPy library}
 
 The scientific Python ecosystem includes many examples
-of domain-specific software libraries that are built 
-on top of SciPy features. The work done on these libraries
-often turns out to be beneficial to the whole ecosystem
-being SciPy the catalyst of this cross-fertilisation process. 
-For example, there are common contributors to the SciPy and 
-Astropy core libraries\cite{astropy-2018}, and what works 
-well for one of the codebases, infrastructures, or communities 
-is often transferred in some form to the other. At the codebase 
-level, the \texttt{binned\_statistic} functionality
+of domain-specific software libraries building on top
+of SciPy features, and then returning to the base SciPy library
+to suggest and even implement improvements to enable
+more effective science applications. For example, there
+are common contributors to the SciPy and astropy core
+libraries\cite{astropy-2018}, and what works well for 
+one of the code bases, infrastructures, or communities 
+is often transferred in some form to the other. At the code
+base level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution---it was initially
 developed in an Astropy-affiliated package
-and then placed in SciPy afterwards.
+and then placed in SciPy after the fact.
 
@@
 metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
-that roughly two-week period. The developer community at that time consisted of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the codebase
+14 and May 27, 2018, with 721 unique copies (``clones") of the code base over
+that roughly two-week period. The developer community consists of 610 unique
+contributors of source code, with $>19,000$ commits accepted into the code base
 (GitHub page data).
@@
 scientific collaboration that lead to the observation of gravitational waves
-\cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
+\cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly with
+MacOS and in the Intel Distribution for Python\cite{intel-python}. SciPy is used
 by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
+
+\subsection*{Future Work}
 

### 2019-05-13 scipy-1.0/subpackages.tex
minor rewording / style fixes

--- 2019-03-01 scipy-1.0/subpackages.tex
@@
     variates, and more;
-    statistical tests, including Pearson's correlation, Spearman's rank-order
-    correlations, Kendall's tau, chi-squared test and its generalization as the
-    Cressie-Read power divergence, contingency table tests including Fisher's
-    exact test and Mood's median test, and many more;
-    and assorted transformations and statistics of data.
+    statistical tests, including tests on equality of means/medians/variance
+    (such as the t-test) and tests whether a sample is drawn from a
+    certain distribution (such as the Kolmogorov-Smirnov test);
+    measures of correlation, including Pearson's $r$, Kendall's $\tau$, and
+    Spearman's $\rho$ coefficients;
+    descriptive statistics including trimmed values;
+    kernel density estimation;
+    and transformations of data such as the Box-Cox power transformation.
 \end{description}

### 2019-05-13 scipy-1.0/paper.tex
Revised description of stats module (#125)
* revised description of stats module
* add comment on features recently added
* minor rewording / style fixes

@@
 packages, and 28700 dependent repositories. This includes usage of SciPy
-in almost half of all machine learning projects on GitHub.
+in almost half of all machine learning projects on GitHub, and usage by
+high profile projects including LIGO gravitational wave analysis and creation
+of the first-ever image of a black hole (M87).
 The library includes functionality spanning clustering, Fourier transforms,
@@
 These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily in scikit-learn.
-
-Over the last three years our sparse matrix handling internals have been
+and are used heavily throughout SciPy and dependent packages.
+
+Over the last three years, our sparse matrix handling internals have been
 rewritten and performance has been improved. Iterating over and slicing of CSC
 and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
- and the COO / DIA to CSR / CSC matrix format
+ and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
 conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,

### 2019-05-13 scipy-1.0/paper.tex
Attempt to incorporate @ilayn's comments
"I think being a maintainer is a bit more long-term and involved (still of course anyone can become) so we might not want to paint the wrong picture about that here." - @ilayn
"You can start with the definition of a maintainer and then switch to SciPy specific details. Also approximately 15 is in fact exactly 15" @ilayn
I didn't want to lead with a separate definition; I think it is more compact and immediately relevant if the definition is provided in-line. I do prefer a colon over the em dash, though, if only to ovoid overuse of the em dash.

--- 2019-05-06 scipy-1.0/paper.tex
@@
 
-\subsection*{Downstream projects}
+\subsection*{Maintainers and contributors}
+
+The SciPy project has approximately 100 unique contributors
+for every 6-monthly release cycle. Anyone with the interest and
+skills can become a contributor; the SciPy Developer
+Guide\cite{scipy-dev-guide} provides guidance on how to do that.
+In addition, the project currently has 15 active maintainers: people who review
+the contributions of others and do everything else needed to ensure that the
+software and the project move forward. Maintainers are critical to the health
+of the project\cite{eghbal2016}; their skills and efforts largely determine how
+fast the project moves forward, and they enable input from the much
+larger group of contributors. Anyone can become a maintainer, too, as they
+are selected on a rolling basis from contributors with a significant history of
+high-quality contributions.
+
+\subsection*{Community beyond the SciPy library}
 
 The scientific Python ecosystem includes many examples
-of domain-specific software libraries that are built 
-on top of SciPy features. The work done on these libraries
-often turns out to be beneficial to the whole ecosystem
-being SciPy the catalyst of this cross-fertilisation process. 
-For example, there are common contributors to the SciPy and 
-Astropy core libraries\cite{astropy-2018}, and what works 
-well for one of the codebases, infrastructures, or communities 
+of domain-specific software libraries building on top
+of SciPy features, and then returning to the base SciPy library
+to suggest and even implement improvements to enable
+more effective science applications. For example, there
+are common contributors to the SciPy and Astropy core
+libraries\cite{astropy-2018}, and what works well for 
+one of the codebases, infrastructures, or communities 
 is often transferred in some form to the other. At the codebase 
@@
 
-\subsection*{Maintainers and contributors}
-
-The SciPy project has approximately 15 active maintainers---people who review
-others' contributions, and in general do everything needed to ensure that the
-software and the project move forward. Maintainers are critical to the health
-of the project\cite{eghbal2016}; their skills and efforts largely determine how
-fast the project moves forward, and they enable contributions from a much
-larger group. Accordingly, the project also has an additional ~100 unique
-contributors for every 6-monthly release cycle. Anyone with the interest and
-skills can become a contributor or maintainer; the SciPy Developer
-Guide\cite{scipy-dev-guide} provides guidance on how to do that.
-
 \section*{Discussion}
 
-% The Discussion should be succinct and must not contain subheadings.
+\fixme{The Discussion should be succinct and must not contain subheadings.}
+
+\subsection*{Impact now}
+
+% While Paul was one of the early adopters at the national laboratories,
+% Python's presence is now ubiquitous.
 

### 2019-05-14 scipy-1.0/paper.tex
Swap subsections in Project Organization (#128)
* Swap subsections in Project Organization
The "Governance" and "Maintainers and contributors" subsections are focused on SciPy itself (internal). "Commmunity beyond the SciPy library" is about relationships with other projects (external), but its sandwiched between the other two internal sections. I changed the order of subsections to "Governance", "Maintainers and contributors", and "Community beyond the SciPy library"; they are now in order from closest to the core of the project to further away.

* Attempt to incorporate @ilayn's comments

"I

@@
 
+\subsection*{Maintainers and contributors}
+
+The SciPy project has approximately 100 unique contributors
+for every 6-monthly release cycle. Anyone with the interest and
+skills can become a contributor; the SciPy Developer
+Guide\cite{scipy-dev-guide} provides guidance on how to do that.
+In addition, the project currently has 15 active maintainers: people who review
+the contributions of others and do everything else needed to ensure that the
+software and the project move forward. Maintainers are critical to the health
+of the project\cite{eghbal2016}; their skills and efforts largely determine how
+fast the project moves forward, and they enable input from the much
+larger group of contributors. Anyone can become a maintainer, too, as they
+are selected on a rolling basis from contributors with a significant history of
+high-quality contributions.
+
 \subsection*{Community beyond the SciPy library}
@@
 and then placed in SciPy afterwards.
-
-\subsection*{Maintainers and contributors}
-
-The SciPy project has approximately 15 active maintainers---people who review
-others' contributions, and in general do everything needed to ensure that the
-software and the project move forward. Maintainers are critical to the health
-of the project\cite{eghbal2016}; their skills and efforts largely determine how
-fast the project moves forward, and they enable contributions from a much
-larger group. Accordingly, the project also has an additional ~100 unique
-contributors for every 6-monthly release cycle. Anyone with the interest and
-skills can become a contributor or maintainer; the SciPy Developer
-Guide\cite{scipy-dev-guide} provides guidance on how to do that.
 

### 2019-05-14 scipy-1.0/ckdtree.tex

@@
-The \texttt{cKDTree} module was rewritten in C++ with templated classes, and support for
-periodic boundary conditions was added. A periodic boundary condition is typically 
-used in computer simulations of physical processes.
+The \texttt{cKDTree} module, which implements a space-partitioning data structure that
+organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 
+Support was added for periodic boundary conditions, which are often used 
+in simulations of physical processes.
 

### 2019-05-14 scipy-1.0/paper.tex

@@
 \usepackage[multiple]{footmisc}
+\usepackage{booktabs,siunitx}
 
@@
 
-Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
-in SciPy. While it is not surprising that Python is predominant, the usage distribution 
-of the other languages warrants discussion.
-
-\begin{figure}[H]
+Python, Cython, Fortran, C and C++ are the programming languages used to
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library\cite{linguistref} provides a 
+detailed breakdown as \% composition by programming language in 
+SciPy (Table~\ref{tab:linguist}).
+
+\begin{table}
 \centering
-		\includegraphics[width=\textwidth]{static/composition}
-    \caption{The breakdown of programming languages used in the
-             SciPy library determined using the \texttt{linguist} library.
-    	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
-    	 and Makefile are excluded for clarity and mostly
-    	 provide supporting roles in tests, building, and
-    	 documentation.}
-    \label{fig:linguist}
-\end{figure}
+\begin{tabular}{p{3cm}S[table-format = <0.1]}
+\toprule
+Language & {Percent}\\
+\midrule
+Python  & 49.5 \\
+Fortran & 25.6 \\
+C       & 19.5 \\
+Cython  & 3.0  \\
+C++     & 2.3  \\[\defaultaddspace]
+{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
+\bottomrule
+\end{tabular}
+\caption{Language composition of SciPy codebase: lines of code in each
+programming language as determined by the \texttt{linguist} package.
+The last row denotes tools used in supporting roles in tests,
+building, and documentation.}
+\label{tab:linguist}
+\end{table}
 

### 2019-05-14 scipy-1.0/scipy-optimize.tex
Updates to address @apbard's comments

@@
 Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
-All are trust-region methods that build a local model of the objective function based on first- and second-derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
-For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second-derivative Hessian matrix to be stored and factored every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
-On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second-derivative information in a faster, approximate way.
+All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
+For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second derivative Hessian matrix to be stored and factored every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
+On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
 A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}.
@@
 \paragraph{Global Minimization}
-As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+As \texttt{minimize} may return any local minimum, some problems require the use of a global optimization routine.
 The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by evolving a population of candidate solutions.
@@
 If the trial candidates represent an improvement, then the population is updated.
-Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits inclusion in the package.
+Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits their inclusion in the package.

### 2019-05-15 scipy-1.0/paper.tex
reword downstream projects section

--- 2019-05-14 scipy-1.0/paper.tex
@@
 
-Figure~\ref{fig:linguist} provides a breakdown of the programming languages used
-in SciPy. While it is not surprising that Python is predominant, the usage distribution 
-of the other languages warrants discussion.
+Python, Cython, Fortran, C and C++ are the programming languages used to
+implement scientific algorithms in the SciPy library. An analysis of our code
+base using the \texttt{linguist} library\cite{linguistref} provides a 
+detailed breakdown as \% composition by programming language in 
+SciPy (Figure~\ref{fig:linguist}).
 
@@
     \caption{The breakdown of programming languages used in the
-             SciPy library determined using the \texttt{linguist} library.
+             SciPy library determined using the linguist library.
     	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
@@
 
-Fortran, despite its age, is still a high-performance scientific programming language with 
-continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
-libraries in order to provide Python convenience while benefiting from their performance: 
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
-ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
-MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
-ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
-CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
+For implementing new functionality, we have a clear order of language
+preference.  First Python, if performance is not an issue. If it is, then in
+order of preference: Cython, C, C++, Fortran. The main motivation for this is
+maintainability: Cython has the highest abstraction level and most Python
+developers will understand it. C is also widely known, and easier to deal with
+for the core development team than C++ and especially Fortran.
+
+While it is not surprising that Python is heavily used in SciPy, the usage
+distribution of the other languages warrants some discussion. Fortran is an
+extremely well-established scientific programming language, both for historical
+reasons and because of its continued excellent
+performance\cite{Koelbel:1993:HPF:562354}. We wrap the FFTPACK Fortran library
+for performing Fourier transforms\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}
+since this library has been a standard in the field for 33 years and has a
+license that is compatible with our own. Likewise, we wrap the Fortran source
+for ODEPACK\cite{citeulike:2644528} as it has been  trusted for the initial
+value problem for ordinary differential equation systems for 30 years. For
+similar reasons we also wrap the Fortran libraries
+QUADPACK\cite{1983qspa.book.....P} (for numerical integration of
+one-dimensional functions), FITPACK\cite{Dierckx:1993:CSF:151103}
+(curve-fitting / interpolation), ODRPACK\cite{ODRPACK_Boggs} (orthogonal
+distance regression), MINPACK\cite{osti_6997568} (minimization of linear and
+nonlinear equations), ARPACK\cite{leh:sor:yan96} (solving large scale
+eigenvalue problems), ALGORITHM 644\cite{Amos:1986:APP:7921.214331} (handling
+Bessel Functions), and CDFLIB\cite{CDFLIB_site} (evaluation of cumulative
+density functions).
 
@@
 well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. The C libraries that we wrap in SciPy include 
-trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
-SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
-Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry) 
-Cephes\cite{cephes_netlib}. % (mathematics algorithms).
-
-Cython has been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
-as a glue between well-established low-level scientific computing libraries written
-in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
+scientific computing. Alongside its objected-oriented relative C++, C can be
+leveraged in Python after being exposed using the Cython language. Cython has
+been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We thus often use Cython
+as a glue between well-established low-level scientific computing libraries and
+the Python interface offered by SciPy. We also use Cython to enable performance
 enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing.
-
-For implementing new functionality, Python is the still the language of choice. If Python 
-performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
-order). The main motivation for this is maintainability: Cython has the 
-highest abstraction level and most Python developers will understand it. C is 
-also widely known, and easier for the current core development team to manage
-than C++ and especially Fortran.
-
-The distribution of secondary programming languages in SciPy is a compromise between
-a powerful, performance-
-enhancing language that interacts well with Python (i.e. Cython) and the
+loops benefit from a compiled code with static typing. Some of the C libraries
+that we wrap in SciPy include trlib\cite{doi:10.1080/10556788.2018.1449842}
+(iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99} (solution of large, sparse, nonsymmetric
+systems of linear equations), Qhull\cite{Barber:1996:QAC:235815.235821}
+(computational geometry), and Cephes\cite{cephes_netlib} (mathematics
+algorithms).
+
+Therefore, the relative abundance of different programming languages in the
+SciPy library results from a combination of the usage of powerful performance
+enhancing languages that interact well with Python (i.e., Cython), and the
 usage of languages (and their libraries) that have proven reliable and
@@
 
-The \texttt{scipy.stats} module contains more than 100 probability
-distributions: 96 continuous and 13 discrete univariate distributions
-and 10 multivariate distributions. The implementation relies on a
-consistent framework that provides methods to sample random variates,
-to evaluate the cumulative distribution function (cdf) and the probability
-density function (pdf) and to fit parameters for every distribution. 
-Generally, the methods rely on specific implementations for each
-distribution such as a closed form expression of the cdf or a sampling 
-algorithm, if available. Otherwise, default methods are used
-based on generic code, e.g., numerical integration of the pdf to
-obtain the cdf. 
-Key recent distributions added to \texttt{scipy.stats} include the
-histogram-based distribution in \texttt{scipy.stats.rv\_histogram}
-and the multinomial distribution in \texttt{scipy.stats.multinomial}
-(used, for example, in natural language processing, see
-\cite{Griffiths5228}).
+The \texttt{scipy.stats} module provides more than 100 probability
+distributions and a framework to implement additional distributions. Given an
+expression for a probability distribution function (pdf), the framework will
+calculate moments, cumulative distribution function (cdf), differential
+entropy, and more. Those properties are calculated based on generic code, e.g.,
+numerical integration of the pdf to obtain the cdf.  Where possible, the
+generic code is replaced by explicit formulas or more accurate or efficient
+implementations. For example, most distributions have an explicit formula for
+the cdf, so it is used instead of integrating the pdf. The distributions
+framework also allows the user to draw random variates from the distribution.
+Drawing random variates now accepts a \texttt{random\_state} parameter, which
+is either a \texttt{numpy.random.RandomState} object or an integer used to
+generate such an object, to provide fine-grained user-level control over the
+random state of a workflow.  Currently about 80 continuous univariate
+distributions, 12 discrete univariate distributions and 10 multivariate
+distributions are implemented.  Key recent distributions added to
+\texttt{scipy.stats} include the histogram-based distribution in
+\texttt{scipy.stats.rv\_histogram} and the multinomial distribution in
+\texttt{scipy.stats.multinomial} (used, for example, in natural language
+processing\cite{Griffiths5228}). \texttt{multinomial} is the first and so far
+the only multivariate discrete distribution available in SciPy.
 
@@
 
-\subsubsection*{Benchmark suite}
-
-In addition to ensuring that unit tests are passing, it is important to confirm that the
-the performance of the SciPy codebase improves over time. Since February 2015, the
-performance of SciPy has been monitored with Airspeed Velocity (\texttt{asv} \cite{asvref}).
-SciPy's \texttt{run.py} script conveniently wraps \texttt{asv} features such that benchmark 
-results over time can be generated with a single console command. For example,
-Figure~\ref{fig:asvbench} illustrates the improvement of\texttt{scipy.spatial.cKDTree.query}
-over roughly nine years of project history.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
-\label{fig:asvbench}
-\end{figure}
-
 \section*{Project organization and community}
@@
 
-\subsection*{Community beyond the SciPy library}
+\subsection*{Downstream projects}
 
@@
 of domain-specific software libraries building on top
-of SciPy features, and then returning to the base SciPy library
-to suggest and even implement improvements to enable
-more effective science applications. For example, there
-are common contributors to the SciPy and Astropy core
-libraries\cite{astropy-2018}, and what works well for 
-one of the codebases, infrastructures, or communities 
+of SciPy features and then returning to the base SciPy library
+to suggest and even implement improvements. 
+For example, there are common contributors to the SciPy and 
+Astropy core libraries\cite{astropy-2018}, and what works 
+well for one of the codebases, infrastructures, or communities
 is often transferred in some form to the other. At the codebase
-level, the \texttt{binned\_statistic} functionality
-is one such cross-project contribution---it was initially
+level, the \texttt{binned_statistic} functionality
+is one such cross-project contribution: it was initially
 developed in an Astropy-affiliated package
 and then placed in SciPy afterwards. 
+In this perspective, SciPy serves as a catalyst for cross-fertilisation 
+throughout the Python scientific computing community.
 

### 2019-05-15 scipy-1.0/paper.tex
escape underscore

@@
 is often transferred in some form to the other. At the codebase
-level, the \texttt{binned_statistic} functionality
+level, the \texttt{binned\_statistic} functionality
 is one such cross-project contribution: it was initially

### 2019-05-15 scipy-1.0/paper.tex

--- 2019-05-14 scipy-1.0/paper.tex
@@
 of domain-specific software libraries building on top
-of SciPy features, and then returning to the base SciPy library
-to suggest and even implement improvements to enable
-more effective science applications. For example, there
-are common contributors to the SciPy and Astropy core
-libraries\cite{astropy-2018}, and what works well for 
-one of the codebases, infrastructures, or communities 
+of SciPy features and then returning to the base SciPy library
+to suggest and even implement improvements. 
+For example, there are common contributors to the SciPy and 
+Astropy core libraries\cite{astropy-2018}, and what works 
+well for one of the codebases, infrastructures, or communities
 is often transferred in some form to the other. At the codebase
 level, the \texttt{binned\_statistic} functionality
-is one such cross-project contribution---it was initially
+is one such cross-project contribution: it was initially
 developed in an Astropy-affiliated package
 and then placed in SciPy afterwards. 
+In this perspective, SciPy serves as a catalyst for cross-fertilisation 
+throughout the Python scientific computing community.
 

### 2019-05-16 scipy-1.0/scipy-optimize.tex
Implement @apbard suggestions
I also changed the several methods of finding local to several methods for finding local, isn't this more correct?

Do you think something like The \texttt{minimize} provides a common (or shared) interface for several methods of... solves the issue? If not, your version it is fine.

@@
 \paragraph{Local Minimization}
-The \texttt{minimize} function unifies several methods of finding local minima of nonlinear optimization problems in a common interface.
+The \texttt{minimize} function provides a unified interface for finding local minima of nonlinear optimization problems.
 Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.

### 2019-05-16 scipy-1.0/scipy-optimize.tex
Update per @andyfaff comments
@andyfaff "'The new _diffev_ function is a ...'
I think this is a matter of preference, but easy enough to make the change.

@@
 As \texttt{minimize} may return any local minimum, some problems require the use of a global optimization routine.
-The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by evolving a population of candidate solutions.
+The new \texttt{scipy.optimize.differential\textunderscore evolution} function \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by evolving a population of candidate solutions.
 In each iteration, trial candidates are generated by combination of candidates from the existing population.

### 2019-05-16 scipy-1.0/paper.tex
Added citations to SciPy begins

--- 2019-05-15 scipy-1.0/paper.tex
@@
 functions, sparse matrices, quadrature, optimization, Fast Fourier
-Transforms, and more.  One of these packages, Multipack, was a set of
+Transforms, and more.  One of these packages, Multipack\cite{multipack}, was a set of
 extension modules that wrapped Fortran and C libraries such as
-ODEPACK, QUADPACK, and MINPACK, to solve and minimize nonlinear
+ODEPACK\cite{citeulike:2644528}, QUADPACK\cite{1983qspa.book.....P}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
 equations, integrate differential equations, and fit splines (the
@@
 of Chicago, liked the plotting functionality
-available in MATLAB, but had problems accessing their laboratory's
+available in MATLAB\cite{matlab}, but had problems accessing their laboratory's
 license, which was governed by a hardware dongle.  In response, he
@@
 STScI.  With the Numeric maintainer's blessing, the decision was made
-to write NumArray, a library that could handle data on a larger
+to write NumArray\cite{greenfield2003numarray}, a library that could handle data on a larger
 scale.  Unfortunately, NumArray proved inefficient for small arrays,
@@
 
-\subsection*{Downstream projects}
+\subsection*{Community beyond the SciPy library}
 
@@
 of domain-specific software libraries building on top
-of SciPy features and then returning to the base SciPy library
-to suggest and even implement improvements. 
-For example, there are common contributors to the SciPy and 
-Astropy core libraries\cite{astropy-2018}, and what works 
-well for one of the codebases, infrastructures, or communities
+of SciPy features, and then returning to the base SciPy library
+to suggest and even implement improvements to enable
+more effective science applications. For example, there
+are common contributors to the SciPy and Astropy core
+libraries\cite{astropy-2018}, and what works well for 
+one of the codebases, infrastructures, or communities 
 is often transferred in some form to the other. At the codebase 
-level, the \texttt{binned_statistic} functionality
-is one such cross-project contribution: it was initially
+level, the \texttt{binned\_statistic} functionality
+is one such cross-project contribution---it was initially
 developed in an Astropy-affiliated package
 and then placed in SciPy afterwards.
-In this perspective, SciPy serves as a catalyst for cross-fertilisation 
-throughout the Python scientific computing community.
 

### 2019-05-16 scipy-1.0/ckdtree.tex
Attempt to make use of (sub)package/module consistent

@@
-The \texttt{cKDTree} module, which implements a space-partitioning data structure that
+The \texttt{scipy.spatial.cKDTree} subpackage, which implements a space-partitioning data structure that
 organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 

### 2019-05-16 scipy-1.0/paper.tex
Attempt to make use of (sub)package/module consistent

--- 2019-05-15 scipy-1.0/paper.tex
@@
 \cite{PhysRevD.93.122003, abbott2017gw170817} 
-import several submodules of SciPy, and the M87 black
+import several subpackages of SciPy, and the M87 black
 hole imaging project directly cites SciPy\cite{2019ApJ875L3E}.
@@
 
-In addition to the Python API, SciPy has C and Cython interfaces in a number
-of submodules. Therefore, we have to also consider the application binary
+In addition to the Python API, SciPy has C and Cython interfaces.
+Therefore, we have to also consider the application binary
 interface (ABI). This ABI has been stable for a long time, and we aim to
@@
 many BLAS and LAPACK routines (added in 2015) and the special functions 
-provided in the \texttt{scipy.{\allowbreak}special} submodule (added in 2016).  
-These Cython wrappers are available in the modules
+provided in the \texttt{scipy.{\allowbreak}special} subpackage (added in 2016).
+These Cython wrappers are available in 
 \texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},
@@
 the functions in \texttt{scipy.{\allowbreak}special} follow corresponding
-changes to the interface of that submodule.
+changes to the interface of that subpackage.
 
@@
 
-The \texttt{scipy.stats} module contains more than 100 probability
+The \texttt{scipy.stats} subpackage contains more than 100 probability
 distributions: 96 continuous and 13 discrete univariate distributions
@@
 \begin{tabular}{ c|c }
-    SciPy submodule & Summary of Change \\
+    SciPy subpackage & Summary of Change \\
     \hline
@@
 \section*{Numerical Optimization: Additional Details}
-The \texttt{scipy.optimize} module provides functions for the numerical solution of several classes of root finding and optimization problems:
+The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
 \begin{enumerate}

### 2019-05-16 scipy-1.0/scipy-optimize.tex
Attempt to make use of (sub)package/module consistent

--- 2019-05-16 scipy-1.0/scipy-optimize.tex
@@
 \newcommand{\RR}{\ensuremath{\mathbb{R}}}
-The \texttt{scipy.optimize} module provides functions for the numerical
+The \texttt{scipy.optimize} subpackage provides functions for the numerical
 solution of several classes of root finding and optimization problems.
@@
 \paragraph{Local Minimization}
-The \texttt{minimize} function provides a unified interface for finding local minima of nonlinear optimization problems.
-Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
-All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
-For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second derivative Hessian matrix to be stored and factored every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
-On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
-A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}.
+
+The \texttt{minimize} function provides methods for finding local minima of
+nonlinear optimization problems. It unifies several methods with a common 
+interface in order to facilitate interchanging between solvers with different
+characteristics. Characteristics of the different methods are summarized in 
+Table~\ref{tab:minimize-si}.
+%% Currently, this introduces a ton of whitespace around a very simple expression (min f(x)).
+The reader can easily find the standard of meaning of `nonlinear programming', so I believe it should be removed.
+If we decide that it is appropriate to include, then we should add the entire problem statement - including constraints - and we should also include the entire linear programming problem statement in the previous section.
+% summarizes all the methods available for solving minimization problems of the type:
+%\begin{equation}
+%  \label{eq:minimization-prob}
+%  \min_x f(x)
+%\end{equation}
+%where $f$ is a multivariate function $f: \mathbb{R}^n \rightarrow \mathbb{R}$ and $x \in \mathbb{R}^n$.
+%%  Without the problem statement this sentence isn't really appropriate.
+%\texttt{COBYLA} and \texttt{SLSQP} methods support nonlinear constraints, \texttt{L-BFGS-B} and \texttt{TNC} allow box constraints $x^l \le x \le x^u$,
+%and the remaining methods are for unconstrained problems.
+
+% The available methods implement a variety of solution strategies.
+Derivative-free methods (e.g. \texttt{Nelder-Mead}, \texttt{Powell} and \texttt{COBYLA}) are appropriate for minimizing non-differentiable or
+noisy objective functions. When derivatives are available, however, other methods offer faster convergence rates
+and have proven convergence properties. 
+The derivative-based methods start with an initial guess and refine it iteratively, seeking
+a local solution to the optimization problem. Each of these algorithms implements one of two complementary strategies:
+line-search or trust-region. At each iteration, the line-search (LS) methods choose a direction
+and approximate the optimal point in the given direction. Trust-region (TR) methods, on the other
+hand, choose a ``trust-region'' region for which a local model is valid and approximate the best point
+inside.
+
+One recently-added trust-region method is \texttt{trust-exact},
+which achieves fast convergence by solving the trust-region subproblem almost exactly.
+Unfortunately, it requires the Hessian to be stored and factored every iteration, which may preclude
+the solution of large problems (e.g.~problems with more than 1000 variables).
+Other methods are appropriate for large-scale problems, including
+\texttt{CG}, which does not use second derivatives and thus has low memory requirements, and \texttt{L-BFGS-B}, which maintains a simple and compact representation of the Hessian.
+\texttt{Newton-CG}, \texttt{trust-ncg} and \texttt{trust-krylov} use second order derivatives to achieve faster
+convergence, but are still appropriate for large problems because they factorize the Hessian in an inexact and iterative way that does not require the Hessian to be stored.
+The method \texttt{trust-krylov}, introduced in the latest release, is a good compromise between
+\texttt{trust-ncg} and \texttt{trust-exact}: it provides a more accurate solution to the trust-region
+subproblem than \texttt{trust-ncg}, but it does not require the Hessian to be stored and factored as in \texttt{trust-exact}.
+
+While most \texttt{minimize} development prior to SciPy 1.0 has focused on unconstrained minimization, 
+% methods \texttt{SLSQP} and \texttt{COBYLA}\footnote{We note that any equality constraint $c_{\text{eq}}(x) = 0$ can be expressed as two inequality constraints $c_{\text{eq}}(x) \leq 0$ and $-c_{\text{eq}}(x) \leq 0$}
+\texttt{SLSQP} supports fully general nonlinear programming.
+However, it does not provide special treatment for sparsity, limiting its utility to small problems.
+Accordingly, ongoing development efforts for \texttt{minimize} are focused on the addition of nonlinear programming algorithms for large, sparse problems.
 
 \paragraph{Global Minimization}
-As \texttt{minimize} may return any local minimum, some problems require the use of a global optimization routine.
-The new \texttt{scipy.optimize.differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic global optimizer that works by evolving a population of candidate solutions.
-In each iteration, trial candidates are generated by combination of candidates from the existing population.
-If the trial candidates represent an improvement, then the population is updated.
-Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits their inclusion in the package.
+As \texttt{minimize} only seeks a local minimum, some problems require the use of a global optimization routine to find the minimum of the objective function over the set of all decision variable values, especially if the objective function possesses many local minima that other solvers could get stuck in.
+% e.g.
+if there is concern that `minimize` could get stuck in an unacceptable local minimum?
+The \texttt{brute} function performs a grid search over the entire domain, and the best grid point is ``polished'' by a local minimizer.
+However, this approach becomes prohibitively expensive as the problem dimensionality increases.
+% As far as I understand how brute works, the run time does not depend on the number of extrema, as it only polishes the best one
+% , or if the objective function has many extrema. 
+To address this, the
+\texttt{basinhopping} \cite{Wales1997} function uses random perturbation of the decision variables in conjunction with local minimization.
+The acceptance of the perturbed coordinates is dictated by the Metropolis test typically used in Monte Carlo simulations.
+\texttt{differential\textunderscore evolution} \cite{Wormington1999,Storn1997} is a stochastic solver that improves by gradual evolution of a population of candidate solutions.
+In each iteration, trial candidates are generated by combination of candidates from the existing population; if the trial candidates represent an improvement then the population is updated.
+% Is the following a unique feature of differential_evolution?
+If so, state it (although it seems that neither `brute`or `basinhopping` inherently require the a differentiable problem).
+If not, why mention it?
+% \texttt{differential\textunderscore evolution} does not require that the objective function be differentiable.
+%  basinhopping may need differentiable objective for the local minimization.
+Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems
+% They're ``comprehensive''
+% of varying difficulties, 
+% that can be used to evaluate the performance of the solvers. 
+%This suite is used to decide
+for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits inclusion in the package. 
+% have good enough performance to be added to the package.

### 2019-05-16 scipy-1.0/subpackages.tex
Attempt to make use of (sub)package/module consistent

--- 2019-03-01 scipy-1.0/subpackages.tex
@@
 \item[\texttt{cluster}]
-    The \texttt{cluster} subpackage contains two modules:
-    \texttt{cluster.vq} provides vector quantization and $k$-means algorithms,
-    and \texttt{cluster.hierarchy} provides functions for hierarchical and
+    The \texttt{cluster} subpackage contains \texttt{cluster.vq}, which provides vector quantization and $k$-means algorithms, and \texttt{cluster.hierarchy}, which provides functions for hierarchical and
     agglomerative clustering.
@@
 \item[\texttt{ndimage}]
-    This package contains various functions for multi-dimensional image
+    This subpackage contains various functions for multi-dimensional image
     processing, including convolution and assorted linear and nonlinear
@@
     While this subpackage still exists in SciPy 1.0, effort is underway
-    to deprecate or relocate the contents of this subpackage and eventually remove it.
+    to deprecate or relocate the contents of this subpackage and remove it.
 \item[\texttt{odr}]
@@
 \item[\texttt{optimize}]
-    The package includes the following, with additional details in the SI:
+    This subpackage includes the following, with additional details in the SI:
     simplex and interior-point linear programming solvers,
@@
 \item[\texttt{sparse}]
-    This package includes implementations of several representations of
-    sparse matrices.  It contains two modules, 
-    \texttt{scipy.sparse.linalg} and \texttt{scipy.sparse.csgraph}.
-
-    \texttt{scipy.sparse.linalg} provides a collection of linear algebra
-    routines that work with sparse matrices, including linear equation
-    solvers, eigenvalue decomposition, singular value decomposition
-    and LU factorization.
-
+    This subpackage includes implementations of several representations of
+    sparse matrices.\newline % \texttt{scipy.sparse.linalg} extends past margin
+		\texttt{scipy.sparse.linalg} provides a collection of linear algebra routines that work with
+		sparse matrices, including linear equation solvers, eigenvalue decomposition, singular value
+		decomposition and LU factorization.
     \texttt{scipy.sparse.csgraph} provides a collections of graph algorithms
@@
     including the $k$-d tree, Delaunay triangulation, convex hulls and Voronoi
-    diagrams.  The subpackage \texttt{scipy.spatial.distance} provides
+    diagrams. \texttt{scipy.spatial.distance} provides
     a large collection of distance functions, along with functions for

### 2019-05-16 scipy-1.0/paper.tex
Added number of downloads via conda

--- 2019-05-16 scipy-1.0/paper.tex
@@
 functions, sparse matrices, quadrature, optimization, Fast Fourier
-Transforms, and more.  One of these packages, Multipack\cite{multipack}, was a set of
+Transforms, and more.  One of these packages, Multipack, was a set of
 extension modules that wrapped Fortran and C libraries such as
-ODEPACK\cite{citeulike:2644528}, QUADPACK\cite{1983qspa.book.....P}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
+ODEPACK, QUADPACK, and MINPACK, to solve and minimize nonlinear
 equations, integrate differential equations, and fit splines (the
@@
 of Chicago, liked the plotting functionality
-available in MATLAB\cite{matlab}, but had problems accessing their laboratory's
+available in MATLAB, but had problems accessing their laboratory's
 license, which was governed by a hardware dongle.  In response, he
@@
 STScI.  With the Numeric maintainer's blessing, the decision was made
-to write NumArray\cite{greenfield2003numarray}, a library that could handle data on a larger
+to write NumArray, a library that could handle data on a larger
 scale.  Unfortunately, NumArray proved inefficient for small arrays,
@@
 From the user side, there were 13,096,468 downloads of SciPy from the Python
-Packaging Index (PyPI) during the year 2017\cite{pypinfo}, which is a lower
-bound on the number of total downloads by users given that downloading from
-PyPI is only one of several popular methods to install SciPy.  The SciPy
+Packaging Index (PyPI)\cite{pypinfo} and 5,776,017 via the default channel of the
+\texttt{conda}\cite{condainfo} package manager during the year 2017. These numbers establish a lower
+bound on the total number of downloads by users given that
+PyPI and \texttt{conda} are only two of several popular methods for installing SciPy.  The SciPy
 website\cite{SciPylib}, which has been the default citation in the absence of a
-peer-reviewed paper, has been cited $>3000$ times. Some of the most prominent
+peer-reviewed paper, has been cited $>3000$ times\cite{googlescholar}. Some of the most prominent
 usages of or demonstrations of credibility for SciPy include the LIGO / Virgo

### 2019-05-17 scipy-1.0/ckdtree.tex

--- 2019-05-14 scipy-1.0/ckdtree.tex
@@
-The \texttt{cKDTree} module, which implements a space-partitioning data structure that
+The \texttt{scipy.spatial.ckdtree} module, which implements a space-partitioning data structure that
 organizes points in $k$-dimensional space, was rewritten in C++ with templated classes. 

### 2019-05-17 scipy-1.0/paper.tex
Combine potentially redundant information about STScI transition to Python

@@
 
-Initial work on numerical computing in Python was driven by graduate
-students, but soon larger research labs became increasingly engaged.
-For example, Paul Dubois at Lawrence Livermore National Laboratory (LLNL) took over the
-maintenance of Numeric and funded the writing of its
-manual\cite{Numeric-manual}, and
 %in the Program for Climate Model Diagnosis and Intercomparison at
@@
 %first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
-in 1998 the Space Telescope Science Institute (STScI), which was in charge of 
-Hubble Space Telescope science operations, decided to replace their
-custom scripting language with Python\cite{STScI-slither}.
 
@@
 essential infrastructure at national laboratories and research institutes.
-For example, starting in 2000, STScI ported their IDL-based analysis pipeline,
-IRAF, to Python.
+For example, Paul Dubois at Lawrence Livermore National Laboratory (LLNL) took over the
+maintenance of Numeric and funded the writing of its
+manual\cite{Numeric-manual}, and
+the Space Telescope Science Institute (STScI), which was in charge of
+Hubble Space Telescope science operations, decided to replace their
+custom scripting language and analysis pipeline with Python\cite{STScI-slither}.
 

### 2019-05-18 scipy-1.0/paper.tex
Fix tense

@@
 At this point, scientific Python started attracting more serious attention;
-code that started out as side projects by graduate students has grown into
+code that started out as side projects by graduate students had grown into
 essential infrastructure at national laboratories and research institutes.

### 2019-05-25 scipy-1.0/paper.tex
Minor punctuation/reference tweaks

@@
 
-SciPy's arrival at this point is surprising, and somewhat anomalous.
+SciPy's arrival at this point is surprising and somewhat anomalous.
 When started, the library had little funding, and was written mainly
@@
 ``rogue'' student programmers could upend the already well-established
-ecosystem of research software---backed by millions in funding, and
+ecosystem of research software---backed by millions in funding and
 many hundreds of highly qualified engineers\cite{mathworks-globe-97,
@@
 list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
-Hinsen, and Guido van Rossum, led to the creation of a package called Numeric with an array object
+Hinsen, and Guido van Rossum led to the creation of a package called Numeric with an array object
 that supported a high number of dimensions.  Jim Hugunin explained the utility
@@
 narrowed, while the breadth of the ecosystem grew
-through a new type of auxiliary package, the scikit\cite{scikits-general}. 
+through a new type of auxiliary package: the scikit\cite{scikits-general}.
 The tooling, development, and release processes became more professional.
@@
 feature highlights from the three years preceding
-SciPy 1.0 are discussed in Section \ref{sec:technical_improvements}, 
+SciPy 1.0 are discussed in the Key Technical Improvements section below,
 and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
@@
 \caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017. Logos reprinted with permission.}
+the release of SciPy 1.0 in 2017.}
 \label{fig:timeline}
@@
 Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our code
-base using the \texttt{linguist} library\cite{linguistref} provides a 
-detailed breakdown as \% composition by programming language in 
-SciPy (Table~\ref{tab:linguist}).
+implement scientific algorithms in the SciPy library. An analysis of our codebase
+using the \texttt{linguist} library\cite{linguistref} provides a
+detailed breakdown of SciPy's composition (Table~\ref{tab:linguist}).
 
@@
 lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
-as a glue between well-established low-level scientific computing libraries written
+as a glue between well-established, low-level scientific computing libraries written
 in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
@@
 From a new features standpoint, \texttt{scipy.sparse} matrices and linear
-operators now support the Python matrix multiplication (@) operator when
-available. We've added \texttt{scipy.sparse.norm} and
+operators now support the Python matrix multiplication (@) operator.
+We've added \texttt{scipy.sparse.norm} and
 \texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
@@
 the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
+integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
 version of the Python language.

### 2019-05-25 scipy-1.0/subpackages.tex
Minor punctuation/reference tweaks

@@
     multivariate) interpolation classes, Lagrange and Taylor polynomial
-    interpolators, and wrappers for FITPACK and DFITPACK functions.
+    interpolators, and wrappers for FITPACK\cite{Dierckx:1993:CSF:151103} and DFITPACK functions.
 \item[\texttt{io}]
-    A collection of functions and classes for reading and writing MATLAB, IDL,
-    Matrix Market, Fortran, NetCDF, Harwell-Boeing, WAV and ARFF data files. 
+    A collection of functions and classes for reading and writing MATLAB\cite{matlab}, IDL,
+    Matrix Market\cite{boisvert1997matrix}, Fortran, NetCDF\cite{rew1990netcdf},
+		Harwell-Boeing\cite{harwellboeing}, WAV and ARFF data files.
 \item[\texttt{linalg}]
@@
     Orthogonal distance regression, including Python wrappers for the Fortran
-    library ODRPACK.
+    library ODRPACK\cite{ODRPACK_Boggs}.
 \item[\texttt{optimize}]
@@
     basic linear systems theory.  Functionality includes
-    convolution and correlation; splines; filtering and filter design;
-    continuous and discrete time linear systems; waveform generation;
-    window functions; wavelet computations; peak finding; and spectral
+    convolution and correlation, splines, filtering and filter design,
+    continuous and discrete time linear systems, waveform generation,
+    window functions, wavelet computations, peak finding, and spectral
     analysis.  

### 2019-05-25 scipy-1.0/poly.tex

@@
-Traditionally, SciPy relied heavily on the venerable \texttt{FITPACK}
+Traditionally, SciPy relied heavily on the venerable FITPACK
 Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for

### 2019-06-03 scipy-1.0/paper.tex
Integrated @WarrenWeckesser conclusion from #9

@@
 
-The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev} is a
+Nevertheless, SciPy continually strives to improve.
+The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev}, summarized in Table~\ref{table:roadmap}, is a
 continuously-updated document
@@
 require assistance moving forward.
-
-We are still striving to increase the number of SciPy usage tutorials beyond
-our current 15 section offering\cite{SciPy_tutorials}, but our standard
-documentation for library features is already in excellent shape.
-
-The low-level Cython code in our library (which interacts with C-level code and
+In addition to the items on the roadmap, 
+we are still working to increase the number of SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}. 
+Also, the low-level Cython code in our library (which interacts with C-level code and
 exposes it for Python usage) could use some measure of modernization, including
-migration to typed memoryviews to handle NumPy arrays instead of the older
-syntax:
-
-\begin{lstlisting}[language=Python]
-# example of a function that takes a single
-# 3 dimensional typed memoryview as an argument
-# this allows Cython-level handling of a NumPy
-# array object
-cpdef int func_new(int[:, :, :] arr):
-    pass
-
-# example of the old syntax for handling
-# the same scenario
-cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
-    pass
-\end{lstlisting}
-
-Some additional roadmap items are summarized in Table~\ref{table:roadmap}.
+migration to typed memoryviews to handle NumPy arrays.
+
+%\begin{lstlisting}[language=Python]
+%# example of a function that takes a single
+%# 3 dimensional typed memoryview as an argument
+%# this allows Cython-level handling of a NumPy
+%# array object
+%cpdef int func_new(int[:, :, :] arr):
+    %pass
+%
+%# example of the old syntax for handling
+%# the same scenario
+%cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    %pass
+%\end{lstlisting}
 
@@
 \end{table}
+
+A problem faced by many open source projects is attracting and retaining developers.
+While it is normal for some individuals to contribute to a project for a while and then move on, too much turnover can result in the loss of institutional memory, leading to mistakes of the past being repeated, APIs of new code becoming inconsistent with the old code, and a drifting project scope.
+We are fortunate that the SciPy project continues to attract enthusiastic and competent new developers while maintaining the involvement of a small but dedicated ``old guard''.
+Pauli Virtanen, our BDFL, has been with the project for more than 10 years, and is still actively contributing code.
+While the person acting as the release manager might change from one release to the next, Ralf Gommers acts as a general manager, keeping the release process moving smoothly and ensuring that pull requests get reviewed and merged where appropriate.
+Ralf is approaching his tenth year anniversary on the SciPy project and shows no signs of slowing down.
+There are early contributors, such as Robert Kern and St\'{e}fan van der Walt, who contributed much to SciPy over the last 10 or more years and while not currently writing code for SciPy, still contribute to discussions of bug reports and reviews of new code contributions.
+There are an additional half dozen or so currently active developers who have been contributing steadily for five or more years.
+The combination of a committed old guard and a host of new contributors ensures that SciPy will continue to grow while maintaining a high level of quality.
 

### 2019-06-06 scipy-1.0/paper.tex
Replaced benchmarking figure/caption

--- 2019-05-25 scipy-1.0/paper.tex
@@
 \centering
-\includegraphics[width=\textwidth]{static/asv_time_query_ckdtree}
-\caption{Airspeed velocity benchmarks for \texttt{scipy.spatial.cKDTree.query}
-over a roughly nine year commit history time frame. The results are based on
-Python 2.7 performance on the master branch of the project using NumPy 1.8.2
-and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward
-compatibility). Only the $L^2$ (Euclidean) norm is shown here, and to improve
-backward compatibility and sampling of the benchmarks there was no application
-of toroidal topology to the \texttt{KDTree} (boxsize argument was ignored).}
+\includegraphics[width=\textwidth]{static/asv}
+\caption{Results of the \texttt{scipy.spatial.cKDTree.query} benchmark from the introduction of \texttt{cKDTree} to the release of SciPy 1.0.
+The benchmark generates a $k$-d tree from 10,000 uniformly distributed points in an $m$-dimensional unit hypercube, then finds the nearest (Euclidean) neighbor in the tree for each of 1,000 query points.
+Each marker in the figure indicates the execution time of the benchmark for a commit in the master branch of SciPy.
+Substantial performance improvements were realized when \texttt{cKDTree} was fully Cythonized and again when it was rewritten in C++.
+The tree was generated without application of toroidal topology (\texttt{boxsize=None}), and tests were performed by Airspeed Velocity 0.4 using Python 2.7, NumPy 1.8.2, and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward compatibility).}
 \label{fig:asvbench}

### 2019-06-06 scipy-1.0/paper.tex

--- 2019-06-03 scipy-1.0/paper.tex
@@
 We are fortunate that the SciPy project continues to attract enthusiastic and competent new developers while maintaining the involvement of a small but dedicated ``old guard''.
-Pauli Virtanen, our BDFL, has been with the project for more than 10 years, and is still actively contributing code.
-While the person acting as the release manager might change from one release to the next, Ralf Gommers acts as a general manager, keeping the release process moving smoothly and ensuring that pull requests get reviewed and merged where appropriate.
-Ralf is approaching his tenth year anniversary on the SciPy project and shows no signs of slowing down.
-There are early contributors, such as Robert Kern and St\'{e}fan van der Walt, who contributed much to SciPy over the last 10 or more years and while not currently writing code for SciPy, still contribute to discussions of bug reports and reviews of new code contributions.
-There are an additional half dozen or so currently active developers who have been contributing steadily for five or more years.
+There are contributors who were present in the early years of the project who still contribute to discussions of bug reports and reviews of new code contributions.
+Our BDFL has been with the project for more than 10 years and is still actively contributing code, and the head of our Steering Council, who also acts as a general manager, is approaching his eleventh anniversary.
+An additional half dozen or so active developers have been contributing steadily for five or more years.
 The combination of a committed old guard and a host of new contributors ensures that SciPy will continue to grow while maintaining a high level of quality.

### 2019-06-07 scipy-1.0/paper.tex
Pruned end material

@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
+and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
 and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
-conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
+conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
+Importantly,
 SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
@@
 
-\section*{Acknowledgments}
-
-\fixme{TODO}
-
-%Acknowledgements should be brief, and should not include thanks to anonymous
-%referees and editors, or effusive comments. Grant or contribution numbers may
-%be acknowledged.
-
 \section*{Author contributions statement}
 
-\fixme{TODO}
+M.H. edited the manuscript, created figures, compiled the author list and contributions statements, and drafted the \texttt{linprog} technical improvements section.
 
@@
 
-\section*{Additional information}
-
-\fixme{TODO}
+%\section*{Additional information}
+%
+%\fixme{TODO}
 
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
-\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
+%\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn.
 The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
-\label{fig:knn-complexity}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
-\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
+%\label{fig:knn-complexity}
+%\end{figure}
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
+%\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0.
 Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used.
@@
 The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
-\label{fig:sparse-iter}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
-\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0.
+%\label{fig:sparse-iter}
+%\end{figure}
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
+%\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0.
 Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used.
 The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
-\label{fig:sparse-conv}
-\end{figure}
-
-\section{Polynomial Interpolators: Additional Details}
+%\label{fig:sparse-conv}
+%\end{figure}
+
+\subsection*{Polynomial Interpolators: Additional Details}
 FITPACK provides a collection of routines for calculating
@@
 
-\section*{Numerical Optimization: Additional Details}
+\subsection*{Numerical Optimization: Additional Details}
 The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
@@
 
+\setlength{\tabcolsep}{3pt}
 \begin{table}[H]

### 2019-06-10 scipy-1.0/paper.tex
Removed SI, moved optimization table, added required sections

--- 2019-06-07 scipy-1.0/paper.tex
@@
 
+\section*{Data Availability}
+% The statement should be provided as a separate section (titled 'Data Availability') at the end of the main text, before the 'References' section. 
+All SciPy library source code and all data generated for the current study are available in the SciPy GitHub repository, \url{https://github.com/scipy}.
+
 \bibliography{references}
 
-\section*{Author contributions statement}
-
-M.H. edited the manuscript, created figures, compiled the author list and contributions statements, and drafted the \texttt{linprog} technical improvements section.
+\section*{Author Contributions Statement}
+
+All authors have contributed code to SciPy, and all authors reviewed the manuscript.
+In the author list, the names of authors who contributed text to the manuscript are indicated with $\dagger$, and the primary manuscript editors are denoted with $\ddagger$.
 
@@
 
-%\section*{Additional information}
-%
-%\fixme{TODO}
-
-\section*{Supporting Information}
+\section*{Competing Interests}
+
+The authors declare no competing interests.
+
+%\section*{Supporting Information}
 
 % separate Figure / Table numbering for SI
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}
-\setcounter{table}{0}
+%\renewcommand{\thefigure}{S\arabic{figure}}
+%\renewcommand{\thetable}{S\arabic{table}}
+%\setcounter{figure}{0}
+%\setcounter{table}{0}
 
@@
 
-\subsection*{Polynomial Interpolators: Additional Details}
-FITPACK provides a collection of routines for calculating
-smoothing splines and automatic knot selection. One issue 
-with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
-and the \code{splrep}/\code{splev} combination, is the monolithic design:
-specific algorithms for construction of interpolating and 
-smoothing splines and knot selection are tightly coupled 
-at the Fortran level. The library offers a very
-limited degree of user control over the details 
-of the spline fitting it performs. 
-%
-Routines for manipulating piecewise polynomials and b-splines were either
-not available or lacking in ease of use and/or performance characteristics.
-
-An alternative set of spline interpolation routines was implemented in the
-(now deprecated) \code{splmake}/\code{spleval} combination, and used as a
-backend of the \code{interp1d} routine. To further complicate matters,
-the spline format used by \code{splmake}/\code{spleval} was incompatible 
-with \code{splrep}/\code{splev}.
-
-Using \code{UnivariateSpline} in new code is not recommended. Given data points,
-the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
-instance, representing a spline which interpolates the data. Interpolated values
-at intermediate points can be computed either via its \code{\_\_call\_\_} method,
-or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
-accepts both \code{tck} tuples and \code{Bspline} instances.
-
-Fitting \emph{smoothing} splines to data can currently only be done with 
-FITPACK. Building alternative algorithms is certainly possible,
-but is still to be implemented.
-
-Interpolation with known or estimated derivatives, i.e., cubic or
-higher-order Hermite splines, can be constructed by either
-subclassing \code{PPoly} or \code{BPoly}, or using the factory
-method \code{BPoly.from\_derivatives}.
-
-The spline representation used by \code{BSpline} is directly
-compatible with FITPACK. We recommend that users work with
-\code{BSpline} objects instead of directly manipulating \code{tck}
-tuples of knots, coefficients and the spline degree.
-
-\subsection*{Numerical Optimization: Additional Details}
-The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
-\begin{enumerate}
-\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
-\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
-\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
-\item general, linear optimization problems (\texttt{linprog}),
-\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
-\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
-\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
-\end{enumerate}
-
-\setlength{\tabcolsep}{3pt}
-\begin{table}[H]
-  \centering
-  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
-The field \textit{version added} specifies the algorithm's first appearance in SciPy.
-Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
-    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
-    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
-    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
-    with $\sim$ the algorithm does not requires second-order derivatives from
-    the user; it computes an approximation internally and uses it to accelerate method convergence.
-    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
-    which does not require explicit matrix factorization or storage of the Hessian.
-    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
-    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
-    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
-    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
-    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
-  \begin{tabular}{cccccccccccccc}
-      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
-    \hline
-    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
-    \hline
-    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
-    \hline
-    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
-    \hline
-    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
-    \hline
-    \makecell{Iterative Hessian \\
-    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
-    \hline
-    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
-    \hline
-    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
-    \hline
-    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
-    \hline
-    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
-    \hline
-    Equality constraints &&&&&&&\cmark&&&&& \\
-    \hline
-    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
-    \hline
-    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
-      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
-      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
-      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
-      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
-      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
-      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
-    \hline
-  \end{tabular}
-  \label{tab:minimize-si}
-\end{table}
+%\subsection*{Polynomial Interpolators: Additional Details}
+%FITPACK provides a collection of routines for calculating
+%smoothing splines and automatic knot selection. One issue 
+%with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
+%and the \code{splrep}/\code{splev} combination, is the monolithic design:
+%specific algorithms for construction of interpolating and 
+%smoothing splines and knot selection are tightly coupled 
+%at the Fortran level. The library offers a very
+%limited degree of user control over the details 
+%of the spline fitting it performs. 
+%%
+%Routines for manipulating piecewise polynomials and b-splines were either
+%not available or lacking in ease of use and/or performance characteristics.
+%
+%An alternative set of spline interpolation routines was implemented in the
+%(now deprecated) \code{splmake}/\code{spleval} combination, and used as a
+%backend of the \code{interp1d} routine. To further complicate matters,
+%the spline format used by \code{splmake}/\code{spleval} was incompatible 
+%with \code{splrep}/\code{splev}.
+%
+%Using \code{UnivariateSpline} in new code is not recommended. Given data points,
+%the helper function \code{make\_interp\_spline} constructs a \code{BSpline}
+%instance, representing a spline which interpolates the data. Interpolated values
+%at intermediate points can be computed either via its \code{\_\_call\_\_} method,
+%or by calling \code{splev(xnew, BSpline\_object)} --- the \code{splev} routine
+%accepts both \code{tck} tuples and \code{Bspline} instances.
+%
+%Fitting \emph{smoothing} splines to data can currently only be done with 
+%FITPACK. Building alternative algorithms is certainly possible,
+%but is still to be implemented.
+%
+%Interpolation with known or estimated derivatives, i.e., cubic or
+%higher-order Hermite splines, can be constructed by either
+%subclassing \code{PPoly} or \code{BPoly}, or using the factory
+%method \code{BPoly.from\_derivatives}.
+%
+%The spline representation used by \code{BSpline} is directly
+%compatible with FITPACK. We recommend that users work with
+%\code{BSpline} objects instead of directly manipulating \code{tck}
+%tuples of knots, coefficients and the spline degree.
+%
+%\subsection*{Numerical Optimization: Additional Details}
+%The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
+%\begin{enumerate}
+%\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
+%\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
+%\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
+%\item general, linear optimization problems (\texttt{linprog}),
+%\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
+%\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
+%\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
+%\end{enumerate}
 

### 2019-06-11 scipy-1.0/subpackages.tex
Pruned end material
removed reference to SI

@@
 \item[\texttt{optimize}]
-    This subpackage includes the following, with additional details in the SI:
-    simplex and interior-point linear programming solvers,
-		implementations of many nonlinear minimization algorithms,  
+		This subpackage includes simplex and interior-point linear programming
+		solvers, implementations of many nonlinear minimization algorithms,
 		a routine for least-squares curve fitting, and 

### 2019-06-11 scipy-1.0/ckdtree.tex
Minor punctuation/reference tweaks

@@
 multiple threads may be used\cite{gh-4374}. This generally improved
-performance on any given problem (Figure~\ref{fig:asvbench}) while
-preserving the asymptotic complexity (Figure~\ref{fig:knn-complexity}).
+performance on any given problem (Figure~\ref{fig:asvbench})
+while preserving the asymptotic complexity.
+%(Figure~\ref{fig:knn-complexity}).
 
@@
 limited to the conventional L2 (Euclidean) norm but supports any Minkowski 
-p-norm between 1 and infinity. By default, the returned data structure is a 
+$p$-norm between 1 and infinity. By default, the returned data structure is a
 Dictionary Of Keys (DOK) based sparse matrix, which is very efficient for matrix 

### 2019-06-11 scipy-1.0/scipy-optimize.tex

@@
 On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
-A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}.
+A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}; it illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
 
@@
 Most recently, the SciPy benchmark suite gained a comprehensive set of 196 global optimization problems for tracking the performance of existing solvers over time and for evaluating whether the performance of new solvers merits their inclusion in the package.
+
+\setlength{\tabcolsep}{3pt}
+\begin{table}[H]
+  \centering
+  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
+    \hline
+    Equality constraints &&&&&&&\cmark&&&&& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize-si}
+\end{table}

### 2019-06-19 scipy-1.0/paper.tex
Remove 'seemed to' — too vacillating

@@
 Yet, the philosophical motivations behind a fully open toolstack, and
-an excited, friendly community with a singular focus, seemed to have
+an excited, friendly community with a singular focus, have
 proven auspicious in the long run.  It led not only to the library

### 2019-06-19 scipy-1.0/paper.tex
Minor language tweak

@@
 \end{quote}
-This is still a distinguishing feature of Python for science, and one of the
+This remains a distinguishing feature of Python for science, and one of the
 reasons why it has been so successful in the realm of data science: instead of

### 2019-06-19 scipy-1.0/paper.tex
Grammar and typos

--- 2019-06-11 scipy-1.0/paper.tex
@@
 scikit-image\cite{vanderwalt2014scikit} provides image processing
-capabilities beyond \texttt{scipy.ndimage}, sympy\cite{meurer2017sympy}
+capabilities beyond SciPy's \texttt{ndimage}, sympy\cite{meurer2017sympy}
 provides a Python interface for symbolic computation, and
@@
 graphs and networks in a more specialized way than SciPy components
-like \texttt{sparse.csgraph} or \texttt{spatial}.
+like \texttt{sparse.csgraph} or \texttt{spatial} offer.
 
@@
 range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) is explicitly out of scope.
+for graphics processing units (GPUs) are explicitly out of scope.
 

### 2019-06-20 scipy-1.0/paper.tex
paraphrase the networkx sentence

@@
 provides a Python interface for symbolic computation, and
-NetworkX\cite{hagberg2008networkx} is a Python library for probing
-graphs and networks in a more specialized way than SciPy components
-like \texttt{sparse.csgraph} or \texttt{spatial} offer.
+while \texttt{sparse.csgraph} or \texttt{spatial} offers basic tools
+for working with graphs and networks compared to a more specialized
+Python library like NetworkX\cite{hagberg2008networkx}.
 

### 2019-06-20 scipy-1.0/paper.tex
remove space before % (convention + internal consistency)

@@
 unit tests in the test suite, which is written for usage with the \texttt{pytest}
-framework, and with 87 \% and 45 \% line coverages for Python and compiled
+framework, and with 87\% and 45\% line coverages for Python and compiled
 code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by

### 2019-06-20 scipy-1.0/paper.tex
more " \%" -> "\%"

@@
 covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
-compiled code with test coverage near $45 \%$ and $106,878$ lines of 
-Python code with test coverage near $87 \%$.\\
+compiled code with test coverage near $45\%$ and $106,878$ lines of 
+Python code with test coverage near $87\%$.\\
 The reconstitution of historical
@@
 MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
-by 47 \% of all machine learning projects on GitHub\cite{octoverse-scipy}.
+by 47\% of all machine learning projects on GitHub\cite{octoverse-scipy}.
 

### 2019-06-20 scipy-1.0/poly.tex

@@
-Traditionally, SciPy relied heavily on the venerable FITPACK
+Historically, SciPy relied heavily on the venerable FITPACK
 Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for

### 2019-06-20 scipy-1.0/paper.tex
Circle publishes but does not host I think
More language reworking
@mdhaber pointed out [1] that the comma after yet was redundant [2], and
the commas surrounding "an excited, friendly community" were too. The
sentence became too long to parse clearly, so I reworked it a little to
make the commas (and thus breaks) fit more nicely.

..[1]: https://github.com/scipy/scipy-articles/pull/152#discussion_r295550199
..[2]: https://english.stackexchange.com/questions/106901/yet-at-the-beginning-of-a-sentence

--- 2019-06-20 scipy-1.0/paper.tex
@@
 
-Yet, the philosophical motivations behind a fully open toolstack, and
-an excited, friendly community with a singular focus, seemed to have
-proven auspicious in the long run.  It led not only to the library
+Yet the philosophical motivations behind a fully open toolstack, combined
+with an excited, friendly community with a singular focus, have
+proven auspicious in the long run.  They led not only to the library
 described in this paper, but also to an entire ecosystem of related
@@
 scikit-image\cite{vanderwalt2014scikit} provides image processing
-capabilities beyond SciPy's \texttt{ndimage}, sympy\cite{meurer2017sympy}
+capabilities beyond \texttt{scipy.ndimage}, sympy\cite{meurer2017sympy}
 provides a Python interface for symbolic computation, and
-while \texttt{sparse.csgraph} or \texttt{spatial} offers basic tools
-for working with graphs and networks compared to a more specialized
-Python library like NetworkX\cite{hagberg2008networkx}.
+NetworkX\cite{hagberg2008networkx} is a Python library for probing
+graphs and networks in a more specialized way than SciPy components
+like \texttt{sparse.csgraph} or \texttt{spatial}.
 
@@
 range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) are explicitly out of scope.
+for graphics processing units (GPUs) is explicitly out of scope.
 

### 2019-06-20 scipy-1.0/paper.tex
Minor language tweaks (#152)
* Remove 'seemed to' — too vacillating
* Minor language tweak
* More language reworking

@mdhaber pointed out [1] that the comma after yet was redundant [2], and
the commas surrounding "an excited, friendly community" were too. The
sentence became too long to parse clearly, so I reworked it a little to
make the commas (and thus breaks) fit more nicely.

..[1]: https://github.com/scipy/scipy-articles/pull/152#discussion_r295550199
..[2]: https://english.stackexchange.com/questions/106901/yet-at-the-beginning-of-a-sentence

--- 2019-06-20 scipy-1.0/paper.tex
@@
 framework, and with 87\% and 45\% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 
-% NOTE: is there a citation for PyPy?
-
 \subsubsection*{Benchmark suite}

### 2019-06-21 scipy-1.0/scipy-optimize.tex

@@
 However, the interior point method typically returns a solution near the center of an optimal face, yet basic solutions are often preferred for sensitivity analysis and for use in mixed integer programming algorithms.
-This motivates the need for a crossover routine or a new implementation of the simplex method for sparse problems in a future release, either of which would require an improved sparse linear system solver with efficient support for rank-one updates
+This motivates the need for a crossover routine or a new implementation of the simplex method for sparse problems in a future release, either of which would require an improved sparse linear system solver with efficient support for rank-one updates.
 

### 2019-06-21 scipy-1.0/paper.tex
AUTH: add affiliation
Add subsection on documentation.
This text was proposed via email by Joe Harrington.

@@
 \author[5]{Anne Archibald}
-\author[6]{Andrew Nelson}
+\author[6]{Andrew R. J. Nelson}
 \author[7]{Charles Harris}
@@
 \affil[5]{Affiliation, department, city, postcode, country}
-\affil[6]{Affiliation, department, city, postcode, country}
+\affil[6]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
 \affil[7]{Affiliation, department, city, postcode, country}
@@
 called PyData.
+
+\subsection{*Documentation}
+Early versions of SciPy had minimal documentation.  Travis
+Oliphant wrote \textit{A guide to NumPy}, an 8000 word e-book, which
+documented the underpinnings and API for compiled components, but did
+not provide reference documentation for the $\sim$2000 routines then in
+NumPy, nor said much about SciPy.  Realizing that the package had gotten too large for a single
+person to document, Joseph Harrington founded the SciPy Documentation
+Project in late Spring 2008, hiring Stefan van der Walt as documentation
+editor.  Pauli Virtanen, Emmanuelle Gouillart, and Gael Varoquaux wrote
+numpydoc, a system for collecting, reviewing, and safely integrating
+documentation with the Python sources.  The system, based on Sphinx,
+produced text, hypertext, and PDF-formatted documents from a single
+source document.  Harrington proposed and they established a
+documentation template and a review system.  Using the SciPy mailing
+list, they solicited community help, and by the end of the summer had
+high-quality documentation for the most frequently used NumPy and SciPy
+routines.  Each contributor of 1000 words or more received a T-shirt
+reading "SciPy Documentation Marathon: I WTFM", emphasizing that the
+project was no mere coding sprint.  In 2009, van der Walt became a core
+developer and David Goldsmith stepped up as lead editor.  The lead
+editor's job included maintaining discussion and interest on the mailing
+lists, which led to high participation.  By 2010, over 75 people had
+contributed significantly to over 1000 printed pages of reference
+documentation\cite{harrington2008scipy, harrington2009scipy}.
+At this point, most user-facing routines had documentation, and the doc project took its place in the
+normal development cycle.  The SciPy project remains committed to
+high-quality, current user documentation.
 

### 2019-06-25 scipy-1.0/paper.tex

--- 2019-06-20 scipy-1.0/paper.tex
@@
 \affil[24]{University of Texas at Austin,
-           Institute for Computational Engineering and Sciences,
+           Oden Institute for Computational Engineering and Sciences,
 	   Austin, TX, 78712, USA}

### 2019-06-28 scipy-1.0/paper.tex
Added my name to authors list
DOC: Add affiliation
Correctly capitalize SymPy.

@@
 \author[3]{St\'efan J. van der Walt}
+\author[26]{\.{I}lhan Polat}
 
@@
 \affil[9]{Affiliation, department, city, postcode, country}
-\affil[10]{Affiliation, department, city, postcode, country}
+\affil[10]{University of Washington, Institute for Learning and Brain Sciences, Seattle, WA, 98195, USA}
 \affil[11]{Affiliation, department, city, postcode, country}
@@
   Berkeley, CA, 94720, USA}
+\affil[26]{Affiliation, department, city, postcode, country}
 
@@
 scikit-image\cite{vanderwalt2014scikit} provides image processing
-capabilities beyond SciPy's \texttt{ndimage}, sympy\cite{meurer2017sympy}
+capabilities beyond SciPy's \texttt{ndimage}, SymPy\cite{meurer2017sympy}
 provides a Python interface for symbolic computation, and

### 2019-06-28 scipy-1.0/paper.tex
Remove a stray comma.

@@
 The follow-up CiSE March/April 2011 Python for Scientists and Engineers
-special issue\cite{millman2011python}, focused more on the core parts
+special issue\cite{millman2011python} focused more on the core parts
 of the scientific Python ecosystem\cite{perez2011python} including
@@
 consistent framework that provides methods to sample random variates,
-to evaluate the cumulative distribution function (cdf) and the probability
-density function (pdf) and to fit parameters for every distribution. 
+to evaluate the cumulative distribution function (CDF) and the probability
+density function (PDF) and to fit parameters for every distribution.
 Generally, the methods rely on specific implementations for each
-distribution such as a closed form expression of the cdf or a sampling 
+distribution such as a closed form expression of the CDF or a sampling
 algorithm, if available. Otherwise, default methods are used
-based on generic code, e.g., numerical integration of the pdf to
-obtain the cdf. 
+based on generic code, e.g., numerical integration of the PDF to
+obtain the CDF.
 Key recent distributions added to \texttt{scipy.stats} include the

### 2019-06-28 scipy-1.0/paper.tex
Consistently use the thousand seperator.

@@
 unique code contributors, millions of downloads per year, 161 dependent
-packages, and 28700 dependent repositories. This includes usage of SciPy
+packages, and 28,700 dependent repositories. This includes usage of SciPy
 in almost half of all machine learning projects on GitHub, and usage by
@@
 600,000 lines of code organized in 16 subpackages. 
-Over 85,000 GitHub repositories and almost 5000 packages depend
+Over 85,000 GitHub repositories and almost 5,000 packages depend
 on SciPy. Some of the major
@@
 The application programming interface (API) for SciPy consists of approximately
-1500 functions and classes.  Our policy for evolving the API over time is that
+1,500 functions and classes.  Our policy for evolving the API over time is that
 new functionality can be added, while removing or changing existing
@@
 website\cite{SciPylib}, which has been the default citation in the absence of a
-peer-reviewed paper, has been cited $>3000$ times\cite{googlescholar}. Some of the most prominent
+peer-reviewed paper, has been cited $>3,000$ times\cite{googlescholar}. Some of the most prominent
 usages of or demonstrations of credibility for SciPy include the LIGO / Virgo

### 2019-06-28 scipy-1.0/paper.tex
Capitalize CDF/PDF.

--- 2019-06-28 scipy-1.0/paper.tex
@@
 The follow-up CiSE March/April 2011 Python for Scientists and Engineers
-special issue\cite{millman2011python} focused more on the core parts
+special issue\cite{millman2011python}, focused more on the core parts
 of the scientific Python ecosystem\cite{perez2011python} including

### 2019-06-28 scipy-1.0/paper.tex
Converted figures to vector format

@@
 \usepackage{pgfplotstable}
-\usepackage[multiple]{footmisc}
+% \usepackage[multiple]{footmisc}
 \usepackage{booktabs,siunitx}
@@
 scikit-image\cite{vanderwalt2014scikit} provides image processing
-capabilities beyond SciPy's \texttt{ndimage}, SymPy\cite{meurer2017sympy}
+capabilities beyond SciPy's \texttt{ndimage}, sympy\cite{meurer2017sympy}
 provides a Python interface for symbolic computation, and
@@
 consistent framework that provides methods to sample random variates,
-to evaluate the cumulative distribution function (CDF) and the probability
-density function (PDF) and to fit parameters for every distribution.
+to evaluate the cumulative distribution function (cdf) and the probability
+density function (pdf) and to fit parameters for every distribution. 
 Generally, the methods rely on specific implementations for each
-distribution such as a closed form expression of the CDF or a sampling
+distribution such as a closed form expression of the cdf or a sampling 
 algorithm, if available. Otherwise, default methods are used
-based on generic code, e.g., numerical integration of the PDF to
-obtain the CDF.
+based on generic code, e.g., numerical integration of the pdf to
+obtain the cdf. 
 Key recent distributions added to \texttt{scipy.stats} include the

### 2019-06-28 scipy-1.0/paper.tex
Moved all figures and tables to end

@@
 \usepackage{pgfplotstable}
-% \usepackage[multiple]{footmisc}
+\usepackage[multiple]{footmisc}
 \usepackage{booktabs,siunitx}
@@
 routines\cite{Strangman-modules}.
-
-\begin{figure}
-\begin{verbatim}
-SciPy is an open source package that builds on the strengths of Python and
-Numeric providing a wide range of fast scientific and numeric functionality.
-SciPy's current module set includes the following:
-
-    Special Functions (Bessel, hanker, Airy, etc.)
-    Signal/Image Processing
-    2D Plotting capabilities
-    Integration
-    ODE solvers
-    Optimization (simplex, BFGS, Netwon-CG, etc.)
-    Genetic Algorithms
-    Numeric -> C++ expression compiler
-    Parallel programming tools
-    Splines and Interpolation
-    And other stuff.
-\end{verbatim}
-\caption{Excerpt from SciPy 0.1 release announcement (typos included).}\label{fig:announce-0.1}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
-\caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017.}
-\label{fig:timeline}
-\end{figure}
-
 % Release management
@@
 
-\begin{table}
+Fortran, despite its age, is still a high-performance scientific programming language with 
+continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
+libraries in order to provide Python convenience while benefiting from their performance: 
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
+ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
+QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
+FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
+ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
+MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
+ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
+CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
+
+Rounding out the top three languages in SciPy is C, which is also extremely
+well-established over several decades\cite{Kernighan:1988:CPL:576122} of
+scientific computing. The C libraries that we wrap in SciPy include 
+trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry) 
+Cephes\cite{cephes_netlib}. % (mathematics algorithms).
+
+Cython has been described as a creole language that mixes the best parts of Python and
+lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
+as a glue between well-established, low-level scientific computing libraries written
+in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
+enhancements in Python code, especially for cases where heavily used inner
+loops benefit from a compiled code with static typing.
+
+For implementing new functionality, Python is the still the language of choice. If Python 
+performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
+order). The main motivation for this is maintainability: Cython has the 
+highest abstraction level and most Python developers will understand it. C is 
+also widely known, and easier for the current core development team to manage
+than C++ and especially Fortran.
+
+The distribution of secondary programming languages in SciPy is a compromise between
+a powerful, performance-
+enhancing language that interacts well with Python (i.e. Cython) and the
+usage of languages (and their libraries) that have proven reliable and
+performant over many decades. The position that SciPy occupies near the
+foundation of the scientific Python ecosystem is such that adoption of new
+languages or major dependencies is generally unlikely---our choices are strongly
+driven by long-term stability. GPU acceleration, new transpiling libraries, and
+the latest JIT compilation approaches (i.e.,
+Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall
+outside the remit of the main SciPy library. That said, we have recently
+increased our efforts to support compatibility with some of these options, and
+having our full test suite pass with the PyPy JIT
+compiler\cite{Bolz:2009:TMP:1565824.1565827} is now a requirement in our
+development workflow.
+
+\subsection*{API and ABI evolution}
+
+The application programming interface (API) for SciPy consists of approximately
+1500 functions and classes.  Our policy for evolving the API over time is that
+new functionality can be added, while removing or changing existing
+functionality can only be done if the benefits of that exceeds the (often
+significant) costs to users, \textit{and} only after giving clear deprecation
+warnings to those users for at least one year. In general, we encourage
+changes that improve clarity in the API of the library but strongly discourage
+breaking backwards compatibility given our position near the base of the 
+scientific Python computing stack.
+
+In addition to the Python API, SciPy has C and Cython interfaces.
+Therefore, we have to also consider the application binary
+interface (ABI). This ABI has been stable for a long time, and we aim to
+evolve it only in a backwards compatible way.
+
+\section*{Key technical improvements}
+\label{sec:technical_improvements}
+Here we describe key technical improvements made in the last three years.
+
+\subsection*{Data structures}
+
+\subsubsection*{Sparse matrices}
+
+\texttt{scipy.sparse} offers seven sparse matrix data structures,
+also known as sparse formats. The most important ones are the row- 
+and column-compressed formats (CSR and CSC, respectively). 
+These offer fast major-axis indexing and fast matrix-vector multiplication,
+and are used heavily throughout SciPy and dependent packages.
+
+Over the last three years, our sparse matrix handling internals have been
+rewritten and performance has been improved. Iterating over and slicing of CSC
+and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
+and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
+conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
+Importantly,
+SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
+low-level implementations leveraged by a subset of our \texttt{sparse}
+offerings.
+
+From a new features standpoint, \texttt{scipy.sparse} matrices and linear
+operators now support the Python matrix multiplication (@) operator.
+We've added \texttt{scipy.sparse.norm} and
+\texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
+random variates from arbitrary distributions, respectively. Also, we've made a
+concerted effort to bring the \texttt{scipy.sparse} API into line with the
+equivalent NumPy API where possible.
+
+\subsubsection*{\texttt{cKDTree}}
+
+\input{ckdtree}
+
+\subsection*{Unified bindings to compiled code}
+
+\subsubsection*{LowLevelCallable}
+
+As of SciPy version 0.19, it is possible for users to wrap low-level functions
+in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
+calling compiled C functions, such as those generated using \texttt{numba}
+or Cython, directly from Python.
+Supported low-level functions include \texttt{PyCapsule}
+objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
+Furthermore, it is even possible to generate a low-level callback function
+automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
+
+\subsection*{Cython bindings for BLAS, LAPACK, and special}
+
+SciPy has provided special functions and leveraged BLAS and
+LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
+includes Cython\cite{behnel2011cython} wrappers for
+many BLAS and LAPACK routines (added in 2015) and the special functions 
+provided in the \texttt{scipy.{\allowbreak}special} subpackage (added in 2016).
+These Cython wrappers are available in 
+\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},
+\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and
+\texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
+When writing algorithms in Cython, it is typically more efficient to call
+directly into the libraries SciPy wraps rather than indirectly, using SciPy's
+Python APIs.  These low-level interfaces for Cython can also be used outside of
+the SciPy codebase to gain access to the functions in the wrapped libraries
+while avoiding the overhead of Python function calls.  This can give
+performance gains of one or two orders of magnitude for many use cases.
+
+Developers can also use the low-level Cython interfaces without linking against
+the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.  This lets other
+extensions avoid the complexity of finding and using the correct libraries.
+Avoiding this complexity is especially important when wrapping libraries
+written in Fortran.  Not only can these low-level wrappers be used without a
+Fortran compiler, they can also be used without having to handle all the
+different Fortran compiler ABIs and name mangling schemes.
+
+Most of these low-level Cython wrappers are generated automatically to help
+with both correctness and ease of maintenance.  The wrappers for BLAS and
+LAPACK are primarily generated using type information that is parsed from the
+BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small
+number of routines use hand-written type signatures instead.  The input and
+output types of each routine are saved in a data file that is read at build
+time and used to generate the corresponding Cython wrapper files.  The wrappers
+in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also
+generated from a data file containing type information for the wrapped
+routines.
+
+Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only
+provided for the routines that maintain a consistent interface across all
+supported LAPACK versions.  The standard BLAS interface provided by the various
+existing BLAS libraries is not currently changing, so changes are not generally
+needed in the wrappers provided by SciPy.  Changes to the Cython wrappers for
+the functions in \texttt{scipy.{\allowbreak}special} follow corresponding
+changes to the interface of that subpackage.
+
+\subsection*{Numerical optimization}
+
+\input{scipy-optimize}
+
+\subsection*{Statistical distributions}
+
+The \texttt{scipy.stats} subpackage contains more than 100 probability
+distributions: 96 continuous and 13 discrete univariate distributions
+and 10 multivariate distributions. The implementation relies on a
+consistent framework that provides methods to sample random variates,
+to evaluate the cumulative distribution function (cdf) and the probability
+density function (pdf) and to fit parameters for every distribution. 
+Generally, the methods rely on specific implementations for each
+distribution such as a closed form expression of the cdf or a sampling 
+algorithm, if available. Otherwise, default methods are used
+based on generic code, e.g., numerical integration of the pdf to
+obtain the cdf. 
+Key recent distributions added to \texttt{scipy.stats} include the
+histogram-based distribution in \texttt{scipy.stats.rv\_histogram}
+and the multinomial distribution in \texttt{scipy.stats.multinomial}
+(used, for example, in natural language processing, see
+\cite{Griffiths5228}).
+
+\subsection*{Polynomial interpolators}
+
+\input{poly}
+
+\subsection*{Test and benchmark suite}
+
+\subsubsection*{Test suite}
+
+The SciPy test suite is orchestrated by a continuous integration matrix that
+includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
+AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which is written for usage with the \texttt{pytest}
+framework, and with 87\% and 45\% line coverages for Python and compiled
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
+the CircleCI service to facilitate evaluation of documentation changes /
+integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
+version of the Python language.
+
+\subsubsection*{Benchmark suite}
+
+In addition to ensuring that unit tests are passing, it is important to confirm that the
+the performance of the SciPy codebase improves over time. Since February 2015, the
+performance of SciPy has been monitored with Airspeed Velocity (\texttt{asv} \cite{asvref}).
+SciPy's \texttt{run.py} script conveniently wraps \texttt{asv} features such that benchmark 
+results over time can be generated with a single console command. For example,
+Figure~\ref{fig:asvbench} illustrates the improvement of\texttt{scipy.spatial.cKDTree.query}
+over roughly nine years of project history.
+
+\section*{Project organization and community}
+
+\subsection*{Governance}
+
+SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
+A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
+Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
+The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community.
+The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.
+
+SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
+\emph{be open} to everyone participating in our community;
+\emph{be empathetic and patient} in resolving conflicts;
+\emph{be collaborative}, as we depend on each other to build the library;
+\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
+\emph{be careful with wording}.
+The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response.
+Our Diversity Statement ``welcomes and encourages participation by everyone''.
+
+\subsection*{Maintainers and contributors}
+
+The SciPy project has approximately 100 unique contributors
+for every 6-monthly release cycle. Anyone with the interest and
+skills can become a contributor; the SciPy Developer
+Guide\cite{scipy-dev-guide} provides guidance on how to do that.
+In addition, the project currently has 15 active maintainers: people who review
+the contributions of others and do everything else needed to ensure that the
+software and the project move forward. Maintainers are critical to the health
+of the project\cite{eghbal2016}; their skills and efforts largely determine how
+fast the project moves forward, and they enable input from the much
+larger group of contributors. Anyone can become a maintainer, too, as they
+are selected on a rolling basis from contributors with a significant history of
+high-quality contributions.
+
+\subsection*{Downstream projects}
+
+The scientific Python ecosystem includes many examples
+of domain-specific software libraries building on top
+of SciPy features and then returning to the base SciPy library
+to suggest and even implement improvements. 
+For example, there are common contributors to the SciPy and 
+Astropy core libraries\cite{astropy-2018}, and what works 
+well for one of the codebases, infrastructures, or communities
+is often transferred in some form to the other. At the codebase
+level, the \texttt{binned\_statistic} functionality
+is one such cross-project contribution: it was initially
+developed in an Astropy-affiliated package
+and then placed in SciPy afterwards. 
+In this perspective, SciPy serves as a catalyst for cross-fertilisation 
+throughout the Python scientific computing community.
+
+\section*{Discussion}
+
+% The Discussion should be succinct and must not contain subheadings.
+
+SciPy has a strong developer community and a massive user base. GitHub traffic
+metrics report roughly 20,000 unique visitors to the source website between May
+14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
+that roughly two-week period. The developer community at that time consisted of 610 unique
+contributors of source code, with $>19,000$ commits accepted into the codebase
+(GitHub page data).
+
+From the user side, there were 13,096,468 downloads of SciPy from the Python
+Packaging Index (PyPI)\cite{pypinfo} and 5,776,017 via the default channel of the
+\texttt{conda}\cite{condainfo} package manager during the year 2017. These numbers establish a lower
+bound on the total number of downloads by users given that
+PyPI and \texttt{conda} are only two of several popular methods for installing SciPy.  The SciPy
+website\cite{SciPylib}, which has been the default citation in the absence of a
+peer-reviewed paper, has been cited $>3000$ times\cite{googlescholar}. Some of the most prominent
+usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
+scientific collaboration that lead to the observation of gravitational waves
+\cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
+MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
+by 47\% of all machine learning projects on GitHub\cite{octoverse-scipy}.
+
+Nevertheless, SciPy continually strives to improve.
+The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev}, summarized in Table~\ref{table:roadmap}, is a
+continuously-updated document
+maintained by the community that describes some of the major directions for
+improvement for the project, as well as specific limitations and matters that
+require assistance moving forward.
+In addition to the items on the roadmap, 
+we are still working to increase the number of SciPy usage tutorials beyond
+our current 15 section offering\cite{SciPy_tutorials}. 
+Also, the low-level Cython code in our library (which interacts with C-level code and
+exposes it for Python usage) could use some measure of modernization, including
+migration to typed memoryviews to handle NumPy arrays.
+
+%\begin{lstlisting}[language=Python]
+%# example of a function that takes a single
+%# 3 dimensional typed memoryview as an argument
+%# this allows Cython-level handling of a NumPy
+%# array object
+%cpdef int func_new(int[:, :, :] arr):
+    %pass
+%
+%# example of the old syntax for handling
+%# the same scenario
+%cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
+    %pass
+%\end{lstlisting}
+
+A problem faced by many open source projects is attracting and retaining developers.
+While it is normal for some individuals to contribute to a project for a while and then move on, too much turnover can result in the loss of institutional memory, leading to mistakes of the past being repeated, APIs of new code becoming inconsistent with the old code, and a drifting project scope.
+We are fortunate that the SciPy project continues to attract enthusiastic and competent new developers while maintaining the involvement of a small but dedicated ``old guard''.
+There are contributors who were present in the early years of the project who still contribute to discussions of bug reports and reviews of new code contributions.
+Our BDFL has been with the project for more than 10 years and is still actively contributing code, and the head of our Steering Council, who also acts as a general manager, is approaching his eleventh anniversary.
+An additional half dozen or so active developers have been contributing steadily for five or more years.
+The combination of a committed old guard and a host of new contributors ensures that SciPy will continue to grow while maintaining a high level of quality.
+
+\begin{table}[H]
 \centering
@@
 
-Fortran, despite its age, is still a high-performance scientific programming language with 
-continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
-libraries in order to provide Python convenience while benefiting from their performance: 
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
-ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression, 
-MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations, 
-ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and 
-CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
-
-Rounding out the top three languages in SciPy is C, which is also extremely
-well-established over several decades\cite{Kernighan:1988:CPL:576122} of
-scientific computing. The C libraries that we wrap in SciPy include 
-trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
-SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
-Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry) 
-Cephes\cite{cephes_netlib}. % (mathematics algorithms).
-
-Cython has been described as a creole language that mixes the best parts of Python and
-lower-level C / C++ paradigms\cite{behnel2011cython}. We often use Cython
-as a glue between well-established, low-level scientific computing libraries written
-in C / C++ and the Python interface offered by SciPy. We also use Cython to enable performance
-enhancements in Python code, especially for cases where heavily used inner
-loops benefit from a compiled code with static typing.
-
-For implementing new functionality, Python is the still the language of choice. If Python 
-performance is an issue, then we prefer the use of Cython followed by C, C++, or Fortran (in that
-order). The main motivation for this is maintainability: Cython has the 
-highest abstraction level and most Python developers will understand it. C is 
-also widely known, and easier for the current core development team to manage
-than C++ and especially Fortran.
-
-The distribution of secondary programming languages in SciPy is a compromise between
-a powerful, performance-
-enhancing language that interacts well with Python (i.e. Cython) and the
-usage of languages (and their libraries) that have proven reliable and
-performant over many decades. The position that SciPy occupies near the
-foundation of the scientific Python ecosystem is such that adoption of new
-languages or major dependencies is generally unlikely---our choices are strongly
-driven by long-term stability. GPU acceleration, new transpiling libraries, and
-the latest JIT compilation approaches (i.e.,
-Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall
-outside the remit of the main SciPy library. That said, we have recently
-increased our efforts to support compatibility with some of these options, and
-having our full test suite pass with the PyPy JIT
-compiler\cite{Bolz:2009:TMP:1565824.1565827} is now a requirement in our
-development workflow.
-
-\subsection*{API and ABI evolution}
-
-The application programming interface (API) for SciPy consists of approximately
-1500 functions and classes.  Our policy for evolving the API over time is that
-new functionality can be added, while removing or changing existing
-functionality can only be done if the benefits of that exceeds the (often
-significant) costs to users, \textit{and} only after giving clear deprecation
-warnings to those users for at least one year. In general, we encourage
-changes that improve clarity in the API of the library but strongly discourage
-breaking backwards compatibility given our position near the base of the 
-scientific Python computing stack.
-
-In addition to the Python API, SciPy has C and Cython interfaces.
-Therefore, we have to also consider the application binary
-interface (ABI). This ABI has been stable for a long time, and we aim to
-evolve it only in a backwards compatible way.
-
-\section*{Key technical improvements}
-\label{sec:technical_improvements}
-Here we describe key technical improvements made in the last three years.
-
-\subsection*{Data structures}
-
-\subsubsection*{Sparse matrices}
-
-\texttt{scipy.sparse} offers seven sparse matrix data structures,
-also known as sparse formats. The most important ones are the row- 
-and column-compressed formats (CSR and CSC, respectively). 
-These offer fast major-axis indexing and fast matrix-vector multiplication,
-and are used heavily throughout SciPy and dependent packages.
-
-Over the last three years, our sparse matrix handling internals have been
-rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
-and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
-conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
-Importantly,
-SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
-low-level implementations leveraged by a subset of our \texttt{sparse}
-offerings.
-
-From a new features standpoint, \texttt{scipy.sparse} matrices and linear
-operators now support the Python matrix multiplication (@) operator.
-We've added \texttt{scipy.sparse.norm} and
-\texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
-random variates from arbitrary distributions, respectively. Also, we've made a
-concerted effort to bring the \texttt{scipy.sparse} API into line with the
-equivalent NumPy API where possible.
-
-\subsubsection*{\texttt{cKDTree}}
-
-\input{ckdtree}
-
-\subsection*{Unified bindings to compiled code}
-
-\subsubsection*{LowLevelCallable}
-
-As of SciPy version 0.19, it is possible for users to wrap low-level functions
-in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions, such as those generated using \texttt{numba}
-or Cython, directly from Python.
-Supported low-level functions include \texttt{PyCapsule}
-objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
-Furthermore, it is even possible to generate a low-level callback function
-automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
-
-\subsection*{Cython bindings for BLAS, LAPACK, and special}
-
-SciPy has provided special functions and leveraged BLAS and
-LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
-includes Cython\cite{behnel2011cython} wrappers for
-many BLAS and LAPACK routines (added in 2015) and the special functions 
-provided in the \texttt{scipy.{\allowbreak}special} subpackage (added in 2016).
-These Cython wrappers are available in 
-\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_blas},
-\texttt{scipy.{\allowbreak}linalg.{\allowbreak}cython\_lapack}, and
-\texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} respectively.
-When writing algorithms in Cython, it is typically more efficient to call
-directly into the libraries SciPy wraps rather than indirectly, using SciPy's
-Python APIs.  These low-level interfaces for Cython can also be used outside of
-the SciPy codebase to gain access to the functions in the wrapped libraries
-while avoiding the overhead of Python function calls.  This can give
-performance gains of one or two orders of magnitude for many use cases.
-
-Developers can also use the low-level Cython interfaces without linking against
-the wrapped libraries\cite{blas-lapack-wrappers-scipy-2015}.  This lets other
-extensions avoid the complexity of finding and using the correct libraries.
-Avoiding this complexity is especially important when wrapping libraries
-written in Fortran.  Not only can these low-level wrappers be used without a
-Fortran compiler, they can also be used without having to handle all the
-different Fortran compiler ABIs and name mangling schemes.
-
-Most of these low-level Cython wrappers are generated automatically to help
-with both correctness and ease of maintenance.  The wrappers for BLAS and
-LAPACK are primarily generated using type information that is parsed from the
-BLAS and LAPACK source files using F2PY\cite{peterson2009f2py}, though a small
-number of routines use hand-written type signatures instead.  The input and
-output types of each routine are saved in a data file that is read at build
-time and used to generate the corresponding Cython wrapper files.  The wrappers
-in \texttt{scipy.{\allowbreak}special.{\allowbreak}cython\_special} are also
-generated from a data file containing type information for the wrapped
-routines.
-
-Since SciPy can be built with LAPACK 3.4.0 or later, Cython wrappers are only
-provided for the routines that maintain a consistent interface across all
-supported LAPACK versions.  The standard BLAS interface provided by the various
-existing BLAS libraries is not currently changing, so changes are not generally
-needed in the wrappers provided by SciPy.  Changes to the Cython wrappers for
-the functions in \texttt{scipy.{\allowbreak}special} follow corresponding
-changes to the interface of that subpackage.
-
-\subsection*{Numerical optimization}
-
-\input{scipy-optimize}
-
-\subsection*{Statistical distributions}
-
-The \texttt{scipy.stats} subpackage contains more than 100 probability
-distributions: 96 continuous and 13 discrete univariate distributions
-and 10 multivariate distributions. The implementation relies on a
-consistent framework that provides methods to sample random variates,
-to evaluate the cumulative distribution function (cdf) and the probability
-density function (pdf) and to fit parameters for every distribution. 
-Generally, the methods rely on specific implementations for each
-distribution such as a closed form expression of the cdf or a sampling 
-algorithm, if available. Otherwise, default methods are used
-based on generic code, e.g., numerical integration of the pdf to
-obtain the cdf. 
-Key recent distributions added to \texttt{scipy.stats} include the
-histogram-based distribution in \texttt{scipy.stats.rv\_histogram}
-and the multinomial distribution in \texttt{scipy.stats.multinomial}
-(used, for example, in natural language processing, see
-\cite{Griffiths5228}).
-
-\subsection*{Polynomial interpolators}
-
-\input{poly}
-
-\subsection*{Test and benchmark suite}
-
-\subsubsection*{Test suite}
-
-The SciPy test suite is orchestrated by a continuous integration matrix that
-includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
-AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which is written for usage with the \texttt{pytest}
-framework, and with 87\% and 45\% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
-the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
-version of the Python language.
+\setlength{\tabcolsep}{3pt}
+\begin{table}[H]
+  \centering
+  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .
+The field \textit{version added} specifies the algorithm's first appearance in SciPy.
+Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
+    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
+    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
+    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
+    with $\sim$ the algorithm does not requires second-order derivatives from
+    the user; it computes an approximation internally and uses it to accelerate method convergence.
+    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
+    which does not require explicit matrix factorization or storage of the Hessian.
+    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
+    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
+    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
+    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
+    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+  \begin{tabular}{cccccccccccccc}
+      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
+    \hline
+    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
+    \hline
+    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
+    \hline
+    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
+    \hline
+    \makecell{Iterative Hessian \\
+    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
+    \hline
+    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
+    \hline
+    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
+    \hline
+    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
+    \hline
+    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
+    \hline
+    Equality constraints &&&&&&&\cmark&&&&& \\
+    \hline
+    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
+    \hline
+    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
+      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
+      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
+      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+    \hline
+  \end{tabular}
+  \label{tab:minimize-si}
+\end{table}
+
+\begin{table}[H]
+\caption{Summary of SciPy Roadmap items following 1.0 release}
+\begin{tabular}{ c|c }
+    SciPy subpackage & Summary of Change \\
+    \hline
+    \texttt{optimize} & a few more high quality global optimizers\\
+    \texttt{fftpack} & reduce overlap with NumPy equivalent \\
+    \texttt{linalg} & reduce overlap with NumPy equivalent \\
+    \texttt{interpolate} & new spline fitting and arithmetic routines \\
+    \texttt{interpolate} & new transparent tensor-product splines\\
+    \texttt{interpolate} & new Non-Uniform Rational B-Splines\\
+    \texttt{interpolate} & mesh refinement \& coarsening of B-splines and tensor products\\
+    \texttt{signal} & migrate spline functionality to \texttt{interpolate}\\
+    \texttt{signal} & Second Order Sections update to match capabilities in other routines\\
+    \texttt{linalg} & support a more recent version of LAPACK\\
+    \texttt{ndimage} & clarify usage of the ``data point'' coordinate
+    model, and add additional wrapping modes\\
+    \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\
+    \texttt{sparse.linalg} & add PROPACK wrappers for faster SVD\\
+    \texttt{spatial} & add support for (quaternion) rotation matrices\\
+    \texttt{special} & precision improvements for hypergeometric, parabolic cylinder, and spheroidal wave functions\\
+\end{tabular}\label{table:roadmap}
+\end{table}
+
+\begin{figure}[H]
+\begin{verbatim}
+SciPy is an open source package that builds on the strengths of Python and
+Numeric providing a wide range of fast scientific and numeric functionality.
+SciPy's current module set includes the following:
+
+    Special Functions (Bessel, hanker, Airy, etc.)
+    Signal/Image Processing
+    2D Plotting capabilities
+    Integration
+    ODE solvers
+    Optimization (simplex, BFGS, Netwon-CG, etc.)
+    Genetic Algorithms
+    Numeric -> C++ expression compiler
+    Parallel programming tools
+    Splines and Interpolation
+    And other stuff.
+\end{verbatim}
+\caption{Excerpt from SciPy 0.1 release announcement (typos included).}\label{fig:announce-0.1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.95\textwidth]{static/scipy_timeline}
+\caption{Major milestones from SciPy's initial release in 2001 to
+the release of SciPy 1.0 in 2017.}
+\label{fig:timeline}
+\end{figure}
 
@@
 
-\subsubsection*{Benchmark suite}
-
-In addition to ensuring that unit tests are passing, it is important to confirm that the
-the performance of the SciPy codebase improves over time. Since February 2015, the
-performance of SciPy has been monitored with Airspeed Velocity (\texttt{asv} \cite{asvref}).
-SciPy's \texttt{run.py} script conveniently wraps \texttt{asv} features such that benchmark 
-results over time can be generated with a single console command. For example,
-Figure~\ref{fig:asvbench} illustrates the improvement of\texttt{scipy.spatial.cKDTree.query}
-over roughly nine years of project history.
-
 \begin{figure}[H]
@@
 \end{figure}
-
-\section*{Project organization and community}
-
-\subsection*{Governance}
-
-SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
-A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
-Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
-The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community.
-The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.
-
-SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
-\emph{be open} to everyone participating in our community;
-\emph{be empathetic and patient} in resolving conflicts;
-\emph{be collaborative}, as we depend on each other to build the library;
-\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
-\emph{be careful with wording}.
-The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response.
-Our Diversity Statement ``welcomes and encourages participation by everyone''.
-
-\subsection*{Maintainers and contributors}
-
-The SciPy project has approximately 100 unique contributors
-for every 6-monthly release cycle. Anyone with the interest and
-skills can become a contributor; the SciPy Developer
-Guide\cite{scipy-dev-guide} provides guidance on how to do that.
-In addition, the project currently has 15 active maintainers: people who review
-the contributions of others and do everything else needed to ensure that the
-software and the project move forward. Maintainers are critical to the health
-of the project\cite{eghbal2016}; their skills and efforts largely determine how
-fast the project moves forward, and they enable input from the much
-larger group of contributors. Anyone can become a maintainer, too, as they
-are selected on a rolling basis from contributors with a significant history of
-high-quality contributions.
-
-\subsection*{Downstream projects}
-
-The scientific Python ecosystem includes many examples
-of domain-specific software libraries building on top
-of SciPy features and then returning to the base SciPy library
-to suggest and even implement improvements. 
-For example, there are common contributors to the SciPy and 
-Astropy core libraries\cite{astropy-2018}, and what works 
-well for one of the codebases, infrastructures, or communities
-is often transferred in some form to the other. At the codebase
-level, the \texttt{binned\_statistic} functionality
-is one such cross-project contribution: it was initially
-developed in an Astropy-affiliated package
-and then placed in SciPy afterwards. 
-In this perspective, SciPy serves as a catalyst for cross-fertilisation 
-throughout the Python scientific computing community.
-
-\section*{Discussion}
-
-% The Discussion should be succinct and must not contain subheadings.
-
-SciPy has a strong developer community and a massive user base. GitHub traffic
-metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
-that roughly two-week period. The developer community at that time consisted of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the codebase
-(GitHub page data).
-
-From the user side, there were 13,096,468 downloads of SciPy from the Python
-Packaging Index (PyPI)\cite{pypinfo} and 5,776,017 via the default channel of the
-\texttt{conda}\cite{condainfo} package manager during the year 2017. These numbers establish a lower
-bound on the total number of downloads by users given that
-PyPI and \texttt{conda} are only two of several popular methods for installing SciPy.  The SciPy
-website\cite{SciPylib}, which has been the default citation in the absence of a
-peer-reviewed paper, has been cited $>3000$ times\cite{googlescholar}. Some of the most prominent
-usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
-scientific collaboration that lead to the observation of gravitational waves
-\cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
-by 47\% of all machine learning projects on GitHub\cite{octoverse-scipy}.
-
-Nevertheless, SciPy continually strives to improve.
-The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev}, summarized in Table~\ref{table:roadmap}, is a
-continuously-updated document
-maintained by the community that describes some of the major directions for
-improvement for the project, as well as specific limitations and matters that
-require assistance moving forward.
-In addition to the items on the roadmap, 
-we are still working to increase the number of SciPy usage tutorials beyond
-our current 15 section offering\cite{SciPy_tutorials}. 
-Also, the low-level Cython code in our library (which interacts with C-level code and
-exposes it for Python usage) could use some measure of modernization, including
-migration to typed memoryviews to handle NumPy arrays.
-
-%\begin{lstlisting}[language=Python]
-%# example of a function that takes a single
-%# 3 dimensional typed memoryview as an argument
-%# this allows Cython-level handling of a NumPy
-%# array object
-%cpdef int func_new(int[:, :, :] arr):
-    %pass
-%
-%# example of the old syntax for handling
-%# the same scenario
-%cpdef int func_old(object[int, ndim=3, mode='strided'] arr):
-    %pass
-%\end{lstlisting}
-
-\begin{table}[H]
-\caption{Summary of SciPy Roadmap items following 1.0 release}
-\begin{tabular}{ c|c }
-    SciPy subpackage & Summary of Change \\
-    \hline
-    \texttt{optimize} & a few more high quality global optimizers\\
-    \texttt{fftpack} & reduce overlap with NumPy equivalent \\
-    \texttt{linalg} & reduce overlap with NumPy equivalent \\
-    \texttt{interpolate} & new spline fitting and arithmetic routines \\
-    \texttt{interpolate} & new transparent tensor-product splines\\
-    \texttt{interpolate} & new Non-Uniform Rational B-Splines\\
-    \texttt{interpolate} & mesh refinement \& coarsening of B-splines and tensor products\\
-    \texttt{signal} & migrate spline functionality to \texttt{interpolate}\\
-    \texttt{signal} & Second Order Sections update to match capabilities in other routines\\
-    \texttt{linalg} & support a more recent version of LAPACK\\
-    \texttt{ndimage} & clarify usage of the ``data point'' coordinate
-    model, and add additional wrapping modes\\
-    \texttt{sparse} & incorporate sparse arrays from Sparse package\cite{abbasi2018sparse} \\
-    \texttt{sparse.linalg} & add PROPACK wrappers for faster SVD\\
-    \texttt{spatial} & add support for (quaternion) rotation matrices\\
-    \texttt{special} & precision improvements for hypergeometric, parabolic cylinder, and spheroidal wave functions\\
-\end{tabular}\label{table:roadmap}
-\end{table}
-
-A problem faced by many open source projects is attracting and retaining developers.
-While it is normal for some individuals to contribute to a project for a while and then move on, too much turnover can result in the loss of institutional memory, leading to mistakes of the past being repeated, APIs of new code becoming inconsistent with the old code, and a drifting project scope.
-We are fortunate that the SciPy project continues to attract enthusiastic and competent new developers while maintaining the involvement of a small but dedicated ``old guard''.
-There are contributors who were present in the early years of the project who still contribute to discussions of bug reports and reviews of new code contributions.
-Our BDFL has been with the project for more than 10 years and is still actively contributing code, and the head of our Steering Council, who also acts as a general manager, is approaching his eleventh anniversary.
-An additional half dozen or so active developers have been contributing steadily for five or more years.
-The combination of a committed old guard and a host of new contributors ensures that SciPy will continue to grow while maintaining a high level of quality.
 

### 2019-06-28 scipy-1.0/paper.tex

--- 2019-06-28 scipy-1.0/paper.tex
@@
 The application programming interface (API) for SciPy consists of approximately
-1500 functions and classes.  Our policy for evolving the API over time is that
+1,500 functions and classes.  Our policy for evolving the API over time is that
 new functionality can be added, while removing or changing existing
-functionality can only be done if the benefits of that exceeds the (often
+functionality can only be done if the benefits exceed the (often
 significant) costs to users, \textit{and} only after giving clear deprecation
@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
-and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
-conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
+and CSR matrices is now up to 35\% faster, % (Figure~\ref{fig:sparse-iter}),
+and the speed of coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
+conversions has increased. % (Figure~\ref{fig:sparse-conv}).
 Importantly,
@@
 objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
-Furthermore, it is even possible to generate a low-level callback function
+Furthermore, it is possible to generate a low-level callback function
 automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
@@
 of the project\cite{eghbal2016}; their skills and efforts largely determine how
-fast the project moves forward, and they enable input from the much
+fast the project progresses, and they enable input from the much
 larger group of contributors. Anyone can become a maintainer, too, as they

### 2019-06-29 scipy-1.0/paper.tex
removed language distribution table

@@
 
-Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our codebase
-using the \texttt{linguist} library\cite{linguistref} provides a
-detailed breakdown of SciPy's composition (Table~\ref{tab:linguist}).
-
-\begin{table}
-\centering
-\begin{tabular}{p{3cm}S[table-format = <0.1]}
-\toprule
-Language & {Percent}\\
-\midrule
-Python  & 49.5 \\
-Fortran & 25.6 \\
-C       & 19.5 \\
-Cython  & 3.0  \\
-C++     & 2.3  \\[\defaultaddspace]
-{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
-\bottomrule
-\end{tabular}
-\caption{Language composition of SciPy codebase: lines of code in each
-programming language as determined by the \texttt{linguist} package.
-The last row denotes tools used in supporting roles in tests,
-building, and documentation.}
-\label{tab:linguist}
-\end{table}
+According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 5 \% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
+The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.
+Cython) and the usage of languages (and their libraries) that have proven reliable and performant over many decades.
 
@@
 
-The distribution of secondary programming languages in SciPy is a compromise between
-a powerful, performance-
-enhancing language that interacts well with Python (i.e. Cython) and the
-usage of languages (and their libraries) that have proven reliable and
-performant over many decades. The position that SciPy occupies near the
+The position that SciPy occupies near the
 foundation of the scientific Python ecosystem is such that adoption of new

### 2019-06-30 scipy-1.0/paper.tex
5% -> 50%

@@
 
-According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 5 \% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
+According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 50\% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
 The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.

### 2019-06-30 scipy-1.0/paper.tex
Updated dependent package/repo numbers and cited GitHub

--- 2019-06-28 scipy-1.0/paper.tex
@@
 in the Python programming language, with more than 600
-unique code contributors, millions of downloads per year, 161 dependent
-packages, and 28,700 dependent repositories. This includes usage of SciPy
+unique code contributors, thousands of dependent packages,
+over 100,000 dependent repositories, and millions of downloads per year.
+This includes usage of SciPy
 in almost half of all machine learning projects on GitHub, and usage by
@@
 600,000 lines of code organized in 16 subpackages. 
-Over 85,000 GitHub repositories and almost 5,000 packages depend
-on SciPy. Some of the major
+Over 110,000 GitHub repositories and 6,500 packages depend
+on SciPy\cite{dependents}. Some of the major
 feature highlights from the three years preceding

### 2019-07-01 scipy-1.0/paper.tex
Clarified how linguist, gcov, and pytest-cov quantify code

--- 2019-06-30 scipy-1.0/paper.tex
@@
 of the removal of $\approx 61,000$ lines of compiled code for SciPy
-v0.14, the volume of both compiled and Python
+v0.14, the volume of both compiled (C, C++, and Fortran) and Python
 code has increased between releases, as has the number of lines
-covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
-compiled code with test coverage near $45\%$ and $106,878$ lines of 
-Python code with test coverage near $87\%$.\\
-The reconstitution of historical
-build and test environments was performed using a docker-based approach
-\cite{scipy-cov}. Analysis of lines of source code covered is performed 
-automatically in our continuous integration suite using the \texttt{pytest-cov}
-and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.}
+covered by unit tests. In SciPy v1.0.0, tests covered approximately
+45\% of $462,574$ lines of compiled code (many of which
+are automatically generated from Cython source), and nearly 87\% of
+106,878 lines of Python code.\\
+Coverage analysis is performed automatically in our continuous integration
+suite using the \texttt{pytest-cov} and \texttt{gcov} libraries for 
+Python and compiled code, respectively. The
+the reconstitution of historical build and test environments for this
+figure was performed using a Docker-based approach\cite{scipy-cov}.
+}
 \label{fig:coverage}

### 2019-07-01 scipy-1.0/paper.tex
Cited pytest, pytest-cov, gcov

@@
 include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which is written for usage with the \texttt{pytest}
-framework, and with 87\% and 45\% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
+unit tests in the test suite, which is written for usage with \texttt{pytest}\cite{pytest}.
+Test coverage at the SciPy 1.0 release point (Figure~\ref{fig:coverage}) was at
+87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
+45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
+Documentation for the code is automatically built and published by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 suite using the \texttt{pytest-cov} and \texttt{gcov} libraries for 
-Python and compiled code, respectively. The
-the reconstitution of historical build and test environments for this
+Python and compiled code, respectively.
+The reconstitution of historical build and test environments for this
 figure was performed using a Docker-based approach\cite{scipy-cov}.

### 2019-07-01 scipy-1.0/paper.tex
tweaks

--- 2019-06-30 scipy-1.0/paper.tex
@@
 \author[3]{St\'efan J. van der Walt}
-\author[26]{\.{I}lhan Polat}
 
@@
 \affil[9]{Affiliation, department, city, postcode, country}
-\affil[10]{University of Washington, Institute for Learning and Brain Sciences, Seattle, WA, 98195, USA}
+\affil[10]{Affiliation, department, city, postcode, country}
 \affil[11]{Affiliation, department, city, postcode, country}
@@
 \affil[24]{University of Texas at Austin,
-           Oden Institute for Computational Engineering and Sciences,
+           Institute for Computational Engineering and Sciences,
 	   Austin, TX, 78712, USA}
@@
   Berkeley, CA, 94720, USA}
-\affil[26]{Affiliation, department, city, postcode, country}
 
@@
 in the Python programming language, with more than 600
-unique code contributors, thousands of dependent packages,
-over 100,000 dependent repositories, and millions of downloads per year.
-This includes usage of SciPy
+unique code contributors, millions of downloads per year, 161 dependent
+packages, and 28700 dependent repositories. This includes usage of SciPy
 in almost half of all machine learning projects on GitHub, and usage by
@@
 through a new type of auxiliary package: the scikit\cite{scikits-general}.
-The tooling, development, and release processes became more professional.
+The tooling, development, documentation, and release processes became more professional.
 SciPy was expanded carefully, with the patience affordable in open source 
 projects and via best practices common in industry \cite{millman2014developing}.
+
+Early versions of SciPy had minimal documentation, but this began to change
+with the 2006 release of a \emph{Guide to NumPy}\cite{oliphant2006guide}. 
+In 2007, Sphinx\cite{sphinx} made it possible to render hypertext and 
+PDF-formatted documents automatically from plain text (docstrings) 
+interspersed with Python code, and in 2008, \texttt{pydocweb}\cite{pydocweb}
+enabled collaborative documentation development in a wiki-like fashion.
+The SciPy Documentation Project \cite{van2008scipy, harrington2008scipy} 
+used these tools to complete documentation of SciPy's user-facing
+functionality: offering t-shirts to contributors from around the world in 
+exchange for high-quality text, it collected contributions from over 75 people 
+to produce an 884 page manual\cite{harrington2009scipy}. Since then,
+SciPy has remained committed to maintaining high-quality documentation as 
+part of the normal development cycle.
 
@@
 600,000 lines of code organized in 16 subpackages. 
-Over 110,000 GitHub repositories and 6,500 packages depend
-on SciPy\cite{dependents}. Some of the major
+Over 85,000 GitHub repositories and almost 5000 packages depend
+on SciPy. Some of the major
 feature highlights from the three years preceding
@@
 The application programming interface (API) for SciPy consists of approximately
-1,500 functions and classes.  Our policy for evolving the API over time is that
+1500 functions and classes.  Our policy for evolving the API over time is that
 new functionality can be added, while removing or changing existing
-functionality can only be done if the benefits exceed the (often
+functionality can only be done if the benefits of that exceeds the (often
 significant) costs to users, \textit{and} only after giving clear deprecation
@@
 rewritten and performance has been improved. Iterating over and slicing of CSC
-and CSR matrices is now up to 35\% faster, % (Figure~\ref{fig:sparse-iter}),
-and the speed of coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
-conversions has increased. % (Figure~\ref{fig:sparse-conv}).
+and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
+and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
+conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
 Importantly,
@@
 objects, \texttt{ctypes} function pointers, and \texttt{cffi} function pointers.
-Furthermore, it is possible to generate a low-level callback function
+Furthermore, it is even possible to generate a low-level callback function
 automatically from a Cython module using \texttt{scipy.LowLevelCallable.from\_cython}.
@@
 of the project\cite{eghbal2016}; their skills and efforts largely determine how
-fast the project progresses, and they enable input from the much
+fast the project moves forward, and they enable input from the much
 larger group of contributors. Anyone can become a maintainer, too, as they
@@
 website\cite{SciPylib}, which has been the default citation in the absence of a
-peer-reviewed paper, has been cited $>3,000$ times\cite{googlescholar}. Some of the most prominent
+peer-reviewed paper, has been cited $>3000$ times\cite{googlescholar}. Some of the most prominent
 usages of or demonstrations of credibility for SciPy include the LIGO / Virgo

### 2019-07-01 scipy-1.0/paper.tex

--- 2019-07-01 scipy-1.0/paper.tex
@@
 
-According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 50\% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
-The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.
-Cython) and the usage of languages (and their libraries) that have proven reliable and performant over many decades.
+Python, Cython, Fortran, C and C++ are the programming languages used to
+implement scientific algorithms in the SciPy library. An analysis of our codebase
+using the \texttt{linguist} library\cite{linguistref} provides a
+detailed breakdown of SciPy's composition (Table~\ref{tab:linguist}).
+
+\begin{table}
+\centering
+\begin{tabular}{p{3cm}S[table-format = <0.1]}
+\toprule
+Language & {Percent}\\
+\midrule
+Python  & 49.5 \\
+Fortran & 25.6 \\
+C       & 19.5 \\
+Cython  & 3.0  \\
+C++     & 2.3  \\[\defaultaddspace]
+{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
+\bottomrule
+\end{tabular}
+\caption{Language composition of SciPy codebase: lines of code in each
+programming language as determined by the \texttt{linguist} package.
+The last row denotes tools used in supporting roles in tests,
+building, and documentation.}
+\label{tab:linguist}
+\end{table}
 
@@
 
-The position that SciPy occupies near the
+The distribution of secondary programming languages in SciPy is a compromise between
+a powerful, performance-
+enhancing language that interacts well with Python (i.e. Cython) and the
+usage of languages (and their libraries) that have proven reliable and
+performant over many decades. The position that SciPy occupies near the
 foundation of the scientific Python ecosystem is such that adoption of new
@@
 of the removal of $\approx 61,000$ lines of compiled code for SciPy
-v0.14, the volume of both compiled (C, C++, and Fortran) and Python
+v0.14, the volume of both compiled and Python
 code has increased between releases, as has the number of lines
-covered by unit tests. In SciPy v1.0.0, tests covered approximately
-45\% of $462,574$ lines of compiled code (many of which
-are automatically generated from Cython source), and nearly 87\% of
-106,878 lines of Python code.\\
-Coverage analysis is performed automatically in our continuous integration
-suite using the \texttt{pytest-cov} and \texttt{gcov} libraries for 
-Python and compiled code, respectively. The
-the reconstitution of historical build and test environments for this
-figure was performed using a Docker-based approach\cite{scipy-cov}.
-}
+covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
+compiled code with test coverage near $45\%$ and $106,878$ lines of 
+Python code with test coverage near $87\%$.\\
+The reconstitution of historical
+build and test environments was performed using a docker-based approach
+\cite{scipy-cov}. Analysis of lines of source code covered is performed 
+automatically in our continuous integration suite using the \texttt{pytest-cov}
+and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
+inclusion of additional vendored code.}
 \label{fig:coverage}

### 2019-07-02 scipy-1.0/covlett.tex
Early draft of cover letter

% based on template:
% https://tex.stackexchange.com/questions/15532/looking-for-cover-letter-template

\documentclass[10pt,stdletter,dateno]{newlfm}
\usepackage{kpfonts}
\usepackage{url}

\widowpenalty=1000
\clubpenalty=1000

\newlfmP{headermarginskip=10pt}
\newlfmP{sigsize=50pt}
\newlfmP{dateskipafter=20pt}
%\newlfmP{addrfromphone}
%\newlfmP{addrfromemail}
%\PhrPhone{Phone}
%\PhrEmail{Email}

\namefrom{Ralf Gommers,
          Matt Haberland,
          and Tyler Reddy}
\addrfrom{%
    \today\\[10pt]
    Quansight Labs,\\
    The Netherlands\\
    \\
    BioResource and Agricultural Engineering,\\
    California Polytechnic State University,\\
    San Luis Obispo, CA, 93407, USA\\
    \\
    CCS-7 Applied Computer Science Group,\\
    Los Alamos National Laboratory,\\
    Los Alamos, NM, 87545, USA
}

\addrto{%
Prof. Syma Khalid,\\
School of Chemistry,\\
University of Southampton,\\
Highfield Campus,\\
Southampton,\\
SO17 1BJ, U.K.}

\greetto{Dear Prof. Khalid,}
\closeline{Sincerely,}
\begin{document}
\begin{newlfm}

We are submitting the enclosed manuscript describing a milestone
1.0 release of SciPy for consideration to be published in
\emph{Scientific Reports}. SciPy is the fundamental scientific 
algorithm computing library in the Python programming language. 
Hundreds of experts have contributed to our library, which
has been download millions of times and has over 100,000 dependent
code repositories.

We have specifically chosen \emph{Scientific Reports} because
the breadth of acceptable topics matches the breadth of algorithms
our library provides. High-profile manuscripts citing SciPy range
from the biosciences to recent reports on gravitational waves and
black holes. These citations target a website, and we now think
it is well past time that our mature library has a peer-reviewed
journal article describing it.

We suggest the following three peer reviewers as suitable candidates,
and we do not have any specific reviewer exclusion requests:

Thank you for your consideration.  We look forward to
hearing from you.

\end{newlfm}
\end{document}

### 2019-07-02 scipy-1.0/consortium.tex
Updated author list and contribution statement, added consortium

{\bf
Aditya Vijaykumar$^{31,32}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{33}$, Andreas Kloeckner$^{34}$, Antony Lee$^{35}$, Ariel Rokem$^{36}$, C. Nathan Woods$^{9}$, Chad Fulton$^{37}$, Charles Masson$^{38}$, Christian Brueffer$^{39}$, Christian H\"aggstr\"om$^{40}$, Clark Fitzgerald$^{41}$, David A. Nicholson$^{42}$, David R Hagen$^{43}$, Dmitrii V. Pasechnik$^{44}$, Emanuele Olivetti$^{45}$, Eric Martin$^{10}$, Eric Wieser$^{46}$, Fabrice Silva$^{47}$, Felix Lenders$^{48,49,50}$, Florian Wilhelm$^{51}$, G. Young$^{10}$, Gavin A. Price$^{52}$, Gert-Ludwig Ingold$^{53}$, Gregory E. Allen$^{54}$, Gregory R. Lee$^{55,56}$, Herv\'e Audren$^{57}$, Irvin Probst$^{58}$, J\"org P. Dietrich$^{59,60}$, Jacob Silterra$^{61}$, James T Webber$^{62}$, Janko Slavi\v{c}$^{63}$, Joel Nothman$^{64}$, Johannes Buchner$^{65,66}$, Johannes Kulick$^{67}$, Johannes L. Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, Joscha Reimer$^{69}$, Joseph Harrington$^{70}$, Juan Luis Cano Rodr'iguez$^{71}$, Juan Nunez-Iglesias$^{72}$, Justin Kuczynski$^{73}$, Kevin Tritz$^{74}$, Martin Thoma$^{75}$, Matthias K\"ummerer$^{76}$, Maximilian Bolingbroke$^{77}$, Michael Tartre$^{78}$, Mikhail Pak$^{79}$, Nathaniel J. Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{80}$, Oleksandr Pavlyk$^{81}$, Per A. Brodtkorb$^{82}$, Perry Lee$^{10}$, Robert T. McGibbon$^{83}$, Roman Feldbauer$^{84}$, Sam Lewis$^{85}$, Sam Tygier$^{86}$, Scott Sievert$^{87}$, Sebastiano Vigna$^{88}$, Sourav Singh$^{89}$, Stefan Peterson$^{10}$, Surhud More$^{90,91}$, Tadeusz Pudlik$^{92}$, Takuya Oshima$^{93}$, Thomas J. Pingel$^{94}$, Thomas P. Robitaille$^{95}$, Thomas Spura$^{10}$, Thouis R. Jones$^{96}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{97}$, Tom Krauss$^{98}$, Utkarsh Upadhyay$^{99}$, Yaroslav O. Halchenko$^{100}$, Yoshiki V\'azquez-Baeza$^{101}$
}
\newline
\hfill \break
$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, $^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., $^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{42}$Emory University, Atlanta, GA 30322, USA, $^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{49}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, $^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{51}$Independent Researcher, Cologne, Germany, $^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, $^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, $^{60}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, $^{61}$Independent Researcher, USA, $^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, $^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{71}$Independent Researcher, Spain, $^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, $^{75}$Independent Researcher, Munich, Germany, $^{76}$University of T\"ubingen, T\"ubingen, Germany, $^{77}$Independent Researcher, United Kingdom, $^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, $^{80}$Independent Researcher, Berlin, Germany, $^{81}$Intel Corp., Austin, TX 78746, USA, $^{82}$Independent Researcher, Horten, Norway, $^{83}$D. E. Shaw Research, New York, NY 10036, USA, $^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{85}$Independent Researcher, Melbourne, Australia, $^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{89}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{92}$Waymo LLC, Mountain View, CA 94043, USA, $^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{96}$Broad Institute, Cambridge, MA 02142, USA, $^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{99}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA

### 2019-07-02 scipy-1.0/core-dev.tex
Updated author list and contribution statement, added consortium

\author[1]{Pauli Virtanen}
\author[2]{Ralf Gommers}
\author[3,4,5,6,2]{Travis E. Oliphant}
\author[7,8]{Matt Haberland}
\author[9]{Tyler Reddy}
\author[10]{David Cournapeau}
\author[11]{Evgeni Burovski}
\author[12,13]{Pearu Peterson}
\author[10]{Warren Weckesser}
\author[14]{Jonathan Bright}
\author[15]{St\'efan J. van der Walt}
\author[16]{Matthew Brett}
\author[10]{Joshua Wilson}
\author[17,18]{K. Jarrod Millman}
\author[19]{Nikolay Mayorov}
\author[20]{Andrew R. J. Nelson}
\author[5]{Eric Jones}
\author[5]{Robert Kern}
\author[21]{Eric Larson}
\author[22]{CJ Carey}
\author[10]{\.{I}lhan Polat}
\author[23]{Yu Feng}
\author[24]{Eric W. Moore}
\author[25]{Jake VanderPlas}
\author[10]{Denis Laxalde}
\author[10]{Josef Perktold}
\author[26]{Robert Cimrman}
\author[27]{Ian Henriksen}
\author[10]{E. A. Quintero}
\author[10]{Charles R Harris}
\author[28]{Anne M. Archibald}
\author[29]{Ant\^{o}nio H. Ribeiro}
\author[30]{Fabian Pedregosa}
\author[30]{Paul van Mulbregt}
\author[ ]{SciPy Contributors}
\affil[1]{University of Jyv\"askyl\"a, FI-40014 University of Jyv\"askyl\"a, Finland}
\affil[2]{Quansight LLC, Austin, TX, USA}
\affil[3]{Ultrasound Imaging, Mayo Clinic, Rochester, MN 55902, USA}
\affil[4]{Electrical Engineering, Brigham Young University, Provo, UT 84604, USA}
\affil[5]{Enthought, Inc., Austin, TX 78701, USA}
\affil[6]{Anaconda Inc, Austin, TX 78701, USA}
\affil[7]{BioResource and Agricultural Engineering Department, California Polytechnic State University, San Luis Obispo, CA 93407, USA}
\affil[8]{Department of Mathematics, University of California, Los Angeles, CA 90025, USA}
\affil[9]{Los Alamos National Laboratory, Los Alamos, NM 87544}
\affil[10]{Independent Researcher}
\affil[11]{Higher School of Economics, Myasnitskaya 20, Moscow 101000 Russia}
\affil[12]{Independent Researcher, Saue, Estonia}
\affil[13]{Department of Mechanics and Applied Mathematics, Institute of Cybernetics at Tallinn Technical University, Akadeemia Rd 21 Tallinn, Estonia}
\affil[14]{Independent Researcher, NY, USA}
\affil[15]{Berkeley Institute for Data Science, University of California Berkeley, Berkeley, CA 94720 USA}
\affil[16]{School of Psychology, University of Birmingham, Edgbaston, Birmigham B15 2TT, UK}
\affil[17]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720 USA}
\affil[18]{Division of Biostatistics, University of California, Berkeley, CA 94720, USA}
\affil[19]{WayRay LLC, Skolkovo Innovation Center, Moscow, 143026, Russia}
\affil[20]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
\affil[21]{Institute for Learning and Brain Sciences, University of Washington, Seattle, WA 98195, USA}
\affil[22]{College of Information and Computing Sciences, University of Massachusetts Amherst, Amherst, MA 01002, USA}
\affil[23]{Berkeley Center for Cosmological Physics, University of California, Berkeley, CA 94720, USA}
\affil[24]{Bruker Biospin Corp., 15 Fortune Drive, Billerica, MA 01821}
\affil[25]{University of Washington, Seattle, WA 98195, USA}
\affil[26]{New Technologies - Research Centre, University of West Bohemia, Univerzitní 8, 306 14, Plzeň, Czech Republic}
\affil[27]{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin, Austin, Texas 78712-1229, USA}
\affil[28]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, Nederland}
\affil[29]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
\affil[30]{Google, Inc., 355 Main St., Cambridge, MA 02142, USA}
\affil[*]{scipy.articles@gmail.com}

### 2019-07-02 scipy-1.0/paper.tex
Updated author list and contribution statement, added consortium

@@
 
-\author[1]{Pauli Virtanen}
-\author[2,*]{Ralf Gommers}
-\author[3,4]{Tyler Reddy}
-\author[5]{Anne Archibald}
-\author[6]{Andrew R. J. Nelson}
-\author[7]{Charles Harris}
-\author[8]{CJ Carey}
-\author[9]{Denis Laxalde}
-\author[10]{Eric Larson}
-\author[11]{Eric Moore}
-\author[12]{Eric Quintero}
-\author[13]{Evgeni Burovski}
-\author[14]{Jaime Fernández del Río}
-\author[15]{Josef Perktold}
-\author[16]{Josh Wilson}
-\author[17]{Matthew Brett}
-\author[18]{Nikolay Mayorov}
-\author[19]{Warren Weckesser}
-\author[20]{Matt Haberland}
-\author[21]{Scott Sievert}
-\author[22]{Yu Feng}
-\author[23]{Antonio Horta Ribeiro}
-\author[24]{Ian Henriksen}
-\author[3,25]{K. Jarrod Millman}
-\author[3]{St\'efan J. van der Walt}
-\author[26]{\.{I}lhan Polat}
-
-\affil[1]{Affiliation, department, city, postcode, country}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA, 94720, USA}
-\affil[4]{Los Alamos National Laboratory,
-	  Theoretical Division 6,
-          Los Alamos, NM, 87545, USA}
-\affil[5]{Affiliation, department, city, postcode, country}
-\affil[6]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
-\affil[7]{Affiliation, department, city, postcode, country}
-\affil[8]{Affiliation, department, city, postcode, country}
-\affil[9]{Affiliation, department, city, postcode, country}
-\affil[10]{University of Washington, Institute for Learning and Brain Sciences, Seattle, WA, 98195, USA}
-\affil[11]{Affiliation, department, city, postcode, country}
-\affil[12]{Affiliation, department, city, postcode, country}
-\affil[13]{Affiliation, department, city, postcode, country}
-\affil[14]{Affiliation, department, city, postcode, country}
-\affil[15]{Affiliation, department, city, postcode, country}
-\affil[16]{Affiliation, department, city, postcode, country}
-\affil[17]{Affiliation, department, city, postcode, country}
-\affil[18]{Affiliation, department, city, postcode, country}
-\affil[19]{Affiliation, department, city, postcode, country}
-\affil[20]{BioResource and Agricultural Engineering, California Polytechnic State University, San Luis Obispo, CA, 93407, USA}
-\affil[21]{Affiliation, department, city, postcode, country}
-\affil[22]{Affiliation, department, city, postcode, country}
-\affil[23]{Affiliation, department, city, postcode, country}
-\affil[24]{University of Texas at Austin,
-           Oden Institute for Computational Engineering and Sciences,
-	   Austin, TX, 78712, USA}
-\affil[25]{Division of Biostatistics, University of California,
-  Berkeley, CA, 94720, USA}
-\affil[26]{Affiliation, department, city, postcode, country}
-
-\affil[*]{ralf.gommers@gmail.com}
+\input{core-dev}
 
@@
 through a new type of auxiliary package: the scikit\cite{scikits-general}.
-The tooling, development, documentation, and release processes became more professional.
+The tooling, development, and release processes became more professional.
 SciPy was expanded carefully, with the patience affordable in open source 
 projects and via best practices common in industry \cite{millman2014developing}.
-
-Early versions of SciPy had minimal documentation, but this began to change
-with the 2006 release of a \emph{Guide to NumPy}\cite{oliphant2006guide}. 
-In 2007, Sphinx\cite{sphinx} made it possible to render hypertext and 
-PDF-formatted documents automatically from plain text (docstrings) 
-interspersed with Python code, and in 2008, \texttt{pydocweb}\cite{pydocweb}
-enabled collaborative documentation development in a wiki-like fashion.
-The SciPy Documentation Project \cite{van2008scipy, harrington2008scipy} 
-used these tools to complete documentation of SciPy's user-facing
-functionality: offering t-shirts to contributors from around the world in 
-exchange for high-quality text, it collected contributions from over 75 people 
-to produce an 884 page manual\cite{harrington2009scipy}. Since then,
-SciPy has remained committed to maintaining high-quality documentation as 
-part of the normal development cycle.
 
@@
 
-All authors have contributed code to SciPy, and all authors reviewed the manuscript.
-In the author list, the names of authors who contributed text to the manuscript are indicated with $\dagger$, and the primary manuscript editors are denoted with $\ddagger$.
+P.V.
+is SciPy's Benevolent Dictator for Life, R.G.
+is the Chair of the SciPy Steering Council, and together they have led the development of SciPy for over 10 years.
+T.E.O., E.J., and P.P.
+created SciPy.
+T.J.R.
+and M.H.
+composed the manuscript with input from others.
+Other named authors are SciPy Core Developers.
+All authors have contributed significant code, documentation, and/or expertise to the SciPy project.
+All authors reviewed the manuscript.
 
@@
 
+\section*{Consortium}
+\subsection*{SciPy Contributors}
+\input{consortium}
+
 \end{document}

### 2019-07-03 scipy-1.0/covlett.tex
Update scipy-1.0/covlett.tex
Co-Authored-By: Joel Nothman <joel.nothman@gmail.com>
Update scipy-1.0/covlett.tex
Co-Authored-By: Alessandro Pietro Bardelli <apbard@users.noreply.github.com>

@@
 \emph{Scientific Reports}. SciPy is the fundamental scientific 
-algorithm computing library in the Python programming language. 
+computing library in the Python programming language. 
 Hundreds of experts have contributed to our library, which
-has been download millions of times and has over 100,000 dependent
+has been downloaded millions of times and has over 100,000 dependent
 code repositories.

### 2019-07-03 scipy-1.0/covlett.tex
Update scipy-1.0/covlett.tex
Co-Authored-By: Alessandro Pietro Bardelli <apbard@users.noreply.github.com>

@@
 We have specifically chosen \emph{Scientific Reports} because
-the breadth of acceptable topics matches the breadth of algorithms
+the breadth of acceptable topics matches the variety of algorithms
 our library provides. High-profile manuscripts citing SciPy range

### 2019-07-05 scipy-1.0/paper.tex
SciPy Contributors -> SciPy 1.0 Contributors
Added Acknowledgements

--- 2019-07-01 scipy-1.0/paper.tex
@@
 
-Python, Cython, Fortran, C and C++ are the programming languages used to
-implement scientific algorithms in the SciPy library. An analysis of our codebase
-using the \texttt{linguist} library\cite{linguistref} provides a
-detailed breakdown of SciPy's composition (Table~\ref{tab:linguist}).
-
-\begin{table}
-\centering
-\begin{tabular}{p{3cm}S[table-format = <0.1]}
-\toprule
-Language & {Percent}\\
-\midrule
-Python  & 49.5 \\
-Fortran & 25.6 \\
-C       & 19.5 \\
-Cython  & 3.0  \\
-C++     & 2.3  \\[\defaultaddspace]
-{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
-\bottomrule
-\end{tabular}
-\caption{Language composition of SciPy codebase: lines of code in each
-programming language as determined by the \texttt{linguist} package.
-The last row denotes tools used in supporting roles in tests,
-building, and documentation.}
-\label{tab:linguist}
-\end{table}
+According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 50\% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
+The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.
+Cython) and the usage of languages (and their libraries) that have proven reliable and performant over many decades.
 

### 2019-07-05 scipy-1.0/covlett.tex
Draft in 3 prospective reviewers for cover letter
Use more appropriate suggestion for 3rd reviewer; switch to enumerate

@@
 
+\begin{enumerate}
+    \item Alan Edelman, \texttt{edelman@math.mit.edu}, co-creator of Julia
+    programming language and Faculty at MIT.
+    \item William Stein, \texttt{wstein@gmail.com}, originator and
+    lead developer of SageMath.
+    \item Kazunori Akiyama, \texttt{kazu@haystack.mit.edu}, first author
+    of event horizon imaging paper that cites SciPy.
+\end{enumerate}
+
 Thank you for your consideration.  We look forward to

### 2019-07-05 scipy-1.0/paper.tex
MAINT: broaden Data Availability statement to include repos cited by manuscript

@@
 AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
-unit tests in the test suite, which is written for usage with the \texttt{pytest}
-framework, and with 87\% and 45\% line coverages for Python and compiled
-code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There
+are more than $13,000$ unit tests in the test suite, which is written for usage
+with the \texttt{pytest}\cite{pytest} framework.
+Test coverage at the SciPy 1.0 release point (Figure~\ref{fig:coverage}) was at
+87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
+45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
+Documentation for the code is automatically built and published by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 of the removal of $\approx 61,000$ lines of compiled code for SciPy
-v0.14, the volume of both compiled and Python
+v0.14, the volume of both compiled (C, C++, and Fortran) and Python
 code has increased between releases, as has the number of lines
-covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
-compiled code with test coverage near $45\%$ and $106,878$ lines of 
-Python code with test coverage near $87\%$.\\
-The reconstitution of historical
-build and test environments was performed using a docker-based approach
-\cite{scipy-cov}. Analysis of lines of source code covered is performed 
-automatically in our continuous integration suite using the \texttt{pytest-cov}
-and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
-inclusion of additional vendored code.}
+covered by unit tests. In SciPy v1.0.0, tests covered approximately
+45\% of $462,574$ lines of compiled code (many of which
+are automatically generated from Cython source), and nearly 87\% of
+106,878 lines of Python code.\\
+Coverage analysis is performed automatically in our continuous integration
+suite using the \texttt{pytest-cov} and \texttt{gcov} libraries for 
+Python and compiled code, respectively.
+The reconstitution of historical build and test environments for this
+figure was performed using a Docker-based approach\cite{scipy-cov}.
+}
 \label{fig:coverage}
@@
 % The statement should be provided as a separate section (titled 'Data Availability') at the end of the main text, before the 'References' section. 
-All SciPy library source code and all data generated for the current study are available in the SciPy GitHub repository, \url{https://github.com/scipy}.
+All SciPy library source code and most data generated for the current study are available in the SciPy GitHub repository, \url{https://github.com/scipy}.
+Some supporting code and data have also been stored in other public repositories cited by this manuscript.
 

### 2019-07-06 scipy-1.0/scipy-optimize.tex
Reorganizes optimize section

--- 2019-06-21 scipy-1.0/scipy-optimize.tex
@@
 The \texttt{minimize} function provides a unified interface for finding local minima of nonlinear optimization problems.
-Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
+A detailed comparison of the characteristics of all \texttt{minimize} methods is presented with references in Table~\ref{tab:minimize-si}.
+This table illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
+
+In recent SciPy versions, four new methods for unconstrained optimization have been added to \texttt{minimize}: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
 All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
@@
 On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
-A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}; it illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
 

### 2019-07-06 scipy-1.0/core-dev.tex
SciPy Contributors -> SciPy 1.0 Contributors
Mark corresponding authors
MAINT: specify 'the Netherlands' in English for consistency
Updated author list

@@
 \author[10]{Joshua Wilson}
-\author[17,18]{K. Jarrod Millman}
-\author[19]{Nikolay Mayorov}
-\author[20]{Andrew R. J. Nelson}
+\author[15,17]{K. Jarrod Millman}
+\author[18]{Nikolay Mayorov}
+\author[19]{Andrew R. J. Nelson}
 \author[5]{Eric Jones}
 \author[5]{Robert Kern}
-\author[21]{Eric Larson}
-\author[22]{CJ Carey}
+\author[20]{Eric Larson}
+\author[21]{CJ Carey}
 \author[10]{\.{I}lhan Polat}
-\author[23]{Yu Feng}
-\author[24]{Eric W. Moore}
-\author[25]{Jake VanderPlas}
+\author[22]{Yu Feng}
+\author[23]{Eric W. Moore}
+\author[24]{Jake VanderPlas}
 \author[10]{Denis Laxalde}
 \author[10]{Josef Perktold}
-\author[26]{Robert Cimrman}
-\author[27]{Ian Henriksen}
+\author[25]{Robert Cimrman}
+\author[26]{Ian Henriksen}
 \author[10]{E. A. Quintero}
 \author[10]{Charles R Harris}
-\author[28]{Anne M. Archibald}
-\author[29]{Ant\^{o}nio H. Ribeiro}
-\author[30]{Fabian Pedregosa}
-\author[30]{Paul van Mulbregt}
-\author[ ]{SciPy Contributors}
+\author[27]{Anne M. Archibald}
+\author[28]{Ant\^{o}nio H. Ribeiro}
+\author[29]{Fabian Pedregosa}
+\author[ ]{SciPy 1.0 Contributors}
 \affil[1]{University of Jyv\"askyl\"a, FI-40014 University of Jyv\"askyl\"a, Finland}
@@
 \affil[14]{Independent Researcher, NY, USA}
-\affil[15]{Berkeley Institute for Data Science, University of California Berkeley, Berkeley, CA 94720 USA}
+\affil[15]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720, USA}
 \affil[16]{School of Psychology, University of Birmingham, Edgbaston, Birmigham B15 2TT, UK}
-\affil[17]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720 USA}
-\affil[18]{Division of Biostatistics, University of California, Berkeley, CA 94720, USA}
-\affil[19]{WayRay LLC, Skolkovo Innovation Center, Moscow, 143026, Russia}
-\affil[20]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
-\affil[21]{Institute for Learning and Brain Sciences, University of Washington, Seattle, WA 98195, USA}
-\affil[22]{College of Information and Computing Sciences, University of Massachusetts Amherst, Amherst, MA 01002, USA}
-\affil[23]{Berkeley Center for Cosmological Physics, University of California, Berkeley, CA 94720, USA}
-\affil[24]{Bruker Biospin Corp., 15 Fortune Drive, Billerica, MA 01821}
-\affil[25]{University of Washington, Seattle, WA 98195, USA}
-\affil[26]{New Technologies - Research Centre, University of West Bohemia, Univerzitní 8, 306 14, Plzeň, Czech Republic}
-\affil[27]{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin, Austin, Texas 78712-1229, USA}
-\affil[28]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, Nederland}
-\affil[29]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
-\affil[30]{Google, Inc., 355 Main St., Cambridge, MA 02142, USA}
+\affil[17]{Division of Biostatistics, University of California, Berkeley, CA 94720, USA}
+\affil[18]{WayRay LLC, Skolkovo Innovation Center, Moscow, 143026, Russia}
+\affil[19]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
+\affil[20]{Institute for Learning and Brain Sciences, University of Washington, Seattle, WA 98195, USA}
+\affil[21]{College of Information and Computing Sciences, University of Massachusetts Amherst, Amherst, MA 01002, USA}
+\affil[22]{Berkeley Center for Cosmological Physics, University of California, Berkeley, CA 94720, USA}
+\affil[23]{Bruker Biospin Corp., 15 Fortune Drive, Billerica, MA 01821}
+\affil[24]{University of Washington, Seattle, WA 98195, USA}
+\affil[25]{New Technologies - Research Centre, University of West Bohemia, Univerzitní 8, 306 14, Plzeň, Czech Republic}
+\affil[26]{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin, Austin, Texas 78712, USA}
+\affil[27]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, The Netherlands}
+\affil[28]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
+\affil[29]{Google, Inc., 355 Main St., Cambridge, MA 02142, USA}
 \affil[*]{scipy.articles@gmail.com}

### 2019-07-06 scipy-1.0/paper.tex
Updated author list

--- 2019-07-05 scipy-1.0/paper.tex
@@
 
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\@maketitle}
+ {\noindent
+{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\color{color1}\large\sffamily\textbf{ABSTRACT}}}}
+ {\newpage\noindent
+{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\color{color1}\large\sffamily\textbf{ABSTRACT}}}}
+ {}{}
+\makeatother
+
 \title{SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python}
 
-\author[1]{Pauli Virtanen}
-\author[2,*]{Ralf Gommers}
-\author[3,4]{Tyler Reddy}
-\author[5]{Anne Archibald}
-\author[6]{Andrew R. J. Nelson}
-\author[7]{Charles Harris}
-\author[8]{CJ Carey}
-\author[9]{Denis Laxalde}
-\author[10]{Eric Larson}
-\author[11]{Eric Moore}
-\author[12]{Eric Quintero}
-\author[13]{Evgeni Burovski}
-\author[14]{Jaime Fernández del Río}
-\author[15]{Josef Perktold}
-\author[16]{Josh Wilson}
-\author[17]{Matthew Brett}
-\author[18]{Nikolay Mayorov}
-\author[19]{Warren Weckesser}
-\author[20]{Matt Haberland}
-\author[21]{Scott Sievert}
-\author[22]{Yu Feng}
-\author[23]{Antonio Horta Ribeiro}
-\author[24]{Ian Henriksen}
-\author[3,25]{K. Jarrod Millman}
-\author[3]{St\'efan J. van der Walt}
-\author[26]{\.{I}lhan Polat}
-
-\affil[1]{Affiliation, department, city, postcode, country}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA, 94720, USA}
-\affil[4]{Los Alamos National Laboratory,
-	  Theoretical Division 6,
-          Los Alamos, NM, 87545, USA}
-\affil[5]{Affiliation, department, city, postcode, country}
-\affil[6]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
-\affil[7]{Affiliation, department, city, postcode, country}
-\affil[8]{Affiliation, department, city, postcode, country}
-\affil[9]{Affiliation, department, city, postcode, country}
-\affil[10]{University of Washington, Institute for Learning and Brain Sciences, Seattle, WA, 98195, USA}
-\affil[11]{Affiliation, department, city, postcode, country}
-\affil[12]{Affiliation, department, city, postcode, country}
-\affil[13]{Affiliation, department, city, postcode, country}
-\affil[14]{Affiliation, department, city, postcode, country}
-\affil[15]{Affiliation, department, city, postcode, country}
-\affil[16]{Affiliation, department, city, postcode, country}
-\affil[17]{Affiliation, department, city, postcode, country}
-\affil[18]{Affiliation, department, city, postcode, country}
-\affil[19]{Affiliation, department, city, postcode, country}
-\affil[20]{BioResource and Agricultural Engineering, California Polytechnic State University, San Luis Obispo, CA, 93407, USA}
-\affil[21]{Affiliation, department, city, postcode, country}
-\affil[22]{Affiliation, department, city, postcode, country}
-\affil[23]{Affiliation, department, city, postcode, country}
-\affil[24]{University of Texas at Austin,
-           Oden Institute for Computational Engineering and Sciences,
-	   Austin, TX, 78712, USA}
-\affil[25]{Division of Biostatistics, University of California,
-  Berkeley, CA, 94720, USA}
-\affil[26]{Affiliation, department, city, postcode, country}
-
-\affil[*]{ralf.gommers@gmail.com}
+\input{core-dev}
 
@@
 
+\section*{Acknowledgements}
+
+We thank everyone who has contributed to SciPy 1.0, from those who have posted one comment about an issue through those who have made several small patches and beyond!
+
 \section*{Author Contributions Statement}
 
-All authors have contributed code to SciPy, and all authors reviewed the manuscript.
-In the author list, the names of authors who contributed text to the manuscript are indicated with $\dagger$, and the primary manuscript editors are denoted with $\ddagger$.
+P.V.
+is SciPy's Benevolent Dictator for Life, R.G.
+is the Chair of the SciPy Steering Council, and together they have led the development of SciPy for over 10 years.
+T.E.O., E.J., and P.P.
+created SciPy.
+T.J.R.
+and M.H.
+composed the manuscript with input from others.
+Other named authors are SciPy Core Developers.
+All authors have contributed significant code, documentation, and/or expertise to the SciPy project.
+All authors reviewed the manuscript.
 
@@
 
+\section*{Consortium}
+\subsection*{SciPy 1.0 Contributors}
+\input{consortium}
+
 \end{document}

### 2019-07-06 scipy-1.0/consortium.tex

@@
 {\bf
-Aditya Vijaykumar$^{31,32}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{33}$, Andreas Kloeckner$^{34}$, Antony Lee$^{35}$, Ariel Rokem$^{36}$, C.
-Nathan Woods$^{9}$, Chad Fulton$^{37}$, Charles Masson$^{38}$, Christian Brueffer$^{39}$, Christian H\"aggstr\"om$^{40}$, Clark Fitzgerald$^{41}$, David A.
-Nicholson$^{42}$, David R Hagen$^{43}$, Dmitrii V.
-Pasechnik$^{44}$, Emanuele Olivetti$^{45}$, Eric Martin$^{10}$, Eric Wieser$^{46}$, Fabrice Silva$^{47}$, Felix Lenders$^{48,49,50}$, Florian Wilhelm$^{51}$, G.
+Aditya Vijaykumar$^{30,31}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{32}$, Andreas Kloeckner$^{33}$, Anthony Scopatz$^{2}$, Antony Lee$^{34}$, Ariel Rokem$^{35}$, C.
+Nathan Woods$^{9}$, Chad Fulton$^{36}$, Charles Masson$^{37}$, Christian Brueffer$^{38}$, Christian H\"aggstr\"om$^{39}$, Clark Fitzgerald$^{40}$, David A.
+Nicholson$^{41}$, David R Hagen$^{42}$, Dmitrii V.
+Pasechnik$^{43}$, Emanuele Olivetti$^{44}$, Eric Martin$^{10}$, Eric Wieser$^{45}$, Fabrice Silva$^{46}$, Felix Lenders$^{47,48,49}$, Florian Wilhelm$^{50}$, G.
 Young$^{10}$, Gavin A.
-Price$^{52}$, Gert-Ludwig Ingold$^{53}$, Gregory E.
-Allen$^{54}$, Gregory R.
-Lee$^{55,56}$, Herv\'e Audren$^{57}$, Irvin Probst$^{58}$, J\"org P.
-Dietrich$^{59,60}$, Jacob Silterra$^{61}$, James T Webber$^{62}$, Janko Slavi\v{c}$^{63}$, Joel Nothman$^{64}$, Johannes Buchner$^{65,66}$, Johannes Kulick$^{67}$, Johannes L.
-Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, Joscha Reimer$^{69}$, Joseph Harrington$^{70}$, Juan Luis Cano Rodr'iguez$^{71}$, Juan Nunez-Iglesias$^{72}$, Justin Kuczynski$^{73}$, Kevin Tritz$^{74}$, Martin Thoma$^{75}$, Matthias K\"ummerer$^{76}$, Maximilian Bolingbroke$^{77}$, Michael Tartre$^{78}$, Mikhail Pak$^{79}$, Nathaniel J.
-Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{80}$, Oleksandr Pavlyk$^{81}$, Per A.
-Brodtkorb$^{82}$, Perry Lee$^{10}$, Robert T.
-McGibbon$^{83}$, Roman Feldbauer$^{84}$, Sam Lewis$^{85}$, Sam Tygier$^{86}$, Scott Sievert$^{87}$, Sebastiano Vigna$^{88}$, Sourav Singh$^{89}$, Stefan Peterson$^{10}$, Surhud More$^{90,91}$, Tadeusz Pudlik$^{92}$, Takuya Oshima$^{93}$, Thomas J.
-Pingel$^{94}$, Thomas P.
-Robitaille$^{95}$, Thomas Spura$^{10}$, Thouis R.
-Jones$^{96}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{97}$, Tom Krauss$^{98}$, Utkarsh Upadhyay$^{99}$, Yaroslav O.
-Halchenko$^{100}$, Yoshiki V\'azquez-Baeza$^{101}$
+Price$^{51}$, Gert-Ludwig Ingold$^{52}$, Gregory E.
+Allen$^{53}$, Gregory R.
+Lee$^{54,55}$, Herv\'e Audren$^{56}$, Irvin Probst$^{57}$, J\"org P.
+Dietrich$^{58,59}$, Jacob Silterra$^{60}$, James T Webber$^{61}$, Janko Slavi\v{c}$^{62}$, Joel Nothman$^{63}$, Johannes Buchner$^{64,65}$, Johannes Kulick$^{66}$, Johannes L.
+Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{67}$, Joscha Reimer$^{68}$, Joseph Harrington$^{69}$, Juan Luis Cano Rodr'iguez$^{70}$, Juan Nunez-Iglesias$^{71}$, Justin Kuczynski$^{72}$, Kevin Tritz$^{73}$, Martin Thoma$^{74}$, Matthias K\"ummerer$^{75}$, Maximilian Bolingbroke$^{76}$, Michael Tartre$^{77}$, Mikhail Pak$^{78}$, Nathaniel J.
+Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{79}$, Oleksandr Pavlyk$^{80}$, Paul van Mulbregt$^{29}$, Per A.
+Brodtkorb$^{81}$, Perry Lee$^{10}$, Robert T.
+McGibbon$^{82}$, Roman Feldbauer$^{83}$, Sam Lewis$^{84}$, Sam Tygier$^{85}$, Scott Sievert$^{86}$, Sebastiano Vigna$^{87}$, Sourav Singh$^{88}$, Stefan Peterson$^{10}$, Surhud More$^{89,90}$, Tadeusz Pudlik$^{91}$, Takuya Oshima$^{92}$, Thomas J.
+Pingel$^{93}$, Thomas P.
+Robitaille$^{94}$, Thomas Spura$^{10}$, Thouis R.
+Jones$^{95}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{96}$, Tom Krauss$^{97}$, Utkarsh Upadhyay$^{98}$, Yaroslav O.
+Halchenko$^{99}$, Yoshiki V\'azquez-Baeza$^{100}$
 }
@@
 \hfill \break
-$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N.
-Goodwin Ave, Urbana, IL 616801, USA, $^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave.
-NW, Washington, DC 20551, U.S.A., $^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{42}$Emory University, Atlanta, GA 30322, USA, $^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{49}$ABB Corporate Research, ABB AG, Wallstadter Str.
-59 68526 Ladenburg, Germany, $^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{51}$Independent Researcher, Cologne, Germany, $^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O.
-Box 8029 Austin, TX 78713, USA, $^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr.
-1, 81679 M\"unchen, Germany, $^{60}$Excellence Cluster Universe, Boltzmannstr.
+$^{30}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{31}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{32}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{33}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N.
+Goodwin Ave, Urbana, IL 616801, USA, $^{34}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{35}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{36}$Federal Reserve Board of Governors,  20th and Constitution Ave.
+NW, Washington, DC 20551, U.S.A., $^{37}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{38}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{39}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{40}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{41}$Emory University, Atlanta, GA 30322, USA, $^{42}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{43}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{44}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{45}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{46}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{47}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{48}$ABB Corporate Research, ABB AG, Wallstadter Str.
+59 68526 Ladenburg, Germany, $^{49}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{50}$Independent Researcher, Cologne, Germany, $^{51}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{52}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{53}$Applied Research Laboratories, The University of Texas at Austin, P.O.
+Box 8029 Austin, TX 78713, USA, $^{54}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{55}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{56}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{57}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{58}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr.
+1, 81679 M\"unchen, Germany, $^{59}$Excellence Cluster Universe, Boltzmannstr.
 2, 85748 Garching b.
-M\"unchen, Germany, $^{61}$Independent Researcher, USA, $^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr.
-38, 70569 Stuttgart, Germany, $^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{71}$Independent Researcher, Spain, $^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N.
-Charles Street, Baltimore, MD 21218, USA, $^{75}$Independent Researcher, Munich, Germany, $^{76}$University of T\"ubingen, T\"ubingen, Germany, $^{77}$Independent Researcher, United Kingdom, $^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr.
-15, 85748 Garching, Germany, $^{80}$Independent Researcher, Berlin, Germany, $^{81}$Intel Corp., Austin, TX 78746, USA, $^{82}$Independent Researcher, Horten, Norway, $^{83}$D.
+M\"unchen, Germany, $^{60}$Independent Researcher, USA, $^{61}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{62}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{63}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{64}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{65}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{66}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr.
+38, 70569 Stuttgart, Germany, $^{67}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{68}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{69}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{70}$Independent Researcher, Spain, $^{71}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{72}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{73}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N.
+Charles Street, Baltimore, MD 21218, USA, $^{74}$Independent Researcher, Munich, Germany, $^{75}$University of T\"ubingen, T\"ubingen, Germany, $^{76}$Independent Researcher, United Kingdom, $^{77}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{78}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr.
+15, 85748 Garching, Germany, $^{79}$Independent Researcher, Berlin, Germany, $^{80}$Intel Corp., Austin, TX 78746, USA, $^{81}$Independent Researcher, Horten, Norway, $^{82}$D.
 E.
-Shaw Research, New York, NY 10036, USA, $^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{85}$Independent Researcher, Melbourne, Australia, $^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{89}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{92}$Waymo LLC, Mountain View, CA 94043, USA, $^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{96}$Broad Institute, Cambridge, MA 02142, USA, $^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{99}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA
+Shaw Research, New York, NY 10036, USA, $^{83}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{84}$Independent Researcher, Melbourne, Australia, $^{85}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{86}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{87}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{88}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{89}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{90}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{91}$Waymo LLC, Mountain View, CA 94043, USA, $^{92}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{93}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{94}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{95}$Broad Institute, Cambridge, MA 02142, USA, $^{96}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{97}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{98}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{99}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{100}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA

### 2019-07-06 scipy-1.0/paper.tex
Small edit on matmul operator

--- 2019-07-06 scipy-1.0/paper.tex
@@
 From a new features standpoint, \texttt{scipy.sparse} matrices and linear
-operators now support the Python matrix multiplication (@) operator.
+operators now support the Python matrix multiplication operator (\texttt{@}).
 We've added \texttt{scipy.sparse.norm} and
@@
 AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
-include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There
-are more than $13,000$ unit tests in the test suite, which is written for usage
-with the \texttt{pytest}\cite{pytest} framework.
-Test coverage at the SciPy 1.0 release point (Figure~\ref{fig:coverage}) was at
-87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
-45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
-Documentation for the code is automatically built and published by
+include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
+unit tests in the test suite, which is written for usage with the \texttt{pytest}
+framework, and with 87\% and 45\% line coverages for Python and compiled
+code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and published by
 the CircleCI service to facilitate evaluation of documentation changes /
@@
 of the removal of $\approx 61,000$ lines of compiled code for SciPy
-v0.14, the volume of both compiled (C, C++, and Fortran) and Python
+v0.14, the volume of both compiled and Python
 code has increased between releases, as has the number of lines
-covered by unit tests. In SciPy v1.0.0, tests covered approximately
-45\% of $462,574$ lines of compiled code (many of which
-are automatically generated from Cython source), and nearly 87\% of
-106,878 lines of Python code.\\
-Coverage analysis is performed automatically in our continuous integration
-suite using the \texttt{pytest-cov} and \texttt{gcov} libraries for 
-Python and compiled code, respectively.
-The reconstitution of historical build and test environments for this
-figure was performed using a Docker-based approach\cite{scipy-cov}.
-}
+covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
+compiled code with test coverage near $45\%$ and $106,878$ lines of 
+Python code with test coverage near $87\%$.\\
+The reconstitution of historical
+build and test environments was performed using a docker-based approach
+\cite{scipy-cov}. Analysis of lines of source code covered is performed 
+automatically in our continuous integration suite using the \texttt{pytest-cov}
+and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
+inclusion of additional vendored code.}
 \label{fig:coverage}
@@
 % The statement should be provided as a separate section (titled 'Data Availability') at the end of the main text, before the 'References' section. 
-All SciPy library source code and most data generated for the current study are available in the SciPy GitHub repository, \url{https://github.com/scipy}.
-Some supporting code and data have also been stored in other public repositories cited by this manuscript.
+All SciPy library source code and all data generated for the current study are available in the SciPy GitHub repository, \url{https://github.com/scipy}.
 

### 2019-07-06 scipy-1.0/scipy-optimize.tex
Adds references to trust-* paragraph

@@
 
-In recent SciPy versions, four new methods for unconstrained optimization have been added to \texttt{minimize}: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
+In recent SciPy versions, four new methods for unconstrained optimization have been added to \texttt{minimize}: \texttt{dogleg}\cite{powell_new_1970, nocedal_numerical_2006}, \texttt{trust-ncg}\cite{steihaug_conjugate_1983, nocedal_numerical_2006}, \texttt{trust-exact}\cite{conn_trust_2000, more_computing_1983}, and \texttt{trust-krylov}\cite{gould_solving_1999, lenders_trlib:_2016}.
 All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.

### 2019-07-06 scipy-1.0/covlett.tex
MAINT: cover letter spacing improvements
* reduce wasted vertical space in the letter
so that we do not overflow to a second page;
also, include link to insructions for adjusting
this spacing
MAINT: apply Stefan's revision to cover letter

@@
 \documentclass[10pt,stdletter,dateno]{newlfm}
+
+% to adjust vertical spacing properties of
+% newlfm class see: https://tex.stackexchange.com/a/65533/43006
+\topmarginsize{2mm}
+\topmarginskip{0in}
+\headermarginsize{0in}
+\headermarginskip{0in}
+
 \usepackage{kpfonts}
@@
 \newlfmP{headermarginskip=10pt}
-\newlfmP{sigsize=50pt}
+\newlfmP{sigsize=30pt}
 \newlfmP{dateskipafter=20pt}
@@
 from the biosciences to recent reports on gravitational waves and
-black holes. These citations target a website, and we now think
-it is well past time that our mature library has a peer-reviewed
-journal article describing it.
+black holes. Lacking a better alternative, these citations all point
+to the SciPy website; instead, we think the package now merits a
+formal publication in a highly visible journal.
 

### 2019-07-08 scipy-1.0/paper.tex
Small fixes.
- Fix typo.
- Consistently spell "tool stack".
- Use Apple's capitalization of macOS.

--- 2019-07-06 scipy-1.0/paper.tex
@@
 
-Yet the philosophical motivations behind a fully open toolstack, combined
+Yet the philosophical motivations behind a fully open tool stack, combined
 with an excited, friendly community with a singular focus, have
@@
 \cite{PhysRevLett.116.061102}, the fact that SciPy is shipped directly with
-MacOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
+macOS and in the Intel Distribution for Python\cite{intel-python}, and that SciPy is used
 by 47\% of all machine learning projects on GitHub\cite{octoverse-scipy}.

### 2019-07-08 scipy-1.0/scipy-optimize.tex
optimize: fix minor typos

--- 2019-07-06 scipy-1.0/scipy-optimize.tex
@@
 A presolve routine \cite{andersen1995presolving} solves trivial problems and otherwise performs problem simplifications, such as bound tightening and removal of fixed variables, and one of several routines for eliminating redundant equality constraints is automatically chosen to reduce the chance of numerical difficulties caused by singular matrices.
-Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers---often for the same system with multiple right hand sides due to the predictor-corrector approach---provide speed sufficient for problems with tens of thousands of variables and constraints.
+Although the main solver implementation is pure Python, end-to-end sparse matrix support and heavy use of SciPy's compiled linear system solvers---often for the same system with multiple right hand sides due to the predictor-corrector approach---provide sufficient speed for problems with tens of thousands of variables and constraints.
 
-Compared to the previously implemented simplex method, the new interior-point method is faster for all but the smallest problems, and is suitable for solving medium- and large-sized problems on which the existing simplex implementation fails.
+Compared to the previously implemented simplex method, the new interior-point method is faster for all but the smallest problems, and is suitable for solving medium and large-sized problems on which the existing simplex implementation fails.
 However, the interior point method typically returns a solution near the center of an optimal face, yet basic solutions are often preferred for sensitivity analysis and for use in mixed integer programming algorithms.
@@
 The \texttt{minimize} function provides a unified interface for finding local minima of nonlinear optimization problems.
-A detailed comparison of the characteristics of all \texttt{minimize} methods is presented with references in Table~\ref{tab:minimize-si}.
-This table illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
-
-In recent SciPy versions, four new methods for unconstrained optimization have been added to \texttt{minimize}: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
+Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
 All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
-For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second derivative Hessian matrix to be stored and factored every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
+For instance, \texttt{trust-exact} achieves fast convergence by solving the trust-region subproblem almost exactly, but it requires the second derivative Hessian matrix to be stored and factored at every iteration, which may preclude the solution of large problems ($\geq 1000$ variables).
 On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
+A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}; it illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
 

### 2019-07-09 scipy-1.0/scipy-optimize.tex

--- 2019-07-06 scipy-1.0/scipy-optimize.tex
@@
 The \texttt{minimize} function provides a unified interface for finding local minima of nonlinear optimization problems.
-A detailed comparison of the characteristics of all \texttt{minimize} methods is presented with references in Table~\ref{tab:minimize-si}.
-This table illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
-
-In recent SciPy versions, four new methods for unconstrained optimization have been added to \texttt{minimize}: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
+Four new methods for unconstrained optimization were added to \texttt{minimize} in recent versions of SciPy: \texttt{dogleg}, \texttt{trust-ncg}, \texttt{trust-exact}, and \texttt{trust-krylov}.
 All are trust-region methods that build a local model of the objective function based on first and second derivative information, approximate the best point within a local ``trust region'', and iterate until a local minimum of the original objective function is reached, but each has unique characteristics that make it appropriate for certain types of problems.
@@
 On the other hand, \texttt{trust-ncg} and \texttt{trust-krylov} are well-suited to large-scale optimization problems because they do not need to store and factor the Hessian explicitly, instead using second derivative information in a faster, approximate way.
+A detailed comparison of the characteristics of all \texttt{minimize} methods is presented in Table~\ref{tab:minimize-si}; it illustrates the level of completeness that SciPy aims for when covering a numerical method or topic.
 
@@
     indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
-    with $\sim$ the algorithm does not requires second-order derivatives from
+    with $\sim$ the algorithm does not require second-order derivatives from
     the user; it computes an approximation internally and uses it to accelerate method convergence.

### 2019-07-19 scipy-1.0/consortium.tex
Added author
Fix non-ASCII last name

@@
 Dietrich$^{58,59}$, Jacob Silterra$^{60}$, James T Webber$^{61}$, Janko Slavi\v{c}$^{62}$, Joel Nothman$^{63}$, Johannes Buchner$^{64,65}$, Johannes Kulick$^{66}$, Johannes L.
-Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{67}$, Joscha Reimer$^{68}$, Joseph Harrington$^{69}$, Juan Luis Cano Rodr'iguez$^{70}$, Juan Nunez-Iglesias$^{71}$, Justin Kuczynski$^{72}$, Kevin Tritz$^{73}$, Martin Thoma$^{74}$, Matthias K\"ummerer$^{75}$, Maximilian Bolingbroke$^{76}$, Michael Tartre$^{77}$, Mikhail Pak$^{78}$, Nathaniel J.
-Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{79}$, Oleksandr Pavlyk$^{80}$, Paul van Mulbregt$^{29}$, Per A.
+Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{67}$, Joscha Reimer$^{68}$, Joseph Harrington$^{69}$, Juan Luis Cano Rodr\'iguez$^{70}$, Juan Nunez-Iglesias$^{71}$, Justin Kuczynski$^{72}$, Kevin Tritz$^{73}$, Martin Thoma$^{74}$, Matthias K\"ummerer$^{75}$, Maximilian Bolingbroke$^{76}$, Michael Tartre$^{77}$, Mikhail Pak$^{78}$, Nathaniel J.
+Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{79}$, Oleksandr Pavlyk$^{80}$, Per A.
 Brodtkorb$^{81}$, Perry Lee$^{10}$, Robert T.

### 2019-07-21 scipy-1.0/paper.tex
Python matrix multiplaction operator font fix
SciPy provides sparse arrays -> matrices

@@
 classes of problems; it also provides
-specialized data structures, such as sparse arrays
+specialized data structures, such as sparse matrices
 and $k$-dimensional trees. SciPy is built on top of 

### 2019-07-21 scipy-1.0/paper.tex

--- 2019-07-08 scipy-1.0/paper.tex
@@
 and $k$-dimensional trees. SciPy is built on top of 
-NumPy\cite{vanderwalt2011numpy,Oliphant-2015},
+NumPy\cite{oliphant2006guide, vanderwalt2011numpy},
 which provides array data structures and related fast numerical routines, and
@@
 extension modules that wrapped Fortran and C libraries such as
-ODEPACK\cite{citeulike:2644528}, QUADPACK\cite{1983qspa.book.....P}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
+ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
 equations, integrate differential equations, and fit splines (the
@@
 enabled collaborative documentation development in a wiki-like fashion.
-The SciPy Documentation Project \cite{van2008scipy, harrington2008scipy} 
+The SciPy Documentation Project \cite{vanderwalt2008scipy, harrington2008scipy} 
 used these tools to complete documentation of SciPy's user-facing
@@
 FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms, 
-ODEPACK\cite{citeulike:2644528}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{1983qspa.book.....P}, % for numerical integration of one-dimensional functions,
+ODEPACK\cite{hindmarsh1983odepack}, % for solving ordinary differential equation initial value problems,
+QUADPACK\cite{piessens1983quadpack}, % for numerical integration of one-dimensional functions,
 FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,

### 2019-07-21 scipy-1.0/poly.tex

@@
 Fortran library by P.~Dierckx \cite{Dierckx:1993:CSF:151103, FITPACK} for
-univariate interpolation and approximation of data. The original monolithic
+univariate interpolation and approximation of data, but the original monolithic
 design and API for interaction between SciPy and FITPACK was limiting for both
-users and developers, as detailed in the Supporting Information.
+users and developers.
 

### 2019-07-21 scipy-1.0/consortium.tex
Corrected author list

--- 2019-07-02 scipy-1.0/consortium.tex
@@
-{\bf
-Aditya Vijaykumar$^{31,32}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{33}$, Andreas Kloeckner$^{34}$, Antony Lee$^{35}$, Ariel Rokem$^{36}$, C.
-Nathan Woods$^{9}$, Chad Fulton$^{37}$, Charles Masson$^{38}$, Christian Brueffer$^{39}$, Christian H\"aggstr\"om$^{40}$, Clark Fitzgerald$^{41}$, David A.
-Nicholson$^{42}$, David R Hagen$^{43}$, Dmitrii V.
-Pasechnik$^{44}$, Emanuele Olivetti$^{45}$, Eric Martin$^{10}$, Eric Wieser$^{46}$, Fabrice Silva$^{47}$, Felix Lenders$^{48,49,50}$, Florian Wilhelm$^{51}$, G.
-Young$^{10}$, Gavin A.
-Price$^{52}$, Gert-Ludwig Ingold$^{53}$, Gregory E.
-Allen$^{54}$, Gregory R.
-Lee$^{55,56}$, Herv\'e Audren$^{57}$, Irvin Probst$^{58}$, J\"org P.
-Dietrich$^{59,60}$, Jacob Silterra$^{61}$, James T Webber$^{62}$, Janko Slavi\v{c}$^{63}$, Joel Nothman$^{64}$, Johannes Buchner$^{65,66}$, Johannes Kulick$^{67}$, Johannes L.
-Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, Joscha Reimer$^{69}$, Joseph Harrington$^{70}$, Juan Luis Cano Rodr'iguez$^{71}$, Juan Nunez-Iglesias$^{72}$, Justin Kuczynski$^{73}$, Kevin Tritz$^{74}$, Martin Thoma$^{75}$, Matthias K\"ummerer$^{76}$, Maximilian Bolingbroke$^{77}$, Michael Tartre$^{78}$, Mikhail Pak$^{79}$, Nathaniel J.
-Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{80}$, Oleksandr Pavlyk$^{81}$, Per A.
-Brodtkorb$^{82}$, Perry Lee$^{10}$, Robert T.
-McGibbon$^{83}$, Roman Feldbauer$^{84}$, Sam Lewis$^{85}$, Sam Tygier$^{86}$, Scott Sievert$^{87}$, Sebastiano Vigna$^{88}$, Sourav Singh$^{89}$, Stefan Peterson$^{10}$, Surhud More$^{90,91}$, Tadeusz Pudlik$^{92}$, Takuya Oshima$^{93}$, Thomas J.
-Pingel$^{94}$, Thomas P.
-Robitaille$^{95}$, Thomas Spura$^{10}$, Thouis R.
-Jones$^{96}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{97}$, Tom Krauss$^{98}$, Utkarsh Upadhyay$^{99}$, Yaroslav O.
-Halchenko$^{100}$, Yoshiki V\'azquez-Baeza$^{101}$
+{\bfseries
+Aditya Vijaykumar$^{31,32}$, 
+Alessandro Pietro Bardelli$^{10}$, 
+Alex Rothberg$^{10}$, 
+Andreas Hilboll$^{33}$, 
+Andreas Kloeckner$^{34}$, 
+Anthony Scopatz$^{2}$, 
+Antony Lee$^{35}$, 
+Ariel Rokem$^{36}$, 
+C. Nathan Woods$^{9}$, 
+Chad Fulton$^{37}$, 
+Charles Masson$^{38}$, 
+Christian Brueffer$^{39}$, 
+Christian H\"aggstr\"om$^{40}$, 
+Clark Fitzgerald$^{41}$, 
+David A. Nicholson$^{42}$, 
+David R Hagen$^{43}$, 
+Dmitrii V. Pasechnik$^{44}$, 
+Emanuele Olivetti$^{45}$, 
+Eric Martin$^{10}$, 
+Eric Wieser$^{46}$, 
+Fabrice Silva$^{47}$, 
+Felix Lenders$^{48,49,50}$, 
+Florian Wilhelm$^{51}$, 
+G. Young$^{10}$, 
+Gavin A. Price$^{52}$, 
+Gert-Ludwig Ingold$^{53}$, 
+Gregory E. Allen$^{54}$, 
+Gregory R. Lee$^{55,56}$, 
+Herv\'e Audren$^{57}$, 
+Irvin Probst$^{58}$, 
+J\"org P. Dietrich$^{59,60}$, 
+Jacob Silterra$^{61}$, 
+James T Webber$^{62}$, 
+Janko Slavi\v{c}$^{63}$, 
+Joel Nothman$^{64}$, 
+Johannes Buchner$^{65,66}$, 
+Johannes Kulick$^{67}$, 
+Johannes L. Sch\"{o}nberger$^{10}$, 
+Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, 
+Joscha Reimer$^{69}$, 
+Joseph Harrington$^{70}$, 
+Juan Luis Cano Rodr\'iguez$^{71}$, 
+Juan Nunez-Iglesias$^{72}$, 
+Justin Kuczynski$^{73}$, 
+Kevin Tritz$^{74}$, 
+Martin Thoma$^{75}$, 
+Matthias K\"ummerer$^{76}$, 
+Maximilian Bolingbroke$^{77}$, 
+Michael Tartre$^{78}$, 
+Mikhail Pak$^{79}$, 
+Nathaniel J. Smith$^{10}$, 
+Nikolai Nowaczyk$^{10}$, 
+Nikolay Shebanov$^{80}$, 
+Oleksandr Pavlyk$^{81}$, 
+Per A. Brodtkorb$^{82}$, 
+Perry Lee$^{10}$, 
+Robert T. McGibbon$^{83}$, 
+Roman Feldbauer$^{84}$, 
+Sam Lewis$^{85}$, 
+Sam Tygier$^{86}$, 
+Scott Sievert$^{87}$, 
+Sebastiano Vigna$^{88}$, 
+Stefan Peterson$^{10}$, 
+Surhud More$^{89,90}$, 
+Tadeusz Pudlik$^{91}$, 
+Takuya Oshima$^{92}$, 
+Thomas J. Pingel$^{93}$, 
+Thomas P. Robitaille$^{94}$, 
+Thomas Spura$^{10}$, 
+Thouis R. Jones$^{95}$, 
+Tim Cera$^{10}$, 
+Tim Leslie$^{10}$, 
+Tiziano Zito$^{96}$, 
+Tom Krauss$^{97}$, 
+Utkarsh Upadhyay$^{98}$, 
+Yaroslav O. Halchenko$^{99}$, 
+Yoshiki V\'azquez-Baeza$^{100}$
 }
@@
 \hfill \break
-$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N.
-Goodwin Ave, Urbana, IL 616801, USA, $^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave.
-NW, Washington, DC 20551, U.S.A., $^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{42}$Emory University, Atlanta, GA 30322, USA, $^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{49}$ABB Corporate Research, ABB AG, Wallstadter Str.
-59 68526 Ladenburg, Germany, $^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{51}$Independent Researcher, Cologne, Germany, $^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O.
-Box 8029 Austin, TX 78713, USA, $^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr.
-1, 81679 M\"unchen, Germany, $^{60}$Excellence Cluster Universe, Boltzmannstr.
-2, 85748 Garching b.
-M\"unchen, Germany, $^{61}$Independent Researcher, USA, $^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr.
-38, 70569 Stuttgart, Germany, $^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{71}$Independent Researcher, Spain, $^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N.
-Charles Street, Baltimore, MD 21218, USA, $^{75}$Independent Researcher, Munich, Germany, $^{76}$University of T\"ubingen, T\"ubingen, Germany, $^{77}$Independent Researcher, United Kingdom, $^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr.
-15, 85748 Garching, Germany, $^{80}$Independent Researcher, Berlin, Germany, $^{81}$Intel Corp., Austin, TX 78746, USA, $^{82}$Independent Researcher, Horten, Norway, $^{83}$D.
-E.
-Shaw Research, New York, NY 10036, USA, $^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{85}$Independent Researcher, Melbourne, Australia, $^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{89}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{92}$Waymo LLC, Mountain View, CA 94043, USA, $^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{96}$Broad Institute, Cambridge, MA 02142, USA, $^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{99}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA
+$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, 
+$^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, 
+$^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, 
+$^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, 
+$^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, 
+$^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, 
+$^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., 
+$^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, 
+$^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, 
+$^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, 
+$^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, 
+$^{42}$Emory University, Atlanta, GA 30322, USA, 
+$^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, 
+$^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, 
+$^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, 
+$^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, 
+$^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, 
+$^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, 
+$^{49}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, 
+$^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, 
+$^{51}$Independent Researcher, Cologne, Germany, 
+$^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, 
+$^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, 
+$^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, 
+$^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, 
+$^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, 
+$^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, 
+$^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, 
+$^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, 
+$^{60}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, 
+$^{61}$Independent Researcher, USA, 
+$^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, 
+$^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, 
+$^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, 
+$^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, 
+$^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, 
+$^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, 
+$^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, 
+$^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, 
+$^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, 
+$^{71}$Independent Researcher, Spain, 
+$^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, 
+$^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, 
+$^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, 
+$^{75}$Independent Researcher, Munich, Germany, 
+$^{76}$University of T\"ubingen, T\"ubingen, Germany, 
+$^{77}$Independent Researcher, United Kingdom, 
+$^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, 
+$^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, 
+$^{80}$Independent Researcher, Berlin, Germany, 
+$^{81}$Intel Corp., Austin, TX 78746, USA, 
+$^{82}$Independent Researcher, Horten, Norway, 
+$^{83}$D. E. Shaw Research, New York, NY 10036, USA, 
+$^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, 
+$^{85}$Independent Researcher, Melbourne, Australia, 
+$^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, 
+$^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, 
+$^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, 
+$^{89}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, 
+$^{90}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, 
+$^{91}$Waymo LLC, Mountain View, CA 94043, USA, 
+$^{92}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, 
+$^{93}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, 
+$^{94}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, 
+$^{95}$Broad Institute, Cambridge, MA 02142, USA, 
+$^{96}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, 
+$^{97}$Epiq Solutions, Schaumburg, IL 60173, USA, 
+$^{98}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, 
+$^{99}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, 
+$^{100}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA

### 2019-07-21 scipy-1.0/covlett.tex
MAINT: adjust covlett for Nature Methods
* adjust the cover letter in preparation
for submission to Nature Methods after
receiving Editorial bounce-back from
Scientific Reports ("out of scope")

* also included are formatting changes
used for the previous submission cover letter,
so that it fits on a single page,
but phone numbers have again been censored
to protect PII

@@
 % newlfm class see: https://tex.stackexchange.com/a/65533/43006
-\topmarginsize{2mm}
+%\unprtop{0mm}
+\topmarginsize{0mm}
 \topmarginskip{0in}
@@
 \headermarginskip{0in}
+\MinHead{0in}
+%\dateskipafter{0pt}
+%\dateskipbefore{0pt}
+%\sigskipbefore{0pt}
+%\sigskipafter{0pt}
+\closeskipbefore{1pt}
+%\closeskipafter{0pt}
+%\addrfromskipbefore{0pt}
+\addrfromskipafter{0in}
+\addrtoskipbefore{0in}
+%\addrtoskipafter{0pt}
+\newlfmP{Headlinewd=0pt,Footlinewd=0pt}
+\newlfmP{sigsize=10pt}
+\bottommarginskip{0pt}
+\MinFoot{0pt}
 
@@
 
-\newlfmP{headermarginskip=10pt}
-\newlfmP{sigsize=30pt}
-\newlfmP{dateskipafter=20pt}
-%\newlfmP{addrfromphone}
-%\newlfmP{addrfromemail}
-%\PhrPhone{Phone}
-%\PhrEmail{Email}
-
 \namefrom{Ralf Gommers,
@@
     \today\\[10pt]
+    Ralf Gommers,\\
     Quansight Labs,\\
-    The Netherlands\\
+    The Netherlands,\\
+    \texttt{ralf.gommers@gmail.com},\\
+    %phone number\\
     \\
+    Matt Haberland,\\
     BioResource and Agricultural Engineering,\\
     California Polytechnic State University,\\
-    San Luis Obispo, CA, 93407, USA\\
+    San Luis Obispo, CA, 93407, USA,\\
+    \texttt{mhaberla@calpoly.edu},\\
+    %phone number\\
     \\
+    Tyler Reddy,\\
     CCS-7 Applied Computer Science Group,\\
     Los Alamos National Laboratory,\\
-    Los Alamos, NM, 87545, USA
+    Los Alamos, NM, 87545, USA,\\
+    \texttt{treddy@lanl.gov},\\
+    %phone number\\
 }
@@
 \addrto{%
-Prof. Syma Khalid,\\
-School of Chemistry,\\
-University of Southampton,\\
-Highfield Campus,\\
-Southampton,\\
-SO17 1BJ, U.K.}
+Dr. Rita Strack,\\
+Nature Methods Editorial Office,\\
+One New York Plaza Suite 4500,\\
+New York, NY 10004,\\
+USA}
 
-\greetto{Dear Prof. Khalid,}
+\greetto{Dear Dr. Strack,}
 \closeline{Sincerely,}
@@
 1.0 release of SciPy for consideration to be published in
-\emph{Scientific Reports}. SciPy is the fundamental scientific 
+\emph{Nature Methods}, as informally discussed.
+SciPy is the fundamental scientific 
 computing library in the Python programming language. 
@@
 has been downloaded millions of times and has over 100,000 dependent
-code repositories.
+code repositories, including many used in the life sciences.
 
-We have specifically chosen \emph{Scientific Reports} because
-the breadth of acceptable topics matches the variety of algorithms
-our library provides. High-profile manuscripts citing SciPy range
-from the biosciences to recent reports on gravitational waves and
+We have specifically chosen \emph{Nature Methods} because
+SciPy lies at the base of the entire ecosystem of scientific methods
+described in the journal scope, and because SciPy is effectively an
+ever-improving methodological library. High-profile manuscripts citing SciPy 
+extend beyond the biosciences to recent reports on gravitational waves and
 black holes. Lacking a better alternative, these citations all point
 to the SciPy website; instead, we think the package now merits a
-formal publication in a highly visible journal.
+formal publication in a highly visible journal. A pre-print will likely
+soon be made available on arXiv.
 
@@
 Thank you for your consideration.  We look forward to
-hearing from you.
-
+    hearing from you.\vspace*{0.1cm}
 \end{newlfm}

### 2019-07-21 scipy-1.0/paper.tex

@@
 From a new features standpoint, \texttt{scipy.sparse} matrices and linear
-operators now support the Python matrix multiplication (@) operator.
+operators now support the Python matrix multiplication (\texttt{@}) operator.
 We've added \texttt{scipy.sparse.norm} and

### 2019-07-22 scipy-1.0/paper.tex
Fix duplicate words.

@@
 %%   programming background (David), others in varying stages of academic/research careers.
-%% - Over time, the project became more more professional:
+%% - Over time, the project became more professional:
 %%   - Standardized release formats
@@
 signals a library's API (Applications Programming Interface) being
-mature enough to be be trusted in production pipelines. This version 
+mature enough to be trusted in production pipelines. This version
 numbering convention, however, belies the history of a project that has

### 2019-07-26 scipy-1.0/consortium.tex
Correct author list

@@
 Charles Masson$^{38}$, 
-Christian Brueffer$^{39}$, 
-Christian H\"aggstr\"om$^{40}$, 
-Clark Fitzgerald$^{41}$, 
-David A. Nicholson$^{42}$, 
-David R Hagen$^{43}$, 
-Dmitrii V. Pasechnik$^{44}$, 
-Emanuele Olivetti$^{45}$, 
+Christian H\"aggstr\"om$^{39}$, 
+Clark Fitzgerald$^{40}$, 
+David A. Nicholson$^{41}$, 
+David R Hagen$^{42}$, 
+Dmitrii V. Pasechnik$^{43}$, 
+Emanuele Olivetti$^{44}$, 
 Eric Martin$^{10}$, 
-Eric Wieser$^{46}$, 
-Fabrice Silva$^{47}$, 
-Felix Lenders$^{48,49,50}$, 
-Florian Wilhelm$^{51}$, 
+Eric Wieser$^{45}$, 
+Fabrice Silva$^{46}$, 
+Felix Lenders$^{47,48,49}$, 
+Florian Wilhelm$^{50}$, 
 G. Young$^{10}$, 
-Gavin A. Price$^{52}$, 
-Gert-Ludwig Ingold$^{53}$, 
-Gregory E. Allen$^{54}$, 
-Gregory R. Lee$^{55,56}$, 
-Herv\'e Audren$^{57}$, 
-Irvin Probst$^{58}$, 
-J\"org P. Dietrich$^{59,60}$, 
-Jacob Silterra$^{61}$, 
-James T Webber$^{62}$, 
-Janko Slavi\v{c}$^{63}$, 
-Joel Nothman$^{64}$, 
-Johannes Buchner$^{65,66}$, 
-Johannes Kulick$^{67}$, 
+Gavin A. Price$^{51}$, 
+Gert-Ludwig Ingold$^{52}$, 
+Gregory E. Allen$^{53}$, 
+Gregory R. Lee$^{54,55}$, 
+Herv\'e Audren$^{56}$, 
+Irvin Probst$^{57}$, 
+J\"org P. Dietrich$^{58,59}$, 
+Jacob Silterra$^{60}$, 
+James T Webber$^{61}$, 
+Janko Slavi\v{c}$^{62}$, 
+Joel Nothman$^{63}$, 
+Johannes Buchner$^{64,65}$, 
+Johannes Kulick$^{66}$, 
 Johannes L. Sch\"{o}nberger$^{10}$, 
-Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, 
-Joscha Reimer$^{69}$, 
-Joseph Harrington$^{70}$, 
-Juan Luis Cano Rodr\'iguez$^{71}$, 
-Juan Nunez-Iglesias$^{72}$, 
-Justin Kuczynski$^{73}$, 
-Kevin Tritz$^{74}$, 
-Martin Thoma$^{75}$, 
+Jos\'e Vin\'icius de Miranda Cardoso$^{67}$, 
+Joscha Reimer$^{68}$, 
+Joseph Harrington$^{69}$, 
+Juan Luis Cano Rodr\'iguez$^{70}$, 
+Juan Nunez-Iglesias$^{71}$, 
+Justin Kuczynski$^{72}$, 
+Kevin Tritz$^{73}$, 
+Martin Thoma$^{74}$, 
+Matthew Newville$^{75}$, 
 Matthias K\"ummerer$^{76}$, 
@@
 $^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, 
-$^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, 
-$^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, 
-$^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, 
-$^{42}$Emory University, Atlanta, GA 30322, USA, 
-$^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, 
-$^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, 
-$^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, 
-$^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, 
-$^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, 
-$^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, 
-$^{49}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, 
-$^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, 
-$^{51}$Independent Researcher, Cologne, Germany, 
-$^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, 
-$^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, 
-$^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, 
-$^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, 
-$^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, 
-$^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, 
-$^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, 
-$^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, 
-$^{60}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, 
-$^{61}$Independent Researcher, USA, 
-$^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, 
-$^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, 
-$^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, 
-$^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, 
-$^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, 
-$^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, 
-$^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, 
-$^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, 
-$^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, 
-$^{71}$Independent Researcher, Spain, 
-$^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, 
-$^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, 
-$^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, 
-$^{75}$Independent Researcher, Munich, Germany, 
+$^{39}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, 
+$^{40}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, 
+$^{41}$Emory University, Atlanta, GA 30322, USA, 
+$^{42}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, 
+$^{43}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, 
+$^{44}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, 
+$^{45}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, 
+$^{46}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, 
+$^{47}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, 
+$^{48}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, 
+$^{49}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, 
+$^{50}$Independent Researcher, Cologne, Germany, 
+$^{51}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, 
+$^{52}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, 
+$^{53}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, 
+$^{54}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, 
+$^{55}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, 
+$^{56}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, 
+$^{57}$ENSTA Bretagne, 2 rue François Verny, Brest, France, 
+$^{58}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, 
+$^{59}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, 
+$^{60}$Independent Researcher, USA, 
+$^{61}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, 
+$^{62}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, 
+$^{63}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, 
+$^{64}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, 
+$^{65}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, 
+$^{66}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, 
+$^{67}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, 
+$^{68}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, 
+$^{69}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, 
+$^{70}$Independent Researcher, Spain, 
+$^{71}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, 
+$^{72}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, 
+$^{73}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, 
+$^{74}$Independent Researcher, Munich, Germany, 
+$^{75}$Center for Advanced Radiation Sources, The University of Chicago, Chicago, IL 60637, USA, 
 $^{76}$University of T\"ubingen, T\"ubingen, Germany, 

### 2019-07-26 scipy-1.0/paper.tex
Avoid starting a sentence with "And".
Fixes typo (hanker => Hankel)

@@
 it easier to implement the domain-specific functionality they needed
-in a general purpose programming language.  And given Python's innate
+in a general purpose programming language.  Given Python's innate
 ability to function as a systems language, controlling specialized or
@@
 
-    Special Functions (Bessel, hanker, Airy, etc.)
+    Special Functions (Bessel, Hankel, Airy, etc.)
     Signal/Image Processing

### 2019-07-26 scipy-1.0/paper.tex
MAINT: Fix Bezier spelling
* Make sure line breaks / spaces are set correctly
  - See https://tex.stackexchange.com/a/22563/5645
* Use \num from siunitx for number formattingi
* Use \textelp{} from csquotes
* Remove trailing spaces

--- 2019-07-22 scipy-1.0/paper.tex
@@
 \usepackage{booktabs,siunitx}
+\usepackage{csquotes}
 
@@
 unique code contributors, thousands of dependent packages,
-over 100,000 dependent repositories, and millions of downloads per year.
+over \num{100000} dependent repositories, and millions of downloads per year.
 This includes usage of SciPy
@@
 suitable for small arrays, but not for the large images processed by
-STScI.  With the Numeric maintainer's blessing, the decision was made
+STScI\@.  With the Numeric maintainer's blessing, the decision was made
 to write NumArray\cite{greenfield2003numarray}, a library that could handle data on a larger
@@
 LLNL now has many Python-based efforts built from scratch or wrapped around
-legacy codes [. . .] hundreds of thousands
-of lines of C++, Python, and Fortran 95, all working together just as we hoped,
+legacy codes \textelp{} hundreds of thousands
+of lines of C++, Python, and Fortran~95, all working together just as we hoped,
 doing compute-intensive calculations on massively parallel computers.
@@
 At the time of writing, the SciPy library consists of nearly
-600,000 lines of code organized in 16 subpackages. 
-Over 110,000 GitHub repositories and 6,500 packages depend
+\num{600000}~lines of code organized in 16~subpackages.
+Over \num{110000}~GitHub repositories and \num{6500}~packages depend
 on SciPy\cite{dependents}. Some of the major
@@
 According to analysis using the \texttt{linguist} library\cite{linguistref}, SciPy is approximately 50\% Python, 25\% Fortran, 20\% C, 3\% Cython, and 2\% C++, with a dash of \TeX, MATLAB, shell script, and Make.
-The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.
-Cython) and the usage of languages (and their libraries) that have proven reliable and performant over many decades.
+The distribution of secondary programming languages in SciPy is a compromise between a powerful, performance-enhancing language that interacts well with Python (i.e.\ Cython) and the usage of languages (and their libraries) that have proven reliable and performant over many decades.
 
@@
 The application programming interface (API) for SciPy consists of approximately
-1,500 functions and classes.  Our policy for evolving the API over time is that
+\num{1500}~functions and classes.  Our policy for evolving the API over time is that
 new functionality can be added, while removing or changing existing
@@
 based on generic code, e.g., numerical integration of the PDF to
-obtain the CDF.
+obtain the CDF\@.
 Key recent distributions added to \texttt{scipy.stats} include the
@@
 include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There
-are more than $13,000$ unit tests in the test suite, which is written for usage
+are more than \num{13000}~unit tests in the test suite, which is written for usage
 with the \texttt{pytest}\cite{pytest} framework.
-Test coverage at the SciPy 1.0 release point (Figure~\ref{fig:coverage}) was at
+Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
 87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
@@
 light-shaded area represents lines not covered. With the exception
-of the removal of $\approx 61,000$ lines of compiled code for SciPy
+of the removal of $\approx \num{61000}$ lines of compiled code for SciPy
 v0.14, the volume of both compiled (C, C++, and Fortran) and Python
@@
 covered by unit tests. In SciPy v1.0.0, tests covered approximately
-45\% of $462,574$ lines of compiled code (many of which
+45\% of \num{462574}~lines of compiled code (many of which
 are automatically generated from Cython source), and nearly 87\% of
-106,878 lines of Python code.\\
+\num{106878}~lines of Python code.\\
 Coverage analysis is performed automatically in our continuous integration
@@
 \includegraphics[width=\textwidth]{static/asv}
-\caption{Results of the \texttt{scipy.spatial.cKDTree.query} benchmark from the introduction of \texttt{cKDTree} to the release of SciPy 1.0.
-The benchmark generates a $k$-d tree from 10,000 uniformly distributed points in an $m$-dimensional unit hypercube, then finds the nearest (Euclidean) neighbor in the tree for each of 1,000 query points.
+\caption{Results of the \texttt{scipy.spatial.cKDTree.query} benchmark from the introduction of \texttt{cKDTree} to the release of SciPy~1.0.
+The benchmark generates a $k$-d tree from \num{10000} uniformly distributed points in an $m$-dimensional unit hypercube, then finds the nearest (Euclidean) neighbor in the tree for each of \num{1000}~query points.
 Each marker in the figure indicates the execution time of the benchmark for a commit in the master branch of SciPy.
 Substantial performance improvements were realized when \texttt{cKDTree} was fully Cythonized and again when it was rewritten in C++.
-The tree was generated without application of toroidal topology (\texttt{boxsize=None}), and tests were performed by Airspeed Velocity 0.4 using Python 2.7, NumPy 1.8.2, and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward compatibility).}
+The tree was generated without application of toroidal topology (\texttt{boxsize=None}), and tests were performed by Airspeed Velocity 0.4 using Python~2.7, NumPy~1.8.2, and Cython versions 0.27.3, 0.21.1, and 0.18 (for improved backward compatibility).}
 \label{fig:asvbench}
@@
 
-SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}.
-A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community.
+SciPy adopted an official Governance Document on August~3, 2017\cite{SciPyProjectGovernance}.
+A Steering Council, currently composed of 18~members, oversees daily development of the project by contributing code and reviewing contributions from the community.
 Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections.
@@
 SciPy has a strong developer community and a massive user base. GitHub traffic
-metrics report roughly 20,000 unique visitors to the source website between May
-14 and May 27, 2018, with 721 unique copies (``clones") of the codebase over
-that roughly two-week period. The developer community at that time consisted of 610 unique
-contributors of source code, with $>19,000$ commits accepted into the codebase
+metrics report roughly \num{20000}~unique visitors to the source website between May
+14 and May~27, 2018, with 721~unique copies (``clones") of the codebase over
+that roughly two-week period. The developer community at that time consisted of 610~unique
+contributors of source code, with more than \num{19000}~commits accepted into the codebase
 (GitHub page data).
 
-From the user side, there were 13,096,468 downloads of SciPy from the Python
-Packaging Index (PyPI)\cite{pypinfo} and 5,776,017 via the default channel of the
+From the user side, there were \num{13096468}~downloads of SciPy from the Python
+Packaging Index (PyPI)\cite{pypinfo} and \num{5776017} via the default channel of the
 \texttt{conda}\cite{condainfo} package manager during the year 2017. These numbers establish a lower
@@
 website\cite{SciPylib}, which has been the default citation in the absence of a
-peer-reviewed paper, has been cited $>3,000$ times\cite{googlescholar}. Some of the most prominent
+peer-reviewed paper, has been cited over \num{3000}~times\cite{googlescholar}. Some of the most prominent
 usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
@@
 There are contributors who were present in the early years of the project who still contribute to discussions of bug reports and reviews of new code contributions.
-Our BDFL has been with the project for more than 10 years and is still actively contributing code, and the head of our Steering Council, who also acts as a general manager, is approaching his eleventh anniversary.
+Our BDFL has been with the project for more than 10~years and is still actively contributing code, and the head of our Steering Council, who also acts as a general manager, is approaching his eleventh anniversary.
 An additional half dozen or so active developers have been contributing steadily for five or more years.
@@
 
-We thank everyone who has contributed to SciPy 1.0, from those who have posted one comment about an issue through those who have made several small patches and beyond!
+We thank everyone who has contributed to SciPy~1.0, from those who have posted one comment about an issue through those who have made several small patches and beyond!
 
@@
 is SciPy's Benevolent Dictator for Life, R.G.
-is the Chair of the SciPy Steering Council, and together they have led the development of SciPy for over 10 years.
+is the Chair of the SciPy Steering Council, and together they have led the development of SciPy for over 10~years.
 T.E.O., E.J., and P.P.

### 2019-07-26 scipy-1.0/paper.tex
Remove word duplication

@@
 
-In addition to ensuring that unit tests are passing, it is important to confirm that the
-the performance of the SciPy codebase improves over time. Since February 2015, the
+In addition to ensuring that unit tests are passing, it is important to confirm that
+the performance of the SciPy codebase improves over time. Since February~2015, the
 performance of SciPy has been monitored with Airspeed Velocity (\texttt{asv} \cite{asvref}).

### 2019-07-26 scipy-1.0/paper.tex
To avoid having others come and fix this

--- 2019-07-26 scipy-1.0/paper.tex
@@
 
-    Special Functions (Bessel, Hankel, Airy, etc.)
+    Special Functions (Bessel, hanker, Airy, etc.) % Yep, `hanker` is a typo, but it was *in the announcement*, so we leave it here too
     Signal/Image Processing

### 2019-07-26 scipy-1.0/paper.tex
Add comment to avoid fixing this typo too

@@
     ODE solvers
-    Optimization (simplex, BFGS, Netwon-CG, etc.)
+    Optimization (simplex, BFGS, Netwon-CG, etc.) % `Netwon` also a typo. We also leave that as-is.
     Genetic Algorithms

### 2019-07-26 scipy-1.0/covlett.tex
MAINT: revise cover letter
* revise cover letter based on reviewer
feedback in PR 193

@@
 %\dateskipbefore{0pt}
-%\sigskipbefore{0pt}
+\sigskipbefore{40pt}
 %\sigskipafter{0pt}
@@
 \newlfmP{sigsize=10pt}
-\bottommarginskip{0pt}
+\bottommarginskip{60pt}
 \MinFoot{0pt}
@@
 \usepackage{kpfonts}
+\usepackage{hyperref}
 \usepackage{url}
 
+\hypersetup{%
+  linkbordercolor=blue,% hyperlink borders will be blue
+  pdfborderstyle={/S/U/W 1}% border style will be underline of width 1pt
+}
 \widowpenalty=1000
@@
 described in the journal scope, and because SciPy is effectively an
-ever-improving methodological library. High-profile manuscripts citing SciPy 
-extend beyond the biosciences to recent reports on gravitational waves and
-black holes. Lacking a better alternative, these citations all point
+ever-improving methodological library. Among the 3000+ articles that 
+cite the SciPy library are hundreds in the field of biology,
+including several in \emph{Nature Methods}\footnote{\href{
+https://www.nature.com/articles/s41592-018-0114-z}{
+Gruber \emph{et al.} (2018) \textbf{15}: 832-836}
+\href{https://www.nature.com/articles/s41592-019-0403-1}{
+; Moen \emph{et al.} (2019)}}, one of which\footnote{
+\href{https://www.nature.com/articles/nmeth.f.303}{
+Caporaso \emph{et al.} (2010) \textbf{7}: 335-336}} is cited
+15,800 times. Citations in recent reports on gravitational waves and
+black holes are also striking. Lacking a better alternative, these citations all point
 to the SciPy website; instead, we think the package now merits a
-formal publication in a highly visible journal. A pre-print will likely
-soon be made available on arXiv.
+formal publication in a highly visible journal. A preprint was recently
+made \href{https://arxiv.org/abs/1907.10121}{available on arXiv}.
 
@@
 
+\vspace*{1cm}
 Thank you for your consideration.  We look forward to
-    hearing from you.\vspace*{0.1cm}
+    hearing from you.
 \end{newlfm}

### 2019-07-26 scipy-1.0/poly.tex

@@
 At the lowest level of the new design are classes which represent univariate
-piecewise polynomials: \code{PPoly} (SciPy 0.13)\cite{scipy-gh2885},
-\code{BPoly} (SciPy 0.13) and \code{BSpline} (SciPy 0.19)\cite{scipy-gh3174},
+piecewise polynomials: \code{PPoly} (SciPy~0.13)\cite{scipy-gh2885},
+\code{BPoly} (SciPy~0.13) and \code{BSpline} (SciPy~0.19)\cite{scipy-gh3174},
 which allow
@@
 represents piecewise polynomials in the Bernstein basis (which is suitable
-for e.g., constructing Bezier curves). \code{BSpline} represents spline
+for e.g., constructing B\'{e}zier curves). \code{BSpline} represents spline
 curves, i.e., linear combinations of B-spline basis elements.\cite{deBoor1978}
@@
 In the next layer, these polynomial classes are used for constructing several
-common ways of interpolating data: \code{CubicSpline} (SciPy 0.18)
+common ways of interpolating data: \code{CubicSpline} (SciPy~0.18)
 \cite{scipy-gh5653} constructs a twice

### 2019-07-27 scipy-1.0/paper.tex
Use same font with siunitx num

--- 2019-07-26 scipy-1.0/paper.tex
@@
 \usepackage{booktabs,siunitx}
+\sisetup{detect-all}
+
 \usepackage{csquotes}
@@
 
-    Special Functions (Bessel, hanker, Airy, etc.)
+    Special Functions (Bessel, Hankel, Airy, etc.) % hanker -> Hankel
     Signal/Image Processing
@@
     ODE solvers
-    Optimization (simplex, BFGS, Netwon-CG, etc.)
+    Optimization (simplex, BFGS, Newton-CG, etc.) % Netwon -> Newton
     Genetic Algorithms
@@
 \end{verbatim}
-\caption{Excerpt from SciPy 0.1 release announcement (typos included).}\label{fig:announce-0.1}
+\caption{Excerpt from SciPy 0.1 release announcement (typos corrected).}\label{fig:announce-0.1}
 \end{figure}

### 2019-07-27 scipy-1.0/covlett.tex
Update scipy-1.0/covlett.tex
Co-Authored-By: Matt Haberland <mdhaber@mit.edu>
MAINT: PR 193 reviewer comments
* reduce side margins in cover letter

* remove end-of-line commas in address
/ contact blocks & do not place "USA"
on its own line

* reformat justificatio paragraph from
Matt H's suggestion & a few more
adjustments to fit letter on single page

--- 2019-07-21 scipy-1.0/covlett.tex
@@
 \newlfmP{sigsize=10pt}
-\bottommarginskip{0pt}
+\bottommarginskip{10pt}
 \MinFoot{0pt}
+\leftmarginsize{0.95in}
+\rightmarginsize{0.95in}
 
 \usepackage{kpfonts}
+\usepackage{hyperref}
 \usepackage{url}
 
+\hypersetup{%
+  linkbordercolor=blue,% hyperlink borders will be blue
+  pdfborderstyle={/S/U/W 1}% border style will be underline of width 1pt
+}
 \widowpenalty=1000
@@
     \today\\[10pt]
-    Ralf Gommers,\\
-    Quansight Labs,\\
-    The Netherlands,\\
-    \texttt{ralf.gommers@gmail.com},\\
+    Ralf Gommers\\
+    Quansight Labs\\
+    The Netherlands\\
+    \texttt{ralf.gommers@gmail.com}\\
     %phone number\\
     \\
-    Matt Haberland,\\
-    BioResource and Agricultural Engineering,\\
-    California Polytechnic State University,\\
-    San Luis Obispo, CA, 93407, USA,\\
-    \texttt{mhaberla@calpoly.edu},\\
+    Matt Haberland\\
+    BioResource and Agricultural Engineering\\
+    California Polytechnic State University\\
+    San Luis Obispo, CA, 93407, USA\\
+    \texttt{mhaberla@calpoly.edu}\\
     %phone number\\
     \\
-    Tyler Reddy,\\
-    CCS-7 Applied Computer Science Group,\\
-    Los Alamos National Laboratory,\\
-    Los Alamos, NM, 87545, USA,\\
-    \texttt{treddy@lanl.gov},\\
+    Tyler Reddy\\
+    CCS-7 Applied Computer Science Group\\
+    Los Alamos National Laboratory\\
+    Los Alamos, NM, 87545, USA\\
+    \texttt{treddy@lanl.gov}\\
     %phone number\\
@@
 \addrto{%
-Dr. Rita Strack,\\
-Nature Methods Editorial Office,\\
-One New York Plaza Suite 4500,\\
-New York, NY 10004,\\
-USA}
+Dr. Rita Strack\\
+Nature Methods Editorial Office\\
+One New York Plaza Suite 4500\\
+New York, NY 10004, USA}
 
@@
 
-We have specifically chosen \emph{Nature Methods} because
-SciPy lies at the base of the entire ecosystem of scientific methods
-described in the journal scope, and because SciPy is effectively an
-ever-improving methodological library. High-profile manuscripts citing SciPy 
-extend beyond the biosciences to recent reports on gravitational waves and
-black holes. Lacking a better alternative, these citations all point
-to the SciPy website; instead, we think the package now merits a
-formal publication in a highly visible journal. A pre-print will likely
-soon be made available on arXiv.
+SciPy lies at the base of the ecosystem of life science computational
+methods. Many studies, including an extremely influential \emph{Nature 
+Methods} paper (\href{https://www.nature.com/articles/nmeth.f.303}
+{Caporaso \emph{et al.}~2010}), describe software that 
+includes SciPy as a dependency without citing it. Others in \emph{
+Nature Methods} (e.g.,\href{https://www.nature.com/articles/s41592-018-0114-z}{
+Gruber \emph{et al.}~2018};\href{https://www.nature.com/articles/s41592-019-0403-1}{
+Moen \emph{et al.}~2019}) -- and some 3000+ other articles -- 
+cite SciPy's website for lack of
+a better alternative. As SciPy is central to so many biological studies,
+the research community would benefit from an article about SciPy to reference.
+Our preprint \href{https://arxiv.org/abs/1907.10121}{available on arXiv} 
+only partially fulfills this need; we appreciate your consideration of
+the manuscript for publication in \emph{Nature Methods}.
 
@@
 
-Thank you for your consideration.  We look forward to
-    hearing from you.\vspace*{0.1cm}
 \end{newlfm}

### 2019-07-27 scipy-1.0/covlett.tex
MAINT: remove more commas
* remove commas between States and
Zip codes in Cover Letter

@@
     California Polytechnic State University\\
-    San Luis Obispo, CA, 93407, USA\\
+    San Luis Obispo, CA 93407, USA\\
     \texttt{mhaberla@calpoly.edu}\\
@@
     Los Alamos National Laboratory\\
-    Los Alamos, NM, 87545, USA\\
+    Los Alamos, NM 87545, USA\\
     \texttt{treddy@lanl.gov}\\

### 2019-07-27 scipy-1.0/scipy-optimize.tex

@@
 solution of several classes of root finding and optimization problems.
-Here we highlight recent additions through SciPy 1.0.
+Here we highlight recent additions through SciPy~1.0.
 
@@
 
-A new interior-point optimizer for continuous linear programming problems, \texttt{linprog} with \texttt{method='interior-point'}, was released with SciPy 1.0.
+A new interior-point optimizer for continuous linear programming problems, \texttt{linprog} with \texttt{method='interior-point'}, was released with SciPy~1.0.
 Implementing the core algorithm of the commercial solver MOSEK \cite{andersen2000mosek}, it solves all of the 90+ NETLIB LP benchmark problems \cite{netlib} tested.
@@
     \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
-    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
-    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
-    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
+    point (i.e.\ a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
+    can deal with constraints on the variables. We distinguish among \textit{bound constraints} (i.e.\ $x^l \le x \le x^u$),
+    \textit{equality constraints} (i.e.\ $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e.\ $c_{\text{ineq}}(x) \ge 0$).}
   \begin{tabular}{cccccccccccccc}
@@
       \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
-      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
+      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, doi:10.1080/10556788.2018.1449842} \\
     \hline

### 2019-07-28 scipy-1.0/ckdtree.tex
Update ckdtree.tex (#1)
Update ckdtree.tex
avoid colloquial style

@@
 consistent with its formal description \cite{kdtree-search-algo}.
-Since then, we've enhanced \texttt{cKDTree.query} by reimplementing it in
+Since then, we have enhanced \texttt{cKDTree.query} by reimplementing it in
 C++, removing memory leaks, and allowing release of the global interpreter lock (GIL) so that
@@
 approximate sparse distance matrices between \texttt{KDTree} objects by ignoring
-all distances that exceed a user-provided value. Also, this routine is not 
+all distances that exceed a user-provided value. This routine is not
 limited to the conventional L2 (Euclidean) norm but supports any Minkowski

### 2019-07-28 scipy-1.0/paper.tex
avoid colloquial style

@@
 operators now support the Python matrix multiplication (\texttt{@}) operator.
-We've added \texttt{scipy.sparse.norm} and
+We have added \texttt{scipy.sparse.norm} and
 \texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
-random variates from arbitrary distributions, respectively. Also, we've made a
+random variates from arbitrary distributions, respectively. Also, we have made a
 concerted effort to bring the \texttt{scipy.sparse} API into line with the

### 2019-07-28 scipy-1.0/paper.tex
removed texttt and changed case of first letter
This change ensures consistency with page 7.

@@
 in a \texttt{scipy.LowLevelCallable} object that reduces the overhead of
-calling compiled C functions, such as those generated using \texttt{numba}
+calling compiled C functions, such as those generated using Numba
 or Cython, directly from Python.

### 2019-07-28 scipy-1.0/paper.tex
verb corrected

@@
 provides a Python interface for symbolic computation, and
-while \texttt{sparse.csgraph} or \texttt{spatial} offers basic tools
+while \texttt{sparse.csgraph} or \texttt{spatial} offer basic tools
 for working with graphs and networks compared to a more specialized

### 2019-07-28 scipy-1.0/paper.tex
Fixed spatial description in project scope per @charris

--- 2019-07-28 scipy-1.0/paper.tex
@@
 capabilities beyond SciPy's \texttt{ndimage}, SymPy\cite{meurer2017sympy}
-provides a Python interface for symbolic computation, and
-while \texttt{sparse.csgraph} or \texttt{spatial} offers basic tools
-for working with graphs and networks compared to a more specialized
-Python library like NetworkX\cite{hagberg2008networkx}.
+provides a Python interface for symbolic computation, and \texttt{sparse.csgraph}
+and \texttt{spatial} offer basic tools
+for working with graphs and networks compared to more specialized
+libraries like NetworkX\cite{hagberg2008networkx}.
 
@@
 operators now support the Python matrix multiplication (\texttt{@}) operator.
-We have added \texttt{scipy.sparse.norm} and
+We've added \texttt{scipy.sparse.norm} and
 \texttt{scipy.sparse.random} for computing sparse matrix norms and drawing
-random variates from arbitrary distributions, respectively. Also, we have made a
+random variates from arbitrary distributions, respectively. Also, we've made a
 concerted effort to bring the \texttt{scipy.sparse} API into line with the

### 2019-07-28 scipy-1.0/paper.tex
Remove detail from a sentence rather than making it more complex

--- 2019-07-28 scipy-1.0/paper.tex
@@
 ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
-equations, integrate differential equations, and fit splines (the
-latter implemented by Pearu Peterson).  Robert Kern, then an
+equations, integrate differential equations, and fit splines.  Robert Kern, then an
 undergraduate student (and currently a SciPy core developer), provided

### 2019-07-28 scipy-1.0/paper.tex
Address #209

@@
 extension modules that wrapped Fortran and C libraries such as
-ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
-equations, integrate differential equations, and fit splines.  Robert Kern, then an
+ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack},
+and MINPACK\cite{osti_6997568}, to solve nonlinear equations and least
+squares problems, integrate differential equations, and fit splines (the
+latter implemented by Pearu Peterson).  Robert Kern, then an
 undergraduate student (and currently a SciPy core developer), provided
@@
 By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
-growth and adoption.
-The informal workshops grew into international conferences with many
-hundreds of attendees. Special issues were organized and published in a
-leading scientific journal\cite{dubois2007guest}, and the scope of the SciPy library
+growth and adoption. The scope of the SciPy library
 narrowed, while the breadth of the ecosystem grew
@@
 
-In the early workshops, recurrent topics reflected the state of development, with
+In the early SciPy workshops, recurrent topics reflected the state of development, with
 emphasis being placed on the underlying array package, plotting,
 parallel processing, acceleration / wrapping, and user interfaces.  By
-2004, a significant shift occurred towards application of SciPy to
-scientific problems.  The event also started to draw in more keynote
+2004, presentations about the application of SciPy to scientific problems
+began to appear.  The event also started to draw in more keynote
 speakers from outside the community, such as Guido van Rossum (creator
@@
 Martelli (Google, 2008), and Peter Norvig (Google Research, 2009).
-The SciPy conference went from being a small gathering of
-core developers to a multi-location event with increased funding, a
+The informal workshop grew from a small gathering of core developers into an
+international conference with hundreds of attendees, increased funding, a
 published proceedings, and scholarships for attending students.

### 2019-07-28 scipy-1.0/paper.tex
"more specialized" -> "specialized"

--- 2019-07-28 scipy-1.0/paper.tex
@@
 and \texttt{spatial} offer basic tools
-for working with graphs and networks compared to more specialized
+for working with graphs and networks compared to specialized
 libraries like NetworkX\cite{hagberg2008networkx}.

### 2019-07-28 scipy-1.0/paper.tex

@@
 \newcommand{\fixme}[1]{\textcolor{red}{{#1}}}
+\newcommand{\inlinecite}[1]{\footnotesize\citen{#1}}
 
@@
 capabilities beyond SciPy's \texttt{ndimage}, SymPy\cite{meurer2017sympy}
-provides a Python interface for symbolic computation, and \texttt{sparse.csgraph}
-and \texttt{spatial} offer basic tools
-for working with graphs and networks compared to specialized
-libraries like NetworkX\cite{hagberg2008networkx}.
+provides a Python interface for symbolic computation, and
+while \texttt{sparse.csgraph} or \texttt{spatial} offer basic tools
+for working with graphs and networks compared to a more specialized
+Python library like NetworkX\cite{hagberg2008networkx}.
 

### 2019-07-28 scipy-1.0/scipy-optimize.tex
Address #209

@@
     \hline
-    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
-      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
-      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
-      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
-      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  &
-      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
-      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, doi:10.1080/10556788.2018.1449842} \\
+    References &
+      \inlinecite{nelder_simplex_1965, wright_direct_1996} &
+      \inlinecite{powell_efficient_1964} &
+      \inlinecite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
+      \inlinecite{polak_note_1969, nocedal_numerical_2006} &
+      \inlinecite{nocedal_numerical_2006} &
+      \inlinecite{byrd_limited_1995, zhu_algorithm_1997} &
+      \inlinecite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
+      \inlinecite{nash_newton-type_1984} &
+      \inlinecite{nocedal_numerical_2006}  &
+      \inlinecite{powell_new_1970, nocedal_numerical_2006} &
+      \inlinecite{steihaug_conjugate_1983, nocedal_numerical_2006} &
+      \inlinecite{conn_trust_2000, more_computing_1983} &
+      \inlinecite{gould_solving_1999, doi:10.1080/10556788.2018.1449842} \\
     \hline

### 2019-08-12 scipy-1.0/paper.tex
typo: remove hyphen

@@
 The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev}, summarized in Table~\ref{table:roadmap}, is a
-continuously-updated document
+continuously updated document
 maintained by the community that describes some of the major directions for

### 2019-08-13 scipy-1.0/paper.tex
Update scipy-1.0/paper.tex
Co-Authored-By: Matt Haberland <mdhaber@mit.edu>

@@
 The SciPy Roadmap\cite{SciPy_roadmap_1,SciPy_roadmap_dev}, summarized in Table~\ref{table:roadmap}, is a
-continuously updated document
+continually updated document
 maintained by the community that describes some of the major directions for

### 2019-08-19 scipy-1.0/paper.tex

@@
 extension modules that wrapped Fortran and C libraries such as
-ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack}, and MINPACK\cite{osti_6997568}, to solve and minimize nonlinear
-equations, integrate differential equations, and fit splines.  Robert Kern, then an
+ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack},
+and MINPACK\cite{osti_6997568}, to solve nonlinear equations and least
+squares problems, integrate differential equations, and fit splines.
+Robert Kern, then an
 undergraduate student (and currently a SciPy core developer), provided

### 2019-09-06 scipy-1.0/paper.tex
MAINT: review comment re: GitHub
* address reviewer comment from Nature Methods requesting
a brief description of GitHub prior to its first usage
in the text
MAINT: rev comment re: test benefits
* address Nature Methods reviewer comment requesting
a brief explanation of the benefits of unit testing
& test line coverage before we dive into the details
of our testing infrastructure

@@
 
+Test-driven development has been described as a way to manage fear and
+uncertainity when making code changes\cite{TDD-Beck}. We write small executable
+tests that verify the correctness of some component of our code, building this
+up over time to increase our confidence. Monitoring the number of lines
+of code in SciPy that are properly covered by unit tests is one way we maintain
+some certainty that changes and new features are not harmful.
 The SciPy test suite is orchestrated by a continuous integration matrix that

### 2019-09-06 scipy-1.0/paper.tex
Expand description of scikits

--- 2019-08-19 scipy-1.0/paper.tex
@@
 narrowed, while the breadth of the ecosystem grew
-through a new type of auxiliary package: the scikit\cite{scikits-general}.
-The tooling, development, documentation, and release processes became more professional.
-SciPy was expanded carefully, with the patience affordable in open source
+through a new type of auxiliary package: the
+scikit\cite{scikits-general}, a complementary library
+developed outside of SciPy, allowing for more rapid exploration of
+experimental ideas while maintaining familiar style and development methodology.
+In SciPy itself, tooling, development, documentation, and release processes became more professional.
+The library was expanded carefully, with the patience affordable in open source
 projects and via best practices common in industry \cite{millman2014developing}.

### 2019-09-06 scipy-1.0/paper.tex
Describe why code coverage is low for compiled libs

--- 2019-08-19 scipy-1.0/paper.tex
@@
 87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
-45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
+45\% for compiled (C, C++, and Fortran) code according to
+\texttt{gcov}\cite{gcov}, the latter primarily due to the vendoring of several
+large, well-vetted Fortran libraries from repositories such as NETLIB.
 Documentation for the code is automatically built and published by

### 2019-09-06 scipy-1.0/paper.tex
Remove mention of binary coverage.
It's complicated: may also currently include generated code, vendored
libraries.  It just confuses the story, so I removed it until we come up
with an implementation where we better understand what exactly is being
measured.

@@
 Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
-87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
-45\% for compiled (C, C++, and Fortran) code according to
-\texttt{gcov}\cite{gcov}, the latter primarily due to the vendoring of several
-large, well-vetted Fortran libraries from repositories such as NETLIB.
+87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov}.
+Coverage for compiled (C, C++, and Fortran) code is still being implemented.
 Documentation for the code is automatically built and published by

### 2019-09-07 scipy-1.0/paper.tex
Added date and medium to Figure 1 caption
Remove unnecessary mention of CPython

--- 2019-09-06 scipy-1.0/paper.tex
@@
 language, designed by Guido van Rossum in the late 1980s,
-with a dynamic type system and an emphasis on readability and rapid prototyping.
-The reference and most popular implementation of Python is
-CPython\cite{cpython-refman,cpython-source}, which is written
-in the C and Python languages and assumed throughout this paper.
+with a dynamic type system and an emphasis on readability and rapid prototyping\cite{cpython-refman,cpython-source}.
 % https://legacy.python.org/search/hypermail/python-1992/0000.html

### 2019-09-07 scipy-1.0/paper.tex
MAINT: reviewer comments PR 234
* apply the patch provided by Stefan to
improve articles PR 234

--- 2019-09-06 scipy-1.0/paper.tex
@@
 
+Test-driven development has been described as a way to manage fear and
+uncertainity when making code changes\cite{TDD-Beck}. For each component of SciPy, we write multiple small executable
+tests that verify its intended behavior. The collection of these, known as a \emph{test suite}, increases confidence in
+the correctness and accuracy of the library, and lets us make code modifications known not to alter desired behavior. Monitoring the number of lines
+of code in SciPy covered by unit tests is one way we maintain
+some certainty that changes and new features are correctly implemented.
 The SciPy test suite is orchestrated by a continuous integration matrix that
@@
 Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
-87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov}.
-Coverage for compiled (C, C++, and Fortran) code is still being implemented.
+87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
+45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
 Documentation for the code is automatically built and published by

### 2019-09-08 scipy-1.0/paper.tex

--- 2019-09-07 scipy-1.0/paper.tex
@@
 \end{verbatim}
-\caption{Excerpt from SciPy 0.1 release announcement (typos corrected).}\label{fig:announce-0.1}
+\caption{Excerpt from the SciPy 0.1 release announcement (typos corrected) posted August 20, 2001, on the ``Python-list'' mailing list\cite{first-scipy-rel}.}\label{fig:announce-0.1}
 \end{figure}

### 2019-09-08 scipy-1.0/paper.tex
Revised line breaks in Test Suite section
Introduce continuous integration

@@
 Test-driven development has been described as a way to manage fear and
-uncertainity when making code changes\cite{TDD-Beck}. For each component of SciPy, we write multiple small executable
-tests that verify its intended behavior. The collection of these, known as a \emph{test suite}, increases confidence in
-the correctness and accuracy of the library, and lets us make code modifications known not to alter desired behavior. Monitoring the number of lines
-of code in SciPy covered by unit tests is one way we maintain
+uncertainity when making code changes\cite{TDD-Beck}. For each component of 
+SciPy, we write multiple small executable tests that verify its intended
+behavior. The collection of these, known as a \emph{test suite}, increases
+confidence in the correctness and accuracy of the library, and lets us make
+code modifications known not to alter desired behavior.
+According to the practice of continuous integration, all proposed contributions
+to SciPy are temporarily integrated with the master branch of the library
+before the test suite is run, and all tests must pass before the contribution
+is permanently merged.
+Continuously monitoring the number
+of lines of code in SciPy covered by unit tests is one way we maintain
 some certainty that changes and new features are correctly implemented.
+
 The SciPy test suite is orchestrated by a continuous integration matrix that

### 2019-09-08 scipy-1.0/paper.tex
Uncommented descriptions of each Fortran library

--- 2019-09-08 scipy-1.0/paper.tex
@@
 libraries in order to provide Python convenience while benefiting from their performance:
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms,
-ODEPACK\cite{hindmarsh1983odepack}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{piessens1983quadpack}, % for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression,
-MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations,
-ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and
-CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms,
+ODEPACK\cite{hindmarsh1983odepack} for solving ordinary differential equation initial value problems,
+QUADPACK\cite{piessens1983quadpack} for numerical integration of one-dimensional functions,
+FITPACK\cite{Dierckx:1993:CSF:151103} for curve-fitting and interpolation,
+ODRPACK\cite{ODRPACK_Boggs} for orthogonal distance regression,
+MINPACK\cite{osti_6997568} for minimization of linear and nonlinear equations,
+ARPACK\cite{leh:sor:yan96} for solving large scale eigenvalue problems,
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for Bessel Functions, and
+CDFLIB\cite{CDFLIB_site} for evaluation of cumulative density functions.
 

### 2019-09-08 scipy-1.0/paper.tex
Categorized Fortran packages where possible

@@
 libraries in order to provide Python convenience while benefiting from their performance:
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms,
-ODEPACK\cite{hindmarsh1983odepack} for solving ordinary differential equation initial value problems,
-QUADPACK\cite{piessens1983quadpack} for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103} for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs} for orthogonal distance regression,
-MINPACK\cite{osti_6997568} for minimization of linear and nonlinear equations,
-ARPACK\cite{leh:sor:yan96} for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for Bessel Functions, and
-CDFLIB\cite{CDFLIB_site} for evaluation of cumulative density functions.
+QUADPACK\cite{piessens1983quadpack} and ODEPACK\cite{hindmarsh1983odepack} for numerical integration
+and solution of initial value problems;
+FITPACK\cite{Dierckx:1993:CSF:151103}, ODRPACK\cite{ODRPACK_Boggs}, and MINPACK\cite{osti_6997568} for
+curve-fitting and least-squares minimization;
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms;
+ARPACK\cite{leh:sor:yan96} for solving eigenvalue problems;
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for computing Bessel Functions; and
+CDFLIB\cite{CDFLIB_site} for evaluating cumulative density functions.
 

### 2019-09-08 scipy-1.0/paper.tex
Introduced C libraries

@@
 scientific computing. The C libraries that we wrap in SciPy include
-trlib\cite{doi:10.1080/10556788.2018.1449842}, % (iterative solving of the trust region problem),
-SuperLU\cite{li05,superlu_ug99}, % (solution of large, sparse, nonsymmetric systems of linear equations),
-Qhull\cite{Barber:1996:QAC:235815.235821}, and % (computational geometry)
-Cephes\cite{cephes_netlib}. % (mathematics algorithms).
+trlib\cite{doi:10.1080/10556788.2018.1449842} for optimization, % (iterative solving of the trust region problem),
+SuperLU\cite{li05,superlu_ug99} for solving sparse linear systems, % (solution of large, sparse, nonsymmetric systems of linear equations),
+Qhull\cite{Barber:1996:QAC:235815.235821} for computational geometry, and % (computational geometry)
+Cephes\cite{cephes_netlib} for special functions. % (mathematics algorithms).
 

### 2019-09-08 scipy-1.0/paper.tex
Defined BLAS/LAPACK acronyms
Removed specific mention of Fortran libraries in Minpack

@@
 Transforms, and more.  One of these packages, Multipack\cite{multipack}, was a set of
-extension modules that wrapped Fortran and C libraries such as
-ODEPACK\cite{hindmarsh1983odepack}, QUADPACK\cite{piessens1983quadpack},
-and MINPACK\cite{osti_6997568}, to solve nonlinear equations and least
+extension modules that wrapped Fortran and C libraries
+to solve nonlinear equations and least
 squares problems, integrate differential equations, and fit splines.
@@
 
-SciPy has provided special functions and leveraged BLAS and
-LAPACK\cite{LAPACK} routines for many years. SciPy now additionally
+SciPy has provided special functions and leveraged
+Basic Linear Algebra Subprograms (BLAS) and
+Linear Algebra PACKage (LAPACK)\cite{LAPACK} routines for many years.
+SciPy now additionally
 includes Cython\cite{behnel2011cython} wrappers for

### 2019-09-08 scipy-1.0/subpackages.tex
Mentioned tutorials again per reviewer comment

@@
 Here we summarize the scope and capabilities of each subpackage.
+Additional information is available in the SciPy Tutorial\cite{SciPy_tutorials} and
+API Reference\cite{SciPy_API}.
 
@@
 \item[\texttt{special}]
-    The name comes the class of functions traditionally known as ``special
+    The name comes from the class of functions traditionally known as ``special
     functions'', but over time, the subpackage has grown to include functions
@@
     It includes a large collection of the classical special functions
-    such as Airy, Bessel, etc.; orthogonal families of polynomials;
+    such as Airy, Bessel, etc.; families of orthogonal polynomials;
     the Gamma function, and functions related to it;

### 2019-09-08 scipy-1.0/paper.tex
Update scipy-1.0/paper.tex
Co-Authored-By: Matt Haberland <mdhaber@mit.edu>

--- 2019-09-08 scipy-1.0/paper.tex
@@
 language, designed by Guido van Rossum in the late 1980s,
-with a dynamic type system and an emphasis on readability and rapid prototyping\cite{cpython-refman,cpython-source}.
+with a dynamic type system and an emphasis on readability and rapid prototyping.
+The reference and most popular implementation of Python is
+CPython\cite{cpython-refman,cpython-source}, which is written
+in the C and Python languages and assumed throughout this paper.
 % https://legacy.python.org/search/hypermail/python-1992/0000.html
@@
 \end{verbatim}
-\caption{Excerpt from the SciPy 0.1 release announcement (typos corrected) posted August 20, 2001, on the ``Python-list'' mailing list\cite{first-scipy-rel}.}\label{fig:announce-0.1}
+\caption{Excerpt from SciPy 0.1 release announcement (typos corrected).}\label{fig:announce-0.1}
 \end{figure}
@@
 narrowed, while the breadth of the ecosystem grew
-through a new type of auxiliary package: the
-scikit\cite{scikits-general}, a complementary library
-developed outside of SciPy, allowing for more rapid exploration of
-experimental ideas while maintaining familiar style and development methodology.
-In SciPy itself, tooling, development, documentation, and release processes became more professional.
-The library was expanded carefully, with the patience affordable in open source
+through a new type of auxiliary package: the scikit\cite{scikits-general}.
+The tooling, development, documentation, and release processes became more professional.
+SciPy was expanded carefully, with the patience affordable in open source
 projects and via best practices common in industry \cite{millman2014developing}.
@@
 At the time of writing, the SciPy library consists of nearly
-\num{600000}~lines of code organized in 16~subpackages.
+\num{600000}~lines of code organized in 16~subpackages. The development
+team and community currently interact and operate primarily
+on GitHub, an online version control and task management platform.
 Over \num{110000}~GitHub repositories and \num{6500}~packages depend
@@
 libraries in order to provide Python convenience while benefiting from their performance:
-QUADPACK\cite{piessens1983quadpack} and ODEPACK\cite{hindmarsh1983odepack} for numerical integration
-and solution of initial value problems;
-FITPACK\cite{Dierckx:1993:CSF:151103}, ODRPACK\cite{ODRPACK_Boggs}, and MINPACK\cite{osti_6997568} for
-curve-fitting and least-squares minimization;
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms;
-ARPACK\cite{leh:sor:yan96} for solving eigenvalue problems;
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for computing Bessel Functions; and
-CDFLIB\cite{CDFLIB_site} for evaluating cumulative density functions.
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms,
+ODEPACK\cite{hindmarsh1983odepack}, % for solving ordinary differential equation initial value problems,
+QUADPACK\cite{piessens1983quadpack}, % for numerical integration of one-dimensional functions,
+FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
+ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression,
+MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations,
+ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and
+CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
 
@@
 
-Test-driven development has been described as a way to manage fear and
-uncertainity when making code changes\cite{TDD-Beck}. For each component of SciPy, we write multiple small executable
-tests that verify its intended behavior. The collection of these, known as a \emph{test suite}, increases confidence in
-the correctness and accuracy of the library, and lets us make code modifications known not to alter desired behavior. Monitoring the number of lines
-of code in SciPy covered by unit tests is one way we maintain
-some certainty that changes and new features are correctly implemented.
 The SciPy test suite is orchestrated by a continuous integration matrix that

### 2019-09-08 scipy-1.0/paper.tex
Added citation for continuous integration from @tylerjereddy

@@
 At the time of writing, the SciPy library consists of nearly
-\num{600000}~lines of code organized in 16~subpackages. The development
-team and community currently interact and operate primarily
-on GitHub, an online version control and task management platform.
+\num{600000}~lines of code organized in 16~subpackages.
 Over \num{110000}~GitHub repositories and \num{6500}~packages depend
@@
 Test-driven development has been described as a way to manage fear and
-uncertainity when making code changes\cite{TDD-Beck}. For each component of SciPy, we write multiple small executable
-tests that verify its intended behavior. The collection of these, known as a \emph{test suite}, increases confidence in
-the correctness and accuracy of the library, and lets us make code modifications known not to alter desired behavior. Monitoring the number of lines
-of code in SciPy covered by unit tests is one way we maintain
+uncertainity when making code changes\cite{TDD-Beck}. For each component of 
+SciPy, we write multiple small executable tests that verify its intended
+behavior. The collection of these, known as a \emph{test suite}, increases
+confidence in the correctness and accuracy of the library, and lets us make
+code modifications known not to alter desired behavior.
+According to the practice of continuous integration\cite{silver2017collaborative}, all proposed contributions
+to SciPy are temporarily integrated with the master branch of the library
+before the test suite is run, and all tests must pass before the contribution
+is permanently merged.
+Continuously monitoring the number
+of lines of code in SciPy covered by unit tests is one way we maintain
 some certainty that changes and new features are correctly implemented.
+
 The SciPy test suite is orchestrated by a continuous integration matrix that

### 2019-09-08 scipy-1.0/paper.tex
Streamlined SciPy begins section

--- 2019-09-08 scipy-1.0/paper.tex
@@
 custom scripting language and analysis pipeline with Python\cite{STScI-slither}.
-
-As SciPy, the algorithms library of the ecosystem, began attracting the attention
-of large research organizations,
-the next generation of graduate students and postdocs were already exploring
-other aspects of the computational environment.
-In 2000, Prabhu Ramachandran, a PhD student at the Indian Institute of
-Technology, started work on a 3D visualization application, based on
-Kitware's C++ Visualization Toolkit\cite{schroeder:2006:VTK}, called
-Mayavi\cite{mayavi-intro}.
-In 2001, Fernando Pérez, a graduate student at the University of
-Colorado, Boulder, created the IPython interactive shell, which
-eventually blossomed into Project Jupyter\cite{Kluyver:2016aa}.
-% IPython hosted on SciPy:
-% https://mail.python.org/pipermail/scipy-user/2003-March/001455.html
-John Hunter, a postdoc
-%John got his PhD in 2001, then a post-doc with John Milton (dept of
-%Neurology, UChicago?) a couple of years; then research professor
-%under Kurt Hecox, whereafter at Tradelink from 2006
-at the University
-of Chicago, liked the plotting functionality
-available in MATLAB\cite{matlab}, but had problems accessing their laboratory's
-license, which was governed by a hardware dongle.  In response, he
-wrote a plotting library from scratch, and Matplotlib 0.1 was released
-April 2003\cite{matplotlib-rel}.
-
-% https://mail.python.org/pipermail/scipy-dev/2002-June/001007.html
-% https://str.llnl.gov/str/News896.html
-
-%2002—https://web.archive.org/web/20021020215937/http://www.scipy.org:80/site_content/scipy02/scipy02_agenda.htm
-%2003—https://web.archive.org/web/20031207151811/http://www.scipy.org:80/site_content/scipy03/
-%2004—Jim Hugunin
-%https://web.archive.org/web/20040920080225/http://www.scipy.org:80/wikis/scipy04
-%Travis Oliphant 2005—https://web.archive.org/web/20051103155042/http://www.scipy.org:80/wikis/scipy05/FrontPage
-%Guido—2006 https://scipy.github.io/old-wiki/pages/SciPy2006.html
-%Martelli—2008 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2004.html
-%Norvig—2009 https://scipy.github.io/old-wiki/pages/Developer_Zone/Conferences/SciPy2005.html
-
 As STScI continued to use Python for an increasingly large portion
@@
 Travis Oliphant combined the best elements of Numeric and NumArray,
-thereby solving the dilemma and paving the way for Python to become
-a very significant player in the Data Science movement. NumPy
-1.0 was released in October 2006\cite{numpy-1.0-tag}, with about 30
-authors recognized for major contributions in its release notes.
-
-% Future directions for SciPy after meeting at Berkeley
-% https://mail.python.org/pipermail/scipy-user/2005-March/004164.html
-
-%In April 2006, Paul
-%Dubois\footnote{https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html} wrote
-%to the NumPy mailing list:
-%\begin{quote}
-%IEEE's magazine, Computing in Science and Engineering (CiSE), has asked me
-%to put together a theme issue on the use of Python in Science and
-%Engineering. I will write an overview to be accompanied by 3-5 articles [. . .]
-%that show a diverse set of applications or
-%tools, to give our readers a sense of whether or not Pyth%on might be useful
-%in their own work.
-%\end{quote}
-In a May/June 2007 special issue of IEEE Computing in Science and
-Engineering, Paul Dubois wrote\cite{dubois2007guest}:
-\begin{quote}
-LLNL now has many Python-based efforts built from scratch or wrapped around
-legacy codes \textelp{} hundreds of thousands
-of lines of C++, Python, and Fortran~95, all working together just as we hoped,
-doing compute-intensive calculations on massively parallel computers.
-\end{quote}
-While it is now more commonly accepted, Dubois took the time to explain to
-the 2007 reader that ``interpreted doesn't mean slow or only interactive.''
-He also shared his own interest in Python for ``computational steering''
-where ``Python serves as the input language to a scientific application, and the actual
-computations are performed both in Python itself and in compiled extensions.''
-%https://www.computer.org/csdl/mags/cs/2007/03/c3007.html
-Using Python for computational steering was one of the earliest uses of Python
-in large scientific processing pipelines.
-In addition to Dubois' overview there were also articles introducing
-Python and SciPy \cite{oliphant2007python}, IPython \cite{perez2007ipython},
-and  Matplotlib \cite{hunter2007matplotlib}.
-There was also a diverse set of
-research applications including
-systems biology\cite{myers2007python},
-astronomy\cite{greenfield2007reaching},
-robotics\cite{krauss2007python},
-nanophotonics\cite{bienstman2007python},
-partial differential equations\cite{mardal2007using},
-neuroimaging\cite{millman2007analysis},
-geographic information systems\cite{shi2007python}, and
-education\cite{backer2007computational, myers2007pythona}.
+thereby solving the dilemma. NumPy 1.0 was released in October
+2006\cite{numpy-1.0-tag}, paving the way for the reunified scientific
+Python community to mature.
 

### 2019-09-08 scipy-1.0/paper.tex
Removed paragraph before SciPy Begins, as it is mostly covered by SciPy Begins

--- 2019-09-08 scipy-1.0/paper.tex
@@
 community that develops code for the platform.
-
-The availability of a standard numerical array data structure in
-Python led to a sudden, rapid growth in the number of scientific
-packages available for solving common numeric problems.
-A number of these packages were written by graduate students and
-postdoctoral researchers to solve the very practical research problems
-that they faced on a daily basis.  While they had access to and were
-familiar with specialized (often commercial) systems, many found
-it easier to implement the domain-specific functionality they needed
-in a general purpose programming language.  Given Python's innate
-ability to function as a systems language, controlling specialized or
-custom-built hardware was also possible.
 

### 2019-09-08 scipy-1.0/paper.tex
Removed paragraph about founding of NumFocus

@@
 community that develops code for the platform.
+
+The availability of a standard numerical array data structure in
+Python led to a sudden, rapid growth in the number of scientific
+packages available for solving common numeric problems.
+A number of these packages were written by graduate students and
+postdoctoral researchers to solve the very practical research problems
+that they faced on a daily basis.  While they had access to and were
+familiar with specialized (often commercial) systems, many found
+it easier to implement the domain-specific functionality they needed
+in a general purpose programming language.  Given Python's innate
+ability to function as a systems language, controlling specialized or
+custom-built hardware was also possible.
 
@@
 Neuroscience\cite{python-FIN}.
-
-In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
-and Fernando Pérez founded NumFOCUS\cite{numfocus},
- a 501(c)3 public charity with a mission
-``to promote sustainable high-level programming languages, open code development,
-and reproducible scientific research.''
-While NumFOCUS is language agnostic, many of the early sponsored projects
-came from the scientific Python stack.
-Today it has 50 affiliated and sponsored projects including NumPy, SciPy, IPython, and
-Matplotlib.
-Among its other projects, NumFOCUS organizes a global network of
-community driven educational programs, meetups, and conferences
-called PyData.
 

### 2019-09-08 scipy-1.0/paper.tex

@@
 
+In 2012, Perry Greenfield, John Hunter, Jarrod Millman, Travis Oliphant,
+and Fernando Pérez founded NumFOCUS\cite{numfocus},
+ a 501(c)3 public charity with a mission
+``to promote sustainable high-level programming languages, open code development,
+and reproducible scientific research.''
+While NumFOCUS is language agnostic, many of the early sponsored projects
+came from the scientific Python stack.
+Today it has 50 affiliated and sponsored projects including NumPy, SciPy, IPython, and
+Matplotlib.
+Among its other projects, NumFOCUS organizes a global network of
+community driven educational programs, meetups, and conferences
+called PyData.
+
 \subsection*{SciPy Today}

### 2019-09-10 scipy-1.0/paper.tex

@@
 libraries in order to provide Python convenience while benefiting from their performance:
-FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms,
-ODEPACK\cite{hindmarsh1983odepack}, % for solving ordinary differential equation initial value problems,
-QUADPACK\cite{piessens1983quadpack}, % for numerical integration of one-dimensional functions,
-FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
-ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression,
-MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations,
-ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
-ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and
-CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.
+QUADPACK\cite{piessens1983quadpack} and ODEPACK\cite{hindmarsh1983odepack} for numerical integration
+and solution of initial value problems;
+FITPACK\cite{Dierckx:1993:CSF:151103}, ODRPACK\cite{ODRPACK_Boggs}, and MINPACK\cite{osti_6997568} for
+curve-fitting and least-squares minimization;
+FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms;
+ARPACK\cite{leh:sor:yan96} for solving eigenvalue problems;
+ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for computing Bessel Functions; and
+CDFLIB\cite{CDFLIB_site} for evaluating cumulative density functions.
 

### 2019-09-10 scipy-1.0/paper.tex
MAINT: reviewer comment re: GPUs, etc.
* both Nature Methods reviewers requested clarification
regarding SciPy's stance on staying relevant with GPUs
and distributed computing

* briefly mention that while these were out of scope
at the 1.0 release point, we now have an explicit
example of experimental GPU backend support in scipy.fft,
to be detailed in a future report

@@
 range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) are explicitly out of scope.
+for graphics processing units (GPUs) were explicitly out of scope at the 1.0
+release point, but this has been revised in our roadmap at the time of writing
+(see below).
 
@@
 the latest JIT compilation approaches (i.e.,
-Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall
+Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but have traditionally fallen
 outside the remit of the main SciPy library. That said, we have recently
 increased our efforts to support compatibility with some of these options, and
-having our full test suite pass with the PyPy JIT
-compiler\cite{Bolz:2009:TMP:1565824.1565827} is now a requirement in our
-development workflow.
+while the exact approach we will adopt across the entire library remains unclear,
+we do now have a concrete implementation of a submodule that allows for the experimental
+use of multiple backends such as GPU-tractable data structures in the new
+\texttt{scipy.fft}. This will be described in detail in a future report.
 

### 2019-09-11 scipy-1.0/paper.tex
Add paragraph re: funding (w/ Stefan)

--- 2019-09-10 scipy-1.0/paper.tex
@@
 
+\subsection*{Funding}
+
+%100,000 annual salary, 112 person-years, 414,473 lines of code -> 11,192,622 (https://www.openhub.net/p/scipy/estimated\_cost, which uses the Basic COCOMO model).
+
+The development cost of SciPy is estimated in excess of 10M dollars by
+OpenHub.  Yet, the project is largely unfunded, having been
+developed predominantly by graduate students, faculty, and members of industry
+in their free time. Small amounts of funding have been applied with success:
+some meetings were sponsored by universities and industry, Google's Summer of Code
+program supported infrastructure and algorithmic work, and teaching grant funds were
+used early on to develop documentation.  However, these small amounts of
+sporadic funding are not commensurate with the enormous stack of important scientific
+software, like those that are used to plan national space missions, that rely
+on SciPy.  In the traditional academic funding model, it is easy to support
+commercial software, but hard to provide funds for community driven projects,
+and as more science depends on open source software national funding agencies
+urgently need to pay attention to this challenge. And, unlike for commercial
+software where a license seat encompasses all development (and other) costs, agencies will
+have to tolerate the more diverse spending required to do planning and maintain
+infrastructure if these open source projects are to remain a healthy
+underpinning infrastructure for international scientific endeavors.
+
+%largely unfunded
+%- graduate students
+%- faculty
+%- industry program
+%
+%-- free time
+%
+%Small funds:
+%- meetings (eg Berkeley)
+%- GSOC (test infrastructure, recently FFT work, been doing it since, what 2011/2012/2013?)
+%- documentation
+%
+%Problems:
+%- funding agencies (NIH, NSF, etc.)
+%- foundation have few mechanism for support, while ironically able to fund e.g. Matlab licenses
+%- difficulty of hiring full time dev for community projects
+%
+%Rec:
+%- Funders need to address
+%- Funding needs to be carefully used to support the community; not just technical work but also process, roadmap, discussing and deciding for new features, etc.
+%
+%- Many stakeholders, supporting space missions, entire fields, ...
+
 \subsection*{Downstream projects}

### 2019-09-11 scipy-1.0/paper.tex
MAINT: review comments in PR 245
* restore PyPy compatibility statement with 1.0
release point qualifier

* move discussion of GPU support / scipy.fft to final
section of manuscript instead of language section

* describe scipy.fft as a subpackage instead of submodule

--- 2019-09-10 scipy-1.0/paper.tex
@@
 increased our efforts to support compatibility with some of these options, and
-while the exact approach we will adopt across the entire library remains unclear,
-we do now have a concrete implementation of a submodule that allows for the experimental
-use of multiple backends such as GPU-tractable data structures in the new
-\texttt{scipy.fft}. This will be described in detail in a future report.
+our full test suite passed with the PyPy JIT compiler\cite{Bolz:2009:TMP:1565824.1565827}
+at the 1.0 release point.
 
@@
 
+A final important challenge to address is the accommodation of GPU and distributed
+computing without disrupting our conventional and heavily-used algorithm / API
+infrastructure. While the exact approach we will adopt across the entire library
+to leverage these emerging technologies remains unclear, and was not a priority
+at the 1.0 release point, we do now have a concrete implementation of a subpackage
+that allows for the experimental use of multiple backends, such as GPU-tractable
+data structures, in the new \texttt{scipy.fft}. This will be described in detail
+in a future report.
+
 \section*{Data Availability}

### 2019-09-11 scipy-1.0/paper.tex
Include GH feedback

--- 2019-09-11 scipy-1.0/paper.tex
@@
 The development cost of SciPy is estimated in excess of 10M dollars by
-OpenHub.  Yet, the project is largely unfunded, having been
+Open Hub\cite{SciPyValue}.  Yet, the project is largely unfunded, having been
 developed predominantly by graduate students, faculty, and members of industry
@@
 commercial software, but hard to provide funds for community driven projects,
-and as more science depends on open source software national funding agencies
+and as more science depends on open source software, national funding agencies
 urgently need to pay attention to this challenge. And, unlike for commercial

### 2019-09-11 scipy-1.0/paper.tex
Add industry sentence and tighten language

@@
 and as more science depends on open source software, national funding agencies
-urgently need to pay attention to this challenge. And, unlike for commercial
-software where a license seat encompasses all development (and other) costs, agencies will
-have to tolerate the more diverse spending required to do planning and maintain
-infrastructure if these open source projects are to remain a healthy
-underpinning infrastructure for international scientific endeavors.
+urgently need to pay attention to this challenge.
+Moreover, the scientific Python stack plays an increasingly important role
+in industry; however, industry provides little to no funding of community-developed
+software, which they depend on.
+Industry and funding agencies will have to tolerate the more diverse spending
+required to do planning and maintain infrastructure if these open source
+projects are to remain a healthy underpinning infrastructure for international
+scientific and industrially endeavors.
 

### 2019-09-13 scipy-1.0/paper.tex

--- 2019-09-11 scipy-1.0/paper.tex
@@
 release point, but this has been revised in our roadmap at the time of writing
-(see below).
+(see Discussion).
 

### 2019-09-13 scipy-1.0/paper.tex
Reword to soften the tone

--- 2019-09-11 scipy-1.0/paper.tex
@@
 software, like those that are used to plan national space missions, that rely
-on SciPy.  In the traditional academic funding model, it is easy to support
-commercial software, but hard to provide funds for community driven projects,
-and as more science depends on open source software, national funding agencies
-urgently need to pay attention to this challenge. And, unlike for commercial
-software where a license seat encompasses all development (and other) costs, agencies will
-have to tolerate the more diverse spending required to do planning and maintain
-infrastructure if these open source projects are to remain a healthy
-underpinning infrastructure for international scientific endeavors.
+on SciPy.
+With traditional grant mechanisms, it is easier to buy software licenses than to support the diverse development effort needed to develop open source software—a challenge that we hope to see national funding agencies address.
+The scientific Python stack is also playing an increasingly important role in industry, and while there are some encouraging signals such as TideFlow, that sector still spends comparatively little on community-developed software.
+For open source projects to remain a healthy underpinning infrastructure for international scientific and industrial endeavors, it needs more diverse spending to support planning, development, management, and infrastructure.
 

### 2019-09-14 scipy-1.0/paper.tex
Avoid repeating `infrastructure`

@@
 The scientific Python stack is also playing an increasingly important role in industry, and while there are some encouraging signals such as TideFlow, that sector still spends comparatively little on community-developed software.
-For open source projects to remain a healthy underpinning infrastructure for international scientific and industrial endeavors, it needs more diverse spending to support planning, development, management, and infrastructure.
+For open source projects to remain a healthy underpinning of international scientific and industrial endeavors, it needs more diverse spending to support planning, development, management, and infrastructure.
 

### 2019-09-16 scipy-1.0/rebuttal.tex
Added rebuttal letter; drafted response about heterogeneous computing

% based on template:
% https://tex.stackexchange.com/questions/15532/looking-for-cover-letter-template

\documentclass[10pt,stdletter,dateno]{newlfm}

% to adjust vertical spacing properties of
% newlfm class see: https://tex.stackexchange.com/a/65533/43006
%\unprtop{0mm}
\topmarginsize{0mm}
\topmarginskip{0in}
\headermarginsize{0in}
\headermarginskip{0in}
\MinHead{0in}
%\dateskipafter{0pt}
%\dateskipbefore{0pt}
\sigskipbefore{40pt}
%\sigskipafter{0pt}
\closeskipbefore{1pt}
%\closeskipafter{0pt}
%\addrfromskipbefore{0pt}
\addrfromskipafter{0in}
\addrtoskipbefore{0in}
%\addrtoskipafter{0pt}
\newlfmP{Headlinewd=0pt,Footlinewd=0pt}
\newlfmP{sigsize=10pt}
\bottommarginskip{10pt}
\MinFoot{0pt}
\leftmarginsize{0.95in}
\rightmarginsize{0.95in}

\usepackage{kpfonts}
\usepackage{hyperref}
\usepackage{url}

\hypersetup{%
  linkbordercolor=blue,% hyperlink borders will be blue
  pdfborderstyle={/S/U/W 1}% border style will be underline of width 1pt
}
\widowpenalty=1000
\clubpenalty=1000

\namefrom{Ralf Gommers,
          Matt Haberland,
          and Tyler Reddy}
\addrfrom{%
    \today\\[10pt]
    Ralf Gommers\\
    Quansight Labs\\
    The Netherlands\\
    \texttt{ralf.gommers@gmail.com}\\
    %phone number\\
    \\
    Matt Haberland\\
    BioResource and Agricultural Engineering\\
    California Polytechnic State University\\
    San Luis Obispo, CA 93407, USA\\
    \texttt{mhaberla@calpoly.edu}\\
    %phone number\\
    \\
    Tyler Reddy\\
    CCS-7 Applied Computer Science Group\\
    Los Alamos National Laboratory\\
    Los Alamos, NM 87545, USA\\
    \texttt{treddy@lanl.gov}\\
    %phone number\\
}

\addrto{%
Dr. Rita Strack\\
Nature Methods Editorial Office\\
One New York Plaza Suite 4500\\
New York, NY 10004, USA}

\greetto{Dear Dr. Strack,}
\closeline{Sincerely,}
\begin{document}
\begin{newlfm}

We appreciate the careful review and helpful suggestions from you and the reviewers. Each comment is quoted and addressed below; we also link the relevant GitHub issues in which we discussed and implemented each suggestion. As requested, the corresponding changes are underlined in the revised manuscript.

\begin{quote}
I would like a longer discussion about the future of SciPy particularly SciPy’s lack of the two features that are starting to become understood as essential to scientific computing: built-in automatic reverse-mode differentiation and heterogenous computation (e.g. CPU, DSP, GPU, etc.). Will SciPy adapt to support these increasingly vital features? Is it even possible without a substantial rewrite?
\end{quote}

The SciPy maintainers have discussed the inclusion of automatic differentiation capabilities several times in recent years, but the conclusion has been that this need is currently fulfilled by existing libraries such as JAX (\url{https://github.com/google/jax}), a Google research project. JAX can automatically perform AD on generic Python and NumPy code with minimal modifications, so it cooperates well with SciPy. Users who need AD capabilities with a more permissive license can use Autograd (\url{https://github.com/HIPS/autograd}), a core component of JAX, and ad (\url{https://pythonhosted.org/ad/}). However, the latter packages are no longer actively maintained, so we will continue to evaluate the demand to add their capabilities -- and our ability to maintain them -- as part of SciPy.

At the SciPy 1.0 release point, support for heterogeneous computation was explicitly out of scope, but we have since added to the roadmap our intention to support distributed and GPU arrays. NumPy is changing its API such that many parts of SciPy will be able to accept any array that implements the \texttt{ndarray} interface; support for parts that don't immediately work well with distributed/GPU arrays will be added over time. For instance, SciPy 1.4 will feature a backend system in \texttt{scipy.fft} that allows calling SciPy functions on GPU and even multi-GPU systems via CuPy and Dask. This will be extended to most other subpackages in subsequent releases, with \texttt{linalg} and \texttt{special} high on the list. There are not immediate plans to tailor algorithms for distributed or GPU arrays, but in the meantime SciPy provides a reference implementation of algorithms and an API on which downstream libraries can model their code.

To address this comment, we have made changes to the Project Scope and Discussion sections as discussed in GitHub \#245 (\url{https://github.com/scipy/scipy-articles/pull/245}).

\end{newlfm}
\end{document}

### 2019-09-16 scipy-1.0/paper.tex
Addressed low test coverage of compiled code

--- 2019-09-13 scipy-1.0/paper.tex
@@
 Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
-87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
-45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
+87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov}.
+Coverage of compiled (C, C++, and Fortran) code was only 45\% according to
+\texttt{gcov}\cite{gcov}, but the compiled codebase is much more robust than
+this figure would suggest as the figure does not correct for the inclusion of
+reputable vendored code, the original library of which is well-tested;
+generated code, for which full coverage is impractical; and
+deprecated code, which does not require unit tests.
 Documentation for the code is automatically built and published by
 the CircleCI service to facilitate evaluation of documentation changes /
-integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
-version of the Python language.
+integrity. % Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827},
+% a just-in-time compiled version of the Python language.
 

### 2019-09-17 scipy-1.0/paper.tex
Shorten

--- 2019-09-14 scipy-1.0/paper.tex
@@
 program supported infrastructure and algorithmic work, and teaching grant funds were
-used early on to develop documentation.  However, these small amounts of
-sporadic funding are not commensurate with the enormous stack of important scientific
-software, like those that are used to plan national space missions, that rely
-on SciPy.
-With traditional grant mechanisms, it is easier to buy software licenses than to support the diverse development effort needed to develop open source software—a challenge that we hope to see national funding agencies address.
-The scientific Python stack is also playing an increasingly important role in industry, and while there are some encouraging signals such as TideFlow, that sector still spends comparatively little on community-developed software.
-For open source projects to remain a healthy underpinning of international scientific and industrial endeavors, it needs more diverse spending to support planning, development, management, and infrastructure.
+used early on to develop documentation.  However, funding from national agencies,
+foundations, and industry has not been commensurate with the enormous stack of important
+software that rely on SciPy.
+More diverse spending to support planning, development, management, and infrastructure would
+help SciPy remain a healthy underpinning of international scientific
+and industrial endeavors.
 

### 2019-09-17 scipy-1.0/paper.tex

@@
 range of data types and process architectures. Distributed computing and support
-for graphics processing units (GPUs) are explicitly out of scope.
+for graphics processing units (GPUs) were explicitly out of scope at the 1.0
+release point, but this has been revised in our roadmap at the time of writing
+(see Discussion).
 
@@
 the latest JIT compilation approaches (i.e.,
-Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but currently fall
+Numba\cite{Lam:2015:NLP:2833157.2833162}) are very powerful, but have traditionally fallen
 outside the remit of the main SciPy library. That said, we have recently
 increased our efforts to support compatibility with some of these options, and
-having our full test suite pass with the PyPy JIT
-compiler\cite{Bolz:2009:TMP:1565824.1565827} is now a requirement in our
-development workflow.
+our full test suite passed with the PyPy JIT compiler\cite{Bolz:2009:TMP:1565824.1565827}
+at the 1.0 release point.
 

### 2019-09-18 scipy-1.0/paper.tex
MAINT: move CI definition
* expand the Figure 2 caption to direct the reader to
explanations for CI and other technical terms to address
reviewer comment that CI was not explained at this point
of manuscript

@@
 \caption{Major milestones from SciPy's initial release in 2001 to
-the release of SciPy 1.0 in 2017.}
+the release of SciPy 1.0 in 2017. Note that SciKits and GitHub 
+have been introduced in the ``Background'' section;  more 
+information about Cython and SciPy subpackages (e.g. 
+\texttt{scipy.sparse}) is available in ``Architecture 
+and implementation choices'', BLAS/LAPACK support is 
+detailed in ``Key technical improvements'', and continuous
+integration is discussed in ``Test and benchmark suite''.}
 \label{fig:timeline}

### 2019-09-18 scipy-1.0/rebuttal.tex
Completed draft rebuttal letter, incorporating @tylerjereddy suggestions

@@
 
+\begin{quote}
+Page 6.
+Distributed and GPU is considered out of scope but is conversely is believed to be required in the future for data and scientific analysis.
+A discussion of alternatives and how you interact with other ecosystems should be considered (e.g., RAPIDS, Dask, etc).
+How does SciPy progress with the rapidly evolving hardware?
+\end{quote}
+
 The SciPy maintainers have discussed the inclusion of automatic differentiation capabilities several times in recent years, but the conclusion has been that this need is currently fulfilled by existing libraries such as JAX (\url{https://github.com/google/jax}), a Google research project.
@@
 
+\begin{quote}
+Scientific software is depressingly underfunded.
+I imagine the economic cost of SciPy is in the tens (if not hundreds) of millions of dollars yet, as you mention, the actual direct financial contributions were a magnitude (or two) less.
+How did the SciPy community address this disparity?
+Are there lessons to be learned for the larger scientific community?
+\end{quote}
+
+We added a ``Funding'' subsection to the ``Project organization and community'' section on page 13 to address this (GitHub \#246, \url{https://github.com/scipy/scipy-articles/pull/245}).
+
+\begin{quote}
+A clear distinction between CPython and Python is stressed, which might not be relevant to many audiences as alternatives like PyPI are likely not well understood.
+This is also an example of where the paper may delve too much into technical details.
+\end{quote}
+
+We've removed the sentence referring to CPython on page 2 of the manuscript--we agree that this is indeed too technical to be of interest for the likely audience (GitHub \#238, \url{https://github.com/scipy/scipy-articles/pull/238}).
+
+\begin{quote}
+Page 3. Figure 1. There should be a date on when the announcement was initially released and potentially medium used.
+\end{quote}
+
+The exact date of the announcement and the name of the mailing list that was used have been added to the caption to Figure 1 on page 3 (GitHub \#237, \url{https://github.com/scipy/scipy-articles/pull/237}).
+
+\begin{quote}
+Page 5. A very brief mention of what GitHub is should be stated; bench chemist may be unaware of this platform.
+\end{quote}
+
+We've introduced GitHub in a new sentence added to the ``SciPy today'' section, page 5 (GitHub \#233, \url{https://github.com/scipy/scipy-articles/pull/233}).
+
+\begin{quote}
+Page 5.
+Figure 2.
+The term ``SciKit'' is only briefly mentioned before this figure and is never fleshed out.
+With the current phrasing, it is unclear if SciKits are competitors or extensions of SciPy, their relationship should be well defined.
+\end{quote}
+
+The nature of SciKits is now discussed in more detail in a modified sentence on page 4, prior to their mention in Figure 2 (GitHub \#235, \url{https://github.com/scipy/scipy-articles/pull/235}).
+
+\begin{quote}
+Page 5. Figure 2. The term ``continuous integration'' needs to be defined before this image.
+\end{quote}
+
+Continuous integration is now defined in the ``Test and benchmark suite'' section, page 11 (GitHub \#239, \url{https://github.com/scipy/scipy-articles/pull/239}).
+As the definition did not fit in any context before the figure, and because readers might also be unfamiliar with Cython, BLAS/LAPACK, and other terms mentioned in the figure, the figure caption now refers the interested reader to the corresponding sections (GitHub \#253, \url{https://github.com/scipy/scipy-articles/pull/253}).
+
+\begin{quote}
+Page 7.
+Many Fortran packages like ODEPACK and ARPACK are discussed without context.
+Readers are unlikely to be familiar with these libraries (even BLAS is questionable).
+While enumerating the capabilities of each package is clearly beyond scope simply categorizing them in some fashion would be worthwhile.
+\end{quote}
+
+The ``Language choices'' section (page 7) has been revised: the external Fortran packages have been grouped into six distinct categories based on what the libraries provide, and the purpose of each C library used by SciPy is briefly mentioned.
+BLAS and LAPACK are now defined on page 8 (GitHub \#240, \url{https://github.com/scipy/scipy-articles/pull/240}).
+
+\begin{quote}
+Page 11.
+From the text, it is unclear why testing and code coverage are useful metrics.
+Some mention of correctness is likely required and could be a good time to mention that these metrics give both contributors courage that they are not breaking the package and maintainers some faith that a contribution does not harm the package.
+\end{quote}
+
+Page 11 of the revised manuscript now contains a new paragraph outlining the value and purpose of test-driven development, including the suggested mention of increasing confidence in code changes (GitHub \#234, \url{https://github.com/scipy/scipy-articles/pull/234}).
+
+\begin{quote}
+Page 11.
+The compiled coverage of 45\% seems surprisingly low and perhaps should be remarked on.
+The maintainers also appear to have allowed remarkable amounts of untested compiled code in recent years.
+One could make an argument that the correctness of SciPy is much more important than the feature set as it is currently thought of as a dependable resource.
+\end{quote}
+
+The ``Test suite'' subsection on page 11 now explains why the compiled code is more robust than the raw 45 \% coverage figure might suggest (GitHub \#250, \url{https://github.com/scipy/scipy-articles/pull/250}).
+
+\begin{quote}
+Links to SciPy cookbooks or tutorials would be useful to highlight to give readers immediate examples of how SciPy would be helpful.
+These tutorials are only linked in your Discussion section and maybe missed.
+\end{quote}
+
+We've added a mention of the SciPy Tutorial and API Reference much earlier in a new sentence of the ``Package organization'' subsection on page 6 (GitHub \#241, \url{https://github.com/scipy/scipy-articles/pull/241}).
+
+\begin{quote}
+Overall the paper is well written, but some additional attention to presenting this towards a general audience should be exercised.
+In particular, the “Background” section should be greatly minimized and moved to another part of the paper (or removed entirely) so that readers can be first presented with capabilities.
+\end{quote}
+
+Per the editor's recommendation, we have substantially streamlined the Background section from $\approx$2000 words to $\approx$1500 words (GitHub \#242, \url{https://github.com/scipy/scipy-articles/pull/242}; \#243, \url{https://github.com/scipy/scipy-articles/pull/243}; and \#244, \url{https://github.com/scipy/scipy-articles/pull/244}).
+
+Thank you again for your helpful comments and for your consideration of the revised manuscript.
+
 \end{newlfm}

### 2019-09-19 scipy-1.0/rebuttal.tex
Mention order of comments in rebuttal

@@
 We appreciate the careful review and helpful suggestions from you and the reviewers.
-Each comment is quoted and addressed below; we also link the relevant GitHub issues in which we discussed and implemented each suggestion.
+Each comment is quoted and addressed below in the order of the original review, with the exception that the requests from both reviewers to address GPU/distributed computing have been combined.
+We also link the relevant GitHub issues in which we discussed and implemented each suggestion.
 As requested, the corresponding changes are underlined in the revised manuscript.

### 2019-09-19 scipy-1.0/consortium.tex
Change location of MPI-SWS.
Fixed affiliation
updated affiliation

--- 2019-07-21 scipy-1.0/consortium.tex
@@
 {\bfseries
-Aditya Vijaykumar$^{31,32}$, 
+Aditya Vijaykumar$^{32,33}$, 
 Alessandro Pietro Bardelli$^{10}$, 
 Alex Rothberg$^{10}$, 
-Andreas Hilboll$^{33}$, 
-Andreas Kloeckner$^{34}$, 
+Andreas Hilboll$^{34}$, 
+Andreas Kloeckner$^{35}$, 
 Anthony Scopatz$^{2}$, 
-Antony Lee$^{35}$, 
-Ariel Rokem$^{36}$, 
+Antony Lee$^{36}$, 
+Ariel Rokem$^{37}$, 
 C. Nathan Woods$^{9}$, 
-Chad Fulton$^{37}$, 
-Charles Masson$^{38}$, 
-Christian Brueffer$^{39}$, 
+Chad Fulton$^{38}$, 
+Charles Masson$^{39}$, 
 Christian H\"aggstr\"om$^{40}$, 
@@
 Martin Thoma$^{75}$, 
-Matthias K\"ummerer$^{76}$, 
-Maximilian Bolingbroke$^{77}$, 
-Michael Tartre$^{78}$, 
-Mikhail Pak$^{79}$, 
+Matthew Newville$^{76}$, 
+Matthias K\"ummerer$^{77}$, 
+Maximilian Bolingbroke$^{78}$, 
+Michael Tartre$^{79}$, 
+Mikhail Pak$^{80}$, 
 Nathaniel J. Smith$^{10}$, 
 Nikolai Nowaczyk$^{10}$, 
-Nikolay Shebanov$^{80}$, 
-Oleksandr Pavlyk$^{81}$, 
-Per A. Brodtkorb$^{82}$, 
+Nikolay Shebanov$^{81}$, 
+Oleksandr Pavlyk$^{82}$, 
+Per A. Brodtkorb$^{83}$, 
 Perry Lee$^{10}$, 
-Robert T. McGibbon$^{83}$, 
-Roman Feldbauer$^{84}$, 
-Sam Lewis$^{85}$, 
-Sam Tygier$^{86}$, 
-Scott Sievert$^{87}$, 
-Sebastiano Vigna$^{88}$, 
+Robert T. McGibbon$^{84}$, 
+Roman Feldbauer$^{85}$, 
+Sam Lewis$^{86}$, 
+Sam Tygier$^{87}$, 
+Scott Sievert$^{88}$, 
+Sebastiano Vigna$^{89}$, 
 Stefan Peterson$^{10}$, 
-Surhud More$^{89,90}$, 
-Tadeusz Pudlik$^{91}$, 
-Takuya Oshima$^{92}$, 
-Thomas J. Pingel$^{93}$, 
-Thomas P. Robitaille$^{94}$, 
+Surhud More$^{90,91}$, 
+Tadeusz Pudlik$^{92}$, 
+Takuya Oshima$^{93}$, 
+Thomas J. Pingel$^{94}$, 
+Thomas P. Robitaille$^{95}$, 
 Thomas Spura$^{10}$, 
-Thouis R. Jones$^{95}$, 
+Thouis R. Jones$^{96}$, 
 Tim Cera$^{10}$, 
 Tim Leslie$^{10}$, 
-Tiziano Zito$^{96}$, 
-Tom Krauss$^{97}$, 
-Utkarsh Upadhyay$^{98}$, 
-Yaroslav O. Halchenko$^{99}$, 
-Yoshiki V\'azquez-Baeza$^{100}$
+Tiziano Zito$^{97}$, 
+Tom Krauss$^{98}$, 
+Utkarsh Upadhyay$^{99}$, 
+Yaroslav O. Halchenko$^{100}$, 
+Yoshiki V\'azquez-Baeza$^{101}$
 }
@@
 \hfill \break
-$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, 
-$^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, 
-$^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, 
-$^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, 
-$^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, 
-$^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, 
-$^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., 
-$^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, 
-$^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, 
+$^{32}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, 
+$^{33}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, 
+$^{34}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, 
+$^{35}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, 
+$^{36}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, 
+$^{37}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, 
+$^{38}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., 
+$^{39}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, 
 $^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, 
@@
 $^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, 
-$^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, 
+$^{47}$Aix Marseille Univ, CNRS, Centrale Marseille, LMA, Marseille, France, 
 $^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, 
@@
 $^{75}$Independent Researcher, Munich, Germany, 
-$^{76}$University of T\"ubingen, T\"ubingen, Germany, 
-$^{77}$Independent Researcher, United Kingdom, 
-$^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, 
-$^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, 
-$^{80}$Independent Researcher, Berlin, Germany, 
-$^{81}$Intel Corp., Austin, TX 78746, USA, 
-$^{82}$Independent Researcher, Horten, Norway, 
-$^{83}$D. E. Shaw Research, New York, NY 10036, USA, 
-$^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, 
-$^{85}$Independent Researcher, Melbourne, Australia, 
-$^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, 
-$^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, 
-$^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, 
-$^{89}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, 
-$^{90}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, 
-$^{91}$Waymo LLC, Mountain View, CA 94043, USA, 
-$^{92}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, 
-$^{93}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, 
-$^{94}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, 
-$^{95}$Broad Institute, Cambridge, MA 02142, USA, 
-$^{96}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, 
-$^{97}$Epiq Solutions, Schaumburg, IL 60173, USA, 
-$^{98}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, 
-$^{99}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, 
-$^{100}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA
+$^{76}$Center for Advanced Radiation Sources, The University of Chicago, Chicago, IL 60637, USA, 
+$^{77}$University of T\"ubingen, T\"ubingen, Germany, 
+$^{78}$Independent Researcher, United Kingdom, 
+$^{79}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, 
+$^{80}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, 
+$^{81}$Independent Researcher, Berlin, Germany, 
+$^{82}$Intel Corp., Austin, TX 78746, USA, 
+$^{83}$Independent Researcher, Horten, Norway, 
+$^{84}$D. E. Shaw Research, New York, NY 10036, USA, 
+$^{85}$Division of Computational Systems Biology, Department of Microbiology and Ecosystem Science, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, 
+$^{86}$Independent Researcher, Melbourne, Australia, 
+$^{87}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, 
+$^{88}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, 
+$^{89}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, 
+$^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, 
+$^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, 
+$^{92}$Waymo LLC, Mountain View, CA 94043, USA, 
+$^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, 
+$^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, 
+$^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, 
+$^{96}$Broad Institute, Cambridge, MA 02142, USA, 
+$^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, 
+$^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, 
+$^{99}$Max Planck Institute for Software Systems, Kaiserslautern, Germany, 
+$^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, 
+$^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA

### 2019-09-19 scipy-1.0/core-dev.tex
Correct my affiliation (I'm based in Montreal, not Cambridge)
Also corrected Google Inc. to Google LLC
Corrected author list
Spell out author first name.

@@
 \author[1]{Pauli Virtanen}
-\author[2]{Ralf Gommers}
+\author[2,*]{Ralf Gommers}
 \author[3,4,5,6,2]{Travis E. Oliphant}
-\author[7,8]{Matt Haberland}
-\author[9]{Tyler Reddy}
+\author[7,8,*]{Matt Haberland}
+\author[9,*]{Tyler Reddy}
 \author[10]{David Cournapeau}
@@
 \author[10]{E. A. Quintero}
-\author[10]{Charles R Harris}
-\author[27]{Anne M. Archibald}
-\author[28]{Ant\^{o}nio H. Ribeiro}
-\author[29]{Fabian Pedregosa}
+\author[10,27]{Charles R Harris}
+\author[28]{Anne M. Archibald}
+\author[29]{Ant\^{o}nio H. Ribeiro}
+\author[30]{Fabian Pedregosa}
+\author[31]{Paul {van Mulbregt}}
 \author[ ]{SciPy 1.0 Contributors}
@@
 \affil[26]{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin, Austin, Texas 78712, USA}
-\affil[27]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, The Netherlands}
-\affil[28]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
-\affil[29]{Google, Inc., 355 Main St., Cambridge, MA 02142, USA}
+\affil[27]{Space Dynamics Laboratory, 1695 North Research Park Way, North Logan, UT 84341, USA}
+\affil[28]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, The Netherlands}
+\affil[29]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
+\affil[30]{Google LLC, Montreal, QC H3B 2Y5, Canada}
+\affil[31]{Google LLC, 355 Main St., Cambridge, MA 02142, USA}
 \affil[*]{scipy.articles@gmail.com}

### 2019-09-19 scipy-1.0/rebuttal.tex

@@
 
-We added a ``Funding'' subsection to the ``Project organization and community'' section on page 13 to address this (GitHub \#246, \url{https://github.com/scipy/scipy-articles/pull/245}).
+We added a ``Funding'' subsection to the ``Project organization and community'' section on page 13 to address this (GitHub \#246, \url{https://github.com/scipy/scipy-articles/pull/246}).
 
